### 2019-10-31 paper/paper.tex
start document

% This document is part of the EPRVCalibration project
% Copyright 2019 the authors. All rights reserved.

\documentclass[12pt, letterpaper]{article}

% typesetting words
\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\acronym}[1]{{\small{#1}}}
\newcommand{\expres}{\project{\acronym{EXPRES}}}

% margins and page setup, etc
\addtolength{\textheight}{1.00in}
\addtolength{\topmargin}{-0.50in}
\sloppy\sloppypar\raggedbottom\frenchspacing

\begin{document}

\section*{Method}

The idea is that the wavelength solution is going to live in a
low-dimensional space, where the degrees of freedom are set by the
degrees of freedom of the spectrograph hardware.

First, some definitions.
The way I (Hogg) think about this is the following:
Given an exposure $n$, and order $m$, there is a relationship between
the two-dimensional $(x,y)$-position on the detector and the
wavelength $\lambda$
\begin{equation}
\lambda = f(x,y;m,\theta_{n})
\quad ,
\end{equation}
where $\theta_{n}$ is a big blob of parameters for this exposure.
This is not precisely how the \expres\ team thinks about the problem,
I (Hogg) think.
For now I am going to ignore that fact, but I think what I say here
could be adapted to any \expres\ conventions.
Right now in the \expres\ pipeline the $\theta_{n}$ comprises the
amplitudes of some 9th-ish-order polynomial in $x$ and $n$, possibly
ignoring $y$?.

If the space of all calibration possibilities is in fact
$K$-dimensional (where I am thinking of this as being a small integer,
like $K=3$ or $K=8$ or thereabouts), and if the calibration variations are so
small that we can linearize, then the function $f(x,y;m,\theta_{n})$ could
be replaced with a tiny model
\begin{equation}
\lambda = g_0(x,y;m) + \sum_{k=1}^K a_{nk}\,g_k(x,y;m)
\quad ,
\end{equation}
where
$g_0(x,y;m)$ is the fiducial or mean or standard calibration of the
spectrograph,
the $a_{nk}$ are scalar amplitudes,
and the $g_k(x,y;m)$ are basis functions expressing the ``directions''
in calibration space that the spectrograph can depart from the
fiducial calibration.
The challenge is to learn these basis functions from the data, and get
the $K$ amplitudes $a_{nk}$ for every exposure $n$.

\end{document}

### 2019-10-31 paper/paper.tex
adjust conventions
semicolons

@@
 \begin{equation}
-\lambda = f(x,y;m,\theta_{n})
+\lambda = f(x,y,m;\theta_{n})
 \quad ,
@@
 amplitudes of some 9th-ish-order polynomial in $x$ and $n$, possibly
-ignoring $y$?.
+ignoring $y$?
+If so, they have
+\begin{equation}
+\lambda = \sum_{i,j} c_{ij}\, x^i\,m^j
+\quad ,
+\end{equation}
+where the $c_{ij}$ are coefficients.
 
@@
 like $K=3$ or $K=8$ or thereabouts), and if the calibration variations are so
-small that we can linearize, then the function $f(x,y;m,\theta_{n})$ could
+small that we can linearize, then the function $f(x,y,m;\theta_{n})$ could
 be replaced with a tiny model
 \begin{equation}
-\lambda = g_0(x,y;m) + \sum_{k=1}^K a_{nk}\,g_k(x,y;m)
+\lambda = g_0(x,y,m) + \sum_{k=1}^K a_{nk}\,g_k(x,y,m)
 \quad ,
@@
 where
-$g_0(x,y;m)$ is the fiducial or mean or standard calibration of the
+$g_0(x,y,m)$ is the fiducial or mean or standard calibration of the
 spectrograph,
 the $a_{nk}$ are scalar amplitudes,
-and the $g_k(x,y;m)$ are basis functions expressing the ``directions''
+and the $g_k(x,y,m)$ are basis functions expressing the ``directions''
 in calibration space that the spectrograph can depart from the

### 2019-10-31 paper/paper.tex
moar stuff for Lily :

@@
 \begin{equation}
-\lambda = f(x,y,m;\theta_{n})
+\lambda(x,y,m,n) = f(x,y,m;\theta_{n})
 \quad ,
@@
 ignoring $y$?
-If so, they have
+If so, they (may?) have
 \begin{equation}
-\lambda = \sum_{i,j} c_{ij}\, x^i\,m^j
+\lambda(x,y,m,n) = \sum_{i=0}^9\sum_{j=0}^9 c_{nij}\, x^i\,m^j + \mathrm{noise}
 \quad ,
 \end{equation}
-where the $c_{ij}$ are coefficients.
+where the $c_{nij}$ are coefficients unique to exposure $n$, and
+found for exposure $n$ without regard to any other exposure $n'$.
+Further, I think they may be finding the coefficients $c_{nij}$ that
+minimize an objective $Q$ that is something like the L2-norm:
+\begin{equation}
+Q = ||\lambda(x,y,m,n) - \sum_{i=0}^9\sum_{j=0}^9 c_{nij}\, x^i\,m^j||_2^2
+\quad .
+\end{equation}
+
+Even in this simple context, where every calibration image $n$ is
+treated as its own unique flower, I think there are improvements to be
+made.
+For one, the sums should not be from 0 to 9, but instead
+\begin{equation}
+\sum_{i=0}^9\sum_{j=0}^{9-i}
+\quad ,
+\end{equation}
+because that is the definition of 9th order.
+If I am right about this, the fit could be taken to 13th order to obtain
+much more expressiveness with (nearly) the same number of fit coefficients.
+For another, the objective function could be made soft to permit
+catastrophic outliers without destroying the fit.
+I would recommend the iteratively reweighted least squares (IRLS).
+This would make the fitting more robust.
+For yet another, I have ideas about rescaling the products $x^i\,m^j$ to
+protect the fitting from near-singularities or bad conditioning of the
+linear-algebra operators.
+
+Of course the main thing I want to do is go hierarchical!
+In this case, the calibration of every image is informed by the
+calibration of every other image.
+The simplest form of hierarchical model is to use the calibration data
+themselves to develop a low-dimensional basis for expressing the
+calibration data.
 
@@
 \begin{equation}
-\lambda = g_0(x,y,m) + \sum_{k=1}^K a_{nk}\,g_k(x,y,m)
+\lambda(x,y,m,n) = g_0(x,y,m) + \sum_{k=1}^K a_{nk}\,g_k(x,y,m)
 \quad ,
@@
 
+I know how to do this, but the margins here are too small to contain my
+beautiful ideas...
+
 \end{document}

### 2019-11-02 paper/paper.tex
title and abstract

--- 2019-10-31 paper/paper.tex
@@
 \newcommand{\expres}{\project{\acronym{EXPRES}}}
+\newcommand{\name}{\project{NameOfThis}}
 
@@
 
-\section*{Method}
+\section*{\raggedright%
+\name:
+A non-parametric, hierarchical model for a laser-comb-calibrated spectrograph}
+
+\noindent
+\textbf{Lily~Zhao} (Yale) (Flatiron),
+\textbf{David~W~Hogg} (NYU) (MPIA) (Flatiron),
+\ldots and others\ldots
+
+\paragraph{Abstract:}
+There have been many hardware improvements in polygonal-fiber-fed,
+temperature-controlled, laser-frequency-comb or etalon-calibrated,
+high-resolution, extreme-precision radial-velocity spectrographs.
+Have the calibration methodologies improved to match?
+Here are three relevant observations:
+The first is that the calibration lines or spots from an etalon or
+comb fill the spectral range with dense calibration points.
+The second is that the spectrograph lives in a stabilized,
+climate-controlled environment, in which the full optical system and
+detectors will only vary within a tiny range of configurations or
+settings.
+The third is that---given this stability--- every calibration image
+ever taken is relevant to every science exposure ever taken; there is
+no reason to calibrate every exposure independently.
+The calibration methodology we propose here---\name---addresses these
+three problems by going non-parametric (no more polynomials) and then
+reducing dimensionality with a robust principal-component method.
+We demonstrate the success of this method with data from the
+\expres\ spectrograph.
+We find... [results and so on].
+
+\section{Introduction}
+
+\section{Method}
 

### 2019-11-02 paper/paper.tex
less me, more we

@@
 The calibration methodology we propose here---\name---addresses these
-three problems by going non-parametric (no more polynomials) and then
+three problems by going non-parametric (no more polynomials!) and then
 reducing dimensionality with a robust principal-component method.
@@
 First, some definitions.
-The way I (Hogg) think about this is the following:
+The way we think about this is the following:
 Given an exposure $n$, and order $m$, there is a relationship between
@@
 This is not precisely how the \expres\ team thinks about the problem,
-I (Hogg) think.
-For now I am going to ignore that fact, but I think what I say here
+but it isn't far off.
+For now we are going to ignore that fact, but what we say
 could be adapted to any \expres\ conventions.
@@
 found for exposure $n$ without regard to any other exposure $n'$.
-Further, I think they may be finding the coefficients $c_{nij}$ that
+Further, they may be finding the coefficients $c_{nij}$ that
 minimize an objective $Q$ that is something like the L2-norm:
@@
 Even in this simple context, where every calibration image $n$ is
-treated as its own unique flower, I think there are improvements to be
+treated as its own unique flower, there are improvements to be
 made.
@@
 because that is the definition of 9th order.
-If I am right about this, the fit could be taken to 13th order to obtain
+If we are right about this, the fit could be taken to 13th order to obtain
 much more expressiveness with (nearly) the same number of fit coefficients.
@@
 catastrophic outliers without destroying the fit.
-I would recommend the iteratively reweighted least squares (IRLS).
+We might recommend the iteratively reweighted least squares (IRLS).
 This would make the fitting more robust.
-For yet another, I have ideas about rescaling the products $x^i\,m^j$ to
+For yet another, there are rescaling issues for the products $x^i\,m^j$ to
 protect the fitting from near-singularities or bad conditioning of the
@@
 
-Of course the main thing I want to do is go hierarchical!
+Of course the main thing we want to do is go hierarchical!
 In this case, the calibration of every image is informed by the
@@
 
-I know how to do this, but the margins here are too small to contain my
-beautiful ideas...
+...
 

### 2019-11-07 paper/paper.tex
Asymmetric Residuals
Added brief discussion in new discussion section

@@
 
+\section{Discussion} \label{sec:discussion}
+
+Due to the dispersion intrinsic to echelle spectrographs, the wavelength change between pixels is greater at greater wavelengths.
+This means that the function of wavelength given pixel across an order will necessarily be concave down everywhere.
+Consequently, linear interpolation of this function will systematically give erroneously low values.
+
 \end{document}

### 2019-11-08 paper/paper.tex
Choices List
Draft of choice list in paper

@@
 
+\subsection{Choice Choices} \label{sec:choices}
+\begin{itemize}
+	\item Using scipy.interpolate.LSQUnivariateSpline to implement cubic spline between lines (fast and incorporates errors well; is chosen spline of the EXPRES pipeline) (and mostly isn't linear)
+	\item Interpolating each order individually
+	\item Wavelength solutions are delivered on a pixel grid of x and m
+	\item Eigenvector determination restricted to orders that are completely covered by LFC lines, and has enough signal to distinguish their peaks (orders 40-75 absolute)
+	\item Also restricted to pixels between 500 and 7000 out of 7920 for high signal
+	\item Using only exposures from after the replacement of the PCF in July 2019; gave rise to a large change in the nature of the instrument, but still gives us 4 months of data which exhibits non-negligible variation (this may have to be proven?)
+	\item Fitting wavelength as dependent variable?  (Should be line centers, right?)
+\end{itemize}
+
 \section{Discussion} \label{sec:discussion}

### 2019-12-09 paper/paper.tex
Paper Outline and Data Munging
Filling in the paper bit by bitter bit

@@
 \documentclass[12pt, letterpaper]{article}
+\usepackage{xcolor}
+\newcommand{\lz}[1]{\textcolor{orange}{#1}}
 
@@
 \section{Method}
+We are operating on the core idea that the wavelength solution is living in a low-dimensional space, where the degrees of freedom are set by the limited degrees of freedom of the spectrograph hardware.
 
-The idea is that the wavelength solution is going to live in a
-low-dimensional space, where the degrees of freedom are set by the
-degrees of freedom of the spectrograph hardware.
-
-First, some definitions.
 The way we think about this is the following:
 Given an exposure $n$, and order $m$, there is a relationship between
-the two-dimensional $(x,y)$-position on the detector and the
+the two-dimensional $(x,m)$-position on the detector and the
 wavelength $\lambda$
@@
 where $\theta_{n}$ is a big blob of parameters for this exposure.
-This is not precisely how the \expres\ team thinks about the problem,
-but it isn't far off.
-For now we are going to ignore that fact, but what we say
-could be adapted to any \expres\ conventions.
+
+Taking into account the low-degree of variability a stabilized spectrograph strives for, the calibration of every image can be informed by the calibration of every other image.
+To implement this hierarchical model, we use the calibration data themselves to develop a low-dimensional basis for expressing the calibration data.
+
+If the space of all calibration possibilities is in fact $K$-dimensional (where K is a small integer, i.e 2 or 8 or thereabouts thereabouts), and if the calibration variations are so
+small that we can linearize, then the function $f(x,y,m;\theta_{n})$ could
+be replaced with a tiny model
+\begin{equation}
+\lambda(x,y,m,n) = g_0(x,y,m) + \sum_{k=1}^K a_{nk}\,g_k(x,y,m)
+\quad ,
+\end{equation}
+where
+$g_0(x,y,m)$ is the fiducial or mean or standard calibration of the
+spectrograph,
+the $a_{nk}$ are scalar amplitudes,
+and the $g_k(x,y,m)$ are basis functions expressing the ``directions'' in calibration space that the spectrograph can depart from the
+fiducial calibration.
+The challenge is to learn these basis functions from the data, and get
+the $K$ amplitudes $a_{nk}$ for every exposure $n$.
+
+\subsection{Data}
+We present results using data from EXPRES, the EXtreme PRecison Spectrograph.
+EXPRES is an envionmentally stabilized, fiber-fed, $R=137,000$, optical spectrograph (\lz{CJ and RB's papers}).
+EXPRES has two different calibration sources, a ThAr lamp and a Menlo Systems laser frequency comb (LFC, e.g.
+\lz{Wilken+ 2012, Molaro+ 2013, Probst+ 2014, from RB's paper}).
+ThAr exposures are taken at the beginning and end of each night.
+LFC exposures are interspersed between science exposures every ~15 minutes.
+(Discussion of number of orders covered or number of lines in LFC and/or ThAr?)
+
+\lz{Discussion of line fitting?  Though the current line fitting causes me recurring nightmares.  I should really take the time to look into at least the peak fitting more.}
+
+EXPRES exposures are separated into different ``instrumental epochs," which correspond to changes to the position of the echellogram on the detector, the shape of the instrumental PSF, or the calibration sources (\lz{(worth including table that explicitly defines epochs?)}).
+Because significant instrumental changes break the assumption of only low-dimensional variations in the spectrograph hardware, we treat each instrument epoch independently.
+
+\subsection{Hierarchical De-Noising of Calibration Frames}
+let ${exp_n}$ be the set of calibration exposures within an instrument epoch that will be used to construct the basis functions that characterize the potential change in calibration space of each exposure throughout that instrument epoch.
+We first identify the full list of lines we can expect to find in a calibration image from this epoch by iterating through all ${exp_n}$ and constructing a list of all unique lines found in these exposures.
+A line is uniquely defined by a combination of order, $m$, and ``true" wavelength value, $\lambda$.
+
+For each line, we then assign the fitted line center
+
+figuring out which lines have measurements
+
+How is the data cleaned
+Find the lines that cover the range of wavelengths for each order
+identify where this line center is observed for each of the exposures
+Lines that are missing from greater than n percent of exposures are cut
+Exposures that are missing n percent of lines are cut \textcolor{red}{This will miss low signal exposures that suffer most in the first few orders but can recover in later orders; not sure if this is a feature or a bug}
+this n is 50 for LFCs and 30 for ThArs
+\lz{If we get to it, a discussion of what we do when we use both ThArs and LFCs to construct the PCA}
+
+Defining generations (if we want to do sub generations?)
+
+Replacing values (patching)
+
+Summary
+
+\subsection{Non-Parametric Calibration}
+interpolate PCA coefficients over time to ``paint" a wavelength solution onto calibrations outside of the training set
+
+Generate new line list for that time
+
+Interpolate, order-by-order in x direction to fill in an exposure (either to all pixels, or to the line centers of other exposures)
+
+Summary
+
+\subsection{Choice Choices} \label{sec:choices}
+(Before or after results presented in validation test and ThAr sections)
+\begin{itemize}
+	\item Using scipy.interpolate.LSQUnivariateSpline to implement cubic spline between lines (fast and incorporates errors well; is chosen spline of the EXPRES pipeline) (and mostly isn't linear)
+	\item Interpolating each order individually
+	\item Wavelength solutions are delivered on a pixel grid of x and m
+	\item Eigenvector determination restricted to orders that are completely covered by LFC lines, and has enough signal to distinguish their peaks (orders 40-75 absolute)
+	\item Also restricted to pixels between 500 and 7000 out of 7920 for high signal
+	\item Using only exposures from after the replacement of the PCF in July 2019; gave rise to a large change in the nature of the instrument, but still gives us 4 months of data which exhibits non-negligible variation (this may have to be proven?)
+	\item Fitting wavelength as dependent variable?  (Should be line centers, right?)
+\end{itemize}
+
+Which K did we choose and why?
+
+Interpolation scheme (both together or individually and which first?)
+
+Order of interpolation
+
+Due to the dispersion intrinsic to echelle spectrographs, the wavelength change between pixels is greater at greater wavelengths.
+This means that the function of wavelength given pixel across an order will necessarily be concave down everywhere.
+Consequently, linear interpolation of this function will systematically give erroneously low values.
+
+\section{Validation Tests}
+
+\subsection{Comparison to Parametric Model}
 Right now in the \expres\ pipeline the $\theta_{n}$ comprises the
@@
 
-Of course the main thing we want to do is go hierarchical!
-In this case, the calibration of every image is informed by the
-calibration of every other image.
-The simplest form of hierarchical model is to use the calibration data
-themselves to develop a low-dimensional basis for expressing the
-calibration data.
+\section{ThAr Stuff}
 
-If the space of all calibration possibilities is in fact
-$K$-dimensional (where I am thinking of this as being a small integer,
-like $K=3$ or $K=8$ or thereabouts), and if the calibration variations are so
-small that we can linearize, then the function $f(x,y,m;\theta_{n})$ could
-be replaced with a tiny model
-\begin{equation}
-\lambda(x,y,m,n) = g_0(x,y,m) + \sum_{k=1}^K a_{nk}\,g_k(x,y,m)
-\quad ,
-\end{equation}
-where
-$g_0(x,y,m)$ is the fiducial or mean or standard calibration of the
-spectrograph,
-the $a_{nk}$ are scalar amplitudes,
-and the $g_k(x,y,m)$ are basis functions expressing the ``directions''
-in calibration space that the spectrograph can depart from the
-fiducial calibration.
-The challenge is to learn these basis functions from the data, and get
-the $K$ amplitudes $a_{nk}$ for every exposure $n$.
-
-...
-
-\subsection{Choice Choices} \label{sec:choices}
-\begin{itemize}
-	\item Using scipy.interpolate.LSQUnivariateSpline to implement cubic spline between lines (fast and incorporates errors well; is chosen spline of the EXPRES pipeline) (and mostly isn't linear)
-	\item Interpolating each order individually
-	\item Wavelength solutions are delivered on a pixel grid of x and m
-	\item Eigenvector determination restricted to orders that are completely covered by LFC lines, and has enough signal to distinguish their peaks (orders 40-75 absolute)
-	\item Also restricted to pixels between 500 and 7000 out of 7920 for high signal
-	\item Using only exposures from after the replacement of the PCF in July 2019; gave rise to a large change in the nature of the instrument, but still gives us 4 months of data which exhibits non-negligible variation (this may have to be proven?)
-	\item Fitting wavelength as dependent variable?  (Should be line centers, right?)
-\end{itemize}
+\section{Real Data Through to RVs}
 
@@
 
-Due to the dispersion intrinsic to echelle spectrographs, the wavelength change between pixels is greater at greater wavelengths.
-This means that the function of wavelength given pixel across an order will necessarily be concave down everywhere.
-Consequently, linear interpolation of this function will systematically give erroneously low values.
-
 \end{document}

### 2019-12-09 paper/paper.tex
They call it a first draft for a reason
Please don't look at me

@@
 \newcommand{\lz}[1]{\textcolor{orange}{#1}}
+
+\usepackage[ruled,vlined]{algorithm2e}
 
@@
 \subsection{Hierarchical De-Noising of Calibration Frames}
+% Gotta make this whole section more math-y
+% Add variable names for matrix of all lines and exposures, what defines a line, what defines a line measurement, fiducial calibration
 let ${exp_n}$ be the set of calibration exposures within an instrument epoch that will be used to construct the basis functions that characterize the potential change in calibration space of each exposure throughout that instrument epoch.
@@
 
-For each line, we then assign the fitted line center
+For each line, we then find the fitted line center, in pixel space, for that line for each exposure.
+If the line is missing from an exposure, we insert a NaN for the line center instead.
+Lines that are missing from greater than some percent of exposures are cut.
+Exposures that are missing greater than some percent of lines are also cut (\lz{This will miss low signal exposures that suffer most in the first few orders but can recover in later orders; can't decide if this is a feature or a bug}).
+This threshold percentage is 50\% for LFC exposures and 30\% for ThAr exposures.
 
-figuring out which lines have measurements
+\lz{Discuss separating out sub-epochs due to echellogram shifts from focus motor position changes here if relevant}
 
-How is the data cleaned
-Find the lines that cover the range of wavelengths for each order
-identify where this line center is observed for each of the exposures
-Lines that are missing from greater than n percent of exposures are cut
-Exposures that are missing n percent of lines are cut \textcolor{red}{This will miss low signal exposures that suffer most in the first few orders but can recover in later orders; not sure if this is a feature or a bug}
-this n is 50 for LFCs and 30 for ThArs
-\lz{If we get to it, a discussion of what we do when we use both ThArs and LFCs to construct the PCA}
+Missing lines were replaced with denoised estimates using iterative PCA.
+The line center of each missing line are initialized to the mean of the nearest 9 exposures in time.
+We then find a fiducial calibration of the spectrograph, which is the mean line center value for each identified line.
+We subtract this fiducial calibration from each exposure's measured line centers and run PCA on the difference between the measured and fiducial line center for each exposure.
 
-Defining generations (if we want to do sub generations?)
+Using the first K principal components, where K is some small integer like 2, and PCA weights, we reconstruct a de-noised matrix of line centers for each exposure.
+Values from this reconstruction is then used to patch missing line centers.
+This is done iteratively until the value of missing line centers being patched changes by less than 0.01\% of a pixel.
+The last interpolation's principal components and coefficients are returned as well as the fiducial calibration and the de-noised matrix.
 
-Replacing values (patching)
-
-Summary
+\begin{algorithm}
+\SetAlgoLined
+\While{change in bad pixels $>$ 0.01\%}{
+	find fiducial calibration of the spectrograph\;
+	determine PCA of line matrix - fiducial calibration\;
+	reconstruct line matrix to order K\;
+	patch bad pixels with reconstructed line matrix
+	}
+\caption{Hierarchical De-Noising}
+\end{algorithm}
 
 \subsection{Non-Parametric Calibration}
-interpolate PCA coefficients over time to ``paint" a wavelength solution onto calibrations outside of the training set
+We can interpolate the returned PCA coefficients over time to determine the wavelength solution at any given time.
+Using the interpolated PCA coefficients and the returned basis vectors, we can construct a de-noised set of line centers for that point in time.
+We use a cubic spline to interpolate the PCA coefficients.
 
-Generate new line list for that time
+Once we have determined line centers for each line, which has an associated known wavelength, we can interpolate the known wavelengths over line centers to any pixel in a given order.
+We can either interpolate onto every x pixel, and generate a complete wavelength solution for that order, or interpolate onto the measured line centers of a calibration image and compare the return wavelengths as a validation of our method.
+This interpolation is also done using a cubic spline.
 
-Interpolate, order-by-order in x direction to fill in an exposure (either to all pixels, or to the line centers of other exposures)
-
-Summary
+\begin{algorithm}
+\SetAlgoLined
+Inputs: new time, new line centers, new orders\; %command not working for w/e reason
+interpolate PCA coefficients to new time\;
+reconstruct line matrix for that time\;
+\For{each order}{
+	identify the new line centers in that order using new orders\;
+	Interpolate the info from the line matrix to the new x values\;	
+	}
+\caption{Non-Parametric Wavelength Solution}
+\end{algorithm}
 
 \subsection{Choice Choices} \label{sec:choices}
-(Before or after results presented in validation test and ThAr sections)
+(Before or after results presented in validation test and ThAr sections?)
 \begin{itemize}
-	\item Using scipy.interpolate.LSQUnivariateSpline to implement cubic spline between lines (fast and incorporates errors well; is chosen spline of the EXPRES pipeline) (and mostly isn't linear)
-	\item Interpolating each order individually
-	\item Wavelength solutions are delivered on a pixel grid of x and m
-	\item Eigenvector determination restricted to orders that are completely covered by LFC lines, and has enough signal to distinguish their peaks (orders 40-75 absolute)
-	\item Also restricted to pixels between 500 and 7000 out of 7920 for high signal
-	\item Using only exposures from after the replacement of the PCF in July 2019; gave rise to a large change in the nature of the instrument, but still gives us 4 months of data which exhibits non-negligible variation (this may have to be proven?)
+	\item Value of K for denoising
+	\item Type of interpolation (cubic spline vs nearest neighbor or linear)
+	\item Interpolation order (interpolate time and x direction together?  Individually?  Which one first?)
 	\item Fitting wavelength as dependent variable?  (Should be line centers, right?)
 \end{itemize}
-
-Which K did we choose and why?
-
-Interpolation scheme (both together or individually and which first?)
-
-Order of interpolation
 
@@
 
-\section{Validation Tests}
+\section{Results}
+\subsection{Self Tests}
+predict wavelengths for the returned denoised x values, should return identical values to machine error
+
+predict wavelengths for measured x values for an exposure included in the training set
+should return results as good as we can expect given error inherent in measuring line centers?
+isolated test of how well the interpolation of lines in the x-pixel-direction does
+
+\subsection{Training and Validation Tests}
+training with 90\% of calibration exposures in an instrumental epoch
+Using the results of that training to predict the wavelengths of the measured line centers in the 10\% of exposures left out
+Tests how well the interpolation of PCA coefficients does at constructing the line centers AND THEN how well the interpolation of lines in the x-pixel-direction does
+You can use results from self test to inform this test, \lz{or just difference the measured line centers for a more direct answer, which is something I should do}
+
+\subsection{Cross Tests}
+Using PCA trained with LFC to predict ThAr wavelengths
+Like training and validation test, but have to take into account the increased error in measuring ThAr line centers, I think
+
+\subsection{ThAr Stuff?  Or Include it in Previous Subsections?}
 
@@
 
-\section{ThAr Stuff}
+\subsection{Comparison to Non-Parametric, Non-Hierarchical Model}
 

### 2019-12-10 paper/paper.tex
Paper: Summary of Tests

@@
 
-\section{Method}
+Why hierarchical?
+
+Why non-parametric?
+
+\section{Method} \label{sec:method}
 We are operating on the core idea that the wavelength solution is living in a low-dimensional space, where the degrees of freedom are set by the limited degrees of freedom of the spectrograph hardware.
@@
 
-\subsection{Data}
-We present results using data from EXPRES, the EXtreme PRecison Spectrograph.
-EXPRES is an envionmentally stabilized, fiber-fed, $R=137,000$, optical spectrograph (\lz{CJ and RB's papers}).
-EXPRES has two different calibration sources, a ThAr lamp and a Menlo Systems laser frequency comb (LFC, e.g.
+\subsection{Data} \label{sec:data}
+We present results using data from \expres\, the EXtreme PRecison Spectrograph.
+\expres\ is an envionmentally stabilized, fiber-fed, $R=137,000$, optical spectrograph (\lz{CJ and RB's papers}).
+\expres\ has two different calibration sources, a ThAr lamp and a Menlo Systems laser frequency comb (LFC, e.g.
 \lz{Wilken+ 2012, Molaro+ 2013, Probst+ 2014, from RB's paper}).
@@
 
-EXPRES exposures are separated into different ``instrumental epochs," which correspond to changes to the position of the echellogram on the detector, the shape of the instrumental PSF, or the calibration sources (\lz{(worth including table that explicitly defines epochs?)}).
+\expres\ exposures are separated into different ``instrumental epochs," which correspond to changes to the position of the echellogram on the detector, the shape of the instrumental PSF, or the calibration sources (\lz{(worth including table that explicitly defines epochs?)}).
 Because significant instrumental changes break the assumption of only low-dimensional variations in the spectrograph hardware, we treat each instrument epoch independently.
 
-\subsection{Hierarchical De-Noising of Calibration Frames}
+\subsection{Hierarchical De-Noising of Calibration Frames} \label{sec:denoising}
 % Gotta make this whole section more math-y
@@
 
-\subsection{Non-Parametric Calibration}
+\subsection{Non-Parametric Calibration} \label{sec:nonparam}
 We can interpolate the returned PCA coefficients over time to determine the wavelength solution at any given time.
@@
 
-\section{Results}
-\subsection{Self Tests}
-predict wavelengths for the returned denoised x values, should return identical values to machine error
+\section{Results} \label{sec:results}
+\subsection{Self Tests} \label{sec:test-self}
+First, let's have a sanity check.
+Using a cubic spline scheme with no smoothing to predict wavelengths for the returned, denoised line center values should return identical values to the assigned true wavelength for each line within machine error.
+We do see this, for both ThArs and LFCs: look at the figure.
 
-predict wavelengths for measured x values for an exposure included in the training set
-should return results as good as we can expect given error inherent in measuring line centers?
-isolated test of how well the interpolation of lines in the x-pixel-direction does
+As an only slightly less silly sanity check, we can also use the constructed PCA basis to predict wavelengths for the measured line centers for an exposure that was included in the set of exposures used to construct the basis.
+This interpolation should return identical PCA coefficients for that time.
+Differences in wavelengths will arise from a combination of 1) error inherent in measuring line centers and 2) interpolation error of lines to other x positions in the pixel direction.
 
-\subsection{Training and Validation Tests}
-training with 90\% of calibration exposures in an instrumental epoch
-Using the results of that training to predict the wavelengths of the measured line centers in the 10\% of exposures left out
-Tests how well the interpolation of PCA coefficients does at constructing the line centers AND THEN how well the interpolation of lines in the x-pixel-direction does
-You can use results from self test to inform this test, \lz{or just difference the measured line centers for a more direct answer, which is something I should do}
+Difference in denoised x-values to measured ones? (translate into rv shift?)
+median and spread in errors in pixel (like 0.006) and RV (like 3.2 m/s?)
+Histograms
 
-\subsection{Cross Tests}
-Using PCA trained with LFC to predict ThAr wavelengths
-Like training and validation test, but have to take into account the increased error in measuring ThAr line centers, I think
+\subsection{Training and Validation Tests} \label{sec:test-trainNvalid}
+A better test is to leave out some percentage of the calibration images while constructing the basis and weights with PCA.
+We can then use those basis and weights to predict wavelengths for the calibrations images that were left out and assess how accurate the predictions are.
+We will call the used calibration images the ``training set," and the left out exposures the ``validation set."
 
-\subsection{ThAr Stuff?  Or Include it in Previous Subsections?}
+We chose to use 90\% of available exposures to train the PCA, and leave the remaining 10\% for validation.
+Exposures were sorted into either set at random.
+This test will finally incorporate testing how well PCA coefficients are interpolated to other times outside of the training set and therefore how well the line centers for each wavelength can be reconstructed at all times.
+As before, differences in wavelength will also be drawn from how well line centers are measured and error in interpolating to those measured values.
+This second source of error was independently quantified in the second test described in the previous section.
+
+LFC looks good, ThAr is struggling a bit
+Difference in denoised x-values to measured ones? (translate into rv shift?)
+median and spread in errors in pixel (like 0.006) and RV (like 3.2 m/s?)
+Histograms
+
+\subsection{Cross Tests} \label{sec:test-cross}
+Similar to the training and validation tests, we can also use one calibration source to construct the PCA basis and weights and the other to test the results.
+For example, we can use LFC exposures to construct the wavelength solution, and then use it to predict wavelengths for measured line centers of ThAr wavelengths, which will have assigned true wavelengths.
+Likewise, we can use ThAr exposures to predict the wavelengths of LFC lines.
+Note that errors in this test will also draw on the difference in quality of the two calibration sources being compared.
+
+LFC predicts ThAr better than ThAr can predict itself.
+ThAr predicting LFC is really messing up somehow, and I should figure out what it is (sorry!)
+
+%\subsection{ThAr Stuff?  Or Include it in Previous Subsections?}
+
+\section{Comparisons to Other Methods} \label{sec:comparisons}
+Presumably we'll use the training/validation method to compare these?
 
 \subsection{Comparison to Parametric Model}
-Right now in the \expres\ pipeline the $\theta_{n}$ comprises the
+Previously, the \expres\ pipeline the $\theta_{n}$ comprises the
 amplitudes of some 9th-ish-order polynomial in $x$ and $n$, possibly
@@
 
-\section{Real Data Through to RVs}
+\section{Real Data Through to RVs} \label{sec:realdata}
 

### 2019-12-11 paper/paper.tex
Comparison Plots and Code
Paper Method and Test Sections

@@
 \usepackage[ruled,vlined]{algorithm2e}
+\usepackage{graphicx}
 
@@
 
+Why non-parametric?
+(because density of lines given by the LFC)
+
 Why hierarchical?
-
-Why non-parametric?
+(Because stability of new instruments)
 
 \section{Method} \label{sec:method}
-We are operating on the core idea that the wavelength solution is living in a low-dimensional space, where the degrees of freedom are set by the limited degrees of freedom of the spectrograph hardware.
+We operate on the core idea that the wavelength solution is living in a low-dimensional space, where the degrees of freedom are set by the limited degrees of freedom of the spectrograph hardware.
 
-The way we think about this is the following:
-Given an exposure $n$, and order $m$, there is a relationship between
-the two-dimensional $(x,m)$-position on the detector and the
+We will pose the problem in the following way.  Given an exposure $n$, and order $m$, there is a relationship between
+the two-dimensional $(x,y)$-position on the detector and the
 wavelength $\lambda$
@@
 
-Taking into account the low-degree of variability a stabilized spectrograph strives for, the calibration of every image can be informed by the calibration of every other image.
-To implement this hierarchical model, we use the calibration data themselves to develop a low-dimensional basis for expressing the calibration data.
+A stabilized spectrograph should experience only low-degree variability, meaning the calibration of any image can be informed by the calibration of every other image.
+To implement such a hierarchical model, we use the calibration data themselves to develop a low-dimensional basis for expressing the calibration data.
 
-If the space of all calibration possibilities is in fact $K$-dimensional (where K is a small integer, i.e 2 or 8 or thereabouts thereabouts), and if the calibration variations are so
+If the space of all calibration possibilities is in fact $K$-dimensional (where K is a small integer, i.e 2 or 8 or thereabouts), and if the calibration variations are so
 small that we can linearize, then the function $f(x,y,m;\theta_{n})$ could
@@
 fiducial calibration.
-The challenge is to learn these basis functions from the data, and get
+The challenge is to learn these basis functions from the data and get
 the $K$ amplitudes $a_{nk}$ for every exposure $n$.
@@
 \expres\ is an envionmentally stabilized, fiber-fed, $R=137,000$, optical spectrograph (\lz{CJ and RB's papers}).
-\expres\ has two different calibration sources, a ThAr lamp and a Menlo Systems laser frequency comb (LFC, e.g.
+\expres\ has two different wavelength calibration sources, a ThAr lamp and a Menlo Systems laser frequency comb (LFC, e.g.
 \lz{Wilken+ 2012, Molaro+ 2013, Probst+ 2014, from RB's paper}).
@@
 
+For the results presented in this paper, we work with the 1D, optimally extracted data.  This removes the y-dependence from equations 1 and 2.
+
 \expres\ exposures are separated into different ``instrumental epochs," which correspond to changes to the position of the echellogram on the detector, the shape of the instrumental PSF, or the calibration sources (\lz{(worth including table that explicitly defines epochs?)}).
@@
 \subsection{Hierarchical De-Noising of Calibration Frames} \label{sec:denoising}
-% Gotta make this whole section more math-y
-% Add variable names for matrix of all lines and exposures, what defines a line, what defines a line measurement, fiducial calibration
-let ${exp_n}$ be the set of calibration exposures within an instrument epoch that will be used to construct the basis functions that characterize the potential change in calibration space of each exposure throughout that instrument epoch.
-We first identify the full list of lines we can expect to find in a calibration image from this epoch by iterating through all ${exp_n}$ and constructing a list of all unique lines found in these exposures.
-A line is uniquely defined by a combination of order, $m$, and ``true" wavelength value, $\lambda$.
+Let $\{exp_n\}$ be the set of calibration exposures within an instrument epoch used to construct $g_k(x,y,m)$.
+We first identify the full list of lines we can expect to find in a calibration image from this epoch.
+A line is uniquely defined by a combination of order, $m$, and ``true" or theoretical wavelength of that line, $\lambda$.
+Let $\{l_{m,\lambda}\}$ be the complete list of unique lines found in $\{exp_n\}$.
 
-For each line, we then find the fitted line center, in pixel space, for that line for each exposure.
-If the line is missing from an exposure, we insert a NaN for the line center instead.
+For each line, $l_{m,\lambda}$, we then find the fitted line center, $x_{n,l}$, in pixel space, for that line for each exposure.
+If this line is missing from an exposure (e.g.
+it was covered by noise or off the edge of its usual order), we insert a $NaN$ for that line center instead.
+
 Lines that are missing from greater than some percent of exposures are cut.
@@
 
-\lz{Discuss separating out sub-epochs due to echellogram shifts from focus motor position changes here if relevant}
+Missing line center measurements are replaced with denoised estimates using iterative PCA.
+For each missing line center measurement, we initialzie their value to the mean of the nearest $n$ exposures in time, where $n$ is typically 9 for LFC exposures and $15$ for ThAr.
+We then find a fiducial calibration of the spectrograph,  $g_0(x,y,m)$, for which we use the mean line center value for each identified line.
+We subtract this fiducial calibration from each exposure's measured line centers and run PCA on the difference between the measured and fiducial line center for each exposure.
+This PCA gives us $a_{nk}$ and $g_k(x,y,m)$.
 
-Missing lines were replaced with denoised estimates using iterative PCA.
-The line center of each missing line are initialized to the mean of the nearest 9 exposures in time.
-We then find a fiducial calibration of the spectrograph, which is the mean line center value for each identified line.
-We subtract this fiducial calibration from each exposure's measured line centers and run PCA on the difference between the measured and fiducial line center for each exposure.
-
-Using the first K principal components, where K is some small integer like 2, and PCA weights, we reconstruct a de-noised matrix of line centers for each exposure.
-Values from this reconstruction is then used to patch missing line centers.
-This is done iteratively until the value of missing line centers being patched changes by less than 0.01\% of a pixel.
-The last interpolation's principal components and coefficients are returned as well as the fiducial calibration and the de-noised matrix.
+Using those values with equation 2 and some small integer (e.g.
+2) for K, we can reconstruct a de-noised matrix of line centers for each exposure.
+Missing line center measurements are then replaced with values from this reconstruction.
+We do this iteratively until the values of the missing line centers change by less than 0.01\% of a pixel.
+This convergence suggests that the PCA is no longer being swayed by outliers and is truly ``de-noised."
 
@@
 \SetAlgoLined
+\textbf{Inputs}: \;
 \While{change in bad pixels $>$ 0.01\%}{
-	find fiducial calibration of the spectrograph\;
-	determine PCA of line matrix - fiducial calibration\;
-	reconstruct line matrix to order K\;
-	patch bad pixels with reconstructed line matrix
+	$g_0(x,y,m) = \overline{\{x_{n,l}\}}$\;
+	find $U, \Sigma, V$ s.t. $U\Sigma V^* = (x_{l,n}-g_0(x,y,m))$\;
+	let $a_{n,k} = U\cdot \Sigma$ and $V = g_k(x,y,m)$\;
+	$\lambda(x,y,m,n) = g_0(x,y,m) + \sum_{k=1}^K a_{nk}\,g_k(x,y,m)$ where $K=2$\;
+	$\{x_{n,l}\} = \lambda(x,y,m,n)$ where $\{x_{n,l}\}$ is $NaN$
 	}
@@
 \subsection{Non-Parametric Calibration} \label{sec:nonparam}
-We can interpolate the returned PCA coefficients over time to determine the wavelength solution at any given time.
-Using the interpolated PCA coefficients and the returned basis vectors, we can construct a de-noised set of line centers for that point in time.
-We use a cubic spline to interpolate the PCA coefficients.
+We can interpolate $a_{nk}$ with respect to time to determine $\lambda(x,y,m,n)$ at any time.
+Once we have the line center for each line, which has an associated known wavelength, we can interpolate the known wavelengths over line centers to any position in x in a given order.
+For example, interpolating onto every integer x will generate wavelengths for each pixel in an order.
 
-Once we have determined line centers for each line, which has an associated known wavelength, we can interpolate the known wavelengths over line centers to any pixel in a given order.
-We can either interpolate onto every x pixel, and generate a complete wavelength solution for that order, or interpolate onto the measured line centers of a calibration image and compare the return wavelengths as a validation of our method.
-This interpolation is also done using a cubic spline.
+We chose to implement both interpolations with a cubic spline.  Interpolation of wavelengths over line centers is done order by order.
 
@@
 \SetAlgoLined
-Inputs: new time, new line centers, new orders\; %command not working for w/e reason
-interpolate PCA coefficients to new time\;
-reconstruct line matrix for that time\;
-\For{each order}{
-	identify the new line centers in that order using new orders\;
-	Interpolate the info from the line matrix to the new x values\;	
+\textbf{Inputs}: new time, new line centers, new orders\; %command not working for w/e reason
+Let $t_n$ be the time associated with exposure $n$\;
+let $t_n'$, $x_n'$, and $m_n'$ be the new time, line centers, and orders we want wavelengths for\;
+Let $a_{n,k}'$ be the interpolation of $a_{n,k}$ to time $t_n'$\;
+$\lambda(x,y,m,n)' = g_0(x,y,m) + \sum_{k=1}^K a_{nk}'\,g_k(x,y,m)$ where $K=2$\;
+\For{each unique $m_n'$}{
+	${l_n,m}$ from $\lambda(x,y,m,n)'$
+	interpolate ${l_n,m}$ onto $x_n'$\;
 	}
@@
 \end{algorithm}
+\lz{What's fancy math symbol for interpolate?}
 
@@
 (Before or after results presented in validation test and ThAr sections?)
+
+We're still testing most of these (will use that as excuse to write this section later)
 \begin{itemize}
@@
 	\item Interpolation order (interpolate time and x direction together?  Individually?  Which one first?)
-	\item Fitting wavelength as dependent variable?  (Should be line centers, right?)
 \end{itemize}
 
+Reason for not linear interpolation:
 Due to the dispersion intrinsic to echelle spectrographs, the wavelength change between pixels is greater at greater wavelengths.
@@
 First, let's have a sanity check.
-Using a cubic spline scheme with no smoothing to predict wavelengths for the returned, denoised line center values should return identical values to the assigned true wavelength for each line within machine error.
-We do see this, for both ThArs and LFCs: look at the figure.
+Using a cubic spline scheme with no smoothing to predict wavelengths for the returned, denoised line center values, $\lambda(x,y,m,n)$, should return identical values to the assigned true wavelength for each line within machine error.
+This is what happens, for both ThArs and LFCs: look at the figure.
 
-As an only slightly less silly sanity check, we can also use the constructed PCA basis to predict wavelengths for the measured line centers for an exposure that was included in the set of exposures used to construct the basis.
+As an only slightly less silly sanity check, we can also use the constructed PCA basis to predict wavelengths for the measured line centers of an exposure that was included in the set of exposures used to construct the basis.
 This interpolation should return identical PCA coefficients for that time.
-Differences in wavelengths will arise from a combination of 1) error inherent in measuring line centers and 2) interpolation error of lines to other x positions in the pixel direction.
+Differences in wavelengths will arise from a combination of 1) the error inherent in measuring line centers and 2) interpolation error of lines to other x positions in the pixel direction.
 
@@
 \subsection{Training and Validation Tests} \label{sec:test-trainNvalid}
-A better test is to leave out some percentage of the calibration images while constructing the basis and weights with PCA.
-We can then use those basis and weights to predict wavelengths for the calibrations images that were left out and assess how accurate the predictions are.
+A better test is to leave out some percentage of the calibration images from $\{exp_n\}$ while constructing the basis and weights with PCA.
+We can then use the results from that PCA to predict wavelengths for lines in the calibrations images that were left out.
 We will call the used calibration images the ``training set," and the left out exposures the ``validation set."
@@
 Exposures were sorted into either set at random.
-This test will finally incorporate testing how well PCA coefficients are interpolated to other times outside of the training set and therefore how well the line centers for each wavelength can be reconstructed at all times.
+This will test how well PCA coefficients are interpolated to other times outside of the training set and therefore how well line centers can be reconstructed at all times.
 As before, differences in wavelength will also be drawn from how well line centers are measured and error in interpolating to those measured values.
@@
 Similar to the training and validation tests, we can also use one calibration source to construct the PCA basis and weights and the other to test the results.
-For example, we can use LFC exposures to construct the wavelength solution, and then use it to predict wavelengths for measured line centers of ThAr wavelengths, which will have assigned true wavelengths.
+For example, we can use LFC exposures to construct the wavelength solution, and then use it to predict wavelengths for measured line centers of ThAr lines, which will have their own assigned true wavelengths.
 Likewise, we can use ThAr exposures to predict the wavelengths of LFC lines.
@@
 \section{Comparisons to Other Methods} \label{sec:comparisons}
-Presumably we'll use the training/validation method to compare these?
+I think we probably have to use the even/odd method to compare.
 
 \subsection{Comparison to Parametric Model}
-Previously, the \expres\ pipeline the $\theta_{n}$ comprises the
-amplitudes of some 9th-ish-order polynomial in $x$ and $n$, possibly
-ignoring $y$?
-If so, they (may?) have
+Previously, the \expres\ pipeline the $\theta_{n}$ from equation 1 comprises the
+amplitudes of some 9th-ish-order polynomial in $x$ and $m$ for each $n$.
+If so, they have
 \begin{equation}
@@
 \end{equation}
-where the $c_{nij}$ are coefficients unique to exposure $n$, and
-found for exposure $n$ without regard to any other exposure $n'$.
-Further, they may be finding the coefficients $c_{nij}$ that
+where the $c_{nij}$ are coefficients unique to exposure $n$ but smoothed by a third order polynomial over time.
+Further, they were finding the coefficients $c_{nij}$ that
 minimize an objective $Q$ that is something like the L2-norm:
@@
 \end{equation}
-because that is the definition of 9th order.
-If we are right about this, the fit could be taken to 13th order to obtain
-much more expressiveness with (nearly) the same number of fit coefficients.
+because that is the definition of 9th order.  We find that fitting to 8th order is sufficient, faster, and leaves less residual pattern than the original 9th degree fit.
 For another, the objective function could be made soft to permit
-catastrophic outliers without destroying the fit.
+catastrophic outliers without destroying the fit. (\lz{we did not do this})
 We might recommend the iteratively reweighted least squares (IRLS).
@@
 
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.9\textwidth]{Figures/polyval2d.pdf}
+\caption{}
+\label{fig:polyValFit}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.9\textwidth]{Figures/designMatrix.pdf}
+\caption{}
+\label{fig:dsnMFit}
+\end{figure}
+
 \subsection{Comparison to Non-Parametric, Non-Hierarchical Model}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.9\textwidth]{Figures/noHierc.pdf}
+\caption{}
+\label{fig:intpFit}
+\end{figure}
 
@@
 \section{Discussion} \label{sec:discussion}
+Generalization to etalon
+
+Maybe some practical statements?
+How many exposures are needed to implement this method at all.
+How often do exposures have to be taken throughout a night (unless that's too dependent on instrument stability)
 

### 2020-01-10 paper/paper.tex
introduction; discussion

@@
 % Copyright 2019 the authors. All rights reserved.
+
+% style notes
+% -----------
+% - use \acronym and use \eprv, \lfc, and so on.
 
@@
 \newcommand{\name}{\project{NameOfThis}}
+\newcommand{\eprv}{\acronym{EPRV}}
+\newcommand{\lfc}{\acronym{LFC}}
+
+% math shih
+\newcommand{\mps}{\mathrm{m\,s^{-1}}}
 
@@
 \name:
-A non-parametric, hierarchical model for a laser-comb-calibrated spectrograph}
+A non-parametric, hierarchical model for a precision spectrograph}
 
@@
 The first is that the calibration lines or spots from an etalon or
-comb fill the spectral range with dense calibration points.
+comb (or even arc lamp) fill the spectral range with dense calibration points.
 The second is that the spectrograph lives in a stabilized,
@@
 settings.
-The third is that---given this stability--- every calibration image
+The third is that, given this stability, every calibration image
 ever taken is relevant to every science exposure ever taken; there is
@@
 
-Why non-parametric?
-(because density of lines given by the LFC)
+Extreme precision radial-velocity (\eprv) programs have been
+outrageously successful in finding and characterizing extra-solar
+planets (CITE THINGS).
+These programs typically make use of spectrographs with resolutions on
+the order of $10^5$, which correspond to line widths on the order of
+$3000\,\mps$.
+Exoplanet science happens at $3\,\mps$ precisions, and the current
+instruments hope to get to $0.1\,\mps$ precisions (CITE THINGS).
+This means that new spectrographs are expected to be calibrated or
+stabilized to better than $10^{-4}$ of a pixel (assuming that the
+spectrograph is well sampled).
+And indeed, some hardware--software systems seem to be reaching these
+levels (CITE THINGS).
 
-Why hierarchical?
-(Because stability of new instruments)
+Here we propose to simplify and improve calibration programs for
+\eprv\ hardware systems with two very simple but novel ideas.
+The first flows from the observation that calibration sources---which
+include lamps, etalons, and laser-frequency combs (\lfc
+s)---illuminate the spectrograph with dense sets of lines; almost every
+location in the spectrograph image plane is surrounded by nearby,
+useful calibration lines.
+This recommends a calibration methodology that is
+\emph{non-parametric}:
+If every point in the spectrograph detector is surrounded by nearby
+calibration lines, the wavelength solution can be made simply an
+interpolation of the calibration data.
+There is no need to choose any functional form for the wavelength
+solution, such as an eighth-order polynomial.
+Going non-parametric will improve calibration accuracy, because
+choosing a parametric form biases the calibration, and biases it
+badly when the form is inappropriate.
+
+The second idea here flows from the observation that contemporary
+\eprv\ instruments are incredibly stable.
+Temperature-controlled, fiber-fed pectrographs vary only slightly over
+the night or season, and only along a small number of axes in what you
+might call ``calibration space''..... HOGG
 
@@
 \section{Discussion} \label{sec:discussion}
+
 Generalization to etalon
@@
 
+Some ideas about self-calibration. Once you know the space in which the
+spectrograph lives, every single science frame gives you information about
+where you are in that space; you don't need to look at the \lfc.
+
 \end{document}

### 2020-01-10 paper/paper.tex
intro notes

@@
 choosing a parametric form biases the calibration, and biases it
-badly when the form is inappropriate.
+badly when the form is inappropriate (as, for example, polynomials
+will be at the detector edges).
 
@@
 might call ``calibration space''..... HOGG
+
+HOGG: Return to the point that non-parametric \emph{increases} the
+freedom of the model (infinitely in some sense) but hierarchical
+\emph{restricts} the freedom of the model (enormously, but we hope
+appropriately).... HOGG
 

### 2020-01-10 py/waveCal.py
WaveCal Edits
Changing interpolation schemes, removing order list option, adding screens for out of order line and outliers

@@ -7,7 +7,7 @@ import matplotlib.pyplot as plt
 from astropy.io import fits
 from astropy.constants import c
 from astropy.time import Time
-from scipy import interpolate
+from scipy import interpolate, optimize
 from scipy.io import readsav
 from sklearn.decomposition import TruncatedSVD
 
@@ -32,6 +32,7 @@ def readParams(file_name):
     """
     info = np.load(file_name,allow_pickle=True)[()]
     # Assemble information into "fit-able" form
+    num_orders = len(info['params'])
     lines = [p[:,1] for p in info['params'] if p is not None]
     errs = [np.sqrt(cov[:,1,1]) for cov in info['cov'] if cov is not None]
     ordrs = [o for o in np.arange(86) if info['params'][o] is not None]
@@ -243,23 +244,33 @@ def interp_train_and_predict(newx, newm, x, m, data, e=None,
             # Interpolate
             if interp_deg==1:
                 prediction[Inew] = np.interp(newx[Inew], ord_xs, ord_data,
-                                             left=np.nan,right=np.nan)
-            elif interp_deg=='lsq':
-                s = interpolate.UnivariateSpline(ord_xs,ord_xs,s=0)
-                t = s.get_knots()[1:-1]
-                if e is not None:
-                    f = interpolate.LSQUnivariateSpline(ord_xs,ord_data,
-                                            t,w=1/e[I][wave_sort]**2)
-                else:
-                    f = interpolate.LSQUnivariateSpline(ord_xs,ord_data,t)
+                                             left=np.nan,right=np.nan,k=interp_deg)
+            elif interp_deg == 'pchip':
+                f = interpolate.PchipInterpolator(ord_xs,ord_data,extrapolate=False)
                 prediction[Inew] = f(newx[Inew])
-            elif interp_deg==3:
-                f = interpolate.interp1d(ord_xs, ord_data,kind='cubic',
-                                     bounds_error=False,fill_value=np.nan)
-                prediction[Inew] = f(newx[Inew])
+            elif interp_deg == 'inverse':
+                f = interpolate.interp1d(ord_data, ord_xs, kind='cubic',
+                                         bounds_error=False,fill_value=0)
+                inv_f = lambda x, a: f(x)-a
+                
+                f0 = interpolate.UnivariateSpline(ord_xs,ord_data,ext=1)
+                
+                predict = np.zeros(np.sum(Inew),dtype=float)
[…]

### 2020-01-10 paper/paper.tex
Discussion Notes

@@
 
+Chosing interpolations
+monotonic vs. smooth (univariate spline stuff)
+
+perfect degenercy between definition of "Centroid of line" and "wavelength solution"
+cross correlation calls center of line something slightly different than what's happening when lines are being fit
+may lead to systematic differences
+in principle it would be preferable for the calibration source to look like a star
+
 \end{document}

### 2020-01-10 py/waveCal.py

@@ -659,6 +659,7 @@ def patchAndDenoise(file_list, file_times=None, K=2,
     patch_dict['denoised_xs'] = denoised_xs.copy()
     patch_dict['mean_xs']  = mean_x_values.copy()
     patch_dict['bad_mask'] = bad_mask.copy()
+    # PCA Information
     patch_dict['u'] = uu.copy()
     patch_dict['s'] = ss.copy()
     patch_dict['v'] = vv.copy()

### 2020-01-11 paper/paper.tex
intro

--- 2020-01-10 paper/paper.tex
@@
 There is no need to choose any functional form for the wavelength
-solution, such as an eighth-order polynomial.
+solution (such as an eighth-order polynomial, for example).
 Going non-parametric will improve calibration accuracy, because
@@
 the night or season, and only along a small number of axes in what you
-might call ``calibration space''..... HOGG
+might call ``calibration space'', or the (very high dimensional) space
+of all possible wavelength solutions.
+That is, not only are they stable, but they don't have many accessible
+degrees of freedom.
+So it is not appropriate to fit each calibration exposure or calibrate
+each science exposure independently.
+It makes sense to use the calibration data (or all the data) to
+determine the space in which the instrument can and does vary, and
+then in subsequent calibration work only permit the solution to live
+in the accessible part of calibration space.
+This structure is \emph{hierarchical}: The calibration data are used
+not just to determine the wavlength solution, but also to determine
+the possible space of wavelength solutions.
+In the statistics literature this concept is often described as
+\emph{de-noising}:
+Every calibration exposure contains information about every other
+exposure; it doesn't just constrain the instrument locally in time.
+Thus every exposure can be improved (de-noised) with information
+coming from every other exposure.
 
-HOGG: Return to the point that non-parametric \emph{increases} the
-freedom of the model (infinitely in some sense) but hierarchical
-\emph{restricts} the freedom of the model (enormously, but we hope
-appropriately).... HOGG
+The method we propose here---\name---embodies these ideas.
+It is a non-parametric, hierarchical, data-driven model for the
+wavelength solution.
+By being non-parametric, it delivers enormous freedom to the
+wavelength solution to match or adapt to any instrument or detector
+oddities.
+By being hierarchical, it restricts that freedom tremendously, but it
+does so appropriately for the observed variations in the
+spectrography.
+\name\ is designed for temperature-controlled, bench-mounted, fiber-fed
+spectrographs with good calibration sources, such as laser-frequency
+combs, etalons, or good arc lamps.
+We have in mind the \eprv\ instruments and \eprv\ science cases.
+But we expect \name\ to have applications for other kinds of
+spectrographs in other contexts.
 

### 2020-01-11 paper/paper.tex
notes to self

@@
 does so appropriately for the observed variations in the
-spectrography.
+spectrograph.
 \name\ is designed for temperature-controlled, bench-mounted, fiber-fed
@@
 But we expect \name\ to have applications for other kinds of
-spectrographs in other contexts.
+spectrographs in other contexts; the motivating ideas behind this
+project are true for almost every astronomical spectrograph.
 
 \section{Method} \label{sec:method}
-We operate on the core idea that the wavelength solution is living in a low-dimensional space, where the degrees of freedom are set by the limited degrees of freedom of the spectrograph hardware.
+We operate on the two core ideas that the wavelength solution should be given enormous flexibility, but that in fact it will be living in a very low-dimensional space, where the degrees of freedom are set by the limited kinematics of the spectrograph hardware.
 
-We will pose the problem in the following way.  Given an exposure $n$, and order $m$, there is a relationship between
+HOGG: THIS IS NOW ALL WRONG... We will pose the problem in the following way.  Given an exposure $n$, and order $m$, there is a relationship between
 the two-dimensional $(x,y)$-position on the detector and the

### 2020-01-11 paper/paper.tex
notes to self
tweaks

@@
 \newcommand{\project}[1]{\textsl{#1}}
+\newcommand{\name}{\project{Excalibur}}
 \newcommand{\acronym}[1]{{\small{#1}}}
 \newcommand{\expres}{\project{\acronym{EXPRES}}}
-\newcommand{\name}{\project{NameOfThis}}
 \newcommand{\eprv}{\acronym{EPRV}}
@@
 reducing dimensionality with a robust principal-component method.
-We demonstrate the success of this method with data from the
-\expres\ spectrograph.
+We demonstrate the success of this method with comb and arc data
+from the \expres\ spectrograph.
 We find... [results and so on].
@@
 the $K$ amplitudes $a_{nk}$ for every exposure $n$.
+
+HOGG: Say that there are many ways to do this, but we use PCA for now.
+Note that PCA is correct in the limit that everything is very high in SNR.
 

### 2020-01-20 paper/paper.tex
Paper Updates
Some slight editing, added some figures, some text added to discussion

--- 2020-01-11 paper/paper.tex
@@
 Missing line center measurements are replaced with denoised estimates using iterative PCA.
-For each missing line center measurement, we initialzie their value to the mean of the nearest $n$ exposures in time, where $n$ is typically 9 for LFC exposures and $15$ for ThAr.
+For each missing line center measurement, we initialzie their value to the mean of the nearest $n$ exposures in time, where $n$ is typically $9$ exposuresr.
 We then find a fiducial calibration of the spectrograph,  $g_0(x,y,m)$, for which we use the mean line center value for each identified line.
@@
 We can interpolate $a_{nk}$ with respect to time to determine $\lambda(x,y,m,n)$ at any time.
-Once we have the line center for each line, which has an associated known wavelength, we can interpolate the known wavelengths over line centers to any position in x in a given order.
+Once we have the line center for each line, which has an associated, known wavelength, we can interpolate the known wavelengths over line centers to any position in x in a given order.
 For example, interpolating onto every integer x will generate wavelengths for each pixel in an order.
 
-We chose to implement both interpolations with a cubic spline.  Interpolation of wavelengths over line centers is done order by order.
+We implement with respect to time using a linear spline, which is sufficient to capture the linear variations we see over a night (see \ref{fig:nightlyVariation}).
+Interpolation of wavelengths over live centers is done order by order using a Piecewise Cubic Hermite Interpolating Polynomial (PCHIP) interpolator.
+This interpolation incorporates the flexibility needed to model the changing dispersion of the spectrograph across an order while also enforcing monotonicity.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.1\textwidth]{Figures/fruit.jpg}
+\caption{Show amplitudes of principal components w.r.t.
+time to show  that a linear interpolation is reasonable.
+Should be a close up of one night or multiple nights (b/c the larger structure does not make it seem like a linear interpolation is a good idea.}
+\label{fig:nightlyVariation}
+\end{figure}
 
@@
 \subsection{Choice Choices} \label{sec:choices}
-(Before or after results presented in validation test and ThAr sections?)
+(Can use results of training/validation or cross-tests to argue for different tests?  Or should we go all the way through to RVs.
 
-We're still testing most of these (will use that as excuse to write this section later)
-\begin{itemize}
-	\item Value of K for denoising
-	\item Type of interpolation (cubic spline vs nearest neighbor or linear)
-	\item Interpolation order (interpolate time and x direction together?  Individually?  Which one first?)
-\end{itemize}
+\subsection{Value of K}
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.1\textwidth]{Figures/fruit.jpg}
+\caption{SVD eigenvectos showing sharp drop off after 2 (may be different if bluer orders are included}
+\label{fig:eigenvalues}
+\end{figure}
 
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.1\textwidth]{Figures/fruit.jpg}
+\caption{eigenvectors through the sixth one showing mostly noise following the first few}
+\label{fig:eigenvectors}
+\end{figure}
+
+\subsection{Interpolation in x Choice}
 Reason for not linear interpolation:
@@
 
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.1\textwidth]{Figures/fruit.jpg}
+\caption{We could have a figure that compares linear interpolation, cubic interpolation, and monotonic interpolation along with a histogram of the residuals}
+\label{fig:xinterpolation}
+\end{figure}
+
+Allowing for a non-zero smoothing parameter is bad as it allows for smoothness we have no reason to believe is true of the wavelength solution.
+However, restricting to a monotonic wavelength solution follows from how the optics work since we are finding the solution order by order.
+
+\subsection{Order of Interpolation}
+Now doing time then x direction.  No real reason for the switch except this makes more sense (well, to me) and is faster.
+
 \section{Results} \label{sec:results}
@@
 First, let's have a sanity check.
-Using a cubic spline scheme with no smoothing to predict wavelengths for the returned, denoised line center values, $\lambda(x,y,m,n)$, should return identical values to the assigned true wavelength for each line within machine error.
-This is what happens, for both ThArs and LFCs: look at the figure.
+Using a linear interpolator with no smoothing, we predict wavelengths for the denoised line center values, $\lambda(x,y,m,n)$, for an exposure used in the training set.
+This returns identical values to the assigned true wavelength for each line within machine error for both ThArs and LFCs (see Fig.
+\ref{fig:sanity test}).
 
-As an only slightly less silly sanity check, we can also use the constructed PCA basis to predict wavelengths for the measured line centers of an exposure that was included in the set of exposures used to construct the basis.
-This interpolation should return identical PCA coefficients for that time.
-Differences in wavelengths will arise from a combination of 1) the error inherent in measuring line centers and 2) interpolation error of lines to other x positions in the pixel direction.
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.48\textwidth]{Figures/lfcLfcDenoised.pdf}
+\includegraphics[width=0.48\textwidth]{Figures/tharTharDenoised.pdf}
+\caption{results of interpolating to denoised value (probably not necessary, more an internal check of whether the code is at a base level behaving how we would expect.}
+\label{fig:sanityTest}
+\end{figure}
 
-Difference in denoised x-values to measured ones? (translate into rv shift?)
-median and spread in errors in pixel (like 0.006) and RV (like 3.2 m/s?)
-Histograms
+We then predict wavelengths for the measured line centers for an exposure that was used to construct the basis.
+The linear interpolation in time should return identical PCA coefficients for this time.
+Differences in wavelengths will arise from a combination of 1) noise in the measured line centers and 2) any interpolation error of lines to other x positions.
+Figure \ref{fig:selfTest} shows a histogram of the residuals.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.1\textwidth]{Figures/fruit.jpg}
+\caption{include median/spread of residuals?}
+\label{fig:selfTest}
+\end{figure}
 
 \subsection{Training and Validation Tests} \label{sec:test-trainNvalid}
-A better test is to leave out some percentage of the calibration images from $\{exp_n\}$ while constructing the basis and weights with PCA.
+A more realistic test is to leave out some percentage of the calibration images from $\{exp_n\}$ while constructing the basis and weights with PCA.
 We can then use the results from that PCA to predict wavelengths for lines in the calibrations images that were left out.
@@
 
-LFC looks good, ThAr is struggling a bit
-Difference in denoised x-values to measured ones? (translate into rv shift?)
-median and spread in errors in pixel (like 0.006) and RV (like 3.2 m/s?)
-Histograms
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.1\textwidth]{Figures/fruit.jpg}
+\caption{results from training and validation super-imposed over self test probably}
+\label{fig:tNvTest}
+\end{figure}
 
@@
 Likewise, we can use ThAr exposures to predict the wavelengths of LFC lines.
-Note that errors in this test will also draw on the difference in quality of the two calibration sources being compared.
 
-LFC predicts ThAr better than ThAr can predict itself.
-ThAr predicting LFC is really messing up somehow, and I should figure out what it is (sorry!)
-
-%\subsection{ThAr Stuff?  Or Include it in Previous Subsections?}
+\begin{figure}[h]
+\centering
+\includegraphics[width=1\textwidth]{Figures/lfcgood.pdf}
+\caption{}
+\label{fig:crossTest}
+\end{figure}
 
@@
 
+To test each method, we took an LFC exposure and separated the even and odd lines.
+We then used the even lines to predict wavelengths for the odd lines and vice versa.
+For example, for the parametric model, we fit a 2D, 8th-degree polynomial to the odd lines.
+We then evaluated this polynomial at the positions of the even lines to assign them wavelengths.
+We then reversed the process by fitting a polynomial to the even lines and evaluating that polynomial at the positions of the odd lines.
+Differences between the polynomial assigned wavelengths and the assigned wavelengths of each line are shown in Figure \ref{fig:dsnMFit}.
+For the non-parametric model, cubic spline interpolation was used in place of a polynomial fit.
+
 \begin{figure}[h]
 \centering
-\includegraphics[width=0.9\textwidth]{Figures/polyval2d.pdf}
-\caption{}
+\includegraphics[width=\textwidth]{Figures/polyval2d.pdf}
+\caption{\lz{Unclear whether it is worth a direct comparison with the old EXPRES pipeline} Residuals of a parametric model.
+A 2D, 8th-degree polynomial fit was used to predict the wavelengths of each line.
+Left shows the difference between the predicted and true wavelength for each line (color) across each order.
+Right is a histogram of these residuals.
+Residuals have been translated to velocity shifts.}
 \label{fig:polyValFit}
@@
 \centering
-\includegraphics[width=0.9\textwidth]{Figures/designMatrix.pdf}
-\caption{}
+\includegraphics[width=\textwidth]{Figures/designMatrix.pdf}
+\caption{Residuals of a parametric model fit using a design matrix.
+A 2D, 8th-degree polynomial fit was used to predict the wavelengths of each line.
+Otherwise, similar to Figure \ref{fig:polyValFit}.}
 \label{fig:dsnMFit}
@@
 \centering
-\includegraphics[width=0.9\textwidth]{Figures/noHierc.pdf}
-\caption{}
+\includegraphics[width=\textwidth]{Figures/noHierc.pdf}
+\caption{Residuals of a non-parametric model.  A PCHIP cubic spline interpolator was used to predict the wavelengths of each line.  otherwise similar to Figure \ref{fig:dsnMFit}.}
 \label{fig:intpFit}
@@
 
-Generalization to etalon
+Practical statements?  How many exposures are needed to implement this method at all.  How often do exposures have to be taken throughout a night (unless that's too dependent on instrument stability)
 
-Maybe some practical statements?
-How many exposures are needed to implement this method at all.
-How often do exposures have to be taken throughout a night (unless that's too dependent on instrument stability)
+Generalization to etalon:
+All methods described here are easily adapted to use with an etalon or any other arc length for which a list of lines and their wavelengths can be constructed.
 
-Some ideas about self-calibration. Once you know the space in which the
-spectrograph lives, every single science frame gives you information about
-where you are in that space; you don't need to look at the \lfc.
+Self calibration idea:
+Once we define a calibration space that captures all possible degrees of freedom for a stabilized spectrograph, then all exposures taken with the spectrograph contain information about where the spectrograph lies in that space at the time of the exposure.
+Meaning, a science exposure can be used to determine the current state of the instrument in this calibration space.
+It would therefore be possible to self-calibrate science exposures without the use of an additional wavelength calibration source.
+
+Perfect degeneracy problem:
+With any wavelength solution, there is a perfect degeneracy between what is defined as the ``center of the line'' and the resultant wavelength solution.
+If, for example, a cross correlation method is used to extract RVs from the data, a systematic difference may be introduced depending on the definition of the center of the line.
+In principle, the best way to mitigate this error would be to use a calibration source that looks like a star.
+
+How hard would it be to go bayesian?
 

### 2020-01-21 paper/paper.tex
More Paper Updates
Fleshed out the writing some, added real figures where there were fruit bats before, added captions

@@
 
-HOGG: THIS IS NOW ALL WRONG... We will pose the problem in the following way.  Given an exposure $n$, and order $m$, there is a relationship between
+We will pose the problem in the following way.  Given an exposure $n$, and order $m$, there is a relationship between
 the two-dimensional $(x,y)$-position on the detector and the
@@
 
-If the space of all calibration possibilities is in fact $K$-dimensional (where K is a small integer, i.e 2 or 8 or thereabouts), and if the calibration variations are so
-small that we can linearize, then the function $f(x,y,m;\theta_{n})$ could
-be replaced with a tiny model
+If the space of all calibration possibilities is in fact $K$-dimensional (where K is a small integer, i.e 2 or 8 or thereabouts), and if the calibration variations are so small that we can linearize, then the function $f(x,y,m;\theta_{n})$ could be replaced with a tiny model
 \begin{equation}
@@
 where
-$g_0(x,y,m)$ is the fiducial or mean or standard calibration of the
-spectrograph,
+$g_0(x,y,m)$ is the fiducial or mean or standard calibration of the spectrograph,
 the $a_{nk}$ are scalar amplitudes,
-and the $g_k(x,y,m)$ are basis functions expressing the ``directions'' in calibration space that the spectrograph can depart from the
-fiducial calibration.
-The challenge is to learn these basis functions from the data and get
-the $K$ amplitudes $a_{nk}$ for every exposure $n$.
+and the $g_k(x,y,m)$ are basis functions expressing the ``directions'' in calibration space that the spectrograph can depart from the fiducial calibration.
 
-HOGG: Say that there are many ways to do this, but we use PCA for now.
-Note that PCA is correct in the limit that everything is very high in SNR.
+The challenge is to learn these basis functions from the data and get the $K$ amplitudes $a_{nk}$ for every exposure $n$.
+There are many ways to discern these basis functions.
+Here, we present a model using principal component analysis (PCA).
+This is the correct thing to do in the limit where we have very high SNR, as is undeniably the case with calibration images.
 
@@
 We present results using data from \expres\, the EXtreme PRecison Spectrograph.
-\expres\ is an envionmentally stabilized, fiber-fed, $R=137,000$, optical spectrograph (\lz{CJ and RB's papers}).
+\expres\ is an environmentally-stabilized, fiber-fed, $R\sim137,000$, optical spectrograph (\lz{CJ and RB's papers}).
 \expres\ has two different wavelength calibration sources, a ThAr lamp and a Menlo Systems laser frequency comb (LFC, e.g.
@@
 LFC exposures are interspersed between science exposures every ~15 minutes.
-(Discussion of number of orders covered or number of lines in LFC and/or ThAr?)
 
-\lz{Discussion of line fitting?  Though the current line fitting causes me recurring nightmares.  I should really take the time to look into at least the peak fitting more.}
+The method and results discussed here are based off of 1260 LFC exposures and ThAr exposures over 90 days.
+Exposures were taken whenever EXPRES was assigned time and so are randomly distriubted throughout the nine days
 
-For the results presented in this paper, we work with the 1D, optimally extracted data.  This removes the y-dependence from equations 1 and 2.
+\lz{(Discussion of number of orders covered or number of lines in LFC and/or ThAr?)}
 
-\expres\ exposures are separated into different ``instrumental epochs," which correspond to changes to the position of the echellogram on the detector, the shape of the instrumental PSF, or the calibration sources (\lz{(worth including table that explicitly defines epochs?)}).
+For the results presented in this paper, we work with the 1D, optimally extracted data.
+This removes the y-dependence from equations 1 and 2.
+Gaussian profiles are fit to each line in the 1D, extracted data.
+The mean of each fitted Gaussian is taken to be the center of the line.
+This process is described in detail in \lz{RP Paper} and considered out of scope of this paper.
+The method described proceeds from a given list of orders, line wavelengths, and line centers for each exposure.
+
+\expres\ exposures are separated into different ``instrumental epochs," which correspond to atypical, large instrument changes that affected the position of the echellogram on the detector, the shape of the instrumental PSF, or the calibration sources (\lz{can cite either Ryan's paper for more detail?}).
 Because significant instrumental changes break the assumption of only low-dimensional variations in the spectrograph hardware, we treat each instrument epoch independently.
@@
 \subsection{Hierarchical De-Noising of Calibration Frames} \label{sec:denoising}
-Let $\{exp_n\}$ be the set of calibration exposures within an instrument epoch used to construct $g_k(x,y,m)$.
+Let $\{exp_n\}$ be the set of calibration exposures within an instrument epoch.
+These are the exposures to be used in constructing $g_k(x,y,m)$.
 We first identify the full list of lines we can expect to find in a calibration image from this epoch.
 A line is uniquely defined by a combination of order, $m$, and ``true" or theoretical wavelength of that line, $\lambda$.
-Let $\{l_{m,\lambda}\}$ be the complete list of unique lines found in $\{exp_n\}$.
+We define $\{l_{m,\lambda}\}$ to be the complete list of unique lines found in $\{exp_n\}$.
 
-For each line, $l_{m,\lambda}$, we then find the fitted line center, $x_{n,l}$, in pixel space, for that line for each exposure.
+For each line, $l_{m,\lambda}$, we then find the fitted line center, $x_{n,l}$, in pixel space, for that line in each calibration exposure.
 If this line is missing from an exposure (e.g.
@@
 
-Lines that are missing from greater than some percent of exposures are cut.
-Exposures that are missing greater than some percent of lines are also cut (\lz{This will miss low signal exposures that suffer most in the first few orders but can recover in later orders; can't decide if this is a feature or a bug}).
-This threshold percentage is 50\% for LFC exposures and 30\% for ThAr exposures.
+Lines that are missing from greater than some percent of exposures are completely cut from the analysis.
+Exposures that are missing greater than some percent of lines are also cut.
+We use a threshold percentage of 50\% for both lines and exposures.
+This largely affects the first lines on either end of an order and exposures with spuriously low signal.
+Typically, about 0.2\% of lines and cut for LFC based calibration and 0.25\% of files are cut.
 
 Missing line center measurements are replaced with denoised estimates using iterative PCA.
-For each missing line center measurement, we initialzie their value to the mean of the nearest $n$ exposures in time, where $n$ is typically $9$ exposuresr.
+For each missing line center measurement, we initialize their value to the mean of the nearest $n$ exposures in time, where $n$ is typically $9$ exposures.
 We then find a fiducial calibration of the spectrograph,  $g_0(x,y,m)$, for which we use the mean line center value for each identified line.
-We subtract this fiducial calibration from each exposure's measured line centers and run PCA on the difference between the measured and fiducial line center for each exposure.
-This PCA gives us $a_{nk}$ and $g_k(x,y,m)$.
+We subtract this fiducial calibration from each exposure's measured line centers and run PCA on the difference between the measured and mean line center for each exposure.
+This process gives us $a_{nk}$ and $g_k(x,y,m)$.
 
 Using those values with equation 2 and some small integer (e.g.
-2) for K, we can reconstruct a de-noised matrix of line centers for each exposure.
-Missing line center measurements are then replaced with values from this reconstruction.
-We do this iteratively until the values of the missing line centers change by less than 0.01\% of a pixel.
+2) for $K$, we can reconstruct a de-noised matrix of line centers for each exposure.
+Missing line center measurements are replaced with values from this reconstruction.
+We iterate until the values of the missing line centers change by less than 0.01\% of a pixel.
 This convergence suggests that the PCA is no longer being swayed by outliers and is truly ``de-noised."
@@
 \SetAlgoLined
-\textbf{Inputs}: \;
-\While{change in bad pixels $>$ 0.01\%}{
+\textbf{Inputs}: ${m,\lambda}$, $\{l_{m,\lambda}\}$ \;
+\While{change in bad lines $>$ 0.01\%}{
 	$g_0(x,y,m) = \overline{\{x_{n,l}\}}$\;
@@
 	$\lambda(x,y,m,n) = g_0(x,y,m) + \sum_{k=1}^K a_{nk}\,g_k(x,y,m)$ where $K=2$\;
-	$\{x_{n,l}\} = \lambda(x,y,m,n)$ where $\{x_{n,l}\}$ is $NaN$
+	$\{x_{n,l}\} = \lambda(x,y,m,n)$ where $\{x_{n,l}\}$ was initially $NaN$
 	}
@@
 
+Following the de-noising process, we check for deviant line centers.
+This is aimed at catching lines that have been mis-identified or poorly fit, resulting in erroneous line centers.
+Outliers are identified via a simple sigma cut for each line.
+Any line center that is reported at more than $3\sigma$ away from the mean line center value is relabeled a ``bad line,'' and we repeat the process of hierarchical de-noising.
+(\lz{Should probably add something about why the sigma cut works})
+
 \subsection{Non-Parametric Calibration} \label{sec:nonparam}
-We can interpolate $a_{nk}$ with respect to time to determine $\lambda(x,y,m,n)$ at any time.
-Once we have the line center for each line, which has an associated, known wavelength, we can interpolate the known wavelengths over line centers to any position in x in a given order.
+The process of hierarchical de-noising gives us $a_{nk}$, which captures the variability of the instrument.
+We can interpolate $a_{nk}$ with respect to housekeeping data to determine $\lambda(x,y,m,n)$ at any time.
+In this paper, we show results where $a_{nk}$ is interpolated with respect to time, but it is possible to incorporate other information such as temperature, pressure, or telescope position.
+(\lz{Does this belong in the discussion?})
+
+Once we have the line center for every line in $\{l_{m,\lambda}\}$, each of which is associated with a known wavelength, we can interpolate the known wavelengths over line centers to any position in x in a given order.
 For example, interpolating onto every integer x will generate wavelengths for each pixel in an order.
-
-We implement with respect to time using a linear spline, which is sufficient to capture the linear variations we see over a night (see \ref{fig:nightlyVariation}).
-Interpolation of wavelengths over live centers is done order by order using a Piecewise Cubic Hermite Interpolating Polynomial (PCHIP) interpolator.
-This interpolation incorporates the flexibility needed to model the changing dispersion of the spectrograph across an order while also enforcing monotonicity.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.1\textwidth]{Figures/fruit.jpg}
-\caption{Show amplitudes of principal components w.r.t.
-time to show  that a linear interpolation is reasonable.
-Should be a close up of one night or multiple nights (b/c the larger structure does not make it seem like a linear interpolation is a good idea.}
-\label{fig:nightlyVariation}
-\end{figure}
 
@@
 \subsection{Choice Choices} \label{sec:choices}
-(Can use results of training/validation or cross-tests to argue for different tests?  Or should we go all the way through to RVs.
+Implementing \name required choosing some global variables and methods that we believe is optimal for constructing a high-fidelity wavelength solution.
+We discuss these choices here and how we came to each decision.
 
 \subsection{Value of K}
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.1\textwidth]{Figures/fruit.jpg}
-\caption{SVD eigenvectos showing sharp drop off after 2 (may be different if bluer orders are included}
-\label{fig:eigenvalues}
-\end{figure}
+The value of $K$ determines how many principal components are used to reconstruct de-noised line centers.
+This $K$ needs to be large enough that all variability in the spectrograph is captured.
+Too large, and the reconstruction will begin to incorporate noise, defeating the purposes.
+
+We settled on a $K$ value of 2.
+Figure \ref{fig:eigenvectors} shows the eigenvectors of the PCA plotted on a log scale for both a LFC-based calibration scheme (blue), and a ThAr-based one (green).
+In both cases, there is a precipitous drop off at one, indicating that the later principal components have a much smaller contribution.
 
@@
 \centering
-\includegraphics[width=0.1\textwidth]{Figures/fruit.jpg}
-\caption{eigenvectors through the sixth one showing mostly noise following the first few}
+\includegraphics[width=0.5\textwidth]{Figures/ss.pdf}
+\caption{Eigenvectors from PCA as step functions.
+The results for using only LFC exposures (blue) and only using ThAr exposures (green) are both shown on a log scale.
+After 1, there is a order of magnitude drop for both the LFC and the ThAr.}
 \label{fig:eigenvectors}
@@
 
-\subsection{Interpolation in x Choice}
-Reason for not linear interpolation:
+Figures \ref{fig:pcLfc} and \ref{fig:pcThar} shows the first 6 principal components for the LFC and ThAr respectively.
+In both cases there is clear structure in the first and second principal components.
+It is doubly encouraging that this structure appears the same in both the case with the LFC and the ThAr.
+
+The later principal components show mostly noise or are swayed by large noise sources.
+For example, the band on the left side of each order in principal component 3 for the LFCs is likely due to lower signal on the edges of the order from the blaze function, giving more uncertain line center measurements.
+The diagonal banding visible in the sixth principal component can be traced to what we believe are detector defects.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=\textwidth]{Figures/pcsLfc6.png}
+%\includegraphics[width=\textwidth]{Figures/pcsLfc6.pdf}
+\caption{The first six principal components using LFC exposures.
+There is clear structure for the first two principal components.
+\lz{Principal components 3-5 show mostly noise (do they though?)}.
+Principal component 6 shows structure from what we believe to be detector defects.}
+\label{fig:pcLfc}
+\end{figure}
+
+\begin{figure}[t!]
+\centering
+\includegraphics[width=\textwidth]{Figures/pcsThar6.png}
+%\includegraphics[width=\textwidth]{Figures/pcsThar6.pdf}
+\caption{The first six principal components using ThAr exposures.  Note the similarities with the LFC result shown in Figure \ref{fig:pcLfc}.}
+\label{fig:pcThar}
+\end{figure}
+
+\subsection{Interpolation in Time}
+Figure \ref{fig:nightlyVariation} shows the amplitude of the first and second principal component with respect to time on the left.
+Again, it is comoforting to note that the LFC and ThAr versions show the same shape.
+Though there are no large trends over the ninety days of data shown here, there is a clear linear trend within each night (see Fig.
+\ref{fig:nightlyVariation} right).
+We therefore interpolate these amplitudes with respect to time using a simple linear interpolation scheme.
+
+\lz{linear interpolation actually doesn't look super great for the second principal component}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.56\textwidth]{Figures/pcAs.pdf}
+\includegraphics[width=0.42\textwidth]{Figures/pcAs_byDay.pdf}
+\caption{Amplitude of the first two principal components show as a function of time (left) or fraction of a day (right).
+Points mark the times at which we have calibration data.
+Lines show the result of a linear interpolation.
+Left: amplitudes as a function of time for both LFC (blue) and ThAr (green).
+The top plot shows the amplitudes for the first prinicpal component while the bottom plot shows the amplitudes for the second principal component.
+Note that the shape of the curves is similar regardless of calibration source.
+Recall from Figures \ref{fig:pcLfc} and \ref{fig:pcThar} that the corresponding principal components are also fairly similar.
+Right: amplitudes of the first two principal components determined by the LFC exposures only.
+Here exposures are grouped by day and plotted with respect to what time of day the exposure was taken.
+A linear interpolation through all the LFC exposures taken in one day are shown as black line.
+Points are colored by the MJD of the exposure.}
+\label{fig:nightlyVariation}
+\end{figure}
+
+\subsection{Interpolation in Pixel}
+Interpolation of wavelengths over line centers is done order by order using a Piecewise Cubic Hermite Interpolating Polynomial (PCHIP) interpolator.
+This interpolation incorporates the flexibility needed to model the changing dispersion of the spectrograph across an order while also enforcing monotonicity.
+
 Due to the dispersion intrinsic to echelle spectrographs, the wavelength change between pixels is greater at greater wavelengths.
@@
 
+A more classic cubic spline interpolator runs into issues with lines in the ThAr that are blended and therefore very close together.
+Very close lines cause huge deviations from the correct wavelengths, as seen in Figure \ref{fig:xinterp}, which highlights deviations caused by a blend in a specific order of ThAr exposures.
+
 \begin{figure}[h]
 \centering
-\includegraphics[width=0.1\textwidth]{Figures/fruit.jpg}
-\caption{We could have a figure that compares linear interpolation, cubic interpolation, and monotonic interpolation along with a histogram of the residuals}
-\label{fig:xinterpolation}
+\includegraphics[width=\textwidth]{Figures/intpx_tests.pdf}
+\caption{Results from different interpolation schemes in the pixel direction.
+Left: resultant wavelength solution curves when using ThAr lines (green points) to predict wavelengths for LFC lines (black points).
+The green points are being used to predict values for the black points.
+Different interpolation schemes are shown in different colors.
+The largest deviation comes with the classic cubic spline (pink).
+The other models are hard to distinguish.
+Right: histograms of residuals for each method.
+The linear interpolation model (yellow) is asymmetric with disproportionately positive residuals.
+The classic cubic spline residuals (pink) has very spread out residuals, the extremes of which have been cut off in the figure.
+The residuals from an interpolation scheme that incorporates smoothing (blue) are non-Gaussian and rather plateau about zero.
+Residuals from the PCHIP interpolator (orange) has the strongest peak at 0 and is roughly Gaussian.}
+\label{fig:xinterp}
 \end{figure}
 
-Allowing for a non-zero smoothing parameter is bad as it allows for smoothness we have no reason to believe is true of the wavelength solution.
-However, restricting to a monotonic wavelength solution follows from how the optics work since we are finding the solution order by order.
+These huge digressions can be avoided by allowing for some smoothing in the interpolation.
+In Figure \ref{fig:xinterp}, we show an example in blue using scipy's Univarate Spline method.
+While the result appears to follow the calibration lines much better, the smoothing ultimately causes larger residuals that are spatially correlated (Fig.
+\ref{fig:xinterp}, right).
+
+For all orders, the edges will be overestimated while the middle will be underestimated.
+(\lz{Do I need to include a figure for this?
+Or should I just not go into this much detail})  This indicates that the smoothing is causing the resultant wavelength solution to underestimate the curvature of the pixel-wavelength relation.
+It also reduces the benefit we get from moving to a non-parametric model by enforcing a smoothness we have no reason to believe is true in the data.
+
+We instead turn to the PCHIP interpolator, which damps down huge deviations in the traditional cubic spline by requiring the resulting interpolated function to be monotonic.
+This is a constraint we know is true from the optics since interpolation in the pixel direction is carried out order-by-order.
+From Figure \ref{fig:xinterp}, it is clear that the PCHIP interpolator returns the lowest residuals.
 
 \subsection{Order of Interpolation}
-Now doing time then x direction.  No real reason for the switch except this makes more sense (well, to me) and is faster.
+Now doing time then x direction.  \lz{No real reason for the switch except this makes more sense (well, to me) and is faster.}
 
@@
 First, let's have a sanity check.
-Using a linear interpolator with no smoothing, we predict wavelengths for the denoised line center values, $\lambda(x,y,m,n)$, for an exposure used in the training set.
+(\lz{Is it worth mentioning this at all?}) Using a linear interpolator with no smoothing, we predict wavelengths for the denoised line center values, $\lambda(x,y,m,n)$, for an exposure used in the training set.
 This returns identical values to the assigned true wavelength for each line within machine error for both ThArs and LFCs (see Fig.
@@
 \includegraphics[width=0.48\textwidth]{Figures/tharTharDenoised.pdf}
-\caption{results of interpolating to denoised value (probably not necessary, more an internal check of whether the code is at a base level behaving how we would expect.}
+\caption{Residuals for predicted wavelengths for the denoised x-values of an exposure used to train the PCA.
+As expected, the residuals are zero within machine error for the case where LFCs are used and where ThArs are used.}
 \label{fig:sanityTest}
@@
 
-We then predict wavelengths for the measured line centers for an exposure that was used to construct the basis.
+We then predict wavelengths for the \textit{measured} line centers for an exposure that was used to construct the basis.
 The linear interpolation in time should return identical PCA coefficients for this time.
 Differences in wavelengths will arise from a combination of 1) noise in the measured line centers and 2) any interpolation error of lines to other x positions.
-Figure \ref{fig:selfTest} shows a histogram of the residuals.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.1\textwidth]{Figures/fruit.jpg}
-\caption{include median/spread of residuals?}
-\label{fig:selfTest}
-\end{figure}
+Figure \ref{fig:testHists} shows a histogram of the residuals for both the LFC (left) and the ThAr (right) in black.
 
 \subsection{Training and Validation Tests} \label{sec:test-trainNvalid}
-A more realistic test is to leave out some percentage of the calibration images from $\{exp_n\}$ while constructing the basis and weights with PCA.
-We can then use the results from that PCA to predict wavelengths for lines in the calibrations images that were left out.
-We will call the used calibration images the ``training set," and the left out exposures the ``validation set."
+A more realistic test is to leave out some percentage of the calibration images from $\{exp_n\}$ while constructing the basis vectors and weights with PCA.
+We then use the results from this PCA to predict wavelengths for lines in the calibrations images that were left out.
+The used calibration images represent the ``training set," while the left out exposures make up the ``validation set."
 
@@
 Exposures were sorted into either set at random.
-This will test how well PCA coefficients are interpolated to other times outside of the training set and therefore how well line centers can be reconstructed at all times.
+\lz{(Worth including?) Because only two ThAr exposures were taken each night (at the very beginning and end), leaving ThAr exposures out often left just one ThAr exposure per night.
+It was therefore impossible to run a similar test using just ThAr exposures.}
+
+We are now also testing how well PCA coefficients are interpolated to other times outside of the training set, and, consequently, how well line centers can be reconstructed for all times.
 As before, differences in wavelength will also be drawn from how well line centers are measured and error in interpolating to those measured values.
@@
 
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.1\textwidth]{Figures/fruit.jpg}
-\caption{results from training and validation super-imposed over self test probably}
-\label{fig:tNvTest}
-\end{figure}
-
 \subsection{Cross Tests} \label{sec:test-cross}
 Similar to the training and validation tests, we can also use one calibration source to construct the PCA basis and weights and the other to test the results.
-For example, we can use LFC exposures to construct the wavelength solution, and then use it to predict wavelengths for measured line centers of ThAr lines, which will have their own assigned true wavelengths.
+We used LFC exposures to construct a wavelength solution, and then used it to predict wavelengths for measured line centers of ThAr lines.
+We can then use the assigned wavelengths for each ThAr line to assess how well we predicted the wavelengths for these lines.
 Likewise, we can use ThAr exposures to predict the wavelengths of LFC lines.
@@
 \includegraphics[width=1\textwidth]{Figures/lfcgood.pdf}
-\caption{}
-\label{fig:crossTest}
+\caption{Residuals from different wavelength calibration tests.
+Left: residuals of wavelength predictions for LFC lines.
+In black are the results from using all LFC exposures to construct a wavelength solution, and using it to predict wavelengths for the measured line centers of each exposure.
+Over-plotted in blue are the residuals from the training and validation tests.
+Right: residuals of wavelength predictions for ThAr lines or using ThAr lines.
+The black histogram is the same as in the left plot, but with ThAr exposures instead of LFC exposures.
+In blue is the residuals from using the LFC exposures to construct a wavelength solution that is then used to predict ThAr line wavelengths.
+In green is using ThAr exposures to predict LFC line wavelengths.
+}
+\label{fig:testHists}
 \end{figure}
 
-\section{Comparisons to Other Methods} \label{sec:comparisons}
-I think we probably have to use the even/odd method to compare.
+Results from these tests are shown in Figure \ref{fig:testHists}.
+The results of the self tests for LFC (left) and ThAr (right) exposures are shown as black outlines.
+This histogram combines residuals from all exposures.
+The self test represents the limit of how well the line centers at any given time can be predicted, since here we are predicting the line centers for an exposure used in the construction of $a_{nk}$.
+\lz{How to quantify error from measurement of line center}.
 
-\subsection{Comparison to Parametric Model}
-Previously, the \expres\ pipeline the $\theta_{n}$ from equation 1 comprises the
-amplitudes of some 9th-ish-order polynomial in $x$ and $m$ for each $n$.
-If so, they have
+Overplotted in the left plot of Figure \ref{fig:testHists} are the residuals from training the wavelength solution with only 90\% of LFC exposures and predicting the wavelengths for the remaining 10\% of exposures.
+The two histograms match very closely, indicating that very little error is introduced when predicting denoised x values for times not used in the construction of the wavelength solution.
+
+The right plot in \ref{fig:testHists} shows results involving the ThAr exposures.
+Similar to with the LFC, the histogram of the ThAr self tests is shown in black, and provide a limit on how well lines at different times can be predicted.
+If we use the LFCs, all taken at different times, to predict the ThAr lines, we see that again the histograms match very closely.
+This confirms that interpolating $a_{nk}$ w.r.t.
+time is introducing little error.
+
+Overplotted in green is the result from using ThAr lines to predict LFC line wavelengths.
+These residuals show a scatter comparable to other results where the wavelength value of ThAr lines are being predicted.
+This implies that using the ThAr to predict wavelengths is dominated by uncertainties in the measured line centers for the ThAr.
+\lz{why is it skewed left a little??}
+
+\section{Parametric Model} \label{sec:comparisons}
+Classically, pipelines employ polynomials to construct smooth wavelength solutions for each exposure.
+For example, the \expres\ pipeline described in \lz{Petersburg+} found the $\theta_{n}$ from equation 1 to be the amplitudes of some 2D, 9th-order polynomial in $x$ and $m$ for each $n$, making the equivalent to equation 2 the following:
 \begin{equation}
@@
 \end{equation}
-where the $c_{nij}$ are coefficients unique to exposure $n$ but smoothed by a third order polynomial over time.
-Further, they were finding the coefficients $c_{nij}$ that
-minimize an objective $Q$ that is something like the L2-norm:
+where the $c_{nij}$ are coefficients unique to exposure $n$.
+These coefficients are then smoothed by a third-order polynomial over time.
+This same polynomial was used to construct wavelengths for non-calibration exposures.
+Coefficients, $c_{nij}$, were found that minimize an objective $Q$ that is something like the L2-norm:
 \begin{equation}
@@
 
-Even in this simple context, where every calibration image $n$ is
-treated as its own unique flower, there are improvements to be
-made.
+Even in this simple context, where every calibration image $n$ is treated as its own unique flower, there are improvements to be made.
 For one, the sums should not be from 0 to 9, but instead
@@
 because that is the definition of 9th order.  We find that fitting to 8th order is sufficient, faster, and leaves less residual pattern than the original 9th degree fit.
-For another, the objective function could be made soft to permit
-catastrophic outliers without destroying the fit. (\lz{we did not do this})
+
+For another, the objective function could be made soft to permit catastrophic outliers without destroying the fit.
+(\lz{we did not do this}).
 We might recommend the iteratively reweighted least squares (IRLS).
 This would make the fitting more robust.
-For yet another, there are rescaling issues for the products $x^i\,m^j$ to
-protect the fitting from near-singularities or bad conditioning of the
-linear-algebra operators.
 
-To test each method, we took an LFC exposure and separated the even and odd lines.
-We then used the even lines to predict wavelengths for the odd lines and vice versa.
+For yet another, there are rescaling issues for the products $x^i\,m^j$ to protect the fitting from near-singularities or bad conditioning of the linear-algebra operators.
+
+\subsection{Comparisons}
+To test each method, we took an LFC exposure and separated the even and odd lines.  We then used the even lines to predict wavelengths for the odd lines and vice versa.
+
 For example, for the parametric model, we fit a 2D, 8th-degree polynomial to the odd lines.
@@
 Differences between the polynomial assigned wavelengths and the assigned wavelengths of each line are shown in Figure \ref{fig:dsnMFit}.
-For the non-parametric model, cubic spline interpolation was used in place of a polynomial fit.
+
+We run the test for both the original method described in \lz{Petersburg+}, but with an 8th degree polynomial rather than 9th, and changing the polynomial for the revised description of a 2D polynomial given in equation 5.
+In the second case, the polynomial is also fit using a design matrix rather than by minimizing an objective (\lz{What's the fancier way of saying this}).
+For the non-parametric model, we repeat the test using a cubic spline interpolator in place of a polynomial fit.
+
+The results are shown in Figures \ref{fig:polyValFit}, \ref{fig:dsnMFit}, and \ref{fig:intpFit}.
+Residuals for each line plotted with respect to order and x-pixel are shown on the left while the histogram of the residuals are shown on the right.
+Moving from \lz{Petursburg+} to a design matrix polynomial fit with half the terms already reduces some of the correlated structure in the residuals that can be seen on the left in Figure \ref{fig:polyValFit}.
+However, the same vertical stripes can be seen in the residuals for both methods.
+This along with other evidence (outside the scope of this paper) suggests that these vertical stripes are due to defects present in the detector.
+
+Using a non-parametric model, i.e.
+a cubic spline, gets rid of all correlated structure in the residuals and results in smaller residuals that are better centered around zero.
+The non-parametric model was able to account for the high-order detector defects that are smoothed over by a polynomial fit, allowing for a more accurate fit.
+
+\lz{Could (should) combine these three plots better.}
 
@@
 \centering
-\includegraphics[width=\textwidth]{Figures/polyval2d.pdf}
+\includegraphics[width=\textwidth]{Figures/polyval2d.png}
+%\includegraphics[width=\textwidth]{Figures/polyval2d.pdf}
 \caption{\lz{Unclear whether it is worth a direct comparison with the old EXPRES pipeline} Residuals of a parametric model.
@@
 \centering
-\includegraphics[width=\textwidth]{Figures/designMatrix.pdf}
+\includegraphics[width=\textwidth]{Figures/designMatrix.png}
+%\includegraphics[width=\textwidth]{Figures/designMatrix.pdf}
 \caption{Residuals of a parametric model fit using a design matrix.
-A 2D, 8th-degree polynomial fit was used to predict the wavelengths of each line.
+A 2D, 8th-degree polynomial fit of the form given in equation 5 was used to predict the wavelengths of each line.
 Otherwise, similar to Figure \ref{fig:polyValFit}.}
@@
 
-\subsection{Comparison to Non-Parametric, Non-Hierarchical Model}
-
 \begin{figure}[h]
 \centering
-\includegraphics[width=\textwidth]{Figures/noHierc.pdf}
+\includegraphics[width=\textwidth]{Figures/noHierc.png}
+%\includegraphics[width=\textwidth]{Figures/noHierc.pdf}
 \caption{Residuals of a non-parametric model.  A PCHIP cubic spline interpolator was used to predict the wavelengths of each line.  otherwise similar to Figure \ref{fig:dsnMFit}.}

### 2020-04-16 paper/paper.tex
Updated Paper Draft
It was quite remarkable the breadth of change required from what was drafted before.

--- 2020-01-12 paper/paper.tex
@@
 \paragraph{Abstract:}
-There have been many hardware improvements in polygonal-fiber-fed,
-temperature-controlled, laser-frequency-comb or etalon-calibrated,
-high-resolution, extreme-precision radial-velocity spectrographs.
-Have the calibration methodologies improved to match?
-Here are three relevant observations:
-The first is that the calibration lines or spots from an etalon or
-comb (or even arc lamp) fill the spectral range with dense calibration points.
-The second is that the spectrograph lives in a stabilized,
-climate-controlled environment, in which the full optical system and
-detectors will only vary within a tiny range of configurations or
-settings.
-The third is that, given this stability, every calibration image
-ever taken is relevant to every science exposure ever taken; there is
-no reason to calibrate every exposure independently.
-The calibration methodology we propose here---\name---addresses these
-three problems by going non-parametric (no more polynomials!) and then
-reducing dimensionality with a robust principal-component method.
-We demonstrate the success of this method with comb and arc data
-from the \expres\ spectrograph.
-We find... [results and so on].
-
+\name is a hierarchical, non-parametric framework for wavelength calibrating precision spectrographs.
+There have been many recent hardware improvements in highly-stabilized, extreme-precision radial-velocity spectrographs.
+In such stabilized spectrographs, the full optical system and detectors will only vary within a tiny range of configurations.
+Additionally, within a given range of stability, every calibration image ever taken contains information relevant to every science exposure ever taken.
+The development of laser-frequency combs or etalons allows the spectral range of an instrument to be covered with dense calibration points.
+Here, we describe a calibration method---\name---that is designed for this new era of stabilized instruments with high-line-density calibrators.
+We find that a high density of calibration lines allows for the freedom of a non-parametric wavelength solution that can match or adapt to any instrument or detector oddities.
+We can also denoise the available calibration data by using it to determine not just the wavelength solution, but also the possible space of wavelength solutions, which is small for a stabilized instrument.
+We demonstrate the success of this method with data from a Menlo Systems laser frequency comb taken by \expres, the Extreme Precision Spectrograph.
+We find that \name\ returns more accurate wavelength predictions than classic polynomial models.
 \section{Introduction}
@@
 $3000\,\mps$.
-Exoplanet science happens at $3\,\mps$ precisions, and the current
-instruments hope to get to $0.1\,\mps$ precisions (CITE THINGS).
-This means that new spectrographs are expected to be calibrated or
+Exoplanet science happens at $3\,\mps$ precisions, with current
+instruments hoping to reach $0.1\,\mps$ precision (CITE THINGS).
+This requires new spectrographs to be calibrated or
 stabilized to better than $10^{-4}$ of a pixel (assuming that the
 spectrograph is well sampled).
-And indeed, some hardware--software systems seem to be reaching these
-levels (CITE THINGS).
+Some hardware--software systems seem to be reaching these
+levels (CITE THINGS) .
 
@@
 include lamps, etalons, and laser-frequency combs (\lfc
-s)---illuminate the spectrograph with dense sets of lines; almost every
+s)---illuminate the spectrograph with very stable, very dense sets of lines; almost every
 location in the spectrograph image plane is surrounded by nearby,
@@
 If every point in the spectrograph detector is surrounded by nearby
-calibration lines, the wavelength solution can be made simply an
-interpolation of the calibration data.
-There is no need to choose any functional form for the wavelength
-solution (such as an eighth-order polynomial, for example).
-Going non-parametric will improve calibration accuracy, because
-choosing a parametric form biases the calibration, and biases it
-badly when the form is inappropriate (as, for example, polynomials
-will be at the detector edges).
+calibration lines, the wavelength solution can, for example,  be made simply an interpolation of the calibration data.
 
-The second idea here flows from the observation that contemporary
+We expand on the work of Milakovi\'{c}2020, who proposed constructing a wavelength solution as multiple, segmented polynomials across an order by proposing a wavelength solution that is completely non-parametric.
+\lz{Should work in their paper more and praise it more?}
+
+The density of lines removes the need to enforce any functional form for the wavelength solution (such as an eighth-order polynomial, for example).
+Going non-parametric will improve calibration accuracy because
+choosing a parametric form biases the calibration, especially when the chosen function is inappropriate (as, for example, polynomials will be at detector edges).
+
+The second idea flows from the observation that contemporary
 \eprv\ instruments are incredibly stable.
@@
 of all possible wavelength solutions.
-That is, not only are they stable, but they don't have many accessible
-degrees of freedom.
-So it is not appropriate to fit each calibration exposure or calibrate
+That is, not only are the spectrographs stable, but they also have few accessible degrees of freedom.
+This renders it inappropriate to fit each calibration exposure or calibrate
 each science exposure independently.
-It makes sense to use the calibration data (or all the data) to
-determine the space in which the instrument can and does vary, and
-then in subsequent calibration work only permit the solution to live
-in the accessible part of calibration space.
-This structure is \emph{hierarchical}: The calibration data are used
-not just to determine the wavlength solution, but also to determine
+Instead, all the calibration data (or all the data) should be used to
+determine the space in which the instrument can and does vary.  Subsequent calibration work then need only determine where in the accessible part of calibration space the spectrograph existed.
+
+This structure is \emph{hierarchical}: the calibration data are used
+not just to determine the wavelength solution, but also to determine
 the possible space of wavelength solutions.
 In the statistics literature this concept is often described as
-\emph{de-noising}:
+\emph{de-noising}.  
 Every calibration exposure contains information about every other
-exposure; it doesn't just constrain the instrument locally in time.
-Thus every exposure can be improved (de-noised) with information
-coming from every other exposure.
+exposure.  Thus every exposure can be improved (de-noised) with information coming from every other exposure.
 
@@
 spectrograph.
-\name\ is designed for temperature-controlled, bench-mounted, fiber-fed
+\name\ is designed for temperature-controlled, fiber-fed
 spectrographs with good calibration sources, such as laser-frequency
 combs, etalons, or good arc lamps.
-We have in mind the \eprv\ instruments and \eprv\ science cases.
-But we expect \name\ to have applications for other kinds of
-spectrographs in other contexts; the motivating ideas behind this
+We have in mind the \eprv\ instruments and \eprv\ science cases, but we expect \name\ to have applications for other kinds of
+spectrographs in other contexts.  The motivating ideas behind this
 project are true for almost every astronomical spectrograph.
@@
 \section{Method} \label{sec:method}
-We operate on the two core ideas that the wavelength solution should be given enormous flexibility, but that in fact it will be living in a very low-dimensional space, where the degrees of freedom are set by the limited kinematics of the spectrograph hardware.
+We operate on the two core ideas that the wavelength solution should be given enormous flexibility, but that it lives in a very low-dimensional space, where the degrees of freedom are set by the limited kinematics of the stabilized, spectrograph hardware.
 
-HOGG: THIS IS NOW ALL WRONG... We will pose the problem in the following way.  Given an exposure $n$, and order $m$, there is a relationship between
+We will pose the problem in the following way.  Given an exposure $n$, and order $m$, there is a relationship between
 the two-dimensional $(x,y)$-position on the detector and the
@@
 \end{equation}
-where $\theta_{n}$ is a big blob of parameters for this exposure.
+where $\theta_{n}$ represents the parameters describing a given exposure.
 
 A stabilized spectrograph should experience only low-degree variability, meaning the calibration of any image can be informed by the calibration of every other image.
-To implement such a hierarchical model, we use the calibration data themselves to develop a low-dimensional basis for expressing the calibration data.
+To implement such a hierarchical model, we can use the calibration data themselves to develop a low-dimensional basis for expressing the calibration data.
 
-If the space of all calibration possibilities is in fact $K$-dimensional (where K is a small integer, i.e 2 or 8 or thereabouts), and if the calibration variations are so
-small that we can linearize, then the function $f(x,y,m;\theta_{n})$ could
-be replaced with a tiny model
+If the space of all calibration possibilities is in fact $K$-dimensional (where K is a small integer, i.e 2 or 8 or thereabouts), and if the calibration variations are so small that we can linearize, then the function $f(x,y,m;\theta_{n})$ could be replaced with a tiny model
 \begin{equation}
@@
 where
-$g_0(x,y,m)$ is the fiducial or mean or standard calibration of the
-spectrograph,
+$g_0(x,y,m)$ is the fiducial or mean or standard calibration of the spectrograph,
 the $a_{nk}$ are scalar amplitudes,
-and the $g_k(x,y,m)$ are basis functions expressing the ``directions'' in calibration space that the spectrograph can depart from the
-fiducial calibration.
-The challenge is to learn these basis functions from the data and get
-the $K$ amplitudes $a_{nk}$ for every exposure $n$.
+and the $g_k(x,y,m)$ are basis functions expressing the ``directions'' in calibration space that the spectrograph can depart from the fiducial calibration.
 
-HOGG: Say that there are many ways to do this, but we use PCA for now.
-Note that PCA is correct in the limit that everything is very high in SNR.
+The challenge is to learn these basis functions from the data and get the $K$ amplitudes, $a_{nk}$, for every exposure $n$.
+There are many ways to discern these basis functions.
+In this paper, we present a model using principal component analysis (PCA).
+This is the correct thing to do in the limit where we have very high SNR, as is undeniably the case with calibration images.
+
+\lz{Is this the place to mention options other than PCA? Or should we save that for the discussion?}
 
@@
 We present results using data from \expres\, the EXtreme PRecison Spectrograph.
-\expres\ is an envionmentally stabilized, fiber-fed, $R=137,000$, optical spectrograph (\lz{CJ and RB's papers}).
+\expres\ is an environmentally-stabilized, fiber-fed, $R\sim137,000$, optical spectrograph (\lz{Jurgenson2016, Blackman2020}).
 \expres\ has two different wavelength calibration sources, a ThAr lamp and a Menlo Systems laser frequency comb (LFC, e.g.
-\lz{Wilken+ 2012, Molaro+ 2013, Probst+ 2014, from RB's paper}).
+\lz{Wilken2012, Molaro2013, Probst2014}).
 ThAr exposures are taken at the beginning and end of each night.
-LFC exposures are interspersed between science exposures every ~15 minutes.
-(Discussion of number of orders covered or number of lines in LFC and/or ThAr?)
+LFC exposures are interspersed between science exposures every ~15-30 minutes.
 
-\lz{Discussion of line fitting?  Though the current line fitting causes me recurring nightmares.  I should really take the time to look into at least the peak fitting more.}
+The results and discussion presented here are based off of 1227 LFC exposures taken over two months months on 29 unique days.
+LFC lines cover 40 echelle orders, which contain 19203 lines.
+Though we primarily work with LFC data, we at times discuss applications to arc lamps.
+These results use 78 ThAr exposures that were taken over the same range of time.
+ThAr lines cover all 86 extracted orders of \expres\ and includes 5295 lines.
 
-For the results presented in this paper, we work with the 1D, optimally extracted data.  This removes the y-dependence from equations 1 and 2.
+For the results presented in this paper, we work with the 1D, optimally extracted data.
+This removes the y-dependence from equations 1 and 2.
+Gaussian profiles are fit to each line in the 1D, extracted data.
+The mean of each fitted Gaussian is taken to be the center of the line (see \lz{Petersburg2020} for more details).
+While precisely defining and determining line positions is essential to precise wavelength calibration, we consider the task out of scope for this paper.
+\name\ proceeds from a given list of orders, line wavelengths, and pixel positions of each line for each exposure.
 
-\expres\ exposures are separated into different ``instrumental epochs," which correspond to changes to the position of the echellogram on the detector, the shape of the instrumental PSF, or the calibration sources (\lz{(worth including table that explicitly defines epochs?)}).
+\expres\ exposures are separated into different ``instrumental epochs," which correspond to atypical, hardware changes that affected the position of the echellogram on the detector, the shape of the instrumental PSF, or the calibration sources.
 Because significant instrumental changes break the assumption of only low-dimensional variations in the spectrograph hardware, we treat each instrument epoch independently.
@@
 \subsection{Hierarchical De-Noising of Calibration Frames} \label{sec:denoising}
-Let $\{exp_n\}$ be the set of calibration exposures within an instrument epoch used to construct $g_k(x,y,m)$.
+\lz{Concerns: 1) how to best phrase the problem using variables?, 2) how in-depth is too in-depth?}
+
+Let $\{exp_n\}$ be the set of calibration exposures within an instrumental epoch.
+These exposures will be used in constructing $g_0(x,y,m)$ and $g_k(x,y,m)$ in equation 2.
 We first identify the full list of lines we can expect to find in a calibration image from this epoch.
-A line is uniquely defined by a combination of order, $m$, and ``true" or theoretical wavelength of that line, $\lambda$.
+A line is uniquely defined by a combination of order, $m$, and ``true" or theoretical wavelength assigned to that line, $\lambda$.
 Let $\{l_{m,\lambda}\}$ be the complete list of unique lines found in $\{exp_n\}$.
 
-For each line, $l_{m,\lambda}$, we then find the fitted line center, $x_{n,l}$, in pixel space, for that line for each exposure.
+Each line, $l_{m,\lambda}$, has a fitted line center, $x_{n,l}$, in pixel space, for each calibration exposure $exp_n$.
 If this line is missing from an exposure (e.g.
-it was covered by noise or off the edge of its usual order), we insert a $NaN$ for that line center instead.
+the fit failed due to noise, the line isn't in its usual order, etc.), we set $x_{n,l} = NaN$ for that line center in that exposure.
+This allows us to construct a $N \times M$ matrix of line positions for each of $M$ lines for each of $N$ exposures.
 
-Lines that are missing from greater than some percent of exposures are cut.
-Exposures that are missing greater than some percent of lines are also cut (\lz{This will miss low signal exposures that suffer most in the first few orders but can recover in later orders; can't decide if this is a feature or a bug}).
-This threshold percentage is 50\% for LFC exposures and 30\% for ThAr exposures.
+Lines that are missing from greater than some percent of exposures are completely cut from the analysis.
+Exposures that are missing greater than some percent of lines are also cut.
+We use a threshold percentage of 50\% for both lines and exposures.
+This largely affects the first lines on either end of an order and exposures with spuriously low signal.
+Typically, about 0.2\% of lines and cut for LFC based calibration and 0.25\% of files are cut.
 
 Missing line center measurements are replaced with denoised estimates using iterative PCA.
-For each missing line center measurement, we initialzie their value to the mean of the nearest $n$ exposures in time, where $n$ is typically 9 for LFC exposures and $15$ for ThAr.
-We then find a fiducial calibration of the spectrograph,  $g_0(x,y,m)$, for which we use the mean line center value for each identified line.
-We subtract this fiducial calibration from each exposure's measured line centers and run PCA on the difference between the measured and fiducial line center for each exposure.
-This PCA gives us $a_{nk}$ and $g_k(x,y,m)$.
+For each missing line center measurement, we initialize its value to the mean of the nearest $h$ exposures in time, where $h$ is typically $9$ exposures.
+We then find a fiducial calibration of the spectrograph,  $g_0(x,y,m)$, which is set to be the mean line center value for each identified line over all exposures.
+We subtract this fiducial calibration from each exposure's measured line centers and run PCA on the difference between the measured and mean line center for each exposure.
+The resulting principal components are $g_k(x,y,m)$ in equation 2 while the amplitude of each principal component is $a_{nk}$.
 
 Using those values with equation 2 and some small integer (e.g.
-2) for K, we can reconstruct a de-noised matrix of line centers for each exposure.
-Missing line center measurements are then replaced with values from this reconstruction.
-We do this iteratively until the values of the missing line centers change by less than 0.01\% of a pixel.
-This convergence suggests that the PCA is no longer being swayed by outliers and is truly ``de-noised."
+6) for $K$, we can reconstruct a denoised matrix of line centers for each exposure.
+Missing line center measurements are replaced with values from this reconstruction.
+We iterate until the values of the missing line centers change by less than 0.01\% of a pixel.
+This convergence suggests that the PCA is no longer being swayed by outliers and is truly ``de-noised."   We have found this typically takes less than 30 iterations.
 
@@
 \SetAlgoLined
-\textbf{Inputs}: \;
-\While{change in bad pixels $>$ 0.01\%}{
+\textbf{Inputs}: $m$, order of each line; $\lambda$, assigned wavelength of each line; $\{x_{n,l}\} $, matrix of pixel line position of each line $l$ and for each exposure $n$\;
+\While{change in missing or outlier line centers $>$ 0.01\%}{
 	$g_0(x,y,m) = \overline{\{x_{n,l}\}}$\;
@@
 	let $a_{n,k} = U\cdot \Sigma$ and $V = g_k(x,y,m)$\;
-	$\lambda(x,y,m,n) = g_0(x,y,m) + \sum_{k=1}^K a_{nk}\,g_k(x,y,m)$ where $K=2$\;
-	$\{x_{n,l}\} = \lambda(x,y,m,n)$ where $\{x_{n,l}\}$ is $NaN$
+	$\lambda(x,y,m,n) = g_0(x,y,m) + \sum_{k=1}^K a_{nk}\,g_k(x,y,m)$ for $K=6$\;
+	$\{x_{n,l}\} = \lambda(x,y,m,n)$ where $\{x_{n,l}\}$ was initially $NaN$
 	}
@@
 
+Following the de-noising process, we check for deviant line centers.
+This is aimed at catching lines that have been mis-identified or poorly fit, resulting in erroneous line centers.
+Outliers are identified via a simple sigma cut for each line.
+Any line center that is reported at more than $3\sigma$ away from the mean line center value is relabeled a ``bad line.''  We then repeat the process of hierarchical de-noising to also replace the line centers of these bad lines.
+(\lz{Should probably add something about why the sigma cut works?})
+
 \subsection{Non-Parametric Calibration} \label{sec:nonparam}
-We can interpolate $a_{nk}$ with respect to time to determine $\lambda(x,y,m,n)$ at any time.
-Once we have the line center for each line, which has an associated known wavelength, we can interpolate the known wavelengths over line centers to any position in x in a given order.
+The process of hierarchical de-noising gives us $a_{nk}$, which captures the variability of the instrument.
+We can interpolate $a_{nk}$ with respect to housekeeping data to determine line positions, $\lambda(x,y,m,n)$, for any state of the spectrograph.
+In this paper, we show results where $a_{nk}$ is interpolated with respect to time, but it is possible to incorporate other information such as temperature, pressure, or telescope position.
+
+Once we have the line center for every line in $\{l_{m,\lambda}\}$, each of which is associated with a known wavelength, we can interpolate  the known wavelengths over line centers to any position in x in a given order.
 For example, interpolating onto every integer x will generate wavelengths for each pixel in an order.
-
-We chose to implement both interpolations with a cubic spline.  Interpolation of wavelengths over line centers is done order by order.
-
 \begin{algorithm}
 \SetAlgoLined
-\textbf{Inputs}: new time, new line centers, new orders\; %command not working for w/e reason
-Let $t_n$ be the time associated with exposure $n$\;
-let $t_n'$, $x_n'$, and $m_n'$ be the new time, line centers, and orders we want wavelengths for\;
-Let $a_{n,k}'$ be the interpolation of $a_{n,k}$ to time $t_n'$\;
-$\lambda(x,y,m,n)' = g_0(x,y,m) + \sum_{k=1}^K a_{nk}'\,g_k(x,y,m)$ where $K=2$\;
-\For{each unique $m_n'$}{
-	${l_n,m}$ from $\lambda(x,y,m,n)'$
-	interpolate ${l_n,m}$ onto $x_n'$\;
+\textbf{Inputs}: $g_0(x,y,m)$; $a_{n,k}$; $g_k(x,y,m)$; $t_n$, time of each exposure; $t'$, time we want wavelengths for\;
+Interpolate $a_{n,k}$ with respect to time to determine $a_{n,k}'$ for time $t'$\;
+Calculate pixel line positions for time $t'$ by evaluating\;
+$\lambda(x,y,m,n)' = g_0(x,y,m) + \sum_{k=1}^K a_{nk}'\,g_k(x,y,m)$ for $K=6$\;
+\For{each unique $m$}{
+	interpolate ${l_n,m}$ onto pixels that we want wavelengths for\;
 	}
@@
 \end{algorithm}
-\lz{What's fancy math symbol for interpolate?}
 
-\subsection{Choice Choices} \label{sec:choices}
-(Before or after results presented in validation test and ThAr sections?)
+\section{Tests}
+\subsection{Self Tests} \label{sec:test-self}
+For these so-called ``self tests", we use \name\ to  predict wavelengths for the measured line centers of a calibration exposure that was used to construct the basis.
+The linear interpolation in time should return identical PCA coefficients for this time, as the calibration data was used in the PCA.
 
-We're still testing most of these (will use that as excuse to write this section later)
-\begin{itemize}
-	\item Value of K for denoising
-	\item Type of interpolation (cubic spline vs nearest neighbor or linear)
-	\item Interpolation order (interpolate time and x direction together?  Individually?  Which one first?)
-\end{itemize}
+The predicted wavelengths for these measured line centers should match the assigned wavelengths for these calibration lines within the range of error of the measured line center.
+Per the EXPRES pipeline, the LFC line centers used in this analysis are measured by fitting each LFC line to a Gaussian \lz{Peteresburg2020}.
+The mean of the fitted Gaussian is set to be the line center.
+The error in this fitted value is on average 2.1 $\mps$.
+Additional error will arise from interpolation error of wavelengths to other pixel positions.
 
-Reason for not linear interpolation:
-Due to the dispersion intrinsic to echelle spectrographs, the wavelength change between pixels is greater at greater wavelengths.
-This means that the function of wavelength given pixel across an order will necessarily be concave down everywhere.
-Consequently, linear interpolation of this function will systematically give erroneously low values.
+\subsection{Training and Validation Tests} \label{sec:test-trainNvalid}
+A more full-scale test is to leave out some percentage of the calibration images from $\{exp_n\}$ while constructing the basis vectors and weights with PCA.
+We can then use the resulting basis to predict wavelengths for lines in the calibrations images that were left out.
+The used calibration images represent the ``training set," while the left out exposures make up the ``validation set."
+
+For these tests, we used 90\% of available exposures to train the PCA, and left the remaining 10\% for validation.
+Exposures were sorted into either set at random.
+Contrary to the self tests, these results may also accrue errors from interpolating the principal component amplitudes with respect to time, i.e.
+we are now also testing how well we can re-construct line centers for all times.
+As before, differences in wavelength will also be drawn from how well line centers are measured in an exposure and any error in interpolating to those measured values.
+This second source of error was independently quantified in the second test described in the previous section.
+
+\lz{I had some discussion of cross tests (i.e.
+using LFC to predict ThAr and ThAr to predict LFC, but I removed it since we decided to leave ThAr mostly out of this due to its data generally being worse}
 
 \section{Results} \label{sec:results}
-\subsection{Self Tests} \label{sec:test-self}
-First, let's have a sanity check.
-Using a cubic spline scheme with no smoothing to predict wavelengths for the returned, denoised line center values, $\lambda(x,y,m,n)$, should return identical values to the assigned true wavelength for each line within machine error.
-This is what happens, for both ThArs and LFCs: look at the figure.
 
-As an only slightly less silly sanity check, we can also use the constructed PCA basis to predict wavelengths for the measured line centers of an exposure that was included in the set of exposures used to construct the basis.
-This interpolation should return identical PCA coefficients for that time.
-Differences in wavelengths will arise from a combination of 1) the error inherent in measuring line centers and 2) interpolation error of lines to other x positions in the pixel direction.
+\begin{figure}[h]
+\centering
+\includegraphics[width=.5\textwidth]{Figures/lfcgood.png}
+\caption{Residuals of different wavelength calibration tests.
+Shown as a black outline is the distribution of residuals from the self test described in Section \ref{sec:test-self}, which incorporates errors in the fitted line centers and interpolating with respect to pixel.
+Over-plotted in blue are the residuals from the training and validation test described in Section \ref{sec:test-trainNvalid}, which builds on the self tests by additionally incorporating possible errors from interpolating line positions through time.\
+\lz{Considering adding a histogram of line center errors to show it is on par?  i.e. the limitation is the line fit.  What other information is needed?  i.e. the median/mean or spread?}}
+\label{fig:testHists}
+\end{figure}
 
-Difference in denoised x-values to measured ones? (translate into rv shift?)
-median and spread in errors in pixel (like 0.006) and RV (like 3.2 m/s?)
-Histograms
+Results from these tests are shown in Figure \ref{fig:testHists}.
+The result of the self tests for LFC exposures is shown as a black outline.
+This histogram combines residuals from all exposures.
+The self test represents the limit of how well the line centers at any given time can be predicted, since here we are predicting the line centers for an exposure used in the construction of $a_{nk}$.
 
-\subsection{Training and Validation Tests} \label{sec:test-trainNvalid}
-A better test is to leave out some percentage of the calibration images from $\{exp_n\}$ while constructing the basis and weights with PCA.
-We can then use the results from that PCA to predict wavelengths for lines in the calibrations images that were left out.
-We will call the used calibration images the ``training set," and the left out exposures the ``validation set."
+Overplotted in blue are the residuals from training the wavelength solution with only 90\% of LFC exposures and predicting the wavelengths for the remaining 10\% of exposures.
+The two histograms match very closely, indicating that very little error is introduced when predicting denoised x values for times not used in the construction of the wavelength solution.
 
-We chose to use 90\% of available exposures to train the PCA, and leave the remaining 10\% for validation.
-Exposures were sorted into either set at random.
-This will test how well PCA coefficients are interpolated to other times outside of the training set and therefore how well line centers can be reconstructed at all times.
-As before, differences in wavelength will also be drawn from how well line centers are measured and error in interpolating to those measured values.
-This second source of error was independently quantified in the second test described in the previous section.
-
-LFC looks good, ThAr is struggling a bit
-Difference in denoised x-values to measured ones? (translate into rv shift?)
-median and spread in errors in pixel (like 0.006) and RV (like 3.2 m/s?)
-Histograms
-
-\subsection{Cross Tests} \label{sec:test-cross}
-Similar to the training and validation tests, we can also use one calibration source to construct the PCA basis and weights and the other to test the results.
-For example, we can use LFC exposures to construct the wavelength solution, and then use it to predict wavelengths for measured line centers of ThAr lines, which will have their own assigned true wavelengths.
-Likewise, we can use ThAr exposures to predict the wavelengths of LFC lines.
-Note that errors in this test will also draw on the difference in quality of the two calibration sources being compared.
-
-LFC predicts ThAr better than ThAr can predict itself.
-ThAr predicting LFC is really messing up somehow, and I should figure out what it is (sorry!)
-
-%\subsection{ThAr Stuff?  Or Include it in Previous Subsections?}
-
-\section{Comparisons to Other Methods} \label{sec:comparisons}
-I think we probably have to use the even/odd method to compare.
-
-\subsection{Comparison to Parametric Model}
-Previously, the \expres\ pipeline the $\theta_{n}$ from equation 1 comprises the
-amplitudes of some 9th-ish-order polynomial in $x$ and $m$ for each $n$.
-If so, they have
+\section{Comparison to Classic, Parametric Model} \label{sec:comparisons}
+Classically, pipelines employ polynomials to construct smooth wavelength solutions for each exposure.
+For example, the \expres\ pipeline described in \lz{Petersburg2020} prescribed $\theta_{n}$ from equation 1 to be the amplitudes of a 2D, 9th-order polynomial in $x$ and $m$ for each $n$, making the equivalent to equation 2 the following:
 \begin{equation}
@@
 \end{equation}
-where the $c_{nij}$ are coefficients unique to exposure $n$ but smoothed by a third order polynomial over time.
-Further, they were finding the coefficients $c_{nij}$ that
-minimize an objective $Q$ that is something like the L2-norm:
+where $c_{nij}$ are coefficients unique to exposure $n$.
+These coefficients are then smoothed by a third-order polynomial over time.
+This third-order polynomial would be evaluated at the time of non-calibration exposures to re-construct a 2D, 9th-order polynomial wavelength solution for that exposure.
+Coefficients, $c_{nij}$, were found that minimize an objective $Q$ that is the L2-norm:
 \begin{equation}
@@
 because that is the definition of 9th order.  We find that fitting to 8th order is sufficient, faster, and leaves less residual pattern than the original 9th degree fit.
-For another, the objective function could be made soft to permit
-catastrophic outliers without destroying the fit. (\lz{we did not do this})
+
+For another, the objective function could be made soft to permit catastrophic outliers without destroying the fit.
+(\lz{we did not do this}).
 We might recommend the iteratively reweighted least squares (IRLS).
 This would make the fitting more robust.
-For yet another, there are rescaling issues for the products $x^i\,m^j$ to
-protect the fitting from near-singularities or bad conditioning of the
-linear-algebra operators.
+
+For yet another, there are rescaling issues for the products $x^i\,m^j$ to protect the fitting from near-singularities or bad conditioning of the linear-algebra operators.
+
+\subsection{Comparisons}
+To test the parametric vs.
+the non-parametric method, we took an LFC exposure and separated the even and odd lines.
+We then used the even lines to predict wavelengths for the odd lines and vice versa.
+
+For example, for the parametric model, we fit a 2D, 9th-degree polynomial to the odd lines.
+We then evaluated this polynomial at the positions of the even lines and compared those returned wavelengths to the assigned wavelength of each line.
+We then reversed the process by fitting a polynomial to the even lines and evaluating that polynomial at the positions of the odd lines.
+Differences between the polynomial given wavelengths and the assigned wavelengths of each line are shown in Figure \ref{fig:parametricTest}, left.
 
@@
 \centering
-\includegraphics[width=0.9\textwidth]{Figures/polyval2d.pdf}
-\caption{}
-\label{fig:polyValFit}
+\includegraphics[width=\textwidth]{Figures/parametricTest.png}
+%\includegraphics[width=\textwidth]{Figures/parametricTest.pdf}
+\caption{Comparison of residuals to a parametric model (left-most) vs.
+a non-parametric model (right-most).
+In each case, the color bar shows the difference between the predicted wavelength and the true wavelength for each line in units of velocity shift.
+The center plot shows a histogram of the residuals of the polynomial fit (blue) and the non-parametric fit (green).
+Medians of the residuals for each model are given in the top-right corner of the center plot in their corresponding colors.}
+\label{fig:parametricTest}
 \end{figure}
+
+We run the test using a polynomial derived as described in \lz{Petersburg2020} with discussed improvements.
+We repeat the test using a PCHIP cubic-spline as a model rather than a polynomial.
+The results are shown in Figure \ref{fig:parametricTest}.
+For each method, we plot each line with respect to order and x-pixel and colored by the difference between the predicted and assigned wavelength for that line in units of $\mps$.
+
+The residuals of the parametric model are shown in the left-most plot of Figure \ref{fig:parametricTest} while the residuals of the non-parametric model are shown in the right-most plot.
+The center plot is a histogram of the residuals of the two methods.
+The residuals from using the non-parametric model show less structure, are less spread around zero, and more symmetric about zero.
+The non-parametric model was able to account for the high-order detector defects that are smoothed over by a polynomial fit, resulting in a more accurate fit.
 
@@
 \centering
-\includegraphics[width=0.9\textwidth]{Figures/designMatrix.pdf}
-\caption{}
-\label{fig:dsnMFit}
+\includegraphics[width=\textwidth]{Figures/waveResids.png}
+\caption{Comparisons of parametric and non-parametric wavelength solutions using either just ThAr lines or LFC lines.
+Left: the difference between each wavelength solution evaluated at the position of the LFC lines and the assigned wavelength of that LFC line.
+The right plot is simply a zoom in of the left plot.
+Note the large deviation when trying to implement Excalibur with just ThAr lines, which are much fewer in number.
+The classic, polynomial  fit exhibits similar residuals regardless of whether the set of ThAr lines or LFC lines are used.
+Excalibur using LFC lines exhibit the lowest residuals.}
+\label{fig:waveResids}
 \end{figure}
 
-\subsection{Comparison to Non-Parametric, Non-Hierarchical Model}
+Incorporating a hierarchical, de-noising element results in ever better residuals!
+
+\lz{come on, future Lily, add more here}
 
@@
 \centering
-\includegraphics[width=0.9\textwidth]{Figures/noHierc.pdf}
-\caption{}
-\label{fig:intpFit}
+\includegraphics[width=0.5\textwidth]{Figures/allHists.png}
+\caption{Histogram of residuals for different wavelength calibration methods.
+The parametric (blue) and non-parametric (green) histograms are the same as described in Figure \ref{fig:parametricTest}.
+In orange is the histogram of residuals to the full Excalibur model, which includes the hierarchical component.
+The median of the residuals for each model is given in the top-right corner in each model's corresponding color.}
+\label{fig:allHists}
 \end{figure}
@@
 \section{Real Data Through to RVs} \label{sec:realdata}
+\lz{Debra and RP currently working through this.  Results seem good}
+
+\section{Choice Choices} \label{sec:choices}
+The implementation of  \name presented here required determining global variables and methods that we believe are or are close to optimal for constructing a high-fidelity wavelength solution.
+This section will describe each choice and the associated decision-making process.
+
+\subsection{Value of K}
+The value of $K$ determines how many principal components are used to reconstruct the de-noised line centers.
+This $K$ needs to be large enough so that all variability in the spectrograph is captured.
+Too large, however, and the reconstruction will incorporate noise, thereby defeating the purpose.
+
+We settled on a $K$ value of 6.
+Figure \ref{fig:pcLfc} shows the first 6 principal components constructed using LFC lines.
+In both cases there is clear structure in the first and second principal components.
+The diagonal-banding structure seen in components 3, 5, and 6 we believe are capturing detector fabrication defects.
+Components 3 and 4 show aberrant behavior on the edges of the two bluest orders, where lower signal results in more variation in the line fits.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=\textwidth]{Figures/pcsLfc6.png}
+%\includegraphics[width=\textwidth]{Figures/pcsLfc6.pdf}
+\caption{The first six principal components constructed using only LFC lines.
+There is clear structure for the first two principal components, though no order information is given to the PCA.
+Principal components 3, 5 and 6 show some slight diagonal-banding structure, which we believe to be due to detector defects.
+There is aberrant behavior in the bluest lines of the bluest two orders, which will be highlighted in the discussion.}
+\label{fig:pcLfc}
+\end{figure}
+
+In deciding a K value, we ran training-validation tests for K values spanning from 2 to 512.
+One would expect the returned wavelengths to get increasingly more accurate with larger K until components that represent only noise are incorporated.
+We found that the returned wavelengths are most accurate with a K value of 32.
+However, the improvement is slight with K values greater than 6.
+Comparisons of wavelengths returned by a K=6 model vs.
+a K=32 model show significant differences in less than 10, bluer lines.
+
+\subsection{Interpolation of $a_{nk}$ with Respect to Time}
+Figure \ref{fig:nightlyVariation} shows the amplitude of the first and second principal component with respect to time on the left.
+Though the shape of the amplitudes with respect to time, a clear linear trend exists within each night.
+This is shown by the right plots in Figure  \ref{fig:nightlyVariation}.
+As the beginning-of-night and end-of-night calibration sets always include LFC exposures, we use a simple linear interpolation to interpolate principal component amplitudes with respect to time.
+
+\lz{Yeah, the right-figure is kind of a mess, isn't it?}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.56\textwidth]{Figures/pcA_lfc.png}
+\includegraphics[width=0.42\textwidth]{Figures/pcAs_byDay.png}
+\caption{Amplitude of the first two principal components show as a function of time (left) or fraction of a day (right).
+Points mark the times at which we have calibration data.
+Lines show the result of a linear interpolation.
+Left: amplitudes as a function of time for both LFC (blue) and ThAr (green).
+The top plot shows the amplitudes for the first principal component while the bottom plot shows the amplitudes for the second principal component.
+The shape of the curves is similar regardless of calibration source.
+Right: amplitudes of the first two principal components determined by the LFC exposures only.
+Here, exposures are grouped by day and plotted with respect to what time of day the exposure was taken.
+A linear interpolation through all the LFC exposures taken in one day are shown as black line.
+Points are colored by the MJD of the exposure.
+For each day, the amplitude has been artificially offset by the median amplitude for that day so all days are roughly on the same scale.}
+\label{fig:nightlyVariation}
+\end{figure}
+
+\subsection{Interpolation of Wavelengths with Respect to Pixel}
+Interpolation of wavelengths over pixel  is done order by order using a Piecewise Cubic Hermite Interpolating Polynomial (PCHIP) interpolator.
+This interpolation incorporates the flexibility needed to model the changing dispersion of the spectrograph across an order along with any detector defects while also enforcing monotonicity.
+
+Due to the dispersion intrinsic to echelle spectrographs, the wavelength change between pixels is greater at greater wavelengths.
+This means that the function of wavelength vs.
+pixel across an order will always be monotonically increasing and concave down everywhere.
+Consequently, linear interpolation of this function will systematically give erroneously low values.
+
+A more classic cubic spline interpolation can run into issues with arc lines, which are irregularly spaced or even blended and so can appear very close together.
+Very close lines cause huge deviations from the correct wavelengths, as seen in the green line of Figure \ref{fig:xinterp}..
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{Figures/intpx_tests.png}
+\caption{Results from different interpolation schemes over pixels.
+Left: resultant wavelength solutions generated using ThAr lines, shown as green points.
+Different interpolation schemes are shown in different colors.
+The largest deviation comes with the classic cubic spline (green).
+Right: histograms of residuals for each method when used to predict the wavelengths of LFC lines, shown as black, vertical lines.
+The linear interpolation model (yellow) of this concave down relation is asymmetric with disproportionately positive residuals.
+The popular cubic spline residuals (green) has very spread out residuals.
+The residuals from an interpolation scheme that incorporates smoothing (blue) appear non-Gaussian with a plateau about zero.
+Residuals from the PCHIP interpolator (orange) has the strongest peak at 0 and is roughly Gaussian.}
+\label{fig:xinterp}
+\end{figure}
+
+These huge digressions can be avoided by allowing for some smoothing in the interpolation.
+In Figure \ref{fig:xinterp}, we show an example in blue using scipy's Univarate Spline method.
+While the result appears to follow the calibration lines much better, the smoothing ultimately causes larger residuals that are spatially correlated (Fig.
+\ref{fig:xinterp}, right).
+
+For all orders, the edges will be overestimated while the middle will be underestimated.
+This indicates that the smoothing is causing the resultant wavelength solution to underestimate the curvature of the pixel-wavelength relation, resulting in similar issues as with an inflexible, parametric model.
+This enforces a smoothness we have no reason to believe is true for the data.
+
+We instead turn to the PCHIP interpolator, which damps down huge deviations in the traditional cubic spline by requiring the resulting interpolated function to be monotonic.
+This is a constraint we know is true from optics, since interpolation in the pixel direction is carried out order-by-order.
+In Figure \ref{fig:xinterp} we show that using the PCHIP interpolator returns the lowest residuals.
+
+\subsection{Order of Interpolation}
+Why do we interpolate in time first and then in x?
+
+\lz{We talked about including this discussion, but there is no real reason for doing things in this order except it is quite a bit faster.}
 
 \section{Discussion} \label{sec:discussion}
+\lz{I feel like we have a lot of hopes for this discussion section, which is not matched by my understanding of how to structure this discussion section.
+For now, I have just written brief snippets that address each of the notes I have of things we have said we want to include.}
 
-Generalization to etalon
+Generalization to etalon:
+All methods described here are easily adapted to use with an etalon or any other arc lamp for which a list of lines and their assigned wavelengths can be constructed.
 
-Maybe some practical statements?
-How many exposures are needed to implement this method at all.
-How often do exposures have to be taken throughout a night (unless that's too dependent on instrument stability)
+Ability to vet for bad exposures:
+An advantage of using PCA on all LFC exposures and line positions is the ability to isolate exposures that exhibit unexpectedly large variation.
+Such variation is likely to be captured by an early principal component.
+Errant exposures will be singled out by a spike in the amplitude for just that principal component.
+This allowed us to quickly find flawed LFC exposures, which otherwise would have required visual inspection of all 1200+ LFC exposures.
 
-Some ideas about self-calibration. Once you know the space in which the
-spectrograph lives, every single science frame gives you information about
-where you are in that space; you don't need to look at the \lfc.
+Changes in instrument/change point detection:
+On the other hand, this same feature of using PCA can also be seen as a bug.
+PCA is good at picking up on variation regardless of the source of the variation.
+Any variation that breaks our assumption of an instrument that exists in only a low-dimensional calibration space will likely be captured by early principal components, which will no longer serve to denoise the calibration data.
+We found that violations of this stability assumption typically manifests as change points in the principal component amplitudes with respect to time.
 
-Chosing interpolations
-monotonic vs. smooth (univariate spline stuff)
+Sensitivity to line-position fitting:
+The PCA will also be sensitive to variations in line fitting.
+For example, if certain regions have lines of lower signal, variable background, or are otherwise harder to fit, any resultant variation in the returned line positions across different exposures will also be captured by any PCA.
+High-fidelity line positions are essential to ensure the PCA is capturing variations in the spectrograph's calibration rather than changes in how well a line can be fit.
 
-perfect degenercy between definition of "Centroid of line" and "wavelength solution"
-cross correlation calls center of line something slightly different than what's happening when lines are being fit
-may lead to systematic differences
-in principle it would be preferable for the calibration source to look like a star
+Perfect degeneracy problem:
+With any wavelength solution, there is a perfect degeneracy between what is defined as the ``center of the line'' and the resultant wavelength solution.
+If, for example, a cross correlation method is used to extract RVs from the data, a systematic difference may be introduced depending on the definition of the center of the line.
+In principle, the best way to mitigate this error would be to use a calibration source that looks like a star.
+
+Self calibration idea:
+Once we define a calibration space that captures all possible degrees of freedom for a stabilized spectrograph, then all exposures taken with the spectrograph contain information about where the spectrograph lies in that space at the time of the exposure.
+Meaning, a science exposure can be used to determine the current state of the instrument in this calibration space.
+It would therefore be possible to self-calibrate science exposures without the use of an additional wavelength calibration source.
+
+How hard would it be to go bayesian?  Description of other implementations of \name\ than what we did here?
 

### 2020-04-21 py/waveCal.py
Tuned findFiles, findBestK code, interpolation edits

@@ -1,5 +1,6 @@
 # Collection of all wavelength related functions so far
 import os
+from glob import glob
 import numpy as np
 from numpy.polynomial.polynomial import polyvander2d, polyval2d
 import pandas as pd
@@ -9,13 +10,100 @@ from astropy.constants import c
 from astropy.time import Time
 from scipy import interpolate, optimize
 from scipy.io import readsav
-from sklearn.decomposition import TruncatedSVD
+# BUG: SKLEARN NOT INSTALLED IN EXPRES_PIPELINE ENVIRONMENT
+#from sklearn.decomposition import TruncatedSVD
+import pickle
 
+#import pcp
 from tqdm import tqdm
 
 # LFC Constants
 rep_rate = 14e9 # magic
 lfc_offset = 6.19e9 # magic
+
+
+###########################################################
+# Bandaids
+###########################################################
+
+# A function for finding all calibration files between specified dates
+# There is a lot of tweaking to avoid LFCs where the LFC was left on for > ~3 minutes
+def findFiles(start_date='000000', end_date='999999', cal_source='lfc'):
+    line_files = []
+    fits_files = []
+    times = []
+    
+    if cal_source == 'thar':
+        line_dir = '/expres/extracted/thidfiles/ThAr_*.thid'
+    else:
+        line_dir = '/expres/extracted/lfc_cal/lfc_id/LFC_*.npy'
+        #line_dir = '/expres/extracted/lfc_cal/checkpoint/LFC_*.npy'
+    
+    for file_name in tqdm(glob(line_dir)):
+        base = os.path.basename(file_name)
+        if cal_source == 'thar':
+            file_date = base[5:11]
+        else:
+            file_date = base[4:10]
+        if (file_date >= start_date) and (file_date <= end_date):
+            file_year = file_date[:2]
+            if cal_source == 'thar':
+                file_base = base[:-5]
+                _, file_onum = file_base.split('.')
+                if file_date=='191017':
+                    if (int(file_onum) >= 1065) and (int(file_onum) < 1067):
+                        # Engineering
+                        continue
+            else:
+                file_base = base[:-4]
+                _, file_onum = file_base.split('.')
+                # Vet for consecutive and bad LFCs
+                # Will eventually be replaced with a check for database status
+                if file_date=='190828':
+                    if int(file_onum) < 1219:
+                        # consecutive LFCs
+                        continue
+                elif file_date=='190825':
[…]

### 2020-04-23 paper/paper.tex
AAS Paper

@@
 
-\documentclass[12pt, letterpaper]{article}
+\documentclass[twocolumn]{aastex63}
 \usepackage{xcolor}
@@
 \begin{document}
-
-\section*{\raggedright%
-\name:
+\title{\name:
 A non-parametric, hierarchical model for a precision spectrograph}
 
-\noindent
-\textbf{Lily~Zhao} (Yale) (Flatiron),
-\textbf{David~W~Hogg} (NYU) (MPIA) (Flatiron),
-\ldots and others\ldots
+\correspondingauthor{Lily Zhao}
+\email{lily.zhao@yale.edu}
 
-\paragraph{Abstract:}
+\author[0000-0002-3852-3590]{Lily Zhao}
+\affil{Yale University, 52 Hillhouse, New Haven, CT 06511, USA}
+
+\author[0000-0003-2866-9403]{David W. Hogg}
+\affil{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, New York, NY 10003, USA}
+\affil{Center for Data Science, New York University, 60 Fifth Avenue, New York, NY 10011, USA}
+\affil{Max-Planck-Institut für Astronomie, Königstuhl 17, D-69117 Heidelberg, Germany}
+\affil{Flatiron Institute, Simons Foundation, 162 Fifth Avenue, New York, NY 10010, USA}
+
+\author[0000-0001-9907-7742]{Megan Bedell}
+\affil{Flatiron Institute, Simons Foundation, 162 Fifth Avenue, New York, NY 10010, USA}
+
+\author[0000-0003-2221-0861]{Debra A. Fischer}
+\affil{Yale University, 52 Hillhouse, New Haven, CT 06511, USA}
+
+\begin{abstract}
 \name is a hierarchical, non-parametric framework for wavelength calibrating precision spectrographs.
@@
 We find that \name\ returns more accurate wavelength predictions than classic polynomial models.
+\end{abstract}
+
+% Add official keywords
+\keywords{instrumentation: spectrographs -- techniques: spectroscopic -- techniques: radial velocities}
+
 \section{Introduction}
 
-Extreme precision radial-velocity (\eprv) programs have been
-outrageously successful in finding and characterizing extra-solar
-planets (CITE THINGS).
-These programs typically make use of spectrographs with resolutions on
-the order of $10^5$, which correspond to line widths on the order of
-$3000\,\mps$.
-Exoplanet science happens at $3\,\mps$ precisions, with current
-instruments hoping to reach $0.1\,\mps$ precision (CITE THINGS).
-This requires new spectrographs to be calibrated or
-stabilized to better than $10^{-4}$ of a pixel (assuming that the
-spectrograph is well sampled).
-Some hardware--software systems seem to be reaching these
-levels (CITE THINGS) .
+Extreme precision radial-velocity (\eprv) programs have been outrageously successful in finding and characterizing extra-solar planets (CITE THINGS).  
+These programs typically make use of spectrographs with resolutions on the order of $10^5$, which correspond to line widths on the order of $3000\,\mps$.  
+Exoplanet science happens at the level of $3\,\mps$ precision, with current instruments hoping to reach $0.1\,\mps$ precision (CITE THINGS).  
+This requires new spectrographs to be calibrated or stabilized to better than $10^{-4}$ of a pixel (assuming that the spectrograph is well sampled).  
+Some hardware--software systems seem to be reaching these levels (CITE THINGS) .
 
-Here we propose to simplify and improve calibration programs for
-\eprv\ hardware systems with two very simple but novel ideas.
-The first flows from the observation that calibration sources---which
-include lamps, etalons, and laser-frequency combs (\lfc
-s)---illuminate the spectrograph with very stable, very dense sets of lines; almost every
-location in the spectrograph image plane is surrounded by nearby,
+Here we propose to simplify and improve calibration programs for \eprv\ hardware systems with two very simple but novel ideas.  
+The first flows from the observation that calibration sources---which include lamps \lz{(looks like the ThAr isn't quite good enough actually)}, etalons, and laser-frequency combs (\lfc s)---illuminate the spectrograph with very stable, very dense sets of lines; almost every location in the spectrograph image plane is surrounded by nearby,
 useful calibration lines.
-This recommends a calibration methodology that is
-\emph{non-parametric}:
-If every point in the spectrograph detector is surrounded by nearby
-calibration lines, the wavelength solution can, for example,  be made simply an interpolation of the calibration data.
+This recommends a calibration methodology that is \emph{non-parametric}:
+If every point in the spectrograph detector is surrounded by nearby calibration lines, the wavelength solution can, for example, be made simply as an interpolation of the calibration data.
 
+The density of lines removes the need to enforce any functional form for the wavelength solution (such as a continuous eighth-order polynomial, for example).
 We expand on the work of Milakovi\'{c}2020, who proposed constructing a wavelength solution as multiple, segmented polynomials across an order by proposing a wavelength solution that is completely non-parametric.
-\lz{Should work in their paper more and praise it more?}
-
-The density of lines removes the need to enforce any functional form for the wavelength solution (such as an eighth-order polynomial, for example).
-Going non-parametric will improve calibration accuracy because
-choosing a parametric form biases the calibration, especially when the chosen function is inappropriate (as, for example, polynomials will be at detector edges).
+\lz{(Should work in their paper more and praise it more?)}
+Going non-parametric will improve calibration accuracy by not forcing the chose of a parametric form that may bias the calibration, especially when the chosen function is inappropriate (as, for example, polynomials will be at detector edges).
 
@@
 \eprv\ instruments are incredibly stable.
-Temperature-controlled, fiber-fed pectrographs vary only slightly over
-the night or season, and only along a small number of axes in what you
-might call ``calibration space'', or the (very high dimensional) space
-of all possible wavelength solutions.
+Temperature-controlled, fiber-fed pectrographs vary only slightly over the night or season, and only along a small number of axes in what you might call ``calibration space'', or the (very high dimensional) space of all possible wavelength solutions.
 That is, not only are the spectrographs stable, but they also have few accessible degrees of freedom.
-This renders it inappropriate to fit each calibration exposure or calibrate
-each science exposure independently.
-Instead, all the calibration data (or all the data) should be used to
-determine the space in which the instrument can and does vary.  Subsequent calibration work then need only determine where in the accessible part of calibration space the spectrograph existed.
+This renders it inappropriate to fit each calibration exposure or calibrate each science exposure independently.
+Instead, all the calibration data (or all the data) should be used to determine the space in which the instrument can and does vary.
+Subsequent calibration work then need only determine where in the accessible part of calibration space the spectrograph existed for each exposure.
 
-This structure is \emph{hierarchical}: the calibration data are used
-not just to determine the wavelength solution, but also to determine
-the possible space of wavelength solutions.
-In the statistics literature this concept is often described as
-\emph{de-noising}.  
-Every calibration exposure contains information about every other
-exposure.  Thus every exposure can be improved (de-noised) with information coming from every other exposure.
+This structure is \emph{hierarchical}: the calibration data are used not just to determine the wavelength solution, but also to determine the possible space of wavelength solutions.
+In the statistics literature this concept is often described as \emph{de-noising}.  
+Every calibration exposure contains information about every other exposure.  Thus every exposure can be improved (de-noised) with information coming from every other exposure.
 
 The method we propose here---\name---embodies these ideas.
-It is a non-parametric, hierarchical, data-driven model for the
-wavelength solution.
-By being non-parametric, it delivers enormous freedom to the
-wavelength solution to match or adapt to any instrument or detector
-oddities.
-By being hierarchical, it restricts that freedom tremendously, but it
-does so appropriately for the observed variations in the
-spectrograph.
-\name\ is designed for temperature-controlled, fiber-fed
-spectrographs with good calibration sources, such as laser-frequency
-combs, etalons, or good arc lamps.
-We have in mind the \eprv\ instruments and \eprv\ science cases, but we expect \name\ to have applications for other kinds of
-spectrographs in other contexts.  The motivating ideas behind this
-project are true for almost every astronomical spectrograph.
+It is a non-parametric, hierarchical, data-driven model for the wavelength solution.
+By being non-parametric, it delivers enormous freedom to the wavelength solution to match or adapt to any instrument or detector oddities.
+By being hierarchical, it restricts that freedom tremendously, but it does so appropriately for the observed variations in the spectrograph.
+\name\ is designed for temperature-controlled, fiber-fed spectrographs with good calibration sources, such as laser-frequency combs, etalons, or good arc lamps.
+The scope of the method is to take given calibration line positions, and de-noise/interpolate them into a full wavelength solution.
+We have in mind the \eprv\ instruments and \eprv\ science cases, but we expect \name\ to have applications for other kinds of spectrographs in other contexts.
+The motivating ideas behind this project are true for almost every astronomical spectrograph.
 
@@
 Let $\{exp_n\}$ be the set of calibration exposures within an instrumental epoch.
-These exposures will be used in constructing $g_0(x,y,m)$ and $g_k(x,y,m)$ in equation 2.
+These exposures will be used in constructing the fiducial calibration of the spectrograph, $g_0(x,y,m)$, and the basis functions defining the spectrograph's calibration space, $g_k(x,y,m)$ (see equation 2).
 We first identify the full list of lines we can expect to find in a calibration image from this epoch.
@@
 If this line is missing from an exposure (e.g.
-the fit failed due to noise, the line isn't in its usual order, etc.), we set $x_{n,l} = NaN$ for that line center in that exposure.
+the fit failed due to noise, the line isn't in its usual order, etc.), we set the line center $x_{n,l} = NaN$ for that line center in that exposure.
 This allows us to construct a $N \times M$ matrix of line positions for each of $M$ lines for each of $N$ exposures.
@@
 We subtract this fiducial calibration from each exposure's measured line centers and run PCA on the difference between the measured and mean line center for each exposure.
-The resulting principal components are $g_k(x,y,m)$ in equation 2 while the amplitude of each principal component is $a_{nk}$.
+The resulting principal components are used as the basis functions defining the spectrograph's calibration space, $g_k(x,y,m)$ while the amplitude of each principal component is $a_{nk}$.
 
@@
 \subsection{Non-Parametric Calibration} \label{sec:nonparam}
-The process of hierarchical de-noising gives us $a_{nk}$, which captures the variability of the instrument.
-We can interpolate $a_{nk}$ with respect to housekeeping data to determine line positions, $\lambda(x,y,m,n)$, for any state of the spectrograph.
-In this paper, we show results where $a_{nk}$ is interpolated with respect to time, but it is possible to incorporate other information such as temperature, pressure, or telescope position.
+The process of hierarchical de-noising gives us the amplitude for each principal component, $a_{nk}$, which captures the variability of the instrument.
+We can interpolate these amplitudes $a_{nk}$ with respect to housekeeping data to determine line positions, $\lambda(x,y,m,n)$, for any state of the spectrograph.
+In this paper, we show results where these amplitudes $a_{nk}$ are interpolated with respect to time, but it is possible to incorporate other information such as temperature, pressure, or telescope position.
 
-Once we have the line center for every line in $\{l_{m,\lambda}\}$, each of which is associated with a known wavelength, we can interpolate  the known wavelengths over line centers to any position in x in a given order.
+Once we have the line center for every line $\{l_{m,\lambda}\}$, each of which is associated with a known wavelength, we can interpolate  the known wavelengths over line centers to any position in x in a given order.
 For example, interpolating onto every integer x will generate wavelengths for each pixel in an order.
@@
 This histogram combines residuals from all exposures.
-The self test represents the limit of how well the line centers at any given time can be predicted, since here we are predicting the line centers for an exposure used in the construction of $a_{nk}$.
+The self test represents the limit of how well the line centers at any given time can be predicted, since here we are predicting the line centers for an exposure used in the construction of the principal component amplitudes $a_{nk}$.
 
@@
 Classically, pipelines employ polynomials to construct smooth wavelength solutions for each exposure.
-For example, the \expres\ pipeline described in \lz{Petersburg2020} prescribed $\theta_{n}$ from equation 1 to be the amplitudes of a 2D, 9th-order polynomial in $x$ and $m$ for each $n$, making the equivalent to equation 2 the following:
+For example, the \expres\ pipeline described in \lz{Petersburg2020} prescribed $\theta_{n}$ from equation 1 to be the amplitudes of a 2D, 9th-order polynomial in pixel $x$ and order $m$ for each exposure $n$, making the equivalent to equation 2 the following:
 \begin{equation}
@@
 
-\begin{figure}[h]
+\begin{figure*}[h]
 \centering
@@
 \label{fig:parametricTest}
-\end{figure}
+\end{figure*}
 
@@
 \label{fig:waveResids}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}[t]
+\begin{figure*}[t]
 \centering
@@
 \label{fig:pcLfc}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}[h!]
+\begin{figure*}[h!]
 \centering
@@
 \label{fig:nightlyVariation}
-\end{figure}
+\end{figure*}
 
@@
 \label{fig:xinterp}
-\end{figure}
+\end{figure*}
 

### 2020-04-25 paper/paper.tex
Abstract Typo
Paper Re-Framing

@@
 \begin{abstract}
-\name is a hierarchical, non-parametric framework for wavelength calibrating precision spectrographs.
-There have been many recent hardware improvements in highly-stabilized, extreme-precision radial-velocity spectrographs.
-In such stabilized spectrographs, the full optical system and detectors will only vary within a tiny range of configurations.
-Additionally, within a given range of stability, every calibration image ever taken contains information relevant to every science exposure ever taken.
-The development of laser-frequency combs or etalons allows the spectral range of an instrument to be covered with dense calibration points.
-Here, we describe a calibration method---\name---that is designed for this new era of stabilized instruments with high-line-density calibrators.
-We find that a high density of calibration lines allows for the freedom of a non-parametric wavelength solution that can match or adapt to any instrument or detector oddities.
-We can also denoise the available calibration data by using it to determine not just the wavelength solution, but also the possible space of wavelength solutions, which is small for a stabilized instrument.
+\name\ is a hierarchical, non-parametric framework for wavelength calibrating precision spectrographs.
+This method makes use of the many, recent hardware improvements in highly-stabilized, extreme-precision radial-velocity spectrographs.
+With stabilized spectrographs, the full optical system and detectors can only vary within a tiny range of configurations.
+This allows us to use all calibration data to determine the entire space of possible calibrations for a stabilized spectrograph.
+A wavelength solution then need only pinpoint where in this low-dimensional calibration space the spectrograph exists.
+\name\ also takes advantage of the development of laser-frequency combs or etalons, which generate a dense set of stable calibration points.
+This density of points permits the freedom of a non-parametric wavelength solution that can match or adapt to any instrument or detector oddities.
 We demonstrate the success of this method with data from a Menlo Systems laser frequency comb taken by \expres, the Extreme Precision Spectrograph.
-We find that \name\ returns more accurate wavelength predictions than classic polynomial models.
+We find that \name\ returns more accurate wavelength predictions by a factor of 10 than classic polynomial models.
+Employing \name\ wavelengths reduced the RMS of RV measurements to a planet fit by \lz{some percent}.
 \end{abstract}
@@
 \section{Introduction} 
-
-Extreme precision radial-velocity (\eprv) programs have been outrageously successful in finding and characterizing extra-solar planets (CITE THINGS).  
+Extreme precision radial-velocity (\eprv) programs have been incredibly fruitful in finding and characterizing extra-solar planets (CITE THINGS).
 These programs typically make use of spectrographs with resolutions on the order of $10^5$, which correspond to line widths on the order of $3000\,\mps$.
@@
 This requires new spectrographs to be calibrated or stabilized to better than $10^{-4}$ of a pixel (assuming that the spectrograph is well sampled).
-Some hardware--software systems seem to be reaching these levels (CITE THINGS) .
+Some hardware--software systems, such as ESPRESSO and \expres\ are now reaching these levels (CITE THINGS).
 
 Here we propose to simplify and improve calibration programs for \eprv\ hardware systems with two very simple but novel ideas.
-The first flows from the observation that calibration sources---which include lamps \lz{(looks like the ThAr isn't quite good enough actually)}, etalons, and laser-frequency combs (\lfc s)---illuminate the spectrograph with very stable, very dense sets of lines; almost every location in the spectrograph image plane is surrounded by nearby,
-useful calibration lines.
-This recommends a calibration methodology that is \emph{non-parametric}:
-If every point in the spectrograph detector is surrounded by nearby calibration lines, the wavelength solution can, for example, be made simply as an interpolation of the calibration data.
+The first flows from the observation that calibration sources---which include lamps \lz{(looks like the ThAr isn't quite good enough actually)}, etalons, and laser-frequency combs (\lfc s)---illuminate the spectrograph with very stable, very dense sets of lines; almost every location in the spectrograph image plane is surrounded by nearby, useful calibration lines.
+This recommends a calibration methodology that is \emph{non-parametric}:  If every point in the spectrograph detector is surrounded by nearby calibration lines, the wavelength solution can, for example, be made simply as an interpolation of the calibration data.
 
 The density of lines removes the need to enforce any functional form for the wavelength solution (such as a continuous eighth-order polynomial, for example).
-We expand on the work of Milakovi\'{c}2020, who proposed constructing a wavelength solution as multiple, segmented polynomials across an order by proposing a wavelength solution that is completely non-parametric.
-\lz{(Should work in their paper more and praise it more?)}
-Going non-parametric will improve calibration accuracy by not forcing the chose of a parametric form that may bias the calibration, especially when the chosen function is inappropriate (as, for example, polynomials will be at detector edges).
+We expand on the work of Milakovi\'{c}2020, which demonstrated the efficacy of constructing a wavelength solution as multiple, segmented polynomials.
+Going non-parametric will improve calibration accuracy by not forcing the choice of a parametric form that may bias the calibration, especially when the chosen function is inappropriate (as, for example, polynomials will be at detector edges).
 
-The second idea flows from the observation that contemporary
-\eprv\ instruments are incredibly stable.
-Temperature-controlled, fiber-fed pectrographs vary only slightly over the night or season, and only along a small number of axes in what you might call ``calibration space'', or the (very high dimensional) space of all possible wavelength solutions.
+The second idea follows from the observation that contemporary \eprv\ instruments are incredibly stable.
+Temperature-controlled, fiber-fed spectrographs vary only slightly over the night or season, and only along a small number of axes in what you might call ``calibration space'', or the (very high dimensional) space of all possible wavelength solutions.
 That is, not only are the spectrographs stable, but they also have few accessible degrees of freedom.
@@
 In the statistics literature this concept is often described as \emph{de-noising}.
-Every calibration exposure contains information about every other exposure.  Thus every exposure can be improved (de-noised) with information coming from every other exposure.
+Every calibration exposure contains information about every other exposure.
+Thus every exposure can be improved (i.e.
+de-noised) with information from every other exposure.
 
@@
 By being non-parametric, it delivers enormous freedom to the wavelength solution to match or adapt to any instrument or detector oddities.
-By being hierarchical, it restricts that freedom tremendously, but it does so appropriately for the observed variations in the spectrograph.
+By being hierarchical, it restricts that freedom tremendously, but it does so appropriately for the observed variations in the stabilized spectrograph.
 \name\ is designed for temperature-controlled, fiber-fed spectrographs with good calibration sources, such as laser-frequency combs, etalons, or good arc lamps.
-The scope of the method is to take given calibration line positions, and de-noise/interpolate them into a full wavelength solution.
 We have in mind the \eprv\ instruments and \eprv\ science cases, but we expect \name\ to have applications for other kinds of spectrographs in other contexts.
@@
 \section{Method} \label{sec:method}
+\name\ is designed to turn a series of calibration lines with known wavelengths and well-fit detector positions, and de-noise and interpolate them into a full wavelength solution.
 We operate on the two core ideas that the wavelength solution should be given enormous flexibility, but that it lives in a very low-dimensional space, where the degrees of freedom are set by the limited kinematics of the stabilized, spectrograph hardware.
 
-We will pose the problem in the following way.  Given an exposure $n$, and order $m$, there is a relationship between
+We therefore assume dense enough calibration line coverage with well-fit line centers to provide some constraint on an interpolated wavelength solution across an order.
+Additionally, we assume the instrument hardware is stabilized enough that the space of possible calibration states is low-dimensional.
+
+We will pose the question of wavelength calibration in the following way.  Given an exposure $n$, and order $m$, there is a relationship between
 the two-dimensional $(x,y)$-position on the detector and the
@@
 \quad ,
+\label{eq:wsol}
 \end{equation}
@@
 
-A stabilized spectrograph should experience only low-degree variability, meaning the calibration of any image can be informed by the calibration of every other image.
-To implement such a hierarchical model, we can use the calibration data themselves to develop a low-dimensional basis for expressing the calibration data.
+Classically, pipelines employ polynomials to construct smooth wavelength solutions for each exposure.
+For example, the pipeline described in \lz{Petersburg2020} sets the function $f(x,y,m;\theta_{n})$ from Equation \ref{eq:wsol} to be a 2D, 9\textsuperscript{th}-order polynomial where $\theta_{n}$ represents the polynomial coefficients, $c_{nij}$, unique to each exposure $n$.
+\begin{equation}
+\lambda(x,y,m,n) = \sum_{i=0}^9\sum_{j=0}^9 c_{nij}\, x^i\,m^j + \mathrm{noise}
+\quad ,
+\label{eq:poly_wsol}
+\end{equation}
+The coefficients $c_{nij}$ are then smoothed by a third-order polynomial over time.
+This third-order polynomial is evaluated at the time of non-calibration exposures to re-construct a 2D, 9th-order polynomial wavelength solution for that exposure.
+Each calibration image is fit to a 2D polynomial independently.
 
-If the space of all calibration possibilities is in fact $K$-dimensional (where K is a small integer, i.e 2 or 8 or thereabouts), and if the calibration variations are so small that we can linearize, then the function $f(x,y,m;\theta_{n})$ could be replaced with a tiny model
+Given a stabilized instrument, however, the spectrograph should experience only low-degree variability, meaning the calibration of any image can and should be informed by the calibration of every other image.
+The calibration data themselves can be used to develop a low-dimensional basis for expressing the space of possible calibrations for a stabilized spectrograph.
+
+If the space of all calibration possibilities is in fact $K$-dimensional (where $K$ is a small integer, i.e 2 or 8 or thereabouts), and if the calibration variations are so small that we can linearize, then the function $f(x,y,m;\theta_{n})$ from Equation \ref{eq:wsol} could be replaced with a tiny model
 \begin{equation}
@@
 \quad ,
+\label{eq:excl_wsol}
 \end{equation}
@@
 
-\subsection{Data} \label{sec:data}
-We present results using data from \expres\, the EXtreme PRecison Spectrograph.
-\expres\ is an environmentally-stabilized, fiber-fed, $R\sim137,000$, optical spectrograph (\lz{Jurgenson2016, Blackman2020}).
-\expres\ has two different wavelength calibration sources, a ThAr lamp and a Menlo Systems laser frequency comb (LFC, e.g.
-\lz{Wilken2012, Molaro2013, Probst2014}).
-ThAr exposures are taken at the beginning and end of each night.
-LFC exposures are interspersed between science exposures every ~15-30 minutes.
-
-The results and discussion presented here are based off of 1227 LFC exposures taken over two months months on 29 unique days.
-LFC lines cover 40 echelle orders, which contain 19203 lines.
-Though we primarily work with LFC data, we at times discuss applications to arc lamps.
-These results use 78 ThAr exposures that were taken over the same range of time.
-ThAr lines cover all 86 extracted orders of \expres\ and includes 5295 lines.
-
-For the results presented in this paper, we work with the 1D, optimally extracted data.
-This removes the y-dependence from equations 1 and 2.
-Gaussian profiles are fit to each line in the 1D, extracted data.
-The mean of each fitted Gaussian is taken to be the center of the line (see \lz{Petersburg2020} for more details).
-While precisely defining and determining line positions is essential to precise wavelength calibration, we consider the task out of scope for this paper.
-\name\ proceeds from a given list of orders, line wavelengths, and pixel positions of each line for each exposure.
-
-\expres\ exposures are separated into different ``instrumental epochs," which correspond to atypical, hardware changes that affected the position of the echellogram on the detector, the shape of the instrumental PSF, or the calibration sources.
-Because significant instrumental changes break the assumption of only low-dimensional variations in the spectrograph hardware, we treat each instrument epoch independently.
-
-\subsection{Hierarchical De-Noising of Calibration Frames} \label{sec:denoising}
+\subsection{De-Noising of Calibration Frames} \label{sec:denoising}
 \lz{Concerns: 1) how to best phrase the problem using variables?, 2) how in-depth is too in-depth?}
 
-Let $\{exp_n\}$ be the set of calibration exposures within an instrumental epoch.
-These exposures will be used in constructing the fiducial calibration of the spectrograph, $g_0(x,y,m)$, and the basis functions defining the spectrograph's calibration space, $g_k(x,y,m)$ (see equation 2).
-We first identify the full list of lines we can expect to find in a calibration image from this epoch.
-A line is uniquely defined by a combination of order, $m$, and ``true" or theoretical wavelength assigned to that line, $\lambda$.
-Let $\{l_{m,\lambda}\}$ be the complete list of unique lines found in $\{exp_n\}$.
+Within a given range of stability, we can use all calibration images to  1) determine the space in which the instrument varies and 2) where in the accessible calibration space the spectrograph existed for each exposure.
+Say we have $N$ calibration exposures, $exp_n$.
+The first task is to identify the full list of lines we can expect to find in a calibration image.
+A line, $l_{m,\lambda}$ is uniquely defined by a combination of order, $m$, and ``true" or theoretical wavelength, $\lambda$.
 
-Each line, $l_{m,\lambda}$, has a fitted line center, $x_{n,l}$, in pixel space, for each calibration exposure $exp_n$.
-If this line is missing from an exposure (e.g.
-the fit failed due to noise, the line isn't in its usual order, etc.), we set the line center $x_{n,l} = NaN$ for that line center in that exposure.
-This allows us to construct a $N \times M$ matrix of line positions for each of $M$ lines for each of $N$ exposures.
+For each exposure, $exp_n$, every line, $l_{m,\lambda}$, has an associated fitted detector position, $x_{n,l}$, for example x-pixel in an 2D extracted order.
+Fitted line centers that are missing from an exposure (e.g.
+because the fit failed due to noise, the line is not in its usual order, etc.) can be assigned a $NaN$ for that order instead.
+Let there be $M$ lines per exposure.
+We can therefore construct a $N \times M$ matrix of line positions for each of $M$ lines for each of $N$ exposures.
 
-Lines that are missing from greater than some percent of exposures are completely cut from the analysis.
-Exposures that are missing greater than some percent of lines are also cut.
-We use a threshold percentage of 50\% for both lines and exposures.
-This largely affects the first lines on either end of an order and exposures with spuriously low signal.
-Typically, about 0.2\% of lines and cut for LFC based calibration and 0.25\% of files are cut.
+The average line position in every exposure represents the fiducial, or standard calibration of the spectrograph, $g_0(x,y,m)$.
+We apply principal component analysis to the difference between this fiducial calibration and each individual line position.
+The returned principal components serve as basis functions,  $g_k(x,y,m)$, expressing the possible deviations of the spectrograph from this fiducial calibration.
+The magnitude of each principal component for each exposure, $a_{nk}$, represents the scalar amplitude of these deviations for each exposure.
+We can now use a small number, $K$, of principal components to reconstruct a de-noised version of the line positions as formulated in Equation \ref{eq:excl_wsol}.
 
-Missing line center measurements are replaced with denoised estimates using iterative PCA.
-For each missing line center measurement, we initialize its value to the mean of the nearest $h$ exposures in time, where $h$ is typically $9$ exposures.
-We then find a fiducial calibration of the spectrograph,  $g_0(x,y,m)$, which is set to be the mean line center value for each identified line over all exposures.
-We subtract this fiducial calibration from each exposure's measured line centers and run PCA on the difference between the measured and mean line center for each exposure.
-The resulting principal components are used as the basis functions defining the spectrograph's calibration space, $g_k(x,y,m)$ while the amplitude of each principal component is $a_{nk}$.
-
-Using those values with equation 2 and some small integer (e.g.
-6) for $K$, we can reconstruct a denoised matrix of line centers for each exposure.
-Missing line center measurements are replaced with values from this reconstruction.
-We iterate until the values of the missing line centers change by less than 0.01\% of a pixel.
-This convergence suggests that the PCA is no longer being swayed by outliers and is truly ``de-noised."   We have found this typically takes less than 30 iterations.
+Missing line center measurements are replaced with de-noised estimates.
+We do this iteratively until the estimates of missing line centers change by less than 0.01\%.
+This process can be repeated on line centers deemed as outliers by some metric, to account for lines that may have been mis-identified or mis-fit.
+The principal components from the final iteration are used to define the spectrograph's calibration space while the associated magnitudes pinpoint where in that calibration space the spectrograph is for each calibration exposure.
 
@@
 \SetAlgoLined
-\textbf{Inputs}: $m$, order of each line; $\lambda$, assigned wavelength of each line; $\{x_{n,l}\} $, matrix of pixel line position of each line $l$ and for each exposure $n$\;
+\textbf{Inputs}: $m$, order of each line; $\lambda$, assigned wavelength of each line; $[x_{n,l}] $, matrix of pixel line position of each line $l$ and for each exposure $n$\;
 \While{change in missing or outlier line centers $>$ 0.01\%}{
-	$g_0(x,y,m) = \overline{\{x_{n,l}\}}$\;
+	$g_0(x,y,m) = \overline{x_{n,l}]}$\;
 	find $U, \Sigma, V$ s.t. $U\Sigma V^* = (x_{l,n}-g_0(x,y,m))$\;
-	let $a_{n,k} = U\cdot \Sigma$ and $V = g_k(x,y,m)$\;
+	let $a_{n,k} = U\cdot \Sigma$ and $g_k(x,y,m) = V$\;
 	$\lambda(x,y,m,n) = g_0(x,y,m) + \sum_{k=1}^K a_{nk}\,g_k(x,y,m)$ for $K=6$\;
-	$\{x_{n,l}\} = \lambda(x,y,m,n)$ where $\{x_{n,l}\}$ was initially $NaN$
+	$[x_{n,l}] = \lambda(x,y,m,n)$ where $[x_{n,l}]$ was initially $NaN$
 	}
-\caption{Hierarchical De-Noising}
+\caption{De-Noising}
 \end{algorithm}
 
-Following the de-noising process, we check for deviant line centers.
-This is aimed at catching lines that have been mis-identified or poorly fit, resulting in erroneous line centers.
-Outliers are identified via a simple sigma cut for each line.
-Any line center that is reported at more than $3\sigma$ away from the mean line center value is relabeled a ``bad line.''  We then repeat the process of hierarchical de-noising to also replace the line centers of these bad lines.
-(\lz{Should probably add something about why the sigma cut works?})
+\subsection{?Interpolating Calibration Frames?}
+ \label{sec:interp_time}
+The magnitude of each principal component can be interpolated with respect to housekeeping data, such as time, temperature, telescope position, etc.
+to determine different calibration states of the spectrograph.
+The choice of what to interpolate with respect to depends on the dominant contribution to variation in the instrument.
 
-\subsection{Non-Parametric Calibration} \label{sec:nonparam}
-The process of hierarchical de-noising gives us the amplitude for each principal component, $a_{nk}$, which captures the variability of the instrument.
-We can interpolate these amplitudes $a_{nk}$ with respect to housekeeping data to determine line positions, $\lambda(x,y,m,n)$, for any state of the spectrograph.
-In this paper, we show results where these amplitudes $a_{nk}$ are interpolated with respect to time, but it is possible to incorporate other information such as temperature, pressure, or telescope position.
+As an example, we can interpolate the magntiudes of the principal components with respect to time to the midpoint time of non-calibration exposures to determine the calibration state of the spectrograph for that exposure.
+Using interpolated magnitudes, $a'_{nk}$ and the basis vectors, $g_k(x,y,m)$, returned by the de-noising process, we can construct a set of calibration lines, $\lambda(x,y,m,n)$, for any exposure as prescribed in Equation \ref{eq:excl_wsol}.
 
-Once we have the line center for every line $\{l_{m,\lambda}\}$, each of which is associated with a known wavelength, we can interpolate  the known wavelengths over line centers to any position in x in a given order.
-For example, interpolating onto every integer x will generate wavelengths for each pixel in an order.
+\subsection{Interpolating a Wavelength Solution} \label{sec:interp_wsol}
+We now have the power to construct a set of calibration lines, $\{l_{m,\lambda}\}$, each with an assigned, known wavelength for any exposure.
+To construct a wavelength solution, we need only interpolate the known wavelengths over the exposure specific line centers.
+For instance, interpolating the known wavelengths vs.
+line centers onto every integer x will generate wavelengths for each pixel in an order.
+
 \begin{algorithm}
 \SetAlgoLined
-\textbf{Inputs}: $g_0(x,y,m)$; $a_{n,k}$; $g_k(x,y,m)$; $t_n$, time of each exposure; $t'$, time we want wavelengths for\;
+\textbf{Inputs}: $g_0(x,y,m)$, the fiducial calibration of the spectrograph; $a_{n,k}$, magnitudes of the principal components for each exposure; $g_k(x,y,m)$, basis vectors spanning the calibration space of the spectrograph; $t_n$, time of each exposure; $t'$, time we want wavelengths for\;
 Interpolate $a_{n,k}$ with respect to time to determine $a_{n,k}'$ for time $t'$\;
@@
 \For{each unique $m$}{
-	interpolate ${l_n,m}$ onto pixels that we want wavelengths for\;
+	interpolate ${l_n,m}$ onto pixels, $x'$, that we want wavelengths for\;
 	}
-\caption{Non-Parametric Wavelength Solution}
+\caption{Generating a Wavelength Solution}
 \end{algorithm}
 
+\section{Data} \label{sec:data}
+We tested \name\ results using data from \expres\, the EXtreme PRecison Spectrograph.
+\expres\ is an environmentally-stabilized, fiber-fed, $R\sim137,000$, optical spectrograph (\lz{Jurgenson2016, Blackman2020}).
+\expres\ has two different wavelength calibration sources, a ThAr lamp and a Menlo Systems laser frequency comb (LFC, e.g.
+\lz{Wilken2012, Molaro2013, Probst2014}).
+
+Rather than using a simultaneous calibration fiber, two to three LFC exposures are interspersed between science exposures every ~15-30 minutes.
+ThAr exposures are taken at the beginning and end of each night.
+All calibration data is taken through the science fiber, thereby traveling through the same optics and hitting the same pixels as all science observations.
+
+The results and discussion presented here are based off of 1227 LFC exposures taken between October 14 and December 18, 2020 on 29 unique nights.
+LFC lines cover 40 echelle orders, which contain 19203 lines.
+Though we primarily work with LFC data, we at times discuss applications to arc lamps.
+These results use 78 ThAr exposures that were taken on the same nights.
+ThAr lines cover all 86 extracted orders of \expres, which includes 5295 lines.
+
+ \name\ proceeds from a given list of orders, line wavelengths, and pixel positions of each line for each exposure.
+We, therefore, use line positions generated by the pre-existing \expres\ pipeline, as describe in \lz{Petersburg2020}.
+
+ A ThAr wavelength solution is generated from each ThAr exposure using the IDL code \texttt{thid.pro}, developed by Jeff Valenti.
+This code identifies ThAr lines by matching lines in an exposure against a line atlas.
+
+ Flat-relative, optimally extracted LFC data is background-corrected using a univariate spline.
+Each peak in an order is then fit to a Gaussian.
+The mean of the fitted Gaussian is taken to be the center of the line.
+For each line, the ThAr wavelength solution is used to estimate the mode number of a line.
+The precise wavelength is then calculated using
+ \begin{equation}
+ f_n = n \times  f_r + f_0
+ \label{eq:lfc}
+ \end{equation}
+ where the repetition rate, $f_r$, is known from design of the LFC, and the offset frequency, $f_0$, has been determined experimentally.
+the pre-existing \expres\ pipeline then generates wavelength solutions on a night-by-night basis as described in Section \ref{sec:method}.
+
+ \lz{(Mention degeneracy with prescribed line center and wavelength solution here?  Or in discussion?)}
+
+ We satisfy the assumption of stability needed for \name\ by using exposures from only one ``instrumental epoch" of \expres.
+\expres\ exposures are separated into different instrumental epochs that correspond to atypical hardware changes--changes that altered the position of the echellogram on the detector, the shape of the instrumental PSF, or the calibration sources.
+Significant instrumental changes break the assumption of a stabilized instrument with only low-dimensional variations.
+We therefore treat each instrument epoch independently.
+
 \section{Tests}
-\subsection{Self Tests} \label{sec:test-self}
-For these so-called ``self tests", we use \name\ to  predict wavelengths for the measured line centers of a calibration exposure that was used to construct the basis.
-The linear interpolation in time should return identical PCA coefficients for this time, as the calibration data was used in the PCA.
+We performed a series of validation tests to test 1) the de-noising step (\textsection \ref{sec:denoising}, \textsection\ref{sec:interp_time}), 2) the interpolated wavelength solution (\textsection \ref{sec:interp_wsol}), and 3) the line density for which an interpolated wavelength solution is justified.
+These tests center on the idea of predicting wavelengths for calibration lines, and comparing the predicted wavelength for each line to the assigned wavelength.
 
-The predicted wavelengths for these measured line centers should match the assigned wavelengths for these calibration lines within the range of error of the measured line center.
-Per the EXPRES pipeline, the LFC line centers used in this analysis are measured by fitting each LFC line to a Gaussian \lz{Peteresburg2020}.
-The mean of the fitted Gaussian is set to be the line center.
-The error in this fitted value is on average 2.1 $\mps$.
-Additional error will arise from interpolation error of wavelengths to other pixel positions.
+We compare the residuals of an \name -generated wavelength solution to wavelengths generated by a classic, parametric method.
+Lastly, we compare RVs from pipelines employing \name\ vs.
+a classic wavelength calibration method.
 
-\subsection{Training and Validation Tests} \label{sec:test-trainNvalid}
-A more full-scale test is to leave out some percentage of the calibration images from $\{exp_n\}$ while constructing the basis vectors and weights with PCA.
-We can then use the resulting basis to predict wavelengths for lines in the calibrations images that were left out.
-The used calibration images represent the ``training set," while the left out exposures make up the ``validation set."
+\subsection{Test of De-Noising}\label{sec:test-denoising}
+To test how well \name\ recreates calibration lines for non-calibration exposures, we employed \name\ on 90\% of all LFC exposures.
+This means the basis vectors, $g_k(x,y,m)$,  and weights, $a_{nk}$, were constructed using only information from 90\% of all exposures.
+We used the results to predict wavelengths for the lines in the remaining 10\% of calibration exposures.
 
-For these tests, we used 90\% of available exposures to train the PCA, and left the remaining 10\% for validation.
-Exposures were sorted into either set at random.
-Contrary to the self tests, these results may also accrue errors from interpolating the principal component amplitudes with respect to time, i.e.
-we are now also testing how well we can re-construct line centers for all times.
-As before, differences in wavelength will also be drawn from how well line centers are measured in an exposure and any error in interpolating to those measured values.
-This second source of error was independently quantified in the second test described in the previous section.
-
-\lz{I had some discussion of cross tests (i.e.
-using LFC to predict ThAr and ThAr to predict LFC, but I removed it since we decided to leave ThAr mostly out of this due to its data generally being worse}
-
-\section{Results} \label{sec:results}
+\subsection{Test of Interpolation}\label{sec:test-interp}
+To test how well an interpolated wavelength solution can predict wavelengths, we employed \name\ on all LFC exposures with 
+of lines removed.
+The resultant basis vectors, $g_k(x,y,m)$,  and weights, $a_{nk}$, therefore have no information about the excluded lines.
+We then predict wavelengths for the excluded lines and compare these predictions to their assigned wavelengths.
 
@@
 \includegraphics[width=.5\textwidth]{Figures/lfcgood.png}
-\caption{Residuals of different wavelength calibration tests.
-Shown as a black outline is the distribution of residuals from the self test described in Section \ref{sec:test-self}, which incorporates errors in the fitted line centers and interpolating with respect to pixel.
-Over-plotted in blue are the residuals from the training and validation test described in Section \ref{sec:test-trainNvalid}, which builds on the self tests by additionally incorporating possible errors from interpolating line positions through time.\
-\lz{Considering adding a histogram of line center errors to show it is on par?  i.e. the limitation is the line fit.  What other information is needed?  i.e. the median/mean or spread?}}
+\caption{\lz{Results of validation tests.  Histogram of residuals?  Scatter plot colored by residual value?}}
 \label{fig:testHists}
@@
 
-Results from these tests are shown in Figure \ref{fig:testHists}.
-The result of the self tests for LFC exposures is shown as a black outline.
-This histogram combines residuals from all exposures.
-The self test represents the limit of how well the line centers at any given time can be predicted, since here we are predicting the line centers for an exposure used in the construction of the principal component amplitudes $a_{nk}$.
+\subsection{Test of Line Density}\label{sec:test-density}
+The move to an interpolated wavelength solution is driven by the assumption that a high density of calibration lines should allow for more freedom in the resultant wavelength solution.
+We test this assumption by employing \name\ on ThAr exposures.
+We then use the results to predict wavelengths for the completely independent LFC exposures taken during the same range of time.
 
-Overplotted in blue are the residuals from training the wavelength solution with only 90\% of LFC exposures and predicting the wavelengths for the remaining 10\% of exposures.
-The two histograms match very closely, indicating that very little error is introduced when predicting denoised x values for times not used in the construction of the wavelength solution.
+\subsection{Classic Model vs. \name } \label{sec:test-classic}
+To compare the classic, polynomial-driven method of wavelength calibration and an interpolated wavelength solution, we focus on one exposure of LFC lines.
+Within this exposure, we separate the even and odd lines.
+We then construct a wavelength solution using only the odd lines and use it to predict wavelengths for the even lines.
 
-\section{Comparison to Classic, Parametric Model} \label{sec:comparisons}
-Classically, pipelines employ polynomials to construct smooth wavelength solutions for each exposure.
-For example, the \expres\ pipeline described in \lz{Petersburg2020} prescribed $\theta_{n}$ from equation 1 to be the amplitudes of a 2D, 9th-order polynomial in pixel $x$ and order $m$ for each exposure $n$, making the equivalent to equation 2 the following:
-\begin{equation}
-\lambda(x,y,m,n) = \sum_{i=0}^9\sum_{j=0}^9 c_{nij}\, x^i\,m^j + \mathrm{noise}
-\quad ,
-\end{equation}
-where $c_{nij}$ are coefficients unique to exposure $n$.
-These coefficients are then smoothed by a third-order polynomial over time.
-This third-order polynomial would be evaluated at the time of non-calibration exposures to re-construct a 2D, 9th-order polynomial wavelength solution for that exposure.
-Coefficients, $c_{nij}$, were found that minimize an objective $Q$ that is the L2-norm:
-\begin{equation}
-Q = ||\lambda(x,y,m,n) - \sum_{i=0}^9\sum_{j=0}^9 c_{nij}\, x^i\,m^j||_2^2
-\quad .
-\end{equation}
+In the case of the classic model, this meant fitting a polynomial to just the odd lines and evaluating it at the position of the even lines.
+In the case of an interpolated model, this mean interpolating the wavelengths of the odd lines onto the position of the even lines.
+We then constructed a wavelength model using only the even lines and used it to predict wavelengths for the even lines.
 
-Even in this simple context, where every calibration image $n$ is treated as its own unique flower, there are improvements to be made.
-For one, the sums should not be from 0 to 9, but instead
-\begin{equation}
-\sum_{i=0}^9\sum_{j=0}^{9-i}
-\quad ,
-\end{equation}
-because that is the definition of 9th order.  We find that fitting to 8th order is sufficient, faster, and leaves less residual pattern than the original 9th degree fit.
-
-For another, the objective function could be made soft to permit catastrophic outliers without destroying the fit.
-(\lz{we did not do this}).
-We might recommend the iteratively reweighted least squares (IRLS).
-This would make the fitting more robust.
-
-For yet another, there are rescaling issues for the products $x^i\,m^j$ to protect the fitting from near-singularities or bad conditioning of the linear-algebra operators.
-
-\subsection{Comparisons}
-To test the parametric vs.
-the non-parametric method, we took an LFC exposure and separated the even and odd lines.
-We then used the even lines to predict wavelengths for the odd lines and vice versa.
-
-For example, for the parametric model, we fit a 2D, 9th-degree polynomial to the odd lines.
-We then evaluated this polynomial at the positions of the even lines and compared those returned wavelengths to the assigned wavelength of each line.
-We then reversed the process by fitting a polynomial to the even lines and evaluating that polynomial at the positions of the odd lines.
-Differences between the polynomial given wavelengths and the assigned wavelengths of each line are shown in Figure \ref{fig:parametricTest}, left.
-
-\begin{figure*}[h]
+\begin{figure*}[t]
 \centering
@@
 a non-parametric model (right-most).
-In each case, the color bar shows the difference between the predicted wavelength and the true wavelength for each line in units of velocity shift.
+In each case, the color bar shows the difference between the predicted wavelength and the true wavelength for each line in units of $/mps$.
 The center plot shows a histogram of the residuals of the polynomial fit (blue) and the non-parametric fit (green).
 Medians of the residuals for each model are given in the top-right corner of the center plot in their corresponding colors.}
-\label{fig:parametricTest}
+\label{fig:comparison}
 \end{figure*}
 
-We run the test using a polynomial derived as described in \lz{Petersburg2020} with discussed improvements.
-We repeat the test using a PCHIP cubic-spline as a model rather than a polynomial.
-The results are shown in Figure \ref{fig:parametricTest}.
-For each method, we plot each line with respect to order and x-pixel and colored by the difference between the predicted and assigned wavelength for that line in units of $\mps$.
+The results to this test are shown in Figure \ref{fig:comparison}.
+For each method, we plot each line with respect to its order (y-axis) and x-pixel on the detector (x-axis).
+Each marker is colored by the difference between the predicted and assigned wavelength for that line in units of $\mps$.
 
-The residuals of the parametric model are shown in the left-most plot of Figure \ref{fig:parametricTest} while the residuals of the non-parametric model are shown in the right-most plot.
-The center plot is a histogram of the residuals of the two methods.
-The residuals from using the non-parametric model show less structure, are less spread around zero, and more symmetric about zero.
-The non-parametric model was able to account for the high-order detector defects that are smoothed over by a polynomial fit, resulting in a more accurate fit.
+The residuals of the classic model is shown in the left-most plot of Figure \ref{fig:comparison}; the residuals of the interpolated model is shown in the right-most plot.
+The center plot is a histogram of the residuals for both methods.
 
-\begin{figure*}[h]
-\centering
-\includegraphics[width=\textwidth]{Figures/waveResids.png}
-\caption{Comparisons of parametric and non-parametric wavelength solutions using either just ThAr lines or LFC lines.
-Left: the difference between each wavelength solution evaluated at the position of the LFC lines and the assigned wavelength of that LFC line.
-The right plot is simply a zoom in of the left plot.
-Note the large deviation when trying to implement Excalibur with just ThAr lines, which are much fewer in number.
-The classic, polynomial  fit exhibits similar residuals regardless of whether the set of ThAr lines or LFC lines are used.
-Excalibur using LFC lines exhibit the lowest residuals.}
-\label{fig:waveResids}
-\end{figure*}
+The residuals of the interpolated model show much less structure, have a smaller spread, and are more symmetric about zero.
+We believe the flexibility of the interpolated model was able to account for high-order detector defects, which can be seen in the structure in the residuals of the classic, smooth, polynomial model.
 
-Incorporating a hierarchical, de-noising element results in ever better residuals!
-
-\lz{come on, future Lily, add more here}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.5\textwidth]{Figures/allHists.png}
-\caption{Histogram of residuals for different wavelength calibration methods.
-The parametric (blue) and non-parametric (green) histograms are the same as described in Figure \ref{fig:parametricTest}.
-In orange is the histogram of residuals to the full Excalibur model, which includes the hierarchical component.
-The median of the residuals for each model is given in the top-right corner in each model's corresponding color.}
-\label{fig:allHists}
-\end{figure}
-
-\section{Real Data Through to RVs} \label{sec:realdata}
-\lz{Debra and RP currently working through this.  Results seem good}
+\subsection{Impact on Radial Velocity Measurements}\label{sec:test-rv}
+\lz{Debra and RP currently working through this.  Results seem good.  Will ask them at next group meeting what we should include here}
 
 \section{Choice Choices} \label{sec:choices}
-The implementation of  \name presented here required determining global variables and methods that we believe are or are close to optimal for constructing a high-fidelity wavelength solution.
+This implementation of \name incorporated determining various global variables and methods that we believe are or are close to optimal for constructing a high-fidelity wavelength solution.
 This section will describe each choice and the associated decision-making process.
@@
 \subsection{Value of K}
-The value of $K$ determines how many principal components are used to reconstruct the de-noised line centers.
-This $K$ needs to be large enough so that all variability in the spectrograph is captured.
+The value of $K$ represents the dimensionality of the calibration space within which the spectrograph lives.
+In practice, it is the number of principal components, or basis vectors, used to reconstruct the de-noised line centers.
+$K$ needs to be large enough so that all variability in the spectrograph is captured.
 Too large, however, and the reconstruction will incorporate noise, thereby defeating the purpose.
@@
 Figure \ref{fig:pcLfc} shows the first 6 principal components constructed using LFC lines.
-In both cases there is clear structure in the first and second principal components.
+In both cases, there is clear structure in the first and second principal components.
 The diagonal-banding structure seen in components 3, 5, and 6 we believe are capturing detector fabrication defects.
@@
 %\includegraphics[width=\textwidth]{Figures/pcsLfc6.pdf}
-\caption{The first six principal components constructed using only LFC lines.
-There is clear structure for the first two principal components, though no order information is given to the PCA.
+\caption{The first six principal components constructed using LFC lines.
+There is clear structure in the first two principal components, though no order information is given to the PCA.
 Principal components 3, 5 and 6 show some slight diagonal-banding structure, which we believe to be due to detector defects.
-There is aberrant behavior in the bluest lines of the bluest two orders, which will be highlighted in the discussion.}
+There is aberrant behavior in the bluest lines of the bluest two orders, likely from variation introduced by more uncertain line fits.}
 \label{fig:pcLfc}
@@
 
-In deciding a K value, we ran training-validation tests for K values spanning from 2 to 512.
+In deciding a K value, we ran validation tests of denoising (as described in Section \ref{sec:test-denoising}) for K values spanning from 2 to 512.
 One would expect the returned wavelengths to get increasingly more accurate with larger K until components that represent only noise are incorporated.
@@
 Comparisons of wavelengths returned by a K=6 model vs.
-a K=32 model show significant differences in less than 10, bluer lines.
+a K=32 model show significant differences for less than 10, bluer lines.
 
@@
 Figure \ref{fig:nightlyVariation} shows the amplitude of the first and second principal component with respect to time on the left.
-Though the shape of the amplitudes with respect to time, a clear linear trend exists within each night.
+Though there exists a complex overall shape to the amplitudes with respect to time, a clear linear trend exists within each night.
 This is shown by the right plots in Figure  \ref{fig:nightlyVariation}.
@@
 \lz{Yeah, the right-figure is kind of a mess, isn't it?}
+
+We differentiated eras of stability by where the principal component amplitudes showed large variation.
+The analysis was done by eye here, though many algorithms for change-point detection exist in the literature \lz{(Am I going to have to cite things?)}.
+We were very conservative with defining eras of stability to ensure that within each era, the calibration-space of the instrument is low-dimensional as assumed.
 
@@
 \caption{Amplitude of the first two principal components show as a function of time (left) or fraction of a day (right).
+The top plots shows the amplitudes for the first principal component while the bottom plot shows the amplitudes for the second principal component.
 Points mark the times at which we have calibration data.
 Lines show the result of a linear interpolation.
-Left: amplitudes as a function of time for both LFC (blue) and ThAr (green).
-The top plot shows the amplitudes for the first principal component while the bottom plot shows the amplitudes for the second principal component.
-The shape of the curves is similar regardless of calibration source.
-Right: amplitudes of the first two principal components determined by the LFC exposures only.
-Here, exposures are grouped by day and plotted with respect to what time of day the exposure was taken.
-A linear interpolation through all the LFC exposures taken in one day are shown as black line.
-Points are colored by the MJD of the exposure.
-For each day, the amplitude has been artificially offset by the median amplitude for that day so all days are roughly on the same scale.}
+In the right plot, the principal component amplitudes have been artificially offset by the median amplitude for each day so that all days are roughly on the same scale.
+Points are colored by the MJD of each exposure.}
 \label{fig:nightlyVariation}
@@
 Interpolation of wavelengths over pixel  is done order by order using a Piecewise Cubic Hermite Interpolating Polynomial (PCHIP) interpolator.
-This interpolation incorporates the flexibility needed to model the changing dispersion of the spectrograph across an order along with any detector defects while also enforcing monotonicity.
+This interpolation incorporates the flexibility needed to model the changing dispersion of the spectrograph across an order along with any detector defects while also enforcing monotonicity, which we know must be true across any one order.
 
-Due to the dispersion intrinsic to echelle spectrographs, the wavelength change between pixels is greater at greater wavelengths.
+Due to the dispersion intrinsic to echelle spectrographs, the wavelength change between pixels grows greater with greater wavelengths.
 This means that the function of wavelength vs.
@@
 \caption{Results from different interpolation schemes over pixels.
-Left: resultant wavelength solutions generated using ThAr lines, shown as green points.
-Different interpolation schemes are shown in different colors.
-The largest deviation comes with the classic cubic spline (green).
-Right: histograms of residuals for each method when used to predict the wavelengths of LFC lines, shown as black, vertical lines.
+ThAr lines, shown as pink points, are used to construct a wavelength solution that is then evaluated at each LFC line, shown as black lines.
+Histograms of the residuals for each method is shown on the right.
 The linear interpolation model (yellow) of this concave down relation is asymmetric with disproportionately positive residuals.
-The popular cubic spline residuals (green) has very spread out residuals.
-The residuals from an interpolation scheme that incorporates smoothing (blue) appear non-Gaussian with a plateau about zero.
-Residuals from the PCHIP interpolator (orange) has the strongest peak at 0 and is roughly Gaussian.}
+The popular cubic spline residuals (green) has very spread out residuals and shows huge deviations following a ThAr line blend at pixel \lz{(tkWill Find)}.
+The residuals from an interpolation scheme that incorporates smoothing (blue) are non-Gaussian with a plateau about zero, as it chronically over-estimates the middle of the order and under-estimates the edges (not shown).
+Residuals from the PCHIP interpolator (orange) has the strongest peak at 0 and appears Gaussian.}
 \label{fig:xinterp}
@@
 We instead turn to the PCHIP interpolator, which damps down huge deviations in the traditional cubic spline by requiring the resulting interpolated function to be monotonic.
-This is a constraint we know is true from optics, since interpolation in the pixel direction is carried out order-by-order.
+This is a constraint we know is true for any one order.
 In Figure \ref{fig:xinterp} we show that using the PCHIP interpolator returns the lowest residuals.
@@
 
-\lz{We talked about including this discussion, but there is no real reason for doing things in this order except it is quite a bit faster.}
+\lz{We talked about including this discussion, but there is no real reason for doing things in this order except it is a bit faster.}
 
 \section{Discussion} \label{sec:discussion}
-\lz{I feel like we have a lot of hopes for this discussion section, which is not matched by my understanding of how to structure this discussion section.
-For now, I have just written brief snippets that address each of the notes I have of things we have said we want to include.}
+The endeavor to reach $0.1\,\mps$ precision necessitates improvements on all levels of spectroscopy, from the instrumentation to the data analysis.
+In this paper, we present \name\, a hierarchical, non-parametric model for wavelength calibration.
 
-Generalization to etalon:
-All methods described here are easily adapted to use with an etalon or any other arc lamp for which a list of lines and their assigned wavelengths can be constructed.
+Starting with a list of calibration lines with known wavelengths and well-fit line centers for each calibration exposure that faithfully represent the calibration state of the spectrograph, \name\ will de-noise and interpolate the given lines into a full wavelength solution.
+We show that it returns more accurate wavelength predictions by a factor of 10 than classic, parametric methods.
+Using \name\ wavelengths reduced the RMS in a planet fit to derived RVs by \lz{some percent hopefully probs}.
 
-Ability to vet for bad exposures:
-An advantage of using PCA on all LFC exposures and line positions is the ability to isolate exposures that exhibit unexpectedly large variation.
-Such variation is likely to be captured by an early principal component.
-Errant exposures will be singled out by a spike in the amplitude for just that principal component.
-This allowed us to quickly find flawed LFC exposures, which otherwise would have required visual inspection of all 1200+ LFC exposures.
+\name\ levies the incredible stability of contemporary EPRV instruments and high density of lines made available by new calibration sources, such as LFCs and etalons, to achieve more accurate wavelengths.
+Denser calibration lines allow us to move to more flexible, interpolated wavelength solutions.
+Stabilized spectrograph hardware with few degrees of freedom allow us to use all calibration images in a given generation of stability to constrain the accessible calibration space of the spectrograph.
 
-Changes in instrument/change point detection:
-On the other hand, this same feature of using PCA can also be seen as a bug.
-PCA is good at picking up on variation regardless of the source of the variation.
-Any variation that breaks our assumption of an instrument that exists in only a low-dimensional calibration space will likely be captured by early principal components, which will no longer serve to denoise the calibration data.
-We found that violations of this stability assumption typically manifests as change points in the principal component amplitudes with respect to time.
+\lz{Choices?}
 
-Sensitivity to line-position fitting:
-The PCA will also be sensitive to variations in line fitting.
-For example, if certain regions have lines of lower signal, variable background, or are otherwise harder to fit, any resultant variation in the returned line positions across different exposures will also be captured by any PCA.
+An advantage of applying PCA to line positions from all LFC exposures is the ability to isolate exposures that exhibit errant variation, which is typically associated with flawed exposures.
+This allowed us to quickly vet for problematic LFC exposures, which otherwise would have required visual inspection of all 1200+ LFC exposures.
+In a classic framework where each calibration exposure is treated independently, these aberrant exposures would likely have persisted undetected and could potentially sway an entire night's of wavelength solutions.
+
+On the other hand, the ability of PCA to pick up on any and all variation, regardless of source, also makes \name\ very sensitive to errors in line fitting.
+For example, we have seen that lower-signal lines that are harder to fit faithfully will have greater variety in returned line positions, which is in turn captured by the PCA.
 High-fidelity line positions are essential to ensure the PCA is capturing variations in the spectrograph's calibration rather than changes in how well a line can be fit.
 
-Perfect degeneracy problem:
-With any wavelength solution, there is a perfect degeneracy between what is defined as the ``center of the line'' and the resultant wavelength solution.
+\name\ can be applied to any data that contains information about the calibration state of the spectrograph.
+For example, though LFC and ThAr exposures are used as an example in this paper, \name\ would work exactly the same for an etalon or any other arc lamp with a list of lines and assigned wavelengths.
+
+Furthermore, once we have defined a calibration space that captures all possible degrees of freedom for a stabilized spectrograph, then all exposures taken with the spectrograph contains information about where the spectrograph lies within that calibration space.
+This means that science exposures themselves could also be used to determine the calibration state of an instrument, negating the need for any additional wavelength-calibration source beyond defining the calibration space.
+
+We have described only one, fairly simplistic implementation of \name\ here.
+There are many options for both the de-noising and interpolation steps.
+\lz{(Maybe something to discuss at a future meeting; I'm not actually entirely clear on other possibilities)}.
+For instance, a Gaussian process model could be constructed to determine wavelengths in place of a spline.
+\lz{(?)}  \lz{Something about going Bayesian?}
+
+\lz{(Not sure where this belongs, honestly.)}  We caution that with any wavelength solution, there is a perfect degeneracy between what is defined as the ``center of the line'' and the resultant wavelength solution.
 If, for example, a cross correlation method is used to extract RVs from the data, a systematic difference may be introduced depending on the definition of the center of the line.
@@
 
-Self calibration idea:
-Once we define a calibration space that captures all possible degrees of freedom for a stabilized spectrograph, then all exposures taken with the spectrograph contain information about where the spectrograph lies in that space at the time of the exposure.
-Meaning, a science exposure can be used to determine the current state of the instrument in this calibration space.
-It would therefore be possible to self-calibrate science exposures without the use of an additional wavelength calibration source.
+\name\ is designed and optimized for extreme-precision instruments.
+We believe that by implementing \name, the error from wavelength calibration is rendered negligible in the face of other error sources.
 
-How hard would it be to go bayesian?  Description of other implementations of \name\ than what we did here?
+\pagebreak
+\section{Graveyard of Rejected Figures}
+I will mine these for caption material later before their final death.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=.5\textwidth]{Figures/lfcgood.png}
+\caption{\lz{Eventually results from the tests actually described in this paper?}.
+Residuals of different wavelength calibration tests.
+Shown as a black outline is the distribution of residuals from the self test described in Section \ref{sec:test-self}, which incorporates errors in the fitted line centers and interpolating with respect to pixel.
+Over-plotted in blue are the residuals from the training and validation test described in Section \ref{sec:test-trainNvalid}, which builds on the self tests by additionally incorporating possible errors from interpolating line positions through time.\
+\lz{Considering adding a histogram of line center errors to show it is on par?  i.e. the limitation is the line fit.  What other information is needed?  i.e. the median/mean or spread?}}
+\label{fig:testHists}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.5\textwidth]{Figures/allHists.png}
+\caption{\lz{This and the previous histogram are eventually going to be one histogram, right?} Histogram of residuals for different wavelength calibration methods.
+The parametric (blue) and non-parametric (green) histograms are the same as described in Figure \ref{fig:parametricTest}.
+In orange is the histogram of residuals to the full Excalibur model, which includes the hierarchical component.
+The median of the residuals for each model is given in the top-right corner in each model's corresponding color.}
+\label{fig:allHists}
+\end{figure}
+
+\begin{figure*}[h]
+\centering
+\includegraphics[width=\textwidth]{Figures/waveResids.png}
+\caption{\lz{(Good companion figure to the line density discussion, but should also have another figure that has a bigger picture)} Comparisons of parametric and non-parametric wavelength solutions using either just ThAr lines or LFC lines.
+Left: the difference between each wavelength solution evaluated at the position of the LFC lines and the assigned wavelength of that LFC line.
+The right plot is simply a zoom in of the left plot.
+Note the large deviation when trying to implement Excalibur with just ThAr lines, which are much fewer in number.
+The classic, polynomial  fit exhibits similar residuals regardless of whether the set of ThAr lines or LFC lines are used.
+Excalibur using LFC lines exhibit the lowest residuals.}
+\label{fig:waveResids}
+\end{figure*}
 

### 2020-04-30 paper/paper.tex
Results, updated figures, citations

@@
 We demonstrate the success of this method with data from a Menlo Systems laser frequency comb taken by \expres, the Extreme Precision Spectrograph.
-We find that \name\ returns more accurate wavelength predictions by a factor of 10 than classic polynomial models.
+We find that \name\ returns more accurate wavelength predictions by more than a factor of 2 over classic polynomial models.
 Employing \name\ wavelengths reduced the RMS of RV measurements to a planet fit by \lz{some percent}.
@@
 \section{Introduction} 
-Extreme precision radial-velocity (\eprv) programs have been incredibly fruitful in finding and characterizing extra-solar planets (CITE THINGS).
+Precision radial-velocity (\eprv) programs have been incredibly fruitful in finding and characterizing extra-solar planets \citep{mayor2011, bonfils2013, plavchan2015, butler2017}.
 These programs typically make use of spectrographs with resolutions on the order of $10^5$, which correspond to line widths on the order of $3000\,\mps$.
-Exoplanet science happens at the level of $3\,\mps$ precision, with current instruments hoping to reach $0.1\,\mps$ precision (CITE THINGS).
+Exoplanet science happens at the level of $3\,\mps$ precision, with current instruments hoping to reach $0.1\,\mps$ precision \citep{fischer2016}.
 This requires new spectrographs to be calibrated or stabilized to better than $10^{-4}$ of a pixel (assuming that the spectrograph is well sampled).
-Some hardware--software systems, such as ESPRESSO and \expres\ are now reaching these levels (CITE THINGS).
+Some hardware--software systems, such as ESPRESSO and \expres\ are now reaching these levels \citep{pepe2013,  jurgenson2016, neid, blackman2020}.
 
@@
 The density of lines removes the need to enforce any functional form for the wavelength solution (such as a continuous eighth-order polynomial, for example).
-We expand on the work of Milakovi\'{c}2020, which demonstrated the efficacy of constructing a wavelength solution as multiple, segmented polynomials.
+We expand on the work of \cite{milakovic2020}, which demonstrated the efficacy of constructing a wavelength solution as multiple, segmented polynomials.
 Going non-parametric will improve calibration accuracy by not forcing the choice of a parametric form that may bias the calibration, especially when the chosen function is inappropriate (as, for example, polynomials will be at detector edges).
@@
 
-We therefore assume dense enough calibration line coverage with well-fit line centers to provide some constraint on an interpolated wavelength solution across an order.
+We therefore assume dense enough calibration line coverage with well-fit line centers to provide some constraint on an interpolated wavelength solution across an echelle order.
 Additionally, we assume the instrument hardware is stabilized enough that the space of possible calibration states is low-dimensional.
 
-We will pose the question of wavelength calibration in the following way.  Given an exposure $n$, and order $m$, there is a relationship between
+We will pose the question of wavelength calibration in the following way.  Given an exposure $n$, and echelle order $m$, there is a relationship between
 the two-dimensional $(x,y)$-position on the detector and the
@@
 Classically, pipelines employ polynomials to construct smooth wavelength solutions for each exposure.
-For example, the pipeline described in \lz{Petersburg2020} sets the function $f(x,y,m;\theta_{n})$ from Equation \ref{eq:wsol} to be a 2D, 9\textsuperscript{th}-order polynomial where $\theta_{n}$ represents the polynomial coefficients, $c_{nij}$, unique to each exposure $n$.
+For example, the pipeline described in \cite{petersburg2020} sets the function $f(x,y,m;\theta_{n})$ from Equation \ref{eq:wsol} to be a 2D, 9\textsuperscript{th}-order polynomial where $\theta_{n}$ represents the polynomial coefficients, $c_{nij}$, unique to each exposure $n$.
 \begin{equation}
@@
 \subsection{De-Noising of Calibration Frames} \label{sec:denoising}
-\lz{Concerns: 1) how to best phrase the problem using variables?, 2) how in-depth is too in-depth?}
-
 Within a given range of stability, we can use all calibration images to  1) determine the space in which the instrument varies and 2) where in the accessible calibration space the spectrograph existed for each exposure.
@@
 The first task is to identify the full list of lines we can expect to find in a calibration image.
-A line, $l_{m,\lambda}$ is uniquely defined by a combination of order, $m$, and ``true" or theoretical wavelength, $\lambda$.
+A line, $l_{m,\lambda}$ is uniquely defined by a combination of echelle order, $m$, and ``true" or theoretical wavelength, $\lambda$.
 
-For each exposure, $exp_n$, every line, $l_{m,\lambda}$, has an associated fitted detector position, $x_{n,l}$, for example x-pixel in an 2D extracted order.
+For each exposure, $exp_n$, every line, $l_{m,\lambda}$, has an associated fitted detector position, $x_{n,l}$, for example x-pixel in an 2D extracted echelle order.
 Fitted line centers that are missing from an exposure (e.g.
-because the fit failed due to noise, the line is not in its usual order, etc.) can be assigned a $NaN$ for that order instead.
+because the fit failed due to noise, the line is not in its usual echelle order, etc.) can be assigned a $NaN$ for that echelle order instead.
 Let there be $M$ lines per exposure.
@@
 \SetAlgoLined
-\textbf{Inputs}: $m$, order of each line; $\lambda$, assigned wavelength of each line; $[x_{n,l}] $, matrix of pixel line position of each line $l$ and for each exposure $n$\;
+\textbf{Inputs}: $m$, echelle order of each line; $\lambda$, assigned wavelength of each line; $[x_{n,l}] $, matrix of pixel line position of each line $l$ and for each exposure $n$\;
 \While{change in missing or outlier line centers $>$ 0.01\%}{
@@
 For instance, interpolating the known wavelengths vs.
-line centers onto every integer x will generate wavelengths for each pixel in an order.
+line centers onto every integer x will generate wavelengths for each pixel in an echelle order.
 
@@
 We tested \name\ results using data from \expres\, the EXtreme PRecison Spectrograph.
-\expres\ is an environmentally-stabilized, fiber-fed, $R\sim137,000$, optical spectrograph (\lz{Jurgenson2016, Blackman2020}).
+\expres\ is an environmentally-stabilized, fiber-fed, $R\sim137,000$, optical spectrograph \citep{jurgenson2016, blackman2020}.
 \expres\ has two different wavelength calibration sources, a ThAr lamp and a Menlo Systems laser frequency comb (LFC, e.g.
-\lz{Wilken2012, Molaro2013, Probst2014}).
+\cite{wilken2012, molaro2013, probst2014}).
 
@@
 
- \name\ proceeds from a given list of orders, line wavelengths, and pixel positions of each line for each exposure.
-We, therefore, use line positions generated by the pre-existing \expres\ pipeline, as describe in \lz{Petersburg2020}.
+ \name\ proceeds from a given list of echelle orders, line wavelengths, and pixel positions of each line for each exposure.
+We, therefore, use line positions generated by the pre-existing \expres\ pipeline, as described in \cite{petersburg2020}.
 
@@
  Flat-relative, optimally extracted LFC data is background-corrected using a univariate spline.
-Each peak in an order is then fit to a Gaussian.
+Each peak in an echelle order is then fit to a Gaussian.
 The mean of the fitted Gaussian is taken to be the center of the line.
@@
 
-\section{Tests}
-We performed a series of validation tests to test 1) the de-noising step (\textsection \ref{sec:denoising}, \textsection\ref{sec:interp_time}), 2) the interpolated wavelength solution (\textsection \ref{sec:interp_wsol}), and 3) the line density for which an interpolated wavelength solution is justified.
-These tests center on the idea of predicting wavelengths for calibration lines, and comparing the predicted wavelength for each line to the assigned wavelength.
+\section{Tests}\label{sec:tests}
+We perform a series of tests to validate the performance of \name\ and benchmark \name -generated wavelengths against wavelengths generated by a classic, parametric method.
+For all these tests, we generate wavelengths for known calibration lines and compare the predicted wavelength to the assigned wavelength of each line.
+This inherently folds in error from the measured line center of each calibration line, but the contribution to the residuals will be the same across all tests.
 
-We compare the residuals of an \name -generated wavelength solution to wavelengths generated by a classic, parametric method.
-Lastly, we compare RVs from pipelines employing \name\ vs.
-a classic wavelength calibration method.
+To assess the classic, polynomial-driven method of wavelength calibration, we take each LFC exposure and separate the lines into even and odd lines.
+We then construct a wavelength solution using only the odd lines and use it to predict wavelengths for the even lines; i.e.
+a polynomial is fit to just the odd lines and then evaluated at the detector position of the even lines.
+We then generate a wavelength solution using only the even lines and use it to predict wavelengths for the odd lines.
 
-\subsection{Test of De-Noising}\label{sec:test-denoising}
-To test how well \name\ recreates calibration lines for non-calibration exposures, we employed \name\ on 90\% of all LFC exposures.
+To test the interpolation step of \name\ (\textsection \ref{sec:interp_wsol}), we employed \name\ on all LFC exposures with odd lines removed.
+The resultant basis vectors, $g_k(x,y,m)$,  and weights, $a_{nk}$, are therefore only informed by the even lines of each LFC exposure.
+We then predict wavelengths for the odd lines that had been excluded and compare these predictions to their assigned wavelengths.
+This allows us to test how accurately an interpolated wavelength solution can predict wavelengths.
+
+To test the denoising step of \name\ (\textsection \ref{sec:denoising}, \textsection\ref{sec:interp_time}), we employed \name\ on 90\% of all LFC exposures.
 This means the basis vectors, $g_k(x,y,m)$,  and weights, $a_{nk}$, were constructed using only information from 90\% of all exposures.
-We used the results to predict wavelengths for the lines in the remaining 10\% of calibration exposures.
+We used the results to predict wavelengths for all the lines in the remaining 10\% of calibration exposures.
+This allows us to test how well we can pinpoint the calibration state of the spectrograph using \name.
 
-\subsection{Test of Interpolation}\label{sec:test-interp}
-To test how well an interpolated wavelength solution can predict wavelengths, we employed \name\ on all LFC exposures with 
-of lines removed.
-The resultant basis vectors, $g_k(x,y,m)$,  and weights, $a_{nk}$, therefore have no information about the excluded lines.
-We then predict wavelengths for the excluded lines and compare these predictions to their assigned wavelengths.
+Histograms of the residuals for each of the above described polynomial, interpolation, and denoising tests (respectively) are shown in Figure \ref{fig:testHists}.
+Note that the spread in residuals is much smaller for both the denoising and interpolation tests relative to the results of the polynomial model.
 
@@
 \centering
-\includegraphics[width=.5\textwidth]{Figures/lfcgood.png}
-\caption{\lz{Results of validation tests.  Histogram of residuals?  Scatter plot colored by residual value?}}
+\includegraphics[width=.5\textwidth]{Figures/all_results.pdf}
+\caption{Difference in predicted and assigned wavelengths for the wavelength calibration tests described in Section \ref{sec:tests}.
+The standard deviation of the residuals for each model is given in the top-right corner in each model's corresponding color.}
 \label{fig:testHists}
@@
 
-\subsection{Test of Line Density}\label{sec:test-density}
-The move to an interpolated wavelength solution is driven by the assumption that a high density of calibration lines should allow for more freedom in the resultant wavelength solution.
-We test this assumption by employing \name\ on ThAr exposures.
-We then use the results to predict wavelengths for the completely independent LFC exposures taken during the same range of time.
-
-\subsection{Classic Model vs. \name } \label{sec:test-classic}
-To compare the classic, polynomial-driven method of wavelength calibration and an interpolated wavelength solution, we focus on one exposure of LFC lines.
-Within this exposure, we separate the even and odd lines.
-We then construct a wavelength solution using only the odd lines and use it to predict wavelengths for the even lines.
-
-In the case of the classic model, this meant fitting a polynomial to just the odd lines and evaluating it at the position of the even lines.
-In the case of an interpolated model, this mean interpolating the wavelengths of the odd lines onto the position of the even lines.
-We then constructed a wavelength model using only the even lines and used it to predict wavelengths for the even lines.
+\name -generated wavelengths also exhibit less structure in the returned residuals.
+For a randomly selected example LFC exposure, Figure \ref{fig:resid2d} plots each line with respect to its echelle order (y-axis) and x-pixel on the detector (x-axis) colored by the difference between the predicted and assigned wavelength for that line in units of $\mps$.
 
@@
 \centering
-\includegraphics[width=\textwidth]{Figures/parametricTest.png}
-%\includegraphics[width=\textwidth]{Figures/parametricTest.pdf}
-\caption{Comparison of residuals to a parametric model (left-most) vs.
-a non-parametric model (right-most).
-In each case, the color bar shows the difference between the predicted wavelength and the true wavelength for each line in units of $/mps$.
-The center plot shows a histogram of the residuals of the polynomial fit (blue) and the non-parametric fit (green).
-Medians of the residuals for each model are given in the top-right corner of the center plot in their corresponding colors.}
-\label{fig:comparison}
+\includegraphics[width=\textwidth]{Figures/lineResids2D.png}
+%\includegraphics[width=\textwidth]{Figures/lineResids2D.pdf}
+\caption{Residuals plotted with respect to detector position (as defined by echelle order and x-pixel) for various wavelength models.
+Each line is colored by the difference between the predicted wavelength and the true wavelength for each line, given in units of $/mps$.
+Residuals to a classic, polynomial fit is shown in the left-most plot.
+The center plot shows residuals to an interpolated wavelength solution.
+The right-most plot shows residuals when a ThAr wavelength solution is interpolated onto LFC line positions.
+The position of ThAr lines are over-plotted in black.}
+\label{fig:resid2d}
 \end{figure*}
 
-The results to this test are shown in Figure \ref{fig:comparison}.
-For each method, we plot each line with respect to its order (y-axis) and x-pixel on the detector (x-axis).
-Each marker is colored by the difference between the predicted and assigned wavelength for that line in units of $\mps$.
+\begin{figure*}[t]
+\centering
+\includegraphics[width=\textwidth]{Figures/lineDensity.png}
+\caption{\lz{(Figure I use for talks, maybe shouldn't be included here)}  Comparison of polynomial-fit and interpolated wavelength solutions using either just ThAr lines or LFC lines.
+Left: the difference between each wavelength solution evaluated at the position of the LFC lines and the assigned wavelength of that LFC line.
+The right plot is simply a zoom in of the left plot.
+Note the large deviation when trying to implement Excalibur with just ThAr lines, which exhibit much large separations than LFC lines.
+The classic, polynomial  fit exhibits similar residuals regardless of whether the set of ThAr lines or LFC lines are used.
+Excalibur using LFC lines exhibit the lowest residuals.}
+\label{fig:waveResids}
+\end{figure*}
 
-The residuals of the classic model is shown in the left-most plot of Figure \ref{fig:comparison}; the residuals of the interpolated model is shown in the right-most plot.
-The center plot is a histogram of the residuals for both methods.
+The residuals of the classic model is shown in the left-most plot of Figure \ref{fig:resid2d}.  There is a lot of vertical structure and some hints of a periodic, diagonal structure as well.
 
-The residuals of the interpolated model show much less structure, have a smaller spread, and are more symmetric about zero.
-We believe the flexibility of the interpolated model was able to account for high-order detector defects, which can be seen in the structure in the residuals of the classic, smooth, polynomial model.
+The residuals of the interpolation test for the same exposure is shown in the center plot of Figure \ref{fig:resid2d}.
+There is no coherent structure here and smaller residuals.
+We believe the flexibility of the interpolated model was able to account for high-order detector defects, which emerged as structure in the residuals of the classic, smooth, polynomial model.
+
+The right-most plot of Figure \ref{fig:resid2d} shows the residuals when ThAr exposures, which have much fewer lines than LFC exposures, are run through \name\ and used to predict wavelengths for the completely independent LFC exposures taken during the same range of time.
+Over-plotted in black are the positions of the ThAr lines.
+
+Though the interpolated wavelength solution returns lower, less-structured residuals than the polynomial model when guided by LFC lines, we see in the right-most plot of Figure \ref{fig:resid2d} that an interpolated wavelength solution can return much worse results in the space between widely separated calibration lines.
+
+The move to an interpolated wavelength solution is driven by the assumption that a high density of calibration lines allows for more freedom in the resultant wavelength solution.
+Introducing this flexibility is no longer justified in the regime where there are large spaces between calibration lines, as is the case with a classic ThAr lamp.
+
+Note that running \name\ informed by only ThAr lines cannot be regarded as a direct comparison to the LFC runs, as the increased error and variability in ThAr line positions alone makes the resultant wavelength predictions an order of magnitude worse, hence the different scale of the colorbar.
+All the same, the residuals are relatively worse where lines are further spaced (for example, in redder echelle orders) than where lines are denser.
 
@@
 \section{Choice Choices} \label{sec:choices}
-This implementation of \name incorporated determining various global variables and methods that we believe are or are close to optimal for constructing a high-fidelity wavelength solution.
+This implementation of \name\ presented here incorporated determining various global variables and methods that we believe are or are close to optimal for constructing a high-fidelity wavelength solution.
 This section will describe each choice and the associated decision-making process.
@@
 $K$ needs to be large enough so that all variability in the spectrograph is captured.
-Too large, however, and the reconstruction will incorporate noise, thereby defeating the purpose.
+Too large, however, and the reconstruction will begin to incorporate noise, thereby defeating the purpose.
 
@@
 In both cases, there is clear structure in the first and second principal components.
-The diagonal-banding structure seen in components 3, 5, and 6 we believe are capturing detector fabrication defects.
-Components 3 and 4 show aberrant behavior on the edges of the two bluest orders, where lower signal results in more variation in the line fits.
+The diagonal-banding structure seen in components 3, 5, and 6 we believe to be capturing detector fabrication defects.
+Components 3 and 4 show aberrant behavior on the edges of the two bluest echelle orders, where lower signal results in more variation in the line fits.
 
@@
 \caption{The first six principal components constructed using LFC lines.
-There is clear structure in the first two principal components, though no order information is given to the PCA.
-Principal components 3, 5 and 6 show some slight diagonal-banding structure, which we believe to be due to detector defects.
-There is aberrant behavior in the bluest lines of the bluest two orders, likely from variation introduced by more uncertain line fits.}
+Each line is plotted according to its echelle order and x-pixel and colored by the value of the principal component for that line.}
 \label{fig:pcLfc}
@@
 
-In deciding a K value, we ran validation tests of denoising (as described in Section \ref{sec:test-denoising}) for K values spanning from 2 to 512.
+In deciding a K value, we ran denoising tests, as described in Section \ref{sec:tests}, for K values spanning from 2 to 512.
 One would expect the returned wavelengths to get increasingly more accurate with larger K until components that represent only noise are incorporated.
-We found that the returned wavelengths are most accurate with a K value of 32.
-However, the improvement is slight with K values greater than 6.
+We found that the returned wavelengths have the lowest residuals with a K value of 32.
+However, the improvement is slight with K values greater than six.
 Comparisons of wavelengths returned by a K=6 model vs.
 a K=32 model show significant differences for less than 10, bluer lines.
+A visual inspection of the principal components, such as the one displayed in Figure \ref{fig:pcLfc}, led us to believe most of the structure is being captured in the first six principal components.
 
@@
 In the right plot, the principal component amplitudes have been artificially offset by the median amplitude for each day so that all days are roughly on the same scale.
-Points are colored by the MJD of each exposure.}
+Here, points are colored by the MJD of each exposure.}
 \label{fig:nightlyVariation}
@@
 Interpolation of wavelengths over pixel  is done order by order using a Piecewise Cubic Hermite Interpolating Polynomial (PCHIP) interpolator.
-This interpolation incorporates the flexibility needed to model the changing dispersion of the spectrograph across an order along with any detector defects while also enforcing monotonicity, which we know must be true across any one order.
+This interpolation incorporates the flexibility needed to model the changing dispersion of the spectrograph across an echelle order along with any detector defects while also enforcing monotonicity, which we know must be true across any one echelle order.
 
@@
 This means that the function of wavelength vs.
-pixel across an order will always be monotonically increasing and concave down everywhere.
-Consequently, linear interpolation of this function will systematically give erroneously low values.
+pixel across an echelle order will always be monotonically increasing and concave down everywhere.
+A linear interpolation would obviously give erroneously low values everywhere.
 
-A more classic cubic spline interpolation can run into issues with arc lines, which are irregularly spaced or even blended and so can appear very close together.
-Very close lines cause huge deviations from the correct wavelengths, as seen in the green line of Figure \ref{fig:xinterp}..
+A more classic cubic spline interpolation can run into issues with arc lines, which are irregularly spaced or even blended and can therefore appear very close together.
+Very close lines cause huge deviations from the correct wavelengths, as seen in the green line of Figure \ref{fig:xinterp}.
 
@@
 Histograms of the residuals for each method is shown on the right.
-The linear interpolation model (yellow) of this concave down relation is asymmetric with disproportionately positive residuals.
-The popular cubic spline residuals (green) has very spread out residuals and shows huge deviations following a ThAr line blend at pixel \lz{(tkWill Find)}.
-The residuals from an interpolation scheme that incorporates smoothing (blue) are non-Gaussian with a plateau about zero, as it chronically over-estimates the middle of the order and under-estimates the edges (not shown).
-Residuals from the PCHIP interpolator (orange) has the strongest peak at 0 and appears Gaussian.}
+Note: there is a blended ThAr line at approximately pixel 5300.}
 \label{fig:xinterp}
@@
 These huge digressions can be avoided by allowing for some smoothing in the interpolation.
-In Figure \ref{fig:xinterp}, we show an example in blue using scipy's Univarate Spline method.
+In Figure \ref{fig:xinterp}, we show an example in blue using SciPy's implementation of a univarate spline.
 While the result appears to follow the calibration lines much better, the smoothing ultimately causes larger residuals that are spatially correlated (Fig.
 \ref{fig:xinterp}, right).
+In all echelle orders, the edges are overestimated while the middle will be underestimated.
+The resultant wavelength solution is underestimating the curvature of the pixel-wavelength relation, giving rise to similar issues as with an inflexible, parametric model.
+Introducing this smoothing parameter is enforcing a smoothness we have no reason to believe is true of the data.
 
-For all orders, the edges will be overestimated while the middle will be underestimated.
-This indicates that the smoothing is causing the resultant wavelength solution to underestimate the curvature of the pixel-wavelength relation, resulting in similar issues as with an inflexible, parametric model.
-This enforces a smoothness we have no reason to believe is true for the data.
-
-We instead turn to the PCHIP interpolator, which damps down huge deviations in the traditional cubic spline by requiring the resulting interpolated function to be monotonic.
-This is a constraint we know is true for any one order.
+We instead turn to the PCHIP interpolator, which damps down huge deviations in the traditional cubic spline by requiring the resulting interpolated function be monotonic.
+This is a constraint we know is true for any one echelle order.
 In Figure \ref{fig:xinterp} we show that using the PCHIP interpolator returns the lowest residuals.
@@
 
-Starting with a list of calibration lines with known wavelengths and well-fit line centers for each calibration exposure that faithfully represent the calibration state of the spectrograph, \name\ will de-noise and interpolate the given lines into a full wavelength solution.
-We show that it returns more accurate wavelength predictions by a factor of 10 than classic, parametric methods.
+Starting with a list of calibration lines with known wavelengths and well-fit line centers for each calibration exposure, \name\ will de-noise and interpolate the given lines into a full wavelength solution.
+We show that \name\ returns residuals with greater than a factor of two less spread than classic, parametric methods.
 Using \name\ wavelengths reduced the RMS in a planet fit to derived RVs by \lz{some percent hopefully probs}.
@@
 This allowed us to quickly vet for problematic LFC exposures, which otherwise would have required visual inspection of all 1200+ LFC exposures.
-In a classic framework where each calibration exposure is treated independently, these aberrant exposures would likely have persisted undetected and could potentially sway an entire night's of wavelength solutions.
+In a classic framework where each calibration exposure is treated independently, these aberrant exposures would likely have persisted undetected and could potentially sway the resultant wavelength solutions.
 
@@
 Furthermore, once we have defined a calibration space that captures all possible degrees of freedom for a stabilized spectrograph, then all exposures taken with the spectrograph contains information about where the spectrograph lies within that calibration space.
-This means that science exposures themselves could also be used to determine the calibration state of an instrument, negating the need for any additional wavelength-calibration source beyond defining the calibration space.
+Theoretically, science exposures themselves could also be used to determine the calibration state of an instrument, negating the need for any additional wavelength-calibration source beyond defining the calibration space.
 
@@
 
-\pagebreak
-\section{Graveyard of Rejected Figures}
-I will mine these for caption material later before their final death.
+\software{SciPy library \citep{scipy}, NumPy \citep{numpy, numpy2}, Astropy \citep{astropy:2013,astropy:2018}.}
 
-\begin{figure}[h]
-\centering
-\includegraphics[width=.5\textwidth]{Figures/lfcgood.png}
-\caption{\lz{Eventually results from the tests actually described in this paper?}.
-Residuals of different wavelength calibration tests.
-Shown as a black outline is the distribution of residuals from the self test described in Section \ref{sec:test-self}, which incorporates errors in the fitted line centers and interpolating with respect to pixel.
-Over-plotted in blue are the residuals from the training and validation test described in Section \ref{sec:test-trainNvalid}, which builds on the self tests by additionally incorporating possible errors from interpolating line positions through time.\
-\lz{Considering adding a histogram of line center errors to show it is on par?  i.e. the limitation is the line fit.  What other information is needed?  i.e. the median/mean or spread?}}
-\label{fig:testHists}
-\end{figure}
+\facilities{LDT}
 
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.5\textwidth]{Figures/allHists.png}
-\caption{\lz{This and the previous histogram are eventually going to be one histogram, right?} Histogram of residuals for different wavelength calibration methods.
-The parametric (blue) and non-parametric (green) histograms are the same as described in Figure \ref{fig:parametricTest}.
-In orange is the histogram of residuals to the full Excalibur model, which includes the hierarchical component.
-The median of the residuals for each model is given in the top-right corner in each model's corresponding color.}
-\label{fig:allHists}
-\end{figure}
-
-\begin{figure*}[h]
-\centering
-\includegraphics[width=\textwidth]{Figures/waveResids.png}
-\caption{\lz{(Good companion figure to the line density discussion, but should also have another figure that has a bigger picture)} Comparisons of parametric and non-parametric wavelength solutions using either just ThAr lines or LFC lines.
-Left: the difference between each wavelength solution evaluated at the position of the LFC lines and the assigned wavelength of that LFC line.
-The right plot is simply a zoom in of the left plot.
-Note the large deviation when trying to implement Excalibur with just ThAr lines, which are much fewer in number.
-The classic, polynomial  fit exhibits similar residuals regardless of whether the set of ThAr lines or LFC lines are used.
-Excalibur using LFC lines exhibit the lowest residuals.}
-\label{fig:waveResids}
-\end{figure*}
+\bibliography{paper.bib}
 

### 2020-05-01 paper/paper.tex
Changes up to section 2.1

@@
 
-\documentclass[twocolumn]{aastex63}
+\documentclass[12pt, onecolumn]{aastex63}
 \usepackage{xcolor}
@@
 \newcommand{\lfc}{\acronym{LFC}}
+\newcommand{\code}[1]{\texttt{#1}}
 
@@
 \title{\name:
-A non-parametric, hierarchical model for a precision spectrograph}
+A Non-Parametric, Hierarchical Wavelength-Calibration Model for a Precision Spectrograph}
 
@@
 \name\ is a hierarchical, non-parametric framework for wavelength calibrating precision spectrographs.
-This method makes use of the many, recent hardware improvements in highly-stabilized, extreme-precision radial-velocity spectrographs.
+This method capitalizes on recent hardware improvements in highly-stabilized, extreme-precision radial-velocity spectrographs.
 With stabilized spectrographs, the full optical system and detectors can only vary within a tiny range of configurations.
@@
 \name\ also takes advantage of the development of laser-frequency combs or etalons, which generate a dense set of stable calibration points.
-This density of points permits the freedom of a non-parametric wavelength solution that can match or adapt to any instrument or detector oddities.
-We demonstrate the success of this method with data from a Menlo Systems laser frequency comb taken by \expres, the Extreme Precision Spectrograph.
-We find that \name\ returns more accurate wavelength predictions by more than a factor of 2 over classic polynomial models.
-Employing \name\ wavelengths reduced the RMS of RV measurements to a planet fit by \lz{some percent}.
+This density of points permits the freedom of a non-parametric wavelength solution that can match or adapt to any instrument or detector oddities, more so than any one polynomial.
+We demonstrate the success of this method with data from the EXtreme PREcision Spectrograph, \expres, which includes a Menlo Systems laser frequency comb.
+We find that \name\ returns more accurate wavelength predictions by more than a factor of 2 over classic, non-hierarchical, polynomial solutions.
+Employing \name\ wavelengths on \expres\ RV measurements on HD 3651 reduced the RMS of a planet fit by \lz{blah}.
 \end{abstract}
@@
 % Add official keywords
-\keywords{instrumentation: spectrographs -- techniques: spectroscopic -- techniques: radial velocities}
+\keywords{instrumentation: spectrographs -- instrumentation: detectors -- techniques: spectroscopic -- techniques: radial velocities -- methods: data analysis -- methods: statistical}
 
 \section{Introduction} 
-Precision radial-velocity (\eprv) programs have been incredibly fruitful in finding and characterizing extra-solar planets \citep{mayor2011, bonfils2013, plavchan2015, butler2017}.
+Extremely precise, radial-velocity (\eprv) programs have been incredibly fruitful in finding and characterizing extra-solar planets \citep{mayor2011, bonfils2013, plavchan2015, butler2017}.
 These programs typically make use of spectrographs with resolutions on the order of $10^5$, which correspond to line widths on the order of $3000\,\mps$.
 Exoplanet science happens at the level of $3\,\mps$ precision, with current instruments hoping to reach $0.1\,\mps$ precision \citep{fischer2016}.
-This requires new spectrographs to be calibrated or stabilized to better than $10^{-4}$ of a pixel (assuming that the spectrograph is well sampled).
-Some hardware--software systems, such as ESPRESSO and \expres\ are now reaching these levels \citep{pepe2013,  jurgenson2016, neid, blackman2020}.
+This requires new spectrographs to be calibrated or stabilized to better than $10^{-4}$ of a pixel (assuming that the spectrographs are well sampled).
+Some hardware--software systems, such as ESPRESSO, \expres\, and NEID were designed to deliver precision at this level \citep{pepe2013,  jurgenson2016, neid}.
+Early data from \expres\ and ESPRESSO suggest that the hardware improvements do propagate through to sub-$\mps$ uncertainties on final, on-sky RV measurements \citep{blackman2020, petersburg2020, pepe2020}.
 
-Here we propose to simplify and improve calibration programs for \eprv\ hardware systems with two very simple but novel ideas.
-The first flows from the observation that calibration sources---which include lamps \lz{(looks like the ThAr isn't quite good enough actually)}, etalons, and laser-frequency combs (\lfc s)---illuminate the spectrograph with very stable, very dense sets of lines; almost every location in the spectrograph image plane is surrounded by nearby, useful calibration lines.
+\lz{review of how wavelength solution has been done in the past.  Make the point that traditional wavelength solutions only make use of each calibration information independently
+i.e. the standard method for wavelength calibration is...citations}
+
+Here we propose to simplify and improve calibration programs for \eprv\ hardware systems with two very simple but innovative ideas.
+The first flows from the observation that calibration sources---which include arc lamps (in some wavelength ranges), etalons, and laser-frequency combs (\lfc s)---illuminate the spectrograph with very stable, very dense sets of lines; almost every location in the spectrograph image plane is surrounded by nearby, useful calibration lines.
+
 This recommends a calibration methodology that is \emph{non-parametric}:  If every point in the spectrograph detector is surrounded by nearby calibration lines, the wavelength solution can, for example, be made simply as an interpolation of the calibration data.
-
 The density of lines removes the need to enforce any functional form for the wavelength solution (such as a continuous eighth-order polynomial, for example).
-We expand on the work of \cite{milakovic2020}, which demonstrated the efficacy of constructing a wavelength solution as multiple, segmented polynomials.
+In some ways, this is a generalization of the recent work that demonstrated the efficacy of constructing a wavelength solution as multiple, segmented polynomials \citep{milakovic2020}.
 Going non-parametric will improve calibration accuracy by not forcing the choice of a parametric form that may bias the calibration, especially when the chosen function is inappropriate (as, for example, polynomials will be at detector edges).
 
-The second idea follows from the observation that contemporary \eprv\ instruments are incredibly stable.
+The second simple idea follows from the observation that contemporary \eprv\ instruments are incredibly stable.
 Temperature-controlled, fiber-fed spectrographs vary only slightly over the night or season, and only along a small number of axes in what you might call ``calibration space'', or the (very high dimensional) space of all possible wavelength solutions.
-That is, not only are the spectrographs stable, but they also have few accessible degrees of freedom.
-This renders it inappropriate to fit each calibration exposure or calibrate each science exposure independently.
+That is, not only are the spectrographs stable, but they also have few environmentally accessible degrees of freedom.
+This renders it inadvisable to fit each calibration exposure or calibrate each science exposure independently.
 Instead, all the calibration data (or all the data) should be used to determine the space in which the instrument can and does vary.
-Subsequent calibration work then need only determine where in the accessible part of calibration space the spectrograph existed for each exposure.
+Subsequent calibration work then need only determine where in the accessible part of calibration space the spectrograph was located for each exposure.
 
-This structure is \emph{hierarchical}: the calibration data are used not just to determine the wavelength solution, but also to determine the possible space of wavelength solutions.
-In the statistics literature this concept is often described as \emph{de-noising}.
-Every calibration exposure contains information about every other exposure.
+In the context of probabilistic models, this structure is \emph{hierarchical}:  The calibration data are used not just to determine the wavelength solution, but also to determine the possible \emph{space} of wavelength solutions.
+In the statistics literature this concept is often described as \emph{de-noising}:  Every calibration exposure contains information about every other exposure.
 Thus every exposure can be improved (i.e.
@@
 By being non-parametric, it delivers enormous freedom to the wavelength solution to match or adapt to any instrument or detector oddities.
-By being hierarchical, it restricts that freedom tremendously, but it does so appropriately for the observed variations in the stabilized spectrograph.
-\name\ is designed for temperature-controlled, fiber-fed spectrographs with good calibration sources, such as laser-frequency combs, etalons, or good arc lamps.
+By being hierarchical, it restricts that freedom tremendously, but it does so appropriately for the empirically exercised variations in the stabilized spectrograph.
+\name\ is designed for temperature-controlled, fiber-fed spectrographs with good calibration sources, such as laser-frequency combs, or etalons.
 We have in mind the \eprv\ instruments and \eprv\ science cases, but we expect \name\ to have applications for other kinds of spectrographs in other contexts.
@@
 \section{Method} \label{sec:method}
-\name\ is designed to turn a series of calibration lines with known wavelengths and well-fit detector positions, and de-noise and interpolate them into a full wavelength solution.
-We operate on the two core ideas that the wavelength solution should be given enormous flexibility, but that it lives in a very low-dimensional space, where the degrees of freedom are set by the limited kinematics of the stabilized, spectrograph hardware.
+\name\ is designed to turn a series of calibration lines with known wavelengths and well-fit detector positions, and de-noise and interpolate them into a full wavelength model applicable to all exposures taken with the instrument.
+\name\ operates on the two core ideas that the wavelength solution should be given enormous flexibility, but that it lives in a very low-dimensional space, where the degrees of freedom are set by the limited kinematics of the stabilized spectrograph hardware.
 
-We therefore assume dense enough calibration line coverage with well-fit line centers to provide some constraint on an interpolated wavelength solution across an echelle order.
-Additionally, we assume the instrument hardware is stabilized enough that the space of possible calibration states is low-dimensional.
+\name\ assumes dense enough calibration line coverage with well-fit line centers to provide some constraint on an interpolated wavelength solution across an echelle order.
+Additionally, \name\ assumes the instrument hardware is stabilized enough that the space of possible calibration states is low-dimensional.
 
-We will pose the question of wavelength calibration in the following way.  Given an exposure $n$, and echelle order $m$, there is a relationship between
+Wavelength calibration is usually posed in the following way.  Given an exposure $n$, and echelle order $m$, there is a relationship between
 the two-dimensional $(x,y)$-position on the detector and the
@@
 Classically, pipelines employ polynomials to construct smooth wavelength solutions for each exposure.
-For example, the pipeline described in \cite{petersburg2020} sets the function $f(x,y,m;\theta_{n})$ from Equation \ref{eq:wsol} to be a 2D, 9\textsuperscript{th}-order polynomial where $\theta_{n}$ represents the polynomial coefficients, $c_{nij}$, unique to each exposure $n$.
+For example, the \expres\ pipeline sets the function $f(x,m;\theta_{n})$ from Equation \ref{eq:wsol} to be a 2D, 9\textsuperscript{th}-order polynomial where $\theta_{n}$ represents the polynomial coefficients, $c_{nij}$, unique to each exposure $n$ \citep{petersburg2020}.
 \begin{equation}
-\lambda(x,y,m,n) = \sum_{i=0}^9\sum_{j=0}^9 c_{nij}\, x^i\,m^j + \mathrm{noise}
+\lambda(x,m,n) = \sum_{i=0}^9\sum_{j=0}^9 c_{nij}\, x^i\,m^j + \mathrm{noise}
 \quad ,
@@
 \end{equation}
-The coefficients $c_{nij}$ are then smoothed by a third-order polynomial over time.
+where the y-dependence is dropped as that dependence is carried by spectral order $m$.
+The coefficients $c_{nij}$ are interpolated to time $t_n$ of exposure $n$ by a third-order polynomial of time.
 This third-order polynomial is evaluated at the time of non-calibration exposures to re-construct a 2D, 9th-order polynomial wavelength solution for that exposure.
-Each calibration image is fit to a 2D polynomial independently.
+Each calibration image obtains in $c_{nij}$ coefficients independently.
 
@@
 
-If the space of all calibration possibilities is in fact $K$-dimensional (where $K$ is a small integer, i.e 2 or 8 or thereabouts), and if the calibration variations are so small that we can linearize, then the function $f(x,y,m;\theta_{n})$ from Equation \ref{eq:wsol} could be replaced with a tiny model
+If the space of all calibration possibilities is in fact $K$-dimensional (where $K$ is a small integer, i.e 2 or 8 or thereabouts), and if the calibration variations are so small that than can be linearized, then the function $f(x,m;\theta_{n})$ from Equation \ref{eq:wsol} should be low-dimensional.
+In \name, we transpose the calibration model---making the position $x$ a function of $\lambda$---into the following form
 \begin{equation}
-\lambda(x,y,m,n) = g_0(x,y,m) + \sum_{k=1}^K a_{nk}\,g_k(x,y,m)
+x(\lambda,m,n) = g_0(\lambda,m) + \sum_{k=1}^K a_{nk}\,g_k(\lambda,m)
 \quad ,
@@
 where
-$g_0(x,y,m)$ is the fiducial or mean or standard calibration of the spectrograph,
-the $a_{nk}$ are scalar amplitudes,
-and the $g_k(x,y,m)$ are basis functions expressing the ``directions'' in calibration space that the spectrograph can depart from the fiducial calibration.
+$g_0(\lambda,m)$ is the fiducial or mean or standard calibration of the spectrograph,
+the $a_{nk}$ are $K$ scalar amplitudes for each exposure $n$,
+and the $g_k(\lambda,m)$ are basis functions expressing the ``directions'' in calibration space that the spectrograph can depart from the fiducial calibration.
+When this calibration structure is used to deliver a wavelength solution, we will have to invert $x(\lambda,m,n)$ into $\lambda(x,m,n)$.
 
@@
 In this paper, we present a model using principal component analysis (PCA).
-This is the correct thing to do in the limit where we have very high SNR, as is undeniably the case with calibration images.
+A PCA is justifiable in the limit where there is very high signal-to-noise ratio, as is usually the case with typical calibration images.
+There are many alternatives to PCA for this dimensionality reduction; we return to this point in the discussion (\textsection \ref{sec:discussion}) below.
 
-\lz{Is this the place to mention options other than PCA? Or should we save that for the discussion?}
+\subsection{Dimensionality Reduction: De-Noising of Calibration Frames} \label{sec:denoising}
+Within a given range of stability, \name\ will use calibration images to  1) determine the space in which the instrument varies and 2) where in the accessible calibration space the spectrograph existed for each exposure.
+For each calibration exposure, $n$, \name\ requires a full list of lines, $(\lambda,m)$ that are expected in each calibration exposure.
+Each line is uniquely defined by a combination of echelle order, $m$, and ``true'' or theoretical wavelength, $\lambda$.
 
-\subsection{De-Noising of Calibration Frames} \label{sec:denoising}
-Within a given range of stability, we can use all calibration images to  1) determine the space in which the instrument varies and 2) where in the accessible calibration space the spectrograph existed for each exposure.
-Say we have $N$ calibration exposures, $exp_n$.
-The first task is to identify the full list of lines we can expect to find in a calibration image.
-A line, $l_{m,\lambda}$ is uniquely defined by a combination of echelle order, $m$, and ``true" or theoretical wavelength, $\lambda$.
+For each exposure, $n$, every line, $(\lambda,m)$, has an associated fitted detector position, $x(\lambda,m,n)$, for example x-pixel in an 2D extracted echelle order.
+Fitted line centers that are missing from an exposure (e.g.
+because the fit failed due to noise, the line is not in its usual echelle order, etc.) can be assigned a \code{NaN} (hardware not-a-number) for that echelle order instead.
+Let there be $M$ lines per exposure.
+\name\ reads in a $N \times M$ matrix of line positions for each of $M$ lines for each of $N$ exposures.
 
-For each exposure, $exp_n$, every line, $l_{m,\lambda}$, has an associated fitted detector position, $x_{n,l}$, for example x-pixel in an 2D extracted echelle order.
-Fitted line centers that are missing from an exposure (e.g.
-because the fit failed due to noise, the line is not in its usual echelle order, etc.) can be assigned a $NaN$ for that echelle order instead.
-Let there be $M$ lines per exposure.
-We can therefore construct a $N \times M$ matrix of line positions for each of $M$ lines for each of $N$ exposures.
-
-The average line position in every exposure represents the fiducial, or standard calibration of the spectrograph, $g_0(x,y,m)$.
-We apply principal component analysis to the difference between this fiducial calibration and each individual line position.
-The returned principal components serve as basis functions,  $g_k(x,y,m)$, expressing the possible deviations of the spectrograph from this fiducial calibration.
+The mean of measured line position over the set of calibration exposures represents the fiducial, or standard calibration of the spectrograph, $g_0(\lambda,m)$.
+In this implementation of \name\, principal component analysis is performed on the difference between this fiducial calibration and each individual line position.
+The returned principal components serve as basis functions,  $g_k(\lambda,m)$, expressing the possible deviations of the spectrograph from this fiducial calibration.
 The magnitude of each principal component for each exposure, $a_{nk}$, represents the scalar amplitude of these deviations for each exposure.
-We can now use a small number, $K$, of principal components to reconstruct a de-noised version of the line positions as formulated in Equation \ref{eq:excl_wsol}.
+\name\ then uses a small number, $K$, of principal components to reconstruct a de-noised version of the line positions as formulated in Equation \ref{eq:excl_wsol}.
 
 Missing line center measurements are replaced with de-noised estimates.
-We do this iteratively until the estimates of missing line centers change by less than 0.01\%.
+This is done iteratively until the estimates of missing line centers change by less than 0.01\%.
 This process can be repeated on line centers deemed as outliers by some metric, to account for lines that may have been mis-identified or mis-fit.
-The principal components from the final iteration are used to define the spectrograph's calibration space while the associated magnitudes pinpoint where in that calibration space the spectrograph is for each calibration exposure.
+The principal components from the final iteration are used to define the spectrograph's calibration space while the associated magnitudes pinpoint where in that calibration space the spectrograph is located for each calibration exposure.
 
@@
 \SetAlgoLined
-\textbf{Inputs}: $m$, echelle order of each line; $\lambda$, assigned wavelength of each line; $[x_{n,l}] $, matrix of pixel line position of each line $l$ and for each exposure $n$\;
-\While{change in missing or outlier line centers $>$ 0.01\%}{
-	$g_0(x,y,m) = \overline{x_{n,l}]}$\;
-	find $U, \Sigma, V$ s.t. $U\Sigma V^* = (x_{l,n}-g_0(x,y,m))$\;
-	let $a_{n,k} = U\cdot \Sigma$ and $g_k(x,y,m) = V$\;
-	$\lambda(x,y,m,n) = g_0(x,y,m) + \sum_{k=1}^K a_{nk}\,g_k(x,y,m)$ for $K=6$\;
-	$[x_{n,l}] = \lambda(x,y,m,n)$ where $[x_{n,l}]$ was initially $NaN$
+\KwData{line positions $x(\lambda,m,n)$ for each exposure $n$, with wavelengths $\lambda$ and echelle orders $m$}
+\KwResult{Low-dimensional calibration space of spectrograph}
+\While{change in missing or outlier line centers $>$ 0.01\%}
+{
+	$g_0(\lambda,m) = \overline{x(\lambda,m,n)}$\;
+	find $U, \Sigma, V$ s.t. $U\Sigma V^* = (x(\lambda,m,n)-g_0(\lambda,m))$\;
+	let $a_{n,k} = U\cdot \Sigma$ and $g_k(\lambda,m) = V$\;
+	$x(\lambda,m,n) = g_0(\lambda,m) + \sum_{k=1}^K a_{nk}\,g_k(\lambda,m)$ where $K=6$ for $x(\lambda,m,n)$ = \code{NaN}
 	}
-\caption{De-Noising}
+\caption{Dimension Reduction and De-Noising}
 \end{algorithm}
@@
 to determine different calibration states of the spectrograph.
-The choice of what to interpolate with respect to depends on the dominant contribution to variation in the instrument.
+The choice of what to interpolate against depends on the dominant contribution to variation in the instrument.
 
-As an example, we can interpolate the magntiudes of the principal components with respect to time to the midpoint time of non-calibration exposures to determine the calibration state of the spectrograph for that exposure.
-Using interpolated magnitudes, $a'_{nk}$ and the basis vectors, $g_k(x,y,m)$, returned by the de-noising process, we can construct a set of calibration lines, $\lambda(x,y,m,n)$, for any exposure as prescribed in Equation \ref{eq:excl_wsol}.
+In the implementation of \name\ presented here, the magnitudes of the principal components are interpolated with respect to time.
+New magnitudes $a_n'k$ values are found for the midpoint time of non-calibration exposures and used to construct the calibration state of the spectrograph for that exposure.
+Using interpolated magnitudes, $a_{n;k}$ and the basis vectors, $g_k(\lambda,m)$, returned by the de-noising process, \name\ can construct a set of calibration lines, $x(\lambda,m,n') $, for any exposure as prescribed in Equation \ref{eq:excl_wsol}.
 
 \subsection{Interpolating a Wavelength Solution} \label{sec:interp_wsol}
-We now have the power to construct a set of calibration lines, $\{l_{m,\lambda}\}$, each with an assigned, known wavelength for any exposure.
-To construct a wavelength solution, we need only interpolate the known wavelengths over the exposure specific line centers.
+From the de-noising step, \name\ can now construct a set of calibration lines, $x(\lambda,m,n')$ for any exposure $n'$.
+To construct a wavelength solution, \name\ then interpolates the known wavelengths over the exposure specific line centers.
 For instance, interpolating the known wavelengths vs.
@@
 \SetAlgoLined
-\textbf{Inputs}: $g_0(x,y,m)$, the fiducial calibration of the spectrograph; $a_{n,k}$, magnitudes of the principal components for each exposure; $g_k(x,y,m)$, basis vectors spanning the calibration space of the spectrograph; $t_n$, time of each exposure; $t'$, time we want wavelengths for\;
-Interpolate $a_{n,k}$ with respect to time to determine $a_{n,k}'$ for time $t'$\;
-Calculate pixel line positions for time $t'$ by evaluating\;
-$\lambda(x,y,m,n)' = g_0(x,y,m) + \sum_{k=1}^K a_{nk}'\,g_k(x,y,m)$ for $K=6$\;
+\KwData{the fiducial calibration of the spectrograph $g_0(\lambda,m)$; magnitudes of the principal components for each exposure $a_{n,k}$; basis vectors spanning the calibration space of the spectrograph $g_k(\lambda,m)$; }
+\KwResult{Wavelengths for detector positions $x'(m,n')$ of exposure $n'$ with time $t_{n'}$}
+
+Find $a_{n',k}$ by interpolating $a_{n,k}$ with to $t_{n'}$\;
+$x(\lambda,m,n') = g_0(\lambda,m) + \sum_{k=1}^K a_{n'k}\,g_k(\lambda,m)$ where $K=6$\;
 \For{each unique $m$}{
-	interpolate ${l_n,m}$ onto pixels, $x'$, that we want wavelengths for\;
+	interpolate $\lambda$ with respect to $x(\lambda,m,n')$ onto pixels $x'(m,n')$\;
 	}
-\caption{Generating a Wavelength Solution}
+\caption{Generating Wavelength Solution}
 \end{algorithm}
@@
 
-Rather than using a simultaneous calibration fiber, two to three LFC exposures are interspersed between science exposures every ~15-30 minutes.
+Rather than using a simultaneous calibration fiber, two to three LFC exposures are interspersed between science exposures every 15-30 minutes.
 ThAr exposures are taken at the beginning and end of each night.
-All calibration data is taken through the science fiber, thereby traveling through the same optics and hitting the same pixels as all science observations.
+All calibration data is taken through the science fiber, meaning they traveling down the same optical pathway and are projected onto the same pixels as all science observations.
 
@@
 LFC lines cover 40 echelle orders, which contain 19203 lines.
-Though we primarily work with LFC data, we at times discuss applications to arc lamps.
+Though the results are primarily based on work with LFC data, there is some discussion of applications to arc lamps.
 These results use 78 ThAr exposures that were taken on the same nights.
 ThAr lines cover all 86 extracted orders of \expres, which includes 5295 lines.
+For both the LFC and ThAr data, lines that appeared in less than 60\% of exposures were not included in the analysis.
+Similarly, files with more than 60\% of expected lines missing were cut from the analysis.
 
  \name\ proceeds from a given list of echelle orders, line wavelengths, and pixel positions of each line for each exposure.
-We, therefore, use line positions generated by the pre-existing \expres\ pipeline, as described in \cite{petersburg2020}.
+We, therefore, use line positions generated by the pre-existing \expres\ pipeline \citep{petersburg2020}.
 
@@
 
- We satisfy the assumption of stability needed for \name\ by using exposures from only one ``instrumental epoch" of \expres.
+In order to satisfy the assumption of stability needed for \name, only exposures from only one ``instrumental epoch" of \expres\ are used.
 \expres\ exposures are separated into different instrumental epochs that correspond to atypical hardware changes--changes that altered the position of the echellogram on the detector, the shape of the instrumental PSF, or the calibration sources.
 Significant instrumental changes break the assumption of a stabilized instrument with only low-dimensional variations.
-We therefore treat each instrument epoch independently.
+Each instrument epoch is therefore treated independently.
 
@@
 For all these tests, we generate wavelengths for known calibration lines and compare the predicted wavelength to the assigned wavelength of each line.
-This inherently folds in error from the measured line center of each calibration line, but the contribution to the residuals will be the same across all tests.
+This inherently folds in errors in the measured line center of each calibration line, but the contribution to the residuals will be the same across all tests.
 
@@
 Histograms of the residuals for each of the above described polynomial, interpolation, and denoising tests (respectively) are shown in Figure \ref{fig:testHists}.
-Note that the spread in residuals is much smaller for both the denoising and interpolation tests relative to the results of the polynomial model.
+Note that the spread in residuals is much smaller for both the denoising and interpolation tests relative to the results of the polynomial wavelength solutionl.
 
@@
 \caption{Difference in predicted and assigned wavelengths for the wavelength calibration tests described in Section \ref{sec:tests}.
-The standard deviation of the residuals for each model is given in the top-right corner in each model's corresponding color.}
+The standard deviation of the residuals for each test is given in the top-right corner in each method's corresponding color.}
 \label{fig:testHists}
@@
 %\includegraphics[width=\textwidth]{Figures/lineResids2D.pdf}
-\caption{Residuals plotted with respect to detector position (as defined by echelle order and x-pixel) for various wavelength models.
+\caption{Residuals plotted with respect to detector position (as defined by echelle order and x-pixel) for various wavelength calibration methods.
 Each line is colored by the difference between the predicted wavelength and the true wavelength for each line, given in units of $/mps$.
@@
 
-The residuals of the classic model is shown in the left-most plot of Figure \ref{fig:resid2d}.  There is a lot of vertical structure and some hints of a periodic, diagonal structure as well.
+The residuals of the classic, polynomial wavelength solution is shown in the left-most plot of Figure \ref{fig:resid2d}.
+There is a lot of vertical structure and some hints of a periodic, diagonal structure as well.
 
@@
 There is no coherent structure here and smaller residuals.
-We believe the flexibility of the interpolated model was able to account for high-order detector defects, which emerged as structure in the residuals of the classic, smooth, polynomial model.
+We believe the flexibility of the interpolated model was able to account for high-order detector defects, which emerged as structure in the residuals of the classic, smooth, polynomial-driven wavelength solution.
 
@@
 
-Though the interpolated wavelength solution returns lower, less-structured residuals than the polynomial model when guided by LFC lines, we see in the right-most plot of Figure \ref{fig:resid2d} that an interpolated wavelength solution can return much worse results in the space between widely separated calibration lines.
+Though the interpolated wavelength solution returns lower, less-structured residuals than the polynomial wavelength solution when guided by LFC lines, we see in the right-most plot of Figure \ref{fig:resid2d} that an interpolated wavelength solution can return much worse results in the space between widely separated calibration lines.
 
@@
 
-Note that running \name\ informed by only ThAr lines cannot be regarded as a direct comparison to the LFC runs, as the increased error and variability in ThAr line positions alone makes the resultant wavelength predictions an order of magnitude worse, hence the different scale of the colorbar.
+Note that running \name\ informed by only ThAr lines cannot be regarded as a direct comparison to the LFC runs, as the increased uncertainty and variability in ThAr line positions alone makes the resultant wavelength predictions an order of magnitude worse, hence the different scale of the colorbar.
 All the same, the residuals are relatively worse where lines are further spaced (for example, in redder echelle orders) than where lines are denser.
@@
 In all echelle orders, the edges are overestimated while the middle will be underestimated.
-The resultant wavelength solution is underestimating the curvature of the pixel-wavelength relation, giving rise to similar issues as with an inflexible, parametric model.
+The resultant wavelength solution is underestimating the curvature of the pixel-wavelength relation, giving rise to similar issues as with an inflexible, parametric wavelength solution.
 Introducing this smoothing parameter is enforcing a smoothness we have no reason to believe is true of the data.
@@
 
-On the other hand, the ability of PCA to pick up on any and all variation, regardless of source, also makes \name\ very sensitive to errors in line fitting.
+On the other hand, the ability of PCA to pick up on any and all variation, regardless of source, also makes \name\ very sensitive to uncertainties in line fitting.
 For example, we have seen that lower-signal lines that are harder to fit faithfully will have greater variety in returned line positions, which is in turn captured by the PCA.
@@
 
-Furthermore, once we have defined a calibration space that captures all possible degrees of freedom for a stabilized spectrograph, then all exposures taken with the spectrograph contains information about where the spectrograph lies within that calibration space.
+Furthermore, once we have defined a calibration space that captures all possible degrees of freedom for a stabilized spectrograph, then all exposures taken with the spectrograph contains information about where the spectrograph is located within that calibration space.
 Theoretically, science exposures themselves could also be used to determine the calibration state of an instrument, negating the need for any additional wavelength-calibration source beyond defining the calibration space.
@@
 If, for example, a cross correlation method is used to extract RVs from the data, a systematic difference may be introduced depending on the definition of the center of the line.
-In principle, the best way to mitigate this error would be to use a calibration source that looks like a star.
+In principle, the best way to mitigate this uncertainty would be to use a calibration source that looks like a star.
 
 \name\ is designed and optimized for extreme-precision instruments.
-We believe that by implementing \name, the error from wavelength calibration is rendered negligible in the face of other error sources.
+We believe that by implementing \name, the uncertainty from wavelength calibration is rendered negligible in the face of other noise sources.
 

### 2020-05-06 paper/paper.tex
Wavelength Calibration Review
It's not a good review, but hopefully good enough to not make anyone mad

@@
 
-\lz{review of how wavelength solution has been done in the past.  Make the point that traditional wavelength solutions only make use of each calibration information independently
-i.e. the standard method for wavelength calibration is...citations}
+Traditionally, wavelength solutions are constructed by fitting a polynomial to lines from a calibration source to describe the relationship between wavelength and pixel for each echelle order \citep{lovis2007, cersullo2019}.
+Each calibration image is treated independently, though the entire set of calibration images has been used to vet incongruous lines \citep{coffinet2019}.
+Just recently, work has been done that suggests a segmented polynomial in the dispersion direction, tuned to detector defects, better describes the wavelength solution than a single, continuous polynomial \citep{milakovic2020}.
 

### 2020-05-08 paper/paper.tex
formatting tweaks; got modern to work
Incorporating MB Text Comments
Still missing some larger-ticket edits and figure changes

@@
 
-\documentclass[12pt, onecolumn]{aastex63}
-\usepackage{xcolor}
-\newcommand{\lz}[1]{\textcolor{orange}{#1}}
+\documentclass[modern]{aastex63}
+\usepackage[ruled,vlined,norelsize]{algorithm2e}
 
-\usepackage[ruled,vlined]{algorithm2e}
-\usepackage{graphicx}
+% hogg dealing with his shit
+\addtolength{\topmargin}{-0.30in}
+\addtolength{\textheight}{0.70in}
+\renewcommand{\twocolumngrid}{} % hahaha!
+\setlength{\parindent}{1.1\baselineskip}
+\sloppy\sloppypar\raggedbottom\frenchspacing
+\shorttitle{excalibur: a wavelength-calibration model}
+\shortauthors{zhao, hogg, bedell, fischer}
 
@@
 \newcommand{\expres}{\project{\acronym{EXPRES}}}
+\newcommand{\espresso}{\project{\acronym{ESPRESSO}}}
+\newcommand{\neid}{\project{\acronym{NEID}}}
 \newcommand{\eprv}{\acronym{EPRV}}
@@
 
+\newcommand{\lz}[1]{\textcolor{orange}{#1}}
+\newcommand{\mb}[1]{\textcolor{violet}{#1}}
+
 % math shih
@@
 
-% margins and page setup, etc
-\addtolength{\textheight}{1.00in}
-\addtolength{\topmargin}{-0.50in}
-\sloppy\sloppypar\raggedbottom\frenchspacing
-
 \begin{document}
 \title{\name:
-A Non-Parametric, Hierarchical Wavelength-Calibration Model for a Precision Spectrograph}
+  A Non-Parametric, Hierarchical \\
+  Wavelength-Calibration Model for a Precision Spectrograph}
 
@@
 
-\begin{abstract}
+\begin{abstract}\noindent%
 \name\ is a hierarchical, non-parametric framework for wavelength calibrating precision spectrographs.
@@
 These programs typically make use of spectrographs with resolutions on the order of $10^5$, which correspond to line widths on the order of $3000\,\mps$.
-Exoplanet science happens at the level of $3\,\mps$ precision, with current instruments hoping to reach $0.1\,\mps$ precision \citep{fischer2016}.
+Exoplanet science happens at the level of $3\,\mps$ precision, with current instruments aiming to reach $0.1\,\mps$ precision \citep{fischer2016}.
 This requires new spectrographs to be calibrated or stabilized to better than $10^{-4}$ of a pixel (assuming that the spectrographs are well sampled).
-Some hardware--software systems, such as ESPRESSO, \expres\, and NEID were designed to deliver precision at this level \citep{pepe2013,  jurgenson2016, neid}.
-Early data from \expres\ and ESPRESSO suggest that the hardware improvements do propagate through to sub-$\mps$ uncertainties on final, on-sky RV measurements \citep{blackman2020, petersburg2020, pepe2020}.
+Some hardware--software systems, such as \espresso, \expres, and \neid\ were designed to deliver precision at this level \citep{pepe2013,  jurgenson2016, neid}.
+Early data from \expres\ and \espresso\ suggest that the hardware improvements do propagate through to sub-$\mps$ uncertainties on final, on-sky RV measurements \citep{blackman2020, petersburg2020, pepe2020}.
 
@@
 Here we propose to simplify and improve calibration programs for \eprv\ hardware systems with two very simple but innovative ideas.
-The first flows from the observation that calibration sources---which include arc lamps (in some wavelength ranges), etalons, and laser-frequency combs (\lfc s)---illuminate the spectrograph with very stable, very dense sets of lines; almost every location in the spectrograph image plane is surrounded by nearby, useful calibration lines.
+The first flows from the fact that calibration sources---which include arc lamps (in some wavelength ranges), etalons, and laser-frequency combs (\lfc s)---illuminate the spectrograph with very stable, very dense sets of lines; almost every location in the spectrograph image plane is surrounded by nearby, useful calibration lines.
 
@@
 The second simple idea follows from the observation that contemporary \eprv\ instruments are incredibly stable.
-Temperature-controlled, fiber-fed spectrographs vary only slightly over the night or season, and only along a small number of axes in what you might call ``calibration space'', or the (very high dimensional) space of all possible wavelength solutions.
+Temperature-controlled, fiber-fed spectrographs vary only slightly over the night or season.
+\lz{This suggests that such stabilized spectrographs vary along only} a small number of axes in what you might call ``calibration space'', or the (very high dimensional) space of all possible wavelength solutions.
 That is, not only are the spectrographs stable, but they also have few environmentally accessible degrees of freedom.
@@
 
-In the context of probabilistic models, this structure is \emph{hierarchical}:  The calibration data are used not just to determine the wavelength solution, but also to determine the possible \emph{space} of wavelength solutions.
+In the context of probabilistic models, this structure is \emph{hierarchical}:  The calibration data are used not just to determine the wavelength solution at one moment, but also to determine the possible \emph{space} of wavelength solutions at all moments.
 In the statistics literature this concept is often described as \emph{de-noising}:  Every calibration exposure contains information about every other exposure.
@@
 By being non-parametric, it delivers enormous freedom to the wavelength solution to match or adapt to any instrument or detector oddities.
-By being hierarchical, it restricts that freedom tremendously, but it does so appropriately for the empirically exercised variations in the stabilized spectrograph.
+By being hierarchical, it restricts that freedom tremendously, but it does so appropriately for the empirically determined variations in the stabilized spectrograph.
 \name\ is designed for temperature-controlled, fiber-fed spectrographs with good calibration sources, such as laser-frequency combs, or etalons.
-We have in mind the \eprv\ instruments and \eprv\ science cases, but we expect \name\ to have applications for other kinds of spectrographs in other contexts.
-The motivating ideas behind this project are true for almost every astronomical spectrograph.
+We have in mind \eprv\ instruments and \eprv\ science cases, but we expect \name\ to have applications for other kinds of spectrographs in other contexts.
+\mb{The motivating ideas behind this project are true for almost every astronomical spectrograph.
+(Is this true?)}
 
 \section{Method} \label{sec:method}
-\name\ is designed to turn a series of calibration lines with known wavelengths and well-fit detector positions, and de-noise and interpolate them into a full wavelength model applicable to all exposures taken with the instrument.
+The \name\ method is designed to turn a series of calibration lines with known wavelengths and well-fit detector positions, and de-noise and interpolate them into a full wavelength model applicable to all exposures taken with the instrument.
 \name\ operates on the two core ideas that the wavelength solution should be given enormous flexibility, but that it lives in a very low-dimensional space, where the degrees of freedom are set by the limited kinematics of the stabilized spectrograph hardware.
@@
 Additionally, \name\ assumes the instrument hardware is stabilized enough that the space of possible calibration states is low-dimensional.
+We revisit and quantify these conditions in Section \ref{sec:discussion}.
 
@@
 \end{equation}
-where the y-dependence is dropped as that dependence is carried by spectral order $m$.
-The coefficients $c_{nij}$ are interpolated to time $t_n$ of exposure $n$ by a third-order polynomial of time.
+Here the y-dependence is dropped as that dependence is carried by spectral order $m$.
+\lz{The line position information is therefore uniquely identified by echelle order $m$ and pixel in the dispersion direction, $x$.}  The coefficients $c_{nij}$ are interpolated to time $t_n$ of exposure $n$ by a third-order polynomial of time.
 This third-order polynomial is evaluated at the time of non-calibration exposures to re-construct a 2D, 9th-order polynomial wavelength solution for that exposure.
-Each calibration image obtains in $c_{nij}$ coefficients independently.
+Each calibration image obtains its $c_{nij}$ coefficients independently.
 
@@
 
-If the space of all calibration possibilities is in fact $K$-dimensional (where $K$ is a small integer, i.e 2 or 8 or thereabouts), and if the calibration variations are so small that than can be linearized, then the function $f(x,m;\theta_{n})$ from Equation \ref{eq:wsol} should be low-dimensional.
+If the space of all calibration possibilities is in fact $K$-dimensional (where $K$ is a small integer, i.e 2 or 8 or thereabouts), and if the calibration variations are so small that they can be linearized, then the function $f(x,m;\theta_{n})$ from Equation \ref{eq:wsol} should be low-dimensional.
 In \name, we transpose the calibration model---making the position $x$ a function of $\lambda$---into the following form
@@
 When this calibration structure is used to deliver a wavelength solution, we will have to invert $x(\lambda,m,n)$ into $\lambda(x,m,n)$.
+\mb{(Explain why this inversion is necessary?)}
 
@@
 
+\begin{figure*}[h]
+\centering
+\includegraphics[width=\textwidth]{Figures/methodCartoon.pdf}
+\caption{}
+\label{fig:testHists}
+\end{figure*}
+
 \subsection{Dimensionality Reduction: De-Noising of Calibration Frames} \label{sec:denoising}
@@
 Each line is uniquely defined by a combination of echelle order, $m$, and ``true'' or theoretical wavelength, $\lambda$.
+\lz{There are many strategies for identifying calibration line positions and matching them to their assigned wavelengths; this problem is left out of scope for this work.}
 
@@
 
-Missing line center measurements are replaced with de-noised estimates.
+Missing line center measurements, which were previously marked by \code{NaN}s, are replaced with de-noised estimates.
 This is done iteratively until the estimates of missing line centers change by less than 0.01\%.
@@
 \KwData{line positions $x(\lambda,m,n)$ for each exposure $n$, with wavelengths $\lambda$ and echelle orders $m$}
-\KwResult{Low-dimensional calibration space of spectrograph}
+\KwResult{Principle vectors of low-dimensional calibration space $g_k(\lambda,m)$, location of exposures in this space $a_{n,k}$}
 \While{change in missing or outlier line centers $>$ 0.01\%}
@@
 
-\subsection{?Interpolating Calibration Frames?}
+\subsection{?Interpolating Calibration Position?}
  \label{sec:interp_time}
@@
 In the implementation of \name\ presented here, the magnitudes of the principal components are interpolated with respect to time.
-New magnitudes $a_n'k$ values are found for the midpoint time of non-calibration exposures and used to construct the calibration state of the spectrograph for that exposure.
-Using interpolated magnitudes, $a_{n;k}$ and the basis vectors, $g_k(\lambda,m)$, returned by the de-noising process, \name\ can construct a set of calibration lines, $x(\lambda,m,n') $, for any exposure as prescribed in Equation \ref{eq:excl_wsol}.
+New magnitudes $a_n'k$ are found for the midpoint time of non-calibration exposures and used to construct the calibration state of the spectrograph for that exposure.
+Using interpolated magnitudes, $a_{n;k}$, and the basis vectors, $g_k(\lambda,m)$, returned by the de-noising process, \name\ can construct a set of calibration lines, $x(\lambda,m,n') $, for any exposure as prescribed in Equation \ref{eq:excl_wsol}.
+\lz{The best interpolation scheme for \expres\' instrument drift is discussed below in Section \ref{sec:choice_avt}, but may vary for different instruments.}
 
@@
 line centers onto every integer x will generate wavelengths for each pixel in an echelle order.
+
+We found that a cubic-spline interpolation that enforces monotonicicity, such as a Piecewise Cubic Hermite Interpolatin Polynoimal (PCHIP) interpolator, works well for interpolating wavelengths onto pixels.
+A cubic spline allows for more flexibility than a continuous function, while the enforced monotonicity prevents large deviations that may befall a cubic spline.
+Choice in interpolation scheme and other tests are further discussed in Section \ref{sec:choice_wvp}.
 
@@
 
-The results and discussion presented here are based off of 1227 LFC exposures taken between October 14 and December 18, 2020 on 29 unique nights.
 LFC lines cover 40 echelle orders, which contain 19203 lines.
 Though the results are primarily based on work with LFC data, there is some discussion of applications to arc lamps.
-These results use 78 ThAr exposures that were taken on the same nights.
 ThAr lines cover all 86 extracted orders of \expres, which includes 5295 lines.
+
 For both the LFC and ThAr data, lines that appeared in less than 60\% of exposures were not included in the analysis.
@@
 
- \name\ proceeds from a given list of echelle orders, line wavelengths, and pixel positions of each line for each exposure.
-We, therefore, use line positions generated by the pre-existing \expres\ pipeline \citep{petersburg2020}.
+ \name\ requests as input a list of echelle orders, line wavelengths, and pixel positions of each line for each exposure.
+We, therefore, use line positions generated by the pre-existing \expres\ pipeline from the following procedure \citep{petersburg2020}.
 
@@
 This code identifies ThAr lines by matching lines in an exposure against a line atlas.
+Lines are matched manually when there is a significant change in the instrument, but is otherwise done using forward-modeling.
+Once each line's position is identified and matched to a wavelength from the line atlas, a sixth-order, 2D polynomial is fit over pixel location $x$, echelle order $m$, and scaled wavelength $m\lambda$ (in order to distinguish lines that may appear in more than one order).
 
@@
  where the repetition rate, $f_r$, is known from design of the LFC, and the offset frequency, $f_0$, has been determined experimentally.
-the pre-existing \expres\ pipeline then generates wavelength solutions on a night-by-night basis as described in Section \ref{sec:method}.
 
@@
 Each instrument epoch is therefore treated independently.
+In the results presented here, we use 1227 LFC exposures and 78 ThAr exposures taken between October 14 and December 18, 2020 on 29 unique nights.
 
@@
 We perform a series of tests to validate the performance of \name\ and benchmark \name -generated wavelengths against wavelengths generated by a classic, parametric method.
-For all these tests, we generate wavelengths for known calibration lines and compare the predicted wavelength to the assigned wavelength of each line.
-This inherently folds in errors in the measured line center of each calibration line, but the contribution to the residuals will be the same across all tests.
+For all these tests, we leave out a subset of calibration lines with known wavelengths as a ``validation'' sample, generate wavelengths for these lines, and compare the predicted wavelength to the assigned wavelength of each line.
+This inherently folds in errors in the measured line center of each calibration line, but this contribution to the residuals will be the same across all tests.
 
@@
 We then construct a wavelength solution using only the odd lines and use it to predict wavelengths for the even lines; i.e.
-a polynomial is fit to just the odd lines and then evaluated at the detector position of the even lines.
+a polynomial is fit to just the odd lines and then evaluated at the detector position of the even lines (see Equation \ref{eq:poly_wsol}).
 We then generate a wavelength solution using only the even lines and use it to predict wavelengths for the odd lines.
@@
 Histograms of the residuals for each of the above described polynomial, interpolation, and denoising tests (respectively) are shown in Figure \ref{fig:testHists}.
-Note that the spread in residuals is much smaller for both the denoising and interpolation tests relative to the results of the polynomial wavelength solutionl.
+Note that the spread in residuals is much smaller for both the denoising and interpolation tests relative to the results of the polynomial wavelength solution.
 
@@
 \includegraphics[width=.5\textwidth]{Figures/all_results.pdf}
-\caption{Difference in predicted and assigned wavelengths for the wavelength calibration tests described in Section \ref{sec:tests}.
+\caption{\lz{(change to be b/w safe)} Difference in predicted and assigned wavelengths for the wavelength calibration tests described in Section \ref{sec:tests}.
 The standard deviation of the residuals for each test is given in the top-right corner in each method's corresponding color.}
@@
 %\includegraphics[width=\textwidth]{Figures/lineResids2D.pdf}
-\caption{Residuals plotted with respect to detector position (as defined by echelle order and x-pixel) for various wavelength calibration methods.
+\caption{Residuals for a single image plotted with respect to detector position (as defined by echelle order and x-pixel) for various wavelength calibration methods.
 Each line is colored by the difference between the predicted wavelength and the true wavelength for each line, given in units of $/mps$.
@@
 There is a lot of vertical structure and some hints of a periodic, diagonal structure as well.
+\lz{(Possibly from detector effects?
+We're double checking this now and thinking about approaching STA about it.
+Debra was considering publishing the results from that discussion in this paper if it seems relevant.)}.
 
@@
 \subsection{Value of K}
+\label{sec:choice_k}
 The value of $K$ represents the dimensionality of the calibration space within which the spectrograph lives.
@@
 One would expect the returned wavelengths to get increasingly more accurate with larger K until components that represent only noise are incorporated.
-We found that the returned wavelengths have the lowest residuals with a K value of 32.
+We found that the returned wavelengths have the lowest residuals with a K value of 32 (see Figure \ref{fig:kvals}).
 However, the improvement is slight with K values greater than six.
@@
 
+\begin{figure}[t]
+\centering
+\includegraphics[width=.1\textwidth]{Figures/pcsLfc6.png}
+\caption{Spread in residuals as a function of $K$.}
+\label{fig:kvals}
+\end{figure}
+
 \subsection{Interpolation of $a_{nk}$ with Respect to Time}
+\label{sec:choice_avt}
 Figure \ref{fig:nightlyVariation} shows the amplitude of the first and second principal component with respect to time on the left.
@@
 We were very conservative with defining eras of stability to ensure that within each era, the calibration-space of the instrument is low-dimensional as assumed.
+
+\mb{(Can we quantify or illustrate how much change is too much change, for use in discussion of other instruments?)}
 
@@
 
+\lz{Add empirical determination of how often EXPRES needs to redo calibration exposures?}
+
 \subsection{Interpolation of Wavelengths with Respect to Pixel}
+\label{sec:choice_wvp}
 Interpolation of wavelengths over pixel  is done order by order using a Piecewise Cubic Hermite Interpolating Polynomial (PCHIP) interpolator.
@@
 
-A more classic cubic spline interpolation can run into issues with arc lines, which are irregularly spaced or even blended and can therefore appear very close together.
-Very close lines cause huge deviations from the correct wavelengths, as seen in the green line of Figure \ref{fig:xinterp}.
+A more classic cubic spline interpolation can run into issues with arc lines, which are irregularly spaced or even blended.
+Close lines appearing at virtually the same pixel location but with different wavelengths could coerce a cubic spline into a very high slope, as with the green curve in Figure \ref{fig:xinterp} at approximately pixel 5300 where the spline is nearly vertical.
+This leads to huge deviations from the correct wavelengths.
 
@@
 Starting with a list of calibration lines with known wavelengths and well-fit line centers for each calibration exposure, \name\ will de-noise and interpolate the given lines into a full wavelength solution.
-We show that \name\ returns residuals with greater than a factor of two less spread than classic, parametric methods.
-Using \name\ wavelengths reduced the RMS in a planet fit to derived RVs by \lz{some percent hopefully probs}.
+We show that \name\ returns a lower RMS than classic, parametric methods by over a factor of two.
+Using \name\ wavelengths also reduced the RMS in a planet fit to derived RVs by \lz{some percent hopefully probs}.
 
@@
 
-\lz{Choices?}
+We have described only one, fairly simplistic implementation of \name\ here.
+There are many options for both the de-noising and interpolation steps.
+\lz{(Maybe something to discuss at a future meeting; I'm not actually entirely clear on other possibilities)}.
+For instance, a Gaussian process model could be constructed to determine wavelengths in place of a spline.
+\lz{(?)}  \lz{Something about going Bayesian?}
 
-An advantage of applying PCA to line positions from all LFC exposures is the ability to isolate exposures that exhibit errant variation, which is typically associated with flawed exposures.
+\mb{(Is PCA a ``choice'' or an ``extension'')}
+An advantage of this implementation, where PCA is applied to line positions from all LFC exposures, is the ability to isolate exposures that exhibit errant variation, which is typically associated with flawed exposures.
 This allowed us to quickly vet for problematic LFC exposures, which otherwise would have required visual inspection of all 1200+ LFC exposures.
@@
 
-We have described only one, fairly simplistic implementation of \name\ here.
-There are many options for both the de-noising and interpolation steps.
-\lz{(Maybe something to discuss at a future meeting; I'm not actually entirely clear on other possibilities)}.
-For instance, a Gaussian process model could be constructed to determine wavelengths in place of a spline.
-\lz{(?)}  \lz{Something about going Bayesian?}
-
 \lz{(Not sure where this belongs, honestly.)}  We caution that with any wavelength solution, there is a perfect degeneracy between what is defined as the ``center of the line'' and the resultant wavelength solution.
@@
 
+\mb{Add discussion of applicability to non-EPRV instruments.  Use ThAr results to comment on instruments w/o very dense calibration sources.}
+\lz{We can maybe use the temperature swing in September to discuss less-stable spectrographs?
+But it will remain hard to properly characterize ``stability'', and I'm not sure how representative that is for other instruments.
+I think I manually did it by warding against high-order changes in the principal component amplitudes?
+But presumably there exists an instrument for which the whole ability to implement PCA won't be true.}
+
 \name\ is designed and optimized for extreme-precision instruments.
+
 We believe that by implementing \name, the uncertainty from wavelength calibration is rendered negligible in the face of other noise sources.

### 2020-05-14 paper/paper.tex
Full Incorporation of Bedell Suggestions
Discussion of applying to other spectrographs, other possible implementations, other things...

@@
 \includegraphics[width=\textwidth]{Figures/methodCartoon.pdf}
-\caption{}
+\caption{A cartoon representation of the \name\ method, as described in Section \ref{sec:method}.
+Variations in line position, amplitude of basis vectors, and interpolation greatly enlarged to show texture.
+In step one, dimensionality reduction and denoising, the complete set of line positions for each exposure is analyzed to return a set of basis vectors, $g_n$, which span the calibration space of the spectrograph.
+In step two, \name\ interpolates the amplitudes of each basis vector, $a_n,k$, to return the calibration state for a specific exposure, given as a set of de-noised lines.
+The assigned wavelengths of these de-noised line positions are then interpolated onto other pixels in step three to construct a full wavelength model.}
 \label{fig:testHists}
@@
 Using interpolated magnitudes, $a_{n;k}$, and the basis vectors, $g_k(\lambda,m)$, returned by the de-noising process, \name\ can construct a set of calibration lines, $x(\lambda,m,n') $, for any exposure as prescribed in Equation \ref{eq:excl_wsol}.
-\lz{The best interpolation scheme for \expres\' instrument drift is discussed below in Section \ref{sec:choice_avt}, but may vary for different instruments.}
+\lz{The best interpolation scheme for \expres\ is discussed below in Section \ref{sec:choice_avt}, but may vary for different instruments.}
 
@@
 
-\begin{figure*}[t]
+\begin{figure}[t]
 \centering
-\includegraphics[width=\textwidth]{Figures/lineDensity.png}
-\caption{\lz{(Figure I use for talks, maybe shouldn't be included here)}  Comparison of polynomial-fit and interpolated wavelength solutions using either just ThAr lines or LFC lines.
-Left: the difference between each wavelength solution evaluated at the position of the LFC lines and the assigned wavelength of that LFC line.
-The right plot is simply a zoom in of the left plot.
-Note the large deviation when trying to implement Excalibur with just ThAr lines, which exhibit much large separations than LFC lines.
-The classic, polynomial  fit exhibits similar residuals regardless of whether the set of ThAr lines or LFC lines are used.
-Excalibur using LFC lines exhibit the lowest residuals.}
+\includegraphics[width=0.5\textwidth]{Figures/lineDensity.png}
+\caption{Comparison of polynomial-fit and interpolated wavelength solutions using either just ThAr lines or LFC lines.
+For a subset of one order, the difference between the assigned and model wavelengths for each LFC line is plotted with respect to pixel.
+The large residuals between two widely-spaced ThAr lines when implementing an interpolated wavelength solution with just ThAr lines (blue curve) extends out of frame.
+The classic, polynomial fit exhibits similar residuals in both amplitude and shape regardless of whether the set of ThAr lines or LFC lines are used.
+An interpolated wavelength solution using LFC lines (orange curve) exhibits the lowest residuals.}
 \label{fig:waveResids}
-\end{figure*}
+\end{figure}
 
@@
 There is a lot of vertical structure and some hints of a periodic, diagonal structure as well.
-\lz{(Possibly from detector effects?
-We're double checking this now and thinking about approaching STA about it.
-Debra was considering publishing the results from that discussion in this paper if it seems relevant.)}.
-
 The residuals of the interpolation test for the same exposure is shown in the center plot of Figure \ref{fig:resid2d}.
@@
 We believe the flexibility of the interpolated model was able to account for high-order detector defects, which emerged as structure in the residuals of the classic, smooth, polynomial-driven wavelength solution.
+In the \expres\ data, for example, the interpolated model accounted for \lz{detector effect STA will explain to us}.
 
@@
 
-\section{Choice Choices} \label{sec:choices}
-This implementation of \name\ presented here incorporated determining various global variables and methods that we believe are or are close to optimal for constructing a high-fidelity wavelength solution.
-This section will describe each choice and the associated decision-making process.
+\section{Choose Your Own Implementation} \label{sec:choices}
+We describe here only one implementation of \name.
+Using PCA and an interpolated wavelength solution is a fairly trivial step towards a complete hierarchical and non-parametric wavelength model.
+It is possible to upgrade both the denoising and wavelength solution steps to true models.
+
+\lz{Hogg: please edit/add to these methods}
+
+For the dimensionality reduction and denoising, the PCA could be replaced by a probabilistic PCA model or other probabilistic linear reduction methods.
+It is also possible to move to non-linear reduction methods, like...?
+
+The wavelength solution could also move past interpolation.
+For example, a Gaussian process could be used that is constrained to ensure monotonicity.
+Replacing each step with a model will allow for full, hierarchical Bayesian inference.
+This means error from wavelength calibration could be marginalized out.
+This is only worth doing, however, if wavelength calibration is a significant fraction of the error budget.
+
+The implementation of \name\ presented here, using PCA for denoising and interpolating a wavelength solution, incorporated various global variables and methods that we believe are or are close to optimal for constructing a high-fidelity wavelength solution.
+The following subsections will describe each choice and the associated decision-making process.
 
@@
 One would expect the returned wavelengths to get increasingly more accurate with larger K until components that represent only noise are incorporated.
-We found that the returned wavelengths have the lowest residuals with a K value of 32 (see Figure \ref{fig:kvals}).
+We found that the returned wavelengths plateau at a K value of 32 (see Figure \ref{fig:kvals}).
 However, the improvement is slight with K values greater than six.
 Comparisons of wavelengths returned by a K=6 model vs.
-a K=32 model show significant differences for less than 10, bluer lines.
-A visual inspection of the principal components, such as the one displayed in Figure \ref{fig:pcLfc}, led us to believe most of the structure is being captured in the first six principal components.
+a K=32 model show significant differences for less than 10 bluer lines, which are know to have greater error in their measured line positions.
+A visual inspection of the principal components, such as the one displayed in Figure \ref{fig:pcLfc}, additionally show most of the structure is being captured in the first six principal components with later components dominated by noise.
 
@@
 \centering
-\includegraphics[width=.1\textwidth]{Figures/pcsLfc6.png}
-\caption{Spread in residuals as a function of $K$.}
+\includegraphics[width=.5\textwidth]{Figures/kvals_all.pdf}
+\caption{Spread in residuals of returned wavelengths for different values of $K$.  There is a dotted, vertical line at $K=6$, our chosen $K$ value.  The dotted, horizontal line is the RMS for $K=6$.}
 \label{fig:kvals}
@@
 
-\subsection{Interpolation of $a_{nk}$ with Respect to Time}
+\subsection{Interpolation of Calibration State with Respect to Time}
 \label{sec:choice_avt}
@@
 
-\lz{Yeah, the right-figure is kind of a mess, isn't it?}
+The choice in interpolation method can help guide how many wavelength calibration images are truly needed.
+It is unnecessary to take calibration images at times where similar information can be filled in by a well-chosen interpolation scheme.
+For example, with the \expres\ data shown here, it is clear that nightly calibration images are needed, but for a linear trend, only two calibration images throughout the night are strictly required.
 
-We differentiated eras of stability by where the principal component amplitudes showed large variation.
-The analysis was done by eye here, though many algorithms for change-point detection exist in the literature \lz{(Am I going to have to cite things?)}.
-We were very conservative with defining eras of stability to ensure that within each era, the calibration-space of the instrument is low-dimensional as assumed.
+The amplitudes $a_{n,k}$ can also be interpolated with respect to any housekeeping data, not just time.
+Best results should come from interpolating with respect to whichever data is most strongly correlated with the calibration state of the spectrograph.
+For example, with \expres, which is passively temperature controlled, the returned amplitudes $a_{n,k}$ were extremely correlated with the enclosure temperature, suggesting it would also be possible to interpolate with respect to temperature.
 
-\mb{(Can we quantify or illustrate how much change is too much change, for use in discussion of other instruments?)}
+Another pertinent example would be a spectrograph that is mounted on the telescope, and therefore moves with the telescope.
+In this case, it may be important to interpolate at least in part with respect to the position of the telescope, which enables the resultant calibration to incorporate the gravitational loading experienced by the spectrograph.
 
@@
 \includegraphics[width=0.42\textwidth]{Figures/pcAs_byDay.png}
-\caption{Amplitude of the first two principal components show as a function of time (left) or fraction of a day (right).
+\caption{Amplitude of the first two principal components show as a function of time (left) or fraction of a day (right) \lz{(Yeah, it's kind of a mess)}.
 The top plots shows the amplitudes for the first principal component while the bottom plot shows the amplitudes for the second principal component.
@@
 \end{figure*}
-
-\lz{Add empirical determination of how often EXPRES needs to redo calibration exposures?}
 
@@
 \includegraphics[width=\textwidth]{Figures/intpx_tests.png}
-\caption{Results from different interpolation schemes over pixels.
-ThAr lines, shown as pink points, are used to construct a wavelength solution that is then evaluated at each LFC line, shown as black lines.
+\caption{Residuals from different interpolation schemes over pixels.
+ThAr lines, shown as black circles, are used to construct a wavelength solution that is then evaluated at each LFC line, shown as black lines.
 Histograms of the residuals for each method is shown on the right.
-Note: there is a blended ThAr line at approximately pixel 5300.}
+Note: there is a blended ThAr line at approximately pixel 5300, the last ThAr line plotted.}
 \label{fig:xinterp}
@@
 \lz{We talked about including this discussion, but there is no real reason for doing things in this order except it is a bit faster.}
+
+\section{Application to Other Spectrographs}
+We use \expres\ data here not because it is the optimal system for a method like \name, but because there is a strong need for good wavelength calibration in the EPRV community.
+The applicability of \name\ to any one instrument is a detailed question of the kind variance experienced by the spectrograph and calibration sources available, but we hope to provide some approximate benchmarks and diagnostics.
+
+To determine the needed line density, we systematically removed LFC lines from the \expres\ data and calculated the RMS of the returned wavelengths.
+Figure \ref{fig:density} plots the increase in the scatter of the wavelengths returned by \name\ as a function of separation between lines.
+However, the stability of the lines and their measured line centers quickly becomes the dominant error source in calibration lamps over the separation between lines.
+
+In terms of stability, \name\ should be adaptable to a wide range of instrument designs.
+In truth, the amplitude of the variation matters little as long as all axes of variation can be captured by PCA or another dimensionality-reducing model.
+Indeed, larger variations from less stabilized instruments might even make it easier for the PCA to identify these axes.
+Most physical systems should have a few dominant axis along which they vary, simply by what is possible with mechanical engineering.
+With PCA, this can be tested by plotting the power of the basis vectors describing the calibration space, which should fall quickly after a few vectors.
+Note, however, that this is only a measure of the variance in the PCA space, not an explicit test of the end-to-end variation in the resulting model.
+This condition is therefore necessary, but not sufficient if implementing \name with PCA.
+
+It could still be possible to run \name\ on a spectrograph that has a high-dimensional calibration space, meaning a large number of basis vectors is required to capture all the variance in the spectrograph.
+In this regime, there is always the risk of over-fitting.
+Regularizing the principal component amplitudes, for example insisting the amplitudes for higher principal components be smaller, can help to return reasonable models \citep{formanmackey2015} .
+Within such a framework, \name\ may still deliver good results.
+\lz{Hogg: My notes said you might want to add more things here?}
+
+For the results presented here, the data was broken up into different eras of stability based on where the principal component amplitudes showed huge deviations.
+This was done visually, though there exists many change-point detection algorithms that could be used.
+There is a trade-off in including more exposures between introducing greater variation, but also more data to offer constraints that may be optimized.
+Here, an era of stability was chosen  manually in order to focus on separating out time-domain intervals in which the data is relatively homogeneous, e.g.
+most exposures show the same lines.
+Homogeneity is, of course, implicitly required when implementing PCA.
 
@@
 
-\mb{Add discussion of applicability to non-EPRV instruments.  Use ThAr results to comment on instruments w/o very dense calibration sources.}
-\lz{We can maybe use the temperature swing in September to discuss less-stable spectrographs?
-But it will remain hard to properly characterize ``stability'', and I'm not sure how representative that is for other instruments.
-I think I manually did it by warding against high-order changes in the principal component amplitudes?
-But presumably there exists an instrument for which the whole ability to implement PCA won't be true.}
-
 \name\ is designed and optimized for extreme-precision instruments.

### 2020-05-14 paper/paper.tex
A bit of clean-up

@@
 Temperature-controlled, fiber-fed spectrographs vary only slightly over the night or season.
-\lz{This suggests that such stabilized spectrographs vary along only} a small number of axes in what you might call ``calibration space'', or the (very high dimensional) space of all possible wavelength solutions.
+This suggests that such stabilized spectrographs vary along only a small number of axes in what you might call ``calibration space'', or the (very high dimensional) space of all possible wavelength solutions.
 That is, not only are the spectrographs stable, but they also have few environmentally accessible degrees of freedom.
@@
 We have in mind \eprv\ instruments and \eprv\ science cases, but we expect \name\ to have applications for other kinds of spectrographs in other contexts.
-\mb{The motivating ideas behind this project are true for almost every astronomical spectrograph.
-(Is this true?)}
+\name\ should be applicable to almost every astronomical spectrograph.
 
@@
 Here the y-dependence is dropped as that dependence is carried by spectral order $m$.
-\lz{The line position information is therefore uniquely identified by echelle order $m$ and pixel in the dispersion direction, $x$.}  The coefficients $c_{nij}$ are interpolated to time $t_n$ of exposure $n$ by a third-order polynomial of time.
+The line position information is therefore uniquely identified by echelle order $m$ and pixel in the dispersion direction, $x$.
+The coefficients $c_{nij}$ are interpolated to time $t_n$ of exposure $n$ by a third-order polynomial of time.
 This third-order polynomial is evaluated at the time of non-calibration exposures to re-construct a 2D, 9th-order polynomial wavelength solution for that exposure.
@@
 When this calibration structure is used to deliver a wavelength solution, we will have to invert $x(\lambda,m,n)$ into $\lambda(x,m,n)$.
-\mb{(Explain why this inversion is necessary?)}
 
@@
 Each line is uniquely defined by a combination of echelle order, $m$, and ``true'' or theoretical wavelength, $\lambda$.
-\lz{There are many strategies for identifying calibration line positions and matching them to their assigned wavelengths; this problem is left out of scope for this work.}
+There are many strategies for identifying calibration line positions and matching them to their assigned wavelengths; this problem is left out of scope for this work.
 
@@
 Using interpolated magnitudes, $a_{n;k}$, and the basis vectors, $g_k(\lambda,m)$, returned by the de-noising process, \name\ can construct a set of calibration lines, $x(\lambda,m,n') $, for any exposure as prescribed in Equation \ref{eq:excl_wsol}.
-\lz{The best interpolation scheme for \expres\ is discussed below in Section \ref{sec:choice_avt}, but may vary for different instruments.}
+The best interpolation scheme for \expres\ is discussed below in Section \ref{sec:choice_avt}, but may vary for different instruments.
 
@@
 
- \lz{(Mention degeneracy with prescribed line center and wavelength solution here?  Or in discussion?)}
-
 In order to satisfy the assumption of stability needed for \name, only exposures from only one ``instrumental epoch" of \expres\ are used.
@@
 \includegraphics[width=.5\textwidth]{Figures/all_results.pdf}
-\caption{\lz{(change to be b/w safe)} Difference in predicted and assigned wavelengths for the wavelength calibration tests described in Section \ref{sec:tests}.
+\caption{Difference in predicted and assigned wavelengths for the wavelength calibration tests described in Section \ref{sec:tests}.
 The standard deviation of the residuals for each test is given in the top-right corner in each method's corresponding color.}
@@
 Stabilized spectrograph hardware with few degrees of freedom allow us to use all calibration images in a given generation of stability to constrain the accessible calibration space of the spectrograph.
+We have described only one, fairly simplistic implementation of \name\ here.
+There are many options for both the de-noising and interpolation steps, as mentioned in Section \ref{sec:choices}.
 
-We have described only one, fairly simplistic implementation of \name\ here.
-There are many options for both the de-noising and interpolation steps.
-\lz{(Maybe something to discuss at a future meeting; I'm not actually entirely clear on other possibilities)}.
-For instance, a Gaussian process model could be constructed to determine wavelengths in place of a spline.
-\lz{(?)}  \lz{Something about going Bayesian?}
-
-\mb{(Is PCA a ``choice'' or an ``extension'')}
 An advantage of this implementation, where PCA is applied to line positions from all LFC exposures, is the ability to isolate exposures that exhibit errant variation, which is typically associated with flawed exposures.

### 2020-05-14 paper/paper.tex
Proofread and New Figures

@@
 \begin{abstract}\noindent%
-\name\ is a hierarchical, non-parametric framework for wavelength calibrating precision spectrographs.
+\name\ is a hierarchical, non-parametric framework for wavelength calibrating spectrographs.
 This method capitalizes on recent hardware improvements in highly-stabilized, extreme-precision radial-velocity spectrographs.
@@
 \name\ also takes advantage of the development of laser-frequency combs or etalons, which generate a dense set of stable calibration points.
-This density of points permits the freedom of a non-parametric wavelength solution that can match or adapt to any instrument or detector oddities, more so than any one polynomial.
+This density of points permits the freedom of a non-parametric wavelength solution that can match or adapt to any instrument or detector oddities, more so than any parametric model.
 We demonstrate the success of this method with data from the EXtreme PREcision Spectrograph, \expres, which includes a Menlo Systems laser frequency comb.
-We find that \name\ returns more accurate wavelength predictions by more than a factor of 2 over classic, non-hierarchical, polynomial solutions.
-Employing \name\ wavelengths on \expres\ RV measurements on HD 3651 reduced the RMS of a planet fit by \lz{blah}.
+We find that \name\ returns more accurate wavelengths by greater than a factor of 2 over classic, non-hierarchical, polynomial solutions.
+Employing \name\ wavelengths on \expres\ RV measurements of HD 3651 reduced the RMS of a planet fit by \lz{blah}.
 \end{abstract}
@@
 Some hardware--software systems, such as \espresso, \expres, and \neid\ were designed to deliver precision at this level \citep{pepe2013,  jurgenson2016, neid}.
-Early data from \expres\ and \espresso\ suggest that the hardware improvements do propagate through to sub-$\mps$ uncertainties on final, on-sky RV measurements \citep{blackman2020, petersburg2020, pepe2020}.
+Early data from \expres\ and \espresso\ suggest that the hardware improvements truly propagate through to sub-$\mps$ uncertainties on final, on-sky RV measurements \citep{blackman2020, petersburg2020, pepe2020}.
 
@@
 
-Here we propose to simplify and improve calibration programs for \eprv\ hardware systems with two very simple but innovative ideas.
+Here, we propose to simplify and improve calibration programs for \eprv\ hardware systems with two very simple but innovative ideas.
 The first flows from the fact that calibration sources---which include arc lamps (in some wavelength ranges), etalons, and laser-frequency combs (\lfc s)---illuminate the spectrograph with very stable, very dense sets of lines; almost every location in the spectrograph image plane is surrounded by nearby, useful calibration lines.
@@
 This recommends a calibration methodology that is \emph{non-parametric}:  If every point in the spectrograph detector is surrounded by nearby calibration lines, the wavelength solution can, for example, be made simply as an interpolation of the calibration data.
-The density of lines removes the need to enforce any functional form for the wavelength solution (such as a continuous eighth-order polynomial, for example).
+The density of lines removes the need to enforce any functional form for the wavelength solution (such as a continuous ninth-order polynomial, for example).
 In some ways, this is a generalization of the recent work that demonstrated the efficacy of constructing a wavelength solution as multiple, segmented polynomials \citep{milakovic2020}.
-Going non-parametric will improve calibration accuracy by not forcing the choice of a parametric form that may bias the calibration, especially when the chosen function is inappropriate (as, for example, polynomials will be at detector edges).
+Going non-parametric will improve calibration accuracy by not forcing the choice of a parametric form that may bias the calibration, especially when the chosen function is inappropriate (as polynomials are at detector edges).
 
@@
 Temperature-controlled, fiber-fed spectrographs vary only slightly over the night or season.
-This suggests that such stabilized spectrographs vary along only a small number of axes in what you might call ``calibration space'', or the (very high dimensional) space of all possible wavelength solutions.
+This suggests that such stabilized spectrographs vary along only a small number of axes in what you might call ``calibration space'', or the (very high-dimensional) space of all possible wavelength solutions.
 That is, not only are the spectrographs stable, but they also have few environmentally accessible degrees of freedom.
@@
 In the context of probabilistic models, this structure is \emph{hierarchical}:  The calibration data are used not just to determine the wavelength solution at one moment, but also to determine the possible \emph{space} of wavelength solutions at all moments.
-In the statistics literature this concept is often described as \emph{de-noising}:  Every calibration exposure contains information about every other exposure.
+In the statistics literature, this concept is often described as \emph{de-noising}:  Every calibration exposure contains information about every other exposure.
 Thus every exposure can be improved (i.e.
@@
 The method we propose here---\name---embodies these ideas.
-It is a non-parametric, hierarchical, data-driven model for the wavelength solution.
+It is a non-parametric, hierarchical, data-driven method to generate a wavelength model.
 By being non-parametric, it delivers enormous freedom to the wavelength solution to match or adapt to any instrument or detector oddities.
@@
 \name\ is designed for temperature-controlled, fiber-fed spectrographs with good calibration sources, such as laser-frequency combs, or etalons.
-We have in mind \eprv\ instruments and \eprv\ science cases, but we expect \name\ to have applications for other kinds of spectrographs in other contexts.
+We have in mind \eprv\ instruments and \eprv\ science cases primarily because the need for good wavelength calibration in this field is so great.
+Irregardless, we expect \name\ to have applications for other kinds of spectrographs in other contexts.
 \name\ should be applicable to almost every astronomical spectrograph.
@@
 \section{Method} \label{sec:method}
-The \name\ method is designed to turn a series of calibration lines with known wavelengths and well-fit detector positions, and de-noise and interpolate them into a full wavelength model applicable to all exposures taken with the instrument.
+The \name\ method is designed to take a series of calibration lines with known wavelengths and well-fit detector positions, and de-noise and interpolate them into a full wavelength model applicable to all exposures taken with the instrument.
 \name\ operates on the two core ideas that the wavelength solution should be given enormous flexibility, but that it lives in a very low-dimensional space, where the degrees of freedom are set by the limited kinematics of the stabilized spectrograph hardware.
@@
 Additionally, \name\ assumes the instrument hardware is stabilized enough that the space of possible calibration states is low-dimensional.
-We revisit and quantify these conditions in Section \ref{sec:discussion}.
+We revisit and quantify these conditions in Section \ref{sec:others}.
 
@@
 \end{equation}
-where $\theta_{n}$ represents the parameters describing a given exposure.
+where $\theta_{n}$ represents the parameters describing the wavelength solution for a given exposure.
 
 Classically, pipelines employ polynomials to construct smooth wavelength solutions for each exposure.
-For example, the \expres\ pipeline sets the function $f(x,m;\theta_{n})$ from Equation \ref{eq:wsol} to be a 2D, 9\textsuperscript{th}-order polynomial where $\theta_{n}$ represents the polynomial coefficients, $c_{nij}$, unique to each exposure $n$ \citep{petersburg2020}.
+For example, the \expres\ pipeline sets the function $f(x,y,m;\theta_{n})$ from Equation \ref{eq:wsol} to be a 2D, 9\textsuperscript{th}-order polynomial where $\theta_{n}$ represents the polynomial coefficients, $c_{nij}$, unique to each exposure $n$ \citep{petersburg2020}.
 \begin{equation}
@@
 A PCA is justifiable in the limit where there is very high signal-to-noise ratio, as is usually the case with typical calibration images.
-There are many alternatives to PCA for this dimensionality reduction; we return to this point in the discussion (\textsection \ref{sec:discussion}) below.
+There are many alternatives to PCA for this dimensionality reduction; we return to this point in Section \ref{sec:choices} below.
 
@@
 \caption{A cartoon representation of the \name\ method, as described in Section \ref{sec:method}.
-Variations in line position, amplitude of basis vectors, and interpolation greatly enlarged to show texture.
-In step one, dimensionality reduction and denoising, the complete set of line positions for each exposure is analyzed to return a set of basis vectors, $g_n$, which span the calibration space of the spectrograph.
-In step two, \name\ interpolates the amplitudes of each basis vector, $a_n,k$, to return the calibration state for a specific exposure, given as a set of de-noised lines.
+We enlarge variations in measured line position, changes in calibration space, and interpolation deviations for clarity.
+In step one, dimensionality reduction and denoising, the complete set of line positions for all exposures is analyzed to return a set of basis vectors, $g_n$, which represent different ways the spectrograph calibration changes, and spans the possible calibration space of the spectrograph.
+In step two, \name\ interpolates the amplitudes of each basis vector, $a_n,k$, to return the calibration state for a specific exposure, returned as a set of de-noised calibration lines.
 The assigned wavelengths of these de-noised line positions are then interpolated onto other pixels in step three to construct a full wavelength model.}
@@
 
-\subsection{?Interpolating Calibration Position?}
+\subsection{Interpolating Calibration Position}
  \label{sec:interp_time}
@@
 
-In the implementation of \name\ presented here, the magnitudes of the principal components are interpolated with respect to time.
-New magnitudes $a_n'k$ are found for the midpoint time of non-calibration exposures and used to construct the calibration state of the spectrograph for that exposure.
-Using interpolated magnitudes, $a_{n;k}$, and the basis vectors, $g_k(\lambda,m)$, returned by the de-noising process, \name\ can construct a set of calibration lines, $x(\lambda,m,n') $, for any exposure as prescribed in Equation \ref{eq:excl_wsol}.
-The best interpolation scheme for \expres\ is discussed below in Section \ref{sec:choice_avt}, but may vary for different instruments.
+In the implementation of \name\ presented here, the magnitudes of the principal components are interpolated linearly with respect to time.
+\name\ finds new magnitudes $a_{n'k}$ for an exposure $n'$ at time $t_{n'}$, which are used to construct the calibration state of the spectrograph for that exposure.
+Using interpolated magnitudes, $a_{n,k}$, and the basis vectors, $g_k(\lambda,m)$, returned by the de-noising process, \name\ can construct a set of calibration lines, $x(\lambda,m,n') $, for any exposure as prescribed in Equation \ref{eq:excl_wsol}.
 
@@
 
-Find $a_{n',k}$ by interpolating $a_{n,k}$ with to $t_{n'}$\;
+Find $a_{n',k}$ by interpolating $a_{n,k}$ with respect to $t_{n'}$\;
 $x(\lambda,m,n') = g_0(\lambda,m) + \sum_{k=1}^K a_{n'k}\,g_k(\lambda,m)$ where $K=6$\;
@@
 ThAr exposures are taken at the beginning and end of each night.
-All calibration data is taken through the science fiber, meaning they traveling down the same optical pathway and are projected onto the same pixels as all science observations.
+All calibration data is taken through the science fiber, so that they down the same optical pathway and are projected onto the same pixels as all science observations.
 
@@
 ThAr lines cover all 86 extracted orders of \expres, which includes 5295 lines.
-
 For both the LFC and ThAr data, lines that appeared in less than 60\% of exposures were not included in the analysis.
@@
 
-In order to satisfy the assumption of stability needed for \name, only exposures from only one ``instrumental epoch" of \expres\ are used.
-\expres\ exposures are separated into different instrumental epochs that correspond to atypical hardware changes--changes that altered the position of the echellogram on the detector, the shape of the instrumental PSF, or the calibration sources.
+In order to satisfy the assumption of only low-order variation needed for \name, we used exposures from only one ``instrumental epoch" of \expres.
+\expres\ exposures are separated into different instrumental epochs that correspond to atypical hardware changes---changes that altered the position of the echellogram on the detector, the shape of the instrumental PSF, or the calibration sources.
 Significant instrumental changes break the assumption of a stabilized instrument with only low-dimensional variations.
@@
 We perform a series of tests to validate the performance of \name\ and benchmark \name -generated wavelengths against wavelengths generated by a classic, parametric method.
-For all these tests, we leave out a subset of calibration lines with known wavelengths as a ``validation'' sample, generate wavelengths for these lines, and compare the predicted wavelength to the assigned wavelength of each line.
+For all tests, we leave out a subset of calibration lines with known wavelengths as a ``validation'' sample, generate wavelengths for these lines using the remaining data, and compare the predicted wavelength to the assigned wavelength of each line.
 This inherently folds in errors in the measured line center of each calibration line, but this contribution to the residuals will be the same across all tests.
@@
 \includegraphics[width=.5\textwidth]{Figures/all_results.pdf}
-\caption{Difference in predicted and assigned wavelengths for the wavelength calibration tests described in Section \ref{sec:tests}.
+\caption{Difference in predicted and theoretical wavelengths for the wavelength calibration tests described in Section \ref{sec:tests}.
 The standard deviation of the residuals for each test is given in the top-right corner in each method's corresponding color.}
@@
 
-\begin{figure*}[t]
+\begin{figure}[t]
 \centering
-\includegraphics[width=\textwidth]{Figures/lineResids2D.png}
-%\includegraphics[width=\textwidth]{Figures/lineResids2D.pdf}
-\caption{Residuals for a single image plotted with respect to detector position (as defined by echelle order and x-pixel) for various wavelength calibration methods.
-Each line is colored by the difference between the predicted wavelength and the true wavelength for each line, given in units of $/mps$.
-Residuals to a classic, polynomial fit is shown in the left-most plot.
+\includegraphics[width=0.5\textwidth]{Figures/lineResids2D_col.png}
+%\includegraphics[width=0.5\textwidth]{Figures/lineResids2D_col.pdf}
+\caption{Residuals for a single LFC exposure plotted with respect to detector position (as defined by echelle order and x-pixel) for various wavelength calibration methods.
+Each line is colored by the difference between the predicted wavelength and the theoretical wavelength for each line, given in units of $/mps$.
+Residuals to a classic, polynomial fit is shown in the top plot.
 The center plot shows residuals to an interpolated wavelength solution.
-The right-most plot shows residuals when a ThAr wavelength solution is interpolated onto LFC line positions.
+The bottom plot shows residuals when a ThAr wavelength model generated by \name\ is interpolated onto LFC line positions.
 The position of ThAr lines are over-plotted in black.}
 \label{fig:resid2d}
-\end{figure*}
+\end{figure}
+
+The residuals of the classic, polynomial wavelength solution is shown in the top plot of Figure \ref{fig:resid2d}.
+There is a lot of vertical structure and some hints of a periodic, diagonal structure as well.
+The residuals of the interpolation test for the same exposure is shown in the center plot of Figure \ref{fig:resid2d}.
+There is no coherent structure here and smaller residuals.
+This shows how the flexibility of an interpolated model can account for high-order detector defects, which emerged as structure in the residuals of the classic, smooth, polynomial-driven wavelength solution.
+In the \expres\ data shown here, for example, the interpolated model accounted for \lz{detector effect STA will explain to us}.
+
+The bottom plot of Figure \ref{fig:resid2d} shows the residuals when ThAr exposures, which have much fewer and less regular lines than LFC exposures, are run through \name\ and used to predict wavelengths for the completely independent LFC exposures taken during the same range of time.
+Over-plotted in black are the positions of the ThAr lines.
+
+Note that running \name\ informed by only ThAr lines cannot be regarded as a direct comparison to the LFC runs, as the increased uncertainty and variability in ThAr line positions alone makes the resultant wavelength predictions an order of magnitude worse, hence the different scale of the colorbar.
+All the same, the residuals are relatively worse where lines are further spaced (for example, in redder echelle orders) than where lines are denser.
+
+Though the interpolated wavelength solution returns lower, less-structured residuals than the polynomial wavelength solution when guided by LFC lines, we see in the bottom plot of Figure \ref{fig:resid2d} that an interpolated wavelength solution can return much worse results in the space between widely separated calibration lines.
+Figure \ref{fig:waveResids} plots the residuals for a subset of one order for both a polynomial-based method and a PCHIP-based method guided by either ThAr lines or LFC lines.
+The PCHIP model with ThAr lines (blue) curve returns huge residuals between two widely-spaced ThAr lines.
+The classic, polynomial fit exhibits similar residuals in both amplitude and shape regardless of whether the set of ThAr lines or LFC lines are used.
+An interpolated wavelength solution using LFC lines (orange curve) exhibits the lowest residuals.
+
+The move to an interpolated wavelength solution is driven by the assumption that a high density of calibration lines allows for more freedom in the resultant wavelength solution.
+This freedom allows the wavelength solution to more accurately ascribe wavelengths.
+This flexibility, however, is no longer justified in the regime where there are large spaces between calibration lines, as is the case with a classic ThAr lamp.
 
@@
 For a subset of one order, the difference between the assigned and model wavelengths for each LFC line is plotted with respect to pixel.
-The large residuals between two widely-spaced ThAr lines when implementing an interpolated wavelength solution with just ThAr lines (blue curve) extends out of frame.
-The classic, polynomial fit exhibits similar residuals in both amplitude and shape regardless of whether the set of ThAr lines or LFC lines are used.
-An interpolated wavelength solution using LFC lines (orange curve) exhibits the lowest residuals.}
+The large residuals between two widely-spaced ThAr lines when implementing an interpolated wavelength solution with just ThAr lines (blue curve) extends out of frame.}
 \label{fig:waveResids}
@@
 
-The residuals of the classic, polynomial wavelength solution is shown in the left-most plot of Figure \ref{fig:resid2d}.
-There is a lot of vertical structure and some hints of a periodic, diagonal structure as well.
-The residuals of the interpolation test for the same exposure is shown in the center plot of Figure \ref{fig:resid2d}.
-There is no coherent structure here and smaller residuals.
-We believe the flexibility of the interpolated model was able to account for high-order detector defects, which emerged as structure in the residuals of the classic, smooth, polynomial-driven wavelength solution.
-In the \expres\ data, for example, the interpolated model accounted for \lz{detector effect STA will explain to us}.
-
-The right-most plot of Figure \ref{fig:resid2d} shows the residuals when ThAr exposures, which have much fewer lines than LFC exposures, are run through \name\ and used to predict wavelengths for the completely independent LFC exposures taken during the same range of time.
-Over-plotted in black are the positions of the ThAr lines.
-
-Though the interpolated wavelength solution returns lower, less-structured residuals than the polynomial wavelength solution when guided by LFC lines, we see in the right-most plot of Figure \ref{fig:resid2d} that an interpolated wavelength solution can return much worse results in the space between widely separated calibration lines.
-
-The move to an interpolated wavelength solution is driven by the assumption that a high density of calibration lines allows for more freedom in the resultant wavelength solution.
-Introducing this flexibility is no longer justified in the regime where there are large spaces between calibration lines, as is the case with a classic ThAr lamp.
-
-Note that running \name\ informed by only ThAr lines cannot be regarded as a direct comparison to the LFC runs, as the increased uncertainty and variability in ThAr line positions alone makes the resultant wavelength predictions an order of magnitude worse, hence the different scale of the colorbar.
-All the same, the residuals are relatively worse where lines are further spaced (for example, in redder echelle orders) than where lines are denser.
-
 \subsection{Impact on Radial Velocity Measurements}\label{sec:test-rv}
-\lz{Debra and RP currently working through this.  Results seem good.  Will ask them at next group meeting what we should include here}
+\lz{Awaiting latest extraction run.  Was supposed to take a weekend, but I'm pretty sure I heard God laugh when RP said that.}
 
 \section{Choose Your Own Implementation} \label{sec:choices}
-We describe here only one implementation of \name.
+We have described and test only one implementation of \name.
 Using PCA and an interpolated wavelength solution is a fairly trivial step towards a complete hierarchical and non-parametric wavelength model.
@@
 Replacing each step with a model will allow for full, hierarchical Bayesian inference.
-This means error from wavelength calibration could be marginalized out.
-This is only worth doing, however, if wavelength calibration is a significant fraction of the error budget.
+This means the uncertainty from wavelength calibration could be completely marginalized out.
+Doing so, however, is only worth it if the wavelength calibration is a significant fraction of the error budget.
 
-The implementation of \name\ presented here, using PCA for denoising and interpolating a wavelength solution, incorporated various global variables and methods that we believe are or are close to optimal for constructing a high-fidelity wavelength solution.
+The implementation of \name\ presented here, using PCA for denoising and interpolating a wavelength solution, uses various global variables and methods that we believe are or are close to optimal for constructing a high-fidelity wavelength solution.
 The following subsections will describe each choice and the associated decision-making process.
@@
 We settled on a $K$ value of 6.
-Figure \ref{fig:pcLfc} shows the first 6 principal components constructed using LFC lines.
-In both cases, there is clear structure in the first and second principal components.
-The diagonal-banding structure seen in components 3, 5, and 6 we believe to be capturing detector fabrication defects.
-Components 3 and 4 show aberrant behavior on the edges of the two bluest echelle orders, where lower signal results in more variation in the line fits.
+Figure \ref{fig:pcLfc} shows the first 9 principal components constructed using LFC lines.
+There is clear structure in the first and second principal components.
+\lz{The diagonal-banding structure seen in components 3, 5, and 6 we believe to be capturing detector fabrication defects.
+(Maybe, awaiting STA response)}  Components 3 and 4 show aberrant behavior on the edges of the two bluest echelle orders, where lower signal results in more variation in the line fits.
+Later principal components become dominated by noise over coherent structure.
 
@@
 \centering
-\includegraphics[width=\textwidth]{Figures/pcsLfc6.png}
+\includegraphics[width=\textwidth]{Figures/pcsLfc9.png}
 %\includegraphics[width=\textwidth]{Figures/pcsLfc6.pdf}
-\caption{The first six principal components constructed using LFC lines.
+\caption{The first nine principal components constructed using LFC lines.
 Each line is plotted according to its echelle order and x-pixel and colored by the value of the principal component for that line.}
@@
 One would expect the returned wavelengths to get increasingly more accurate with larger K until components that represent only noise are incorporated.
-We found that the returned wavelengths plateau at a K value of 32 (see Figure \ref{fig:kvals}).
-However, the improvement is slight with K values greater than six.
-Comparisons of wavelengths returned by a K=6 model vs.
-a K=32 model show significant differences for less than 10 bluer lines, which are know to have greater error in their measured line positions.
+Residuals might then get worse before ultimately starting to get better again with large K, which marks te model over-fitting the data.
+Though the returned RMS never gets worse, we find that the improvement plateaus between $K=6$ and $K=128$.
+Comparisons of wavelengths returned by a $K=6$ model vs.
+a $K=32$ model show significant differences in less than 10 bluer lines, which are known to have greater error and variance in their measured line positions.
 A visual inspection of the principal components, such as the one displayed in Figure \ref{fig:pcLfc}, additionally show most of the structure is being captured in the first six principal components with later components dominated by noise.
@@
 
-The choice in interpolation method can help guide how many wavelength calibration images are truly needed.
-It is unnecessary to take calibration images at times where similar information can be filled in by a well-chosen interpolation scheme.
-For example, with the \expres\ data shown here, it is clear that nightly calibration images are needed, but for a linear trend, only two calibration images throughout the night are strictly required.
-
-The amplitudes $a_{n,k}$ can also be interpolated with respect to any housekeeping data, not just time.
-Best results should come from interpolating with respect to whichever data is most strongly correlated with the calibration state of the spectrograph.
-For example, with \expres, which is passively temperature controlled, the returned amplitudes $a_{n,k}$ were extremely correlated with the enclosure temperature, suggesting it would also be possible to interpolate with respect to temperature.
-
-Another pertinent example would be a spectrograph that is mounted on the telescope, and therefore moves with the telescope.
-In this case, it may be important to interpolate at least in part with respect to the position of the telescope, which enables the resultant calibration to incorporate the gravitational loading experienced by the spectrograph.
-
 \begin{figure*}[h!]
@@
 
+The choice in interpolation method can help guide how many wavelength calibration images are truly needed.
+It is unnecessary to take calibration images at times where the same information can be constructed by a well-chosen interpolation scheme.
+For example, with the \expres\ data shown here, it is clear that nightly calibration images are needed, but for a linear trend, only two calibration images throughout the night are strictly required.
+
+The amplitudes $a_{n,k}$ can also be interpolated with respect to any housekeeping data, not just time.
+Best results will come from interpolating with respect to whichever data is most strongly correlated with the calibration state of the spectrograph.
+For example, with \expres, which is passively temperature controlled, the returned amplitudes $a_{n,k}$ were extremely correlated with the enclosure temperature, suggesting it would also be possible to interpolate with respect to temperature.
+
+Another pertinent example would be a spectrograph that is mounted on the telescope, and therefore moves with the telescope.
+In this case, it may be important to interpolate at least in part with respect to the position of the telescope, which enables the resultant calibration to incorporate the gravitational loading experienced by the spectrograph.
+
 \subsection{Interpolation of Wavelengths with Respect to Pixel}
@@
 
+A simple linear interpolation would obviously give erroneously low values everywhere.
 Due to the dispersion intrinsic to echelle spectrographs, the wavelength change between pixels grows greater with greater wavelengths.
@@
 pixel across an echelle order will always be monotonically increasing and concave down everywhere.
-A linear interpolation would obviously give erroneously low values everywhere.
 
 A more classic cubic spline interpolation can run into issues with arc lines, which are irregularly spaced or even blended.
-Close lines appearing at virtually the same pixel location but with different wavelengths could coerce a cubic spline into a very high slope, as with the green curve in Figure \ref{fig:xinterp} at approximately pixel 5300 where the spline is nearly vertical.
-This leads to huge deviations from the correct wavelengths.
+Close lines appearing at virtually the same pixel location but with different wavelengths could coerce a cubic spline into a very high slope, as with the blue curve in Figure \ref{fig:xinterp}.
+A line blend appears at approximately pixel 5300 causing the spline to twist to nearly vertical to account for both points.
+This leads to huge deviations from the correct wavelengths around this line blend as the extraneously high slope of the spline is accounted for.
 
@@
 \includegraphics[width=\textwidth]{Figures/intpx_tests.png}
-\caption{Residuals from different interpolation schemes over pixels.
+\caption{Residuals from different interpolation schemes over pixels in a single order.
 ThAr lines, shown as black circles, are used to construct a wavelength solution that is then evaluated at each LFC line, shown as black lines.
-Histograms of the residuals for each method is shown on the right.
+The residuals of each wavelength solution for a subset of the order is shown on the left.
+Histograms of the residuals for each method for the complete order is shown on the right.
+Here, the y-axis is zoomed in to show the shape of each histogram more clearly.
 Note: there is a blended ThAr line at approximately pixel 5300, the last ThAr line plotted.}
@@
 These huge digressions can be avoided by allowing for some smoothing in the interpolation.
-In Figure \ref{fig:xinterp}, we show an example in blue using SciPy's implementation of a univarate spline.
+In Figure \ref{fig:xinterp}, we show an example in green using SciPy's implementation of a univarate spline.
 While the result appears to follow the calibration lines much better, the smoothing ultimately causes larger residuals that are spatially correlated (Fig.
 \ref{fig:xinterp}, right).
-In all echelle orders, the edges are overestimated while the middle will be underestimated.
+In all echelle orders, the edges are overestimated while the middle will be underestimated, shown by the flattened shape of the histogram of residuals.
 The resultant wavelength solution is underestimating the curvature of the pixel-wavelength relation, giving rise to similar issues as with an inflexible, parametric wavelength solution.
@@
 This is a constraint we know is true for any one echelle order.
-In Figure \ref{fig:xinterp} we show that using the PCHIP interpolator returns the lowest residuals.
+Though the PCHIP interpolator shows a similar issue as a classic cubic spline around the ThAr line blend at pixel 5300, the effect is much smaller and effects fewer pixels.
+Figure \ref{fig:xinterp} shows that using the PCHIP interpolator returns the lowest residuals.
 
@@
 \section{Application to Other Spectrographs}
+\label{sec:others}
 We use \expres\ data here not because it is the optimal system for a method like \name, but because there is a strong need for good wavelength calibration in the EPRV community.
-The applicability of \name\ to any one instrument is a detailed question of the kind variance experienced by the spectrograph and calibration sources available, but we hope to provide some approximate benchmarks and diagnostics.
+The applicability of \name\ to any one instrument is a detailed question of the kind of variance experienced by the spectrograph and calibration sources available, but we hope to provide some approximate benchmarks and diagnostics.
 
+\lz{placeholder, pending Friday discussion}
 To determine the needed line density, we systematically removed LFC lines from the \expres\ data and calculated the RMS of the returned wavelengths.
 Figure \ref{fig:density} plots the increase in the scatter of the wavelengths returned by \name\ as a function of separation between lines.
-However, the stability of the lines and their measured line centers quickly becomes the dominant error source in calibration lamps over the separation between lines.
+Note, however, that the stability of the lines and their measured line centers quickly becomes the dominant error source in calibration lamps over the separation between lines.
 
@@
 Indeed, larger variations from less stabilized instruments might even make it easier for the PCA to identify these axes.
-Most physical systems should have a few dominant axis along which they vary, simply by what is possible with mechanical engineering.
+
+Most physical systems should have only a few dominant axis along which they vary, simply by what is possible from general mechanical engineering.
 With PCA, this can be tested by plotting the power of the basis vectors describing the calibration space, which should fall quickly after a few vectors.
-Note, however, that this is only a measure of the variance in the PCA space, not an explicit test of the end-to-end variation in the resulting model.
-This condition is therefore necessary, but not sufficient if implementing \name with PCA.
+This only provides a measure of the variance in the PCA space, however, not an explicit test of the end-to-end variation in the resulting model.
+This condition is therefore necessary but not sufficient if implementing \name with PCA.
 
@@
 Furthermore, once we have defined a calibration space that captures all possible degrees of freedom for a stabilized spectrograph, then all exposures taken with the spectrograph contains information about where the spectrograph is located within that calibration space.
-Theoretically, science exposures themselves could also be used to determine the calibration state of an instrument, negating the need for any additional wavelength-calibration source beyond defining the calibration space.
+Theoretically, tellurics, lines in science exposures, or just the trace positions themselves could also be used to determine the calibration state of an instrument, negating the need for any additional wavelength-calibration source beyond defining the calibration space.
 
@@
 
-\name\ is designed and optimized for extreme-precision instruments.
-
-We believe that by implementing \name, the uncertainty from wavelength calibration is rendered negligible in the face of other noise sources.
+\name\ is designed and optimized for extreme-precision, radial-velocity spectrographs.
+The errors returned by \name\ on \expres\ data renders the uncertainty from wavelength calibration negligible in the face of other noise sources, for example, stellar activity.
+Though the focus was on \eprv\ instruments here, \name\ should be largely applicable to nearly any astronomical spectrograph.
 

### 2020-05-15 paper/paper.tex
Line density, simultaneous reference, acknowledgements

@@
 \affil{Yale University, 52 Hillhouse, New Haven, CT 06511, USA}
+\affil{Flatiron Institute, Simons Foundation, 162 Fifth Avenue, New York, NY 10010, USA}
 
@@
 There is no coherent structure here and smaller residuals.
+
 This shows how the flexibility of an interpolated model can account for high-order detector defects, which emerged as structure in the residuals of the classic, smooth, polynomial-driven wavelength solution.
 In the \expres\ data shown here, for example, the interpolated model accounted for \lz{detector effect STA will explain to us}.
+This same flexibility should similarly allow interpolated wavelength solutions to account for position errors in pixel image blocks for different detectors.
+
+Position errors in pixel image blocks.
 
@@
 
-\subsection{Order of Interpolation}
-Why do we interpolate in time first and then in x?
-
-\lz{We talked about including this discussion, but there is no real reason for doing things in this order except it is a bit faster.}
-
 \section{Application to Other Spectrographs}
@@
 
-\lz{placeholder, pending Friday discussion}
-To determine the needed line density, we systematically removed LFC lines from the \expres\ data and calculated the RMS of the returned wavelengths.
-Figure \ref{fig:density} plots the increase in the scatter of the wavelengths returned by \name\ as a function of separation between lines.
+To determine the needed line density, we systematically removed LFC lines from the \expres\ data and calculated the RMS of the returned wavelengths for the removed lines.
+Figure \ref{fig:density} plots the scatter of residuals of the wavelengths returned by \name\ as a function of separation between lines.
+It gives an approximate estimation of the expected error between lines of different separation.
 Note, however, that the stability of the lines and their measured line centers quickly becomes the dominant error source in calibration lamps over the separation between lines.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.5\textwidth]{Figures/lfcDensityTest.pdf}
+\caption{Scatter in residuals as a function of spacing between lines.
+The residuals are the difference between the\name\ wavelengths and each line's theoretical wavelength in units of $\mps$.
+The average line spacing is calculated using the average distance between LFC lines across its wavelength range.}
+\label{fig:density}
+\end{figure}
 
@@
 
-Furthermore, once we have defined a calibration space that captures all possible degrees of freedom for a stabilized spectrograph, then all exposures taken with the spectrograph contains information about where the spectrograph is located within that calibration space.
+Incorporating calibration lines taken through a simultaneous fiber is trivial, especially if the simultaneous reference is also used while taking calibration images.
+The calibration lines through the simultaneous fiber could simply be concatenated to the array of calibration lines run through the dimensionality reduction and de-noising step.
+This not only allows the simultaneous calibration information to contribute to constructing the complete calibration space of the spectrograph, but also allows the simultaneous calibration information to uniquely pinpoint where the spectrograph is in that calibration space for any exposure, removing the need to interpolate the basis vector amplitudes.
+
+Once we have defined a calibration space that captures all possible degrees of freedom for a stabilized spectrograph, then all exposures taken with the spectrograph contains information about where the spectrograph is located within that calibration space.
 Theoretically, tellurics, lines in science exposures, or just the trace positions themselves could also be used to determine the calibration state of an instrument, negating the need for any additional wavelength-calibration source beyond defining the calibration space.
@@
 
+\acknowledgements{We thank the EXPRES team for their work on the instrument, software, and pipeline.
+We thank Dan Forman-Mackey for illuminating discussion.
+The data presented here made use of the Lowell Discovery Telescope at Lowell Observatory.
+Lowell is a private, non-profit institution dedicated to astrophysical research and public appreciation of astronomy and operates the LDT in partnership with Boston University, the University of Maryland, the University of Toledo, Northern Arizona University and Yale University.
+We gratefully acknowledge ongoing support for telescope time from Yale University, the Heising-Simons Foundation, and an anonymous donor in the Yale community.
+We especially thank the NSF for funding that allowed for precise wavelength calibration and software pipelines through NSF ATI-1509436 and NSF AST-1616086 and for the construction of EXPRES through MRI-1429365.
+LLZ gratefully acknowledges support from the National Science Foundation Graduate Research Fellowship under Grant No.
+DGE1122492.}
+
 \bibliography{paper.bib}

### 2020-05-21 paper/paper.tex
Updated Discussion

@@
 There are many ways to discern these basis functions.
-In this paper, we present a model using principal component analysis (PCA).
-A PCA is justifiable in the limit where there is very high signal-to-noise ratio, as is usually the case with typical calibration images.
+In this paper, we present a model using principal component analysis (PCA) \citep{wiki_pca}.
+A PCA is justifiable \lz{(Hogg)} in the limit where there is very high signal-to-nose ratio, as is usually the case with typical calibration images.
 There are many alternatives to PCA for this dimensionality reduction; we return to this point in Section \ref{sec:choices} below.
@@
 \lz{The diagonal-banding structure seen in components 3, 5, and 6 we believe to be capturing detector fabrication defects.
-(Maybe, awaiting STA response)}  Components 3 and 4 show aberrant behavior on the edges of the two bluest echelle orders, where lower signal results in more variation in the line fits.
+in \expres's detector or other optical elements}  Components 3 and 4 show aberrant behavior on the edges of the two bluest echelle orders, where lower signal results in more variation in the line fits.
 Later principal components become dominated by noise over coherent structure.
@@
 \section{Discussion} \label{sec:discussion}
-The endeavor to reach $0.1\,\mps$ precision necessitates improvements on all levels of spectroscopy, from the instrumentation to the data analysis.
-In this paper, we present \name\, a hierarchical, non-parametric model for wavelength calibration.
+We show that \name\ returns a lower RMS than classic, parametric methods by over a factor of two (Section \ref{sec:tests}).
+Using \name\ wavelengths also reduced the RMS in a planet fit to derived RVs by \lz{some percent hopefully probs} (Section \ref{sec:test-rv}).
+The endeavor to reach $0.1\,\mps$ radial-velocity precision necessitates improvements on all levels of spectroscopy, from the instrumentation to the data analysis.
+We believe implementing \name, a hierarchical, non-parametric model for wavelength calibration, is a step towards reducing the error contribution from data reduction pipelines.
 
 Starting with a list of calibration lines with known wavelengths and well-fit line centers for each calibration exposure, \name\ will de-noise and interpolate the given lines into a full wavelength solution.
-We show that \name\ returns a lower RMS than classic, parametric methods by over a factor of two.
-Using \name\ wavelengths also reduced the RMS in a planet fit to derived RVs by \lz{some percent hopefully probs}.
+\name\ levies the incredible stability of contemporary EPRV instruments and high density of lines made available by new calibration sources, such as LFCs and etalons, to achieve more accurate wavelengths.
+\name\ therefore assumes dense enough calibration lines to properly constrain a non-parametric wavelength model, and that the instrument has low degrees of freedom (see Section \ref{sec:others} for further discussion).
 
-\name\ levies the incredible stability of contemporary EPRV instruments and high density of lines made available by new calibration sources, such as LFCs and etalons, to achieve more accurate wavelengths.
-Denser calibration lines allow us to move to more flexible, interpolated wavelength solutions.
-Stabilized spectrograph hardware with few degrees of freedom allow us to use all calibration images in a given generation of stability to constrain the accessible calibration space of the spectrograph.
+Denser calibration lines allow us to move to more flexible wavelength models that can account for non-smooth features in the wavelength solution.
+Stabilized spectrograph hardware insures the calibration state of the instrument is low-dimensional.
+All calibration images in a given generation of stability can therefore be used to constrain the accessible calibration space of the spectrograph as well as where in that calibration space the spectrograph lies for each calibration exposures.
 We have described only one, fairly simplistic implementation of \name\ here.
@@
 This allowed us to quickly vet for problematic LFC exposures, which otherwise would have required visual inspection of all 1200+ LFC exposures.
-In a classic framework where each calibration exposure is treated independently, these aberrant exposures would likely have persisted undetected and could potentially sway the resultant wavelength solutions.
+In a classic framework where each calibration exposure is treated independently, these aberrant exposures would likely have persisted undetected and would likely have swayed the resultant wavelength solutions for at least an entire night.
 
-On the other hand, the ability of PCA to pick up on any and all variation, regardless of source, also makes \name\ very sensitive to uncertainties in line fitting.
+On the other hand, PCA captures all variance, regardless of source.
+Though \name\ only wants to capture variation in the instrument, the PCA is also sensitive to uncertainties and failures in the upstream line position fitting.
 For example, we have seen that lower-signal lines that are harder to fit faithfully will have greater variety in returned line positions, which is in turn captured by the PCA.
-High-fidelity line positions are essential to ensure the PCA is capturing variations in the spectrograph's calibration rather than changes in how well a line can be fit.
+In this sense, \name\ is actually a model of not just the instrument, but all the upstream choices used to drive live positions.
+High-fidelity line positions are essential to ensure the PCA is capturing variations in just the spectrograph's calibration rather than changes in how well a line can be fit or other effects.
+
+Along those lines, we caution that with any wavelength solution, there is a perfect degeneracy between what is defined as the ``position of the line'' and the resultant wavelength solution.
+If, for example, a cross correlation method is used to extract RVs from the data, a systematic difference may be introduced depending on the definition of what exactly is the line position, whether it be the center of a Gaussian fit, the edge of a complicated PSF, etc.
+In principle, the best way to mitigate this uncertainty would be to use a calibration source that looks like a star.
+
+Compared to traditional methods, which involve fitting high-order polynomials, \name\ has several useful statistical properties.
+\name\ is technically robust to localized issues with either the calibration source or the pipeline used to return line positions.
+With an interpolated wavelength model, one errant line position will only effect the resultant wavelength model out to the closest, neighboring lines.
+The effect of an outlier is diminished and kept local.
+In contrast, a parametric fit will change from a single errant line, changing the wavelength solution for the entire echelle order or the entire exposure.
+Through de-noising and outlier rejection, \name\ adds additional robustness against erroneous line positions.
+
+The locality of the interpolation caries other benefits as well.
+Manufacturing artifacts in the detector or other optical elements can lead to non-smooth structure in the wavelength solution that can not be captured by polynomials or other smooth functions (see \figurename~\ref{fig:resid2d}).
+An interpolated model introduces greater flexibility, enabling the model to account for such high-order effects.
+As discussed in Section \ref{sec:choice_wvp}, there are better and worse interpolators for the task, which may differ for different instruments and different calibration sources.
+Instead of using an interpolator at all, there might be better results from implementing something more sophisticated, such as a kernel method or a Gaussian process with a kernel adapted for the specifics of an instrument.
+There is in principle an enormous space of non-parametric methods to explore, which we leave outside the scope of this paper.
+
+Similarly, PCA is just one of many possible dimensionality-reduction methods.
+We chose to implement \name\ using PCA here for simplicity and computational tractability.
+If \name\ was updated to a full probabilistic model, the PCA along with the interpolation model would have to be upgraded to something with better probabilistic properties.
+Further discussion of other implementations of \name\ can be found in Section \ref{sec:choices}.
 
 \name\ can be applied to any data that contains information about the calibration state of the spectrograph.
-For example, though LFC and ThAr exposures are used as an example in this paper, \name\ would work exactly the same for an etalon or any other arc lamp with a list of lines and assigned wavelengths.
+For example, though LFC and ThAr exposures are used as an example in this paper, \name\ would work similarly for an etalon or any other arc lamp with a list of lines and assigned wavelengths.
+Modifications would have to be made with an etalon to include the free spectral range (FSR) as a free parameter that is fit for along with the low-dimensional basis.
 
-Incorporating calibration lines taken through a simultaneous fiber is trivial, especially if the simultaneous reference is also used while taking calibration images.
-The calibration lines through the simultaneous fiber could simply be concatenated to the array of calibration lines run through the dimensionality reduction and de-noising step.
-This not only allows the simultaneous calibration information to contribute to constructing the complete calibration space of the spectrograph, but also allows the simultaneous calibration information to uniquely pinpoint where the spectrograph is in that calibration space for any exposure, removing the need to interpolate the basis vector amplitudes.
+Incorporating calibration lines taken through a simultaneous fiber would also be fairly simple, especially if the simultaneous reference is also used while taking calibration images.
+The calibration lines through the simultaneous fiber could simply be concatenated to the array of calibration lines when the low-dimensional basis is constructed.
+This not only allows the simultaneous calibration information to contribute to constructing the complete calibration space of the spectrograph, but also allows the simultaneous calibration information to uniquely pinpoint where the spectrograph is in that calibration space for any exposure, removing the need to interpolate the basis vector amplitudes from calibration exposures to science exposures.
 
-Once we have defined a calibration space that captures all possible degrees of freedom for a stabilized spectrograph, then all exposures taken with the spectrograph contains information about where the spectrograph is located within that calibration space.
-Theoretically, tellurics, lines in science exposures, or just the trace positions themselves could also be used to determine the calibration state of an instrument, negating the need for any additional wavelength-calibration source beyond defining the calibration space.
-
-\lz{(Not sure where this belongs, honestly.)}  We caution that with any wavelength solution, there is a perfect degeneracy between what is defined as the ``center of the line'' and the resultant wavelength solution.
-If, for example, a cross correlation method is used to extract RVs from the data, a systematic difference may be introduced depending on the definition of the center of the line.
-In principle, the best way to mitigate this uncertainty would be to use a calibration source that looks like a star.
+Once we have defined a calibration space that captures all possible degrees of freedom for a stabilized spectrograph, there are many options for pinpointing where the spectrograph is located within that calibration space.
+Good housekeeping data, such as temperature or pressure could be used in addition to or instead of time (as mentioned in Section \ref{sec:choice_avt}).
+Telemetry that is seen to be correlated with the calibration state of the spectrograph can even be added to the data used to construct the low-dimensional basis.
+Furthermore, all exposures taken with the spectrograph in principle contains information about the calibration state of the spectrograph.
+Theoretically, tellurics, lines in science exposures, or just the trace positions themselves could also be used to determine an instrument's calibration state, negating the need for any additional wavelength-calibration source beyond defining the calibration space.
 
 \name\ is designed and optimized for extreme-precision, radial-velocity spectrographs.
-The errors returned by \name\ on \expres\ data renders the uncertainty from wavelength calibration negligible in the face of other noise sources, for example, stellar activity.
+In the battle to construct a high-fidelity data pipeline for extreme-precision radial-velocity measurements, we believe \name\ represents a step towards mitigating the error from wavelength calibration, as demonstrated by tests using \expres\ data (Section \ref{sec:tests}).
 Though the focus was on \eprv\ instruments here, \name\ should be largely applicable to nearly any astronomical spectrograph.

### 2020-05-21 paper/paper.tex
Megan Edits
including updated figures

@@
 \begin{abstract}\noindent%
-\name\ is a hierarchical, non-parametric framework for wavelength calibrating spectrographs.
-This method capitalizes on recent hardware improvements in highly-stabilized, extreme-precision radial-velocity spectrographs.
+\name\ is a hierarchical, non-parametric framework for wavelength-calibrating spectrographs.
+This method capitalizes on recent hardware improvements in \mb{highly-stabilized, extreme-precision radial-velocity spectrographs.
 With stabilized spectrographs, the full optical system and detectors can only vary within a tiny range of configurations.
 This allows us to use all calibration data to determine the entire space of possible calibrations for a stabilized spectrograph.
-A wavelength solution then need only pinpoint where in this low-dimensional calibration space the spectrograph exists.
-\name\ also takes advantage of the development of laser-frequency combs or etalons, which generate a dense set of stable calibration points.
+A wavelength solution then need only pinpoint where in this low-dimensional calibration space the spectrograph exists.}  \name\ also takes advantage of the development of laser-frequency combs or etalons, which generate a dense set of stable calibration points.
 This density of points permits the freedom of a non-parametric wavelength solution that can match or adapt to any instrument or detector oddities, more so than any parametric model.
@@
 
-The second simple idea follows from the observation that contemporary \eprv\ instruments are incredibly stable.
+\mb{The second simple idea follows from the observation that contemporary \eprv\ instruments are incredibly stable.
 Temperature-controlled, fiber-fed spectrographs vary only slightly over the night or season.
@@
 Instead, all the calibration data (or all the data) should be used to determine the space in which the instrument can and does vary.
-Subsequent calibration work then need only determine where in the accessible part of calibration space the spectrograph was located for each exposure.
+Subsequent calibration work then need only determine where in the accessible part of calibration space the spectrograph was located for each exposure.}
 
@@
 Irregardless, we expect \name\ to have applications for other kinds of spectrographs in other contexts.
-\name\ should be applicable to almost every astronomical spectrograph.
+\name\ should be applicable to almost every astronomical spectrograph, though the precision of the returned wavelengths will depend on the available calibration sources (more discussion in Section \ref{sec:others} below) .
 
@@
 because the fit failed due to noise, the line is not in its usual echelle order, etc.) can be assigned a \code{NaN} (hardware not-a-number) for that echelle order instead.
-Let there be $M$ lines per exposure.
-\name\ reads in a $N \times M$ matrix of line positions for each of $M$ lines for each of $N$ exposures.
+Let there be $P$ lines per exposure.
+\name\ reads in a $N \times P$ matrix of line positions for each of $P$ lines for each of $N$ exposures.
 
@@
 
-We found that a cubic-spline interpolation that enforces monotonicicity, such as a Piecewise Cubic Hermite Interpolatin Polynoimal (PCHIP) interpolator, works well for interpolating wavelengths onto pixels.
+We found that a cubic-spline interpolation that enforces monotonicity, such as a Piecewise Cubic Hermite Interpolating Polynomial (PCHIP) interpolator, works well for interpolating wavelengths onto pixels.
 A cubic spline allows for more flexibility than a continuous function, while the enforced monotonicity prevents large deviations that may befall a cubic spline.
@@
  Flat-relative, optimally extracted LFC data is background-corrected using a univariate spline.
-Each peak in an echelle order is then fit to a Gaussian.
+Each peak in an echelle order is then fit with a Gaussian.
 The mean of the fitted Gaussian is taken to be the center of the line.
@@
 
-To assess the classic, polynomial-driven method of wavelength calibration, we take each LFC exposure and separate the lines into even and odd lines.
+To assess the classic, polynomial-driven method of wavelength calibration, we take each LFC exposure and separate the lines into even- and odd-indexed lines.
 We then construct a wavelength solution using only the odd lines and use it to predict wavelengths for the even lines; i.e.
@@
 
-To test the denoising step of \name\ (\textsection \ref{sec:denoising}, \textsection\ref{sec:interp_time}), we employed \name\ on 90\% of all LFC exposures.
+To test the denoising step of \name\ (\textsection \ref{sec:denoising}, \textsection\ref{sec:interp_time}), we employed \name\ on a randomly selected 90\% of all LFC exposures.
 This means the basis vectors, $g_k(x,y,m)$,  and weights, $a_{nk}$, were constructed using only information from 90\% of all exposures.
@@
 \caption{Residuals for a single LFC exposure plotted with respect to detector position (as defined by echelle order and x-pixel) for various wavelength calibration methods.
-Each line is colored by the difference between the predicted wavelength and the theoretical wavelength for each line, given in units of $/mps$.
-Residuals to a classic, polynomial fit is shown in the top plot.
+Each line is colored by the difference between the predicted wavelength and the theoretical wavelength for each line, given in units of $\mps$.
+Residuals to a classic, polynomial fit are shown in the top plot.
 The center plot shows residuals to an interpolated wavelength solution.
 The bottom plot shows residuals when a ThAr wavelength model generated by \name\ is interpolated onto LFC line positions.
-The position of ThAr lines are over-plotted in black.}
+The positions of ThAr lines are over-plotted in black.}
 \label{fig:resid2d}
@@
 In the \expres\ data shown here, for example, the interpolated model accounted for \lz{detector effect STA will explain to us}.
-This same flexibility should similarly allow interpolated wavelength solutions to account for position errors in pixel image blocks for different detectors.
-
-Position errors in pixel image blocks.
+This same flexibility should similarly allow interpolated wavelength solutions to account for position errors in pixel image blocks for different detectors \citep{fischer2016, milakovic2020}.
 
@@
 This means the uncertainty from wavelength calibration could be completely marginalized out.
-Doing so, however, is only worth it if the wavelength calibration is a significant fraction of the error budget.
+Doing so, however, is only worthwhile if the wavelength calibration is a significant fraction of the error budget.
 
@@
 One would expect the returned wavelengths to get increasingly more accurate with larger K until components that represent only noise are incorporated.
-Residuals might then get worse before ultimately starting to get better again with large K, which marks te model over-fitting the data.
+Residuals might then get worse before ultimately starting to get better again with large K, which marks the model over-fitting the data.
 Though the returned RMS never gets worse, we find that the improvement plateaus between $K=6$ and $K=128$.
@@
 \includegraphics[width=0.42\textwidth]{Figures/pcAs_byDay.png}
-\caption{Amplitude of the first two principal components show as a function of time (left) or fraction of a day (right) \lz{(Yeah, it's kind of a mess)}.
-The top plots shows the amplitudes for the first principal component while the bottom plot shows the amplitudes for the second principal component.
+\caption{Amplitude of the first two principal components shown as a function of time (left) or fraction of a day (right).
+The top row of plots shows the amplitudes for the first principal component while the bottom row shows the amplitudes for the second principal component.
 Points mark the times at which we have calibration data.
@@
 
-A simple linear interpolation would obviously give erroneously low values everywhere.
+A simple linear interpolation would give erroneously low values everywhere.
 Due to the dispersion intrinsic to echelle spectrographs, the wavelength change between pixels grows greater with greater wavelengths.
@@
 
+It is also possible to apply \name\ to etalon data with some modifications.
+If the free spectral range (FSR) and therefor wavelength of each line is well known, then the etalon lines can be treated just like LFC lines with known wavelengths and orders.
+Alternatively, the FSR can be included as a free parameter that is included when constructing the low-dimensional basis.
+\lz{Hogg, I think more detail is needed here that I don't understand but would like to; can we discuss?}
+
 In terms of stability, \name\ should be adaptable to a wide range of instrument designs.
-In truth, the amplitude of the variation matters little as long as all axes of variation can be captured by PCA or another dimensionality-reducing model.
+\mb{In truth, the amplitude of the variation matters little as long as all axes of variation can be captured by PCA or another dimensionality-reducing model.
 Indeed, larger variations from less stabilized instruments might even make it easier for the PCA to identify these axes.
+(Is this true?)}
 
@@
 This only provides a measure of the variance in the PCA space, however, not an explicit test of the end-to-end variation in the resulting model.
-This condition is therefore necessary but not sufficient if implementing \name with PCA.
+This condition is therefore necessary but not sufficient if implementing \name\ with PCA.
 
@@
 For the results presented here, the data was broken up into different eras of stability based on where the principal component amplitudes showed huge deviations.
-This was done visually, though there exists many change-point detection algorithms that could be used.
+This was done visually, though there exists many change-point detection algorithms that could be used \citep{aminikhanghahi2017}.
 There is a trade-off in including more exposures between introducing greater variation, but also more data to offer constraints that may be optimized.
@@
 Homogeneity is, of course, implicitly required when implementing PCA.
+
+Simultaneous calibration data is easily incorporated, especially if the simultaneous reference is also used when taking calibration images.
+The position of calibration lines taken through the simultaneous fiber can be concatenated to the array of line positions when the low-dimensional basis is constructed.
+This not only allows the simultaneous calibration information to contribute to constructing the complete calibration space of the spectrograph, but also allows the simultaneous calibration information to uniquely pinpoint where the spectrograph is in that calibration space for any exposure, thereby removing the need to interpolate the basis vector amplitudes from calibration exposures to science exposures.
 
@@
 
+\lz{Hogg: I'm still not sure I completely grasp the issue, can this be made clearer?}
 Along those lines, we caution that with any wavelength solution, there is a perfect degeneracy between what is defined as the ``position of the line'' and the resultant wavelength solution.
@@
 
-\name\ can be applied to any data that contains information about the calibration state of the spectrograph.
+\name\ can be applied to any data that contains information about the calibration state of the spectrograph (see Section \ref{sec:others}).
 For example, though LFC and ThAr exposures are used as an example in this paper, \name\ would work similarly for an etalon or any other arc lamp with a list of lines and assigned wavelengths.
-Modifications would have to be made with an etalon to include the free spectral range (FSR) as a free parameter that is fit for along with the low-dimensional basis.
-
-Incorporating calibration lines taken through a simultaneous fiber would also be fairly simple, especially if the simultaneous reference is also used while taking calibration images.
-The calibration lines through the simultaneous fiber could simply be concatenated to the array of calibration lines when the low-dimensional basis is constructed.
-This not only allows the simultaneous calibration information to contribute to constructing the complete calibration space of the spectrograph, but also allows the simultaneous calibration information to uniquely pinpoint where the spectrograph is in that calibration space for any exposure, removing the need to interpolate the basis vector amplitudes from calibration exposures to science exposures.
+Simultaneous calibration information can easily be accounted for by simply including the line position information from the simultaneous reference when constructing a low-dimensional basis of the instrument's calibration space.
 
@@
 \acknowledgements{We thank the EXPRES team for their work on the instrument, software, and pipeline.
-We thank Dan Forman-Mackey for illuminating discussion.
+We thank Dan Foreman-Mackey for illuminating discussion.
 The data presented here made use of the Lowell Discovery Telescope at Lowell Observatory.

### 2020-05-25 paper/paper.tex
New Figures and Captions

@@
 We enlarge variations in measured line position, changes in calibration space, and interpolation deviations for clarity.
-In step one, dimensionality reduction and denoising, the complete set of line positions for all exposures is analyzed to return a set of basis vectors, $g_n$, which represent different ways the spectrograph calibration changes, and spans the possible calibration space of the spectrograph.
-In step two, \name\ interpolates the amplitudes of each basis vector, $a_n,k$, to return the calibration state for a specific exposure, returned as a set of de-noised calibration lines.
-The assigned wavelengths of these de-noised line positions are then interpolated onto other pixels in step three to construct a full wavelength model.}
+In step one, dimensionality reduction and denoising (\textsection \ref{sec:denoising}), the complete set of line positions for all exposures is analyzed to return a set of basis vectors, $g_n$, which represent different ways the spectrograph calibration changes.
+These basis vectors span the possible calibration space of the spectrograph.
+In step two (\textsection \ref{sec:interp_time}), \name\ interpolates the amplitudes of each basis vector, $a_n,k$, to return the calibration state for a specific science exposure, returned as a set of de-noised calibration lines.
+The assigned wavelengths of these de-noised line positions are then interpolated onto other pixels in step three (\textsection \ref{sec:interp_wsol}) to construct a full wavelength model that returns wavelength as a function of detector position $x$ and echelle order $m$.}
 \label{fig:testHists}
@@
 \caption{Difference in predicted and theoretical wavelengths for the wavelength calibration tests described in Section \ref{sec:tests}.
-The standard deviation of the residuals for each test is given in the top-right corner in each method's corresponding color.}
+The standard deviation of the per line residuals for each test, or the per line RMS as defined in Equation \ref{eq:rms} is given in the top-right corner in each method's corresponding color.
+\name\ returns the smallest spread in residuals.}
 \label{fig:testHists}
@@
 %\includegraphics[width=0.5\textwidth]{Figures/lineResids2D_col.pdf}
-\caption{Residuals for a single LFC exposure plotted with respect to detector position (as defined by echelle order and x-pixel) for various wavelength calibration methods.
+\caption{Residuals for a single LFC exposure plotted with respect to detector position (as defined by echelle order and x-pixel) for parametric and non-parametric wavelength calibration methods.
 Each line is colored by the difference between the predicted wavelength and the theoretical wavelength for each line, given in units of $\mps$.
 Residuals to a classic, polynomial fit are shown in the top plot.
-The center plot shows residuals to an interpolated wavelength solution.
-The bottom plot shows residuals when a ThAr wavelength model generated by \name\ is interpolated onto LFC line positions.
-The positions of ThAr lines are over-plotted in black.}
+The bottom plot shows residuals to an interpolated wavelength solution.
+High-order structure, i.e.
+vertical stripes and patchiness, is apparent in the residuals to a polynomial wavelength solution, which assumes smoothness.
+The increased flexibility of the interpolated wavelength solution can account for this high-order structure, resulting in much smaller residuals with less spatial coherence.}
 \label{fig:resid2d}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{Figures/lineDensity.png}
-\caption{Comparison of polynomial-fit and interpolated wavelength solutions using either just ThAr lines or LFC lines.
-For a subset of one order, the difference between the assigned and model wavelengths for each LFC line is plotted with respect to pixel.
-The large residuals between two widely-spaced ThAr lines when implementing an interpolated wavelength solution with just ThAr lines (blue curve) extends out of frame.}
+\includegraphics[width=\textwidth]{Figures/lineDensity.png}
+\caption{Residuals when using ThAr lines to predict wavelengths for LFC lines.
+\textbf{Right}: residuals for a single LFC exposure plotted with respect to detector position as in Figure \ref{fig:resid2d}.
+Each line is colored by the difference between the predicted wavelength and the theoretical wavelength for each line, given in units of $\mps$.
+The positions of ThAr lines are over-plotted in yellow.
+In general, residuals are greater between ThAr lines with greater separation.
+\textbf{Left}: Comparison of polynomial-fit and interpolated wavelength solutions using either just ThAr lines or LFC lines.
+For a subset of echelle order 94, the difference in Angstroms between the assigned and model wavelengths for each LFC line is plotted with respect to pixel.
+The large residuals between two widely-spaced ThAr lines when implementing an interpolated wavelength solution with just ThAr lines (orange curve) extends out of frame.
+Note how the shape of the residuals from a polynomial fit are similar whether using ThAr lines or LFC lines, suggesting that little is added with the denser LFC lines.
+The LFC lines only serve to make the residuals more centered around zero.
+A PCHIP interpolated wavelength model guided by LFC lines returns the smallest residuals.}
 \label{fig:waveResids}
@@
 %\includegraphics[width=\textwidth]{Figures/pcsLfc6.pdf}
-\caption{The first nine principal components constructed using LFC lines.
-Each line is plotted according to its echelle order and x-pixel and colored by the value of the principal component for that line.}
+\caption{\textbf{Top}:  The fiducial calibration of the spectrograph, i.e.
+the mean line positions for each line throughout the epoch of stability.
+The following plots show the first nine principal components constructed using LFC lines.
+These principal components represent the basis vectors along which the calibration of the spectrograph can deviate from the fiducial calibration.
+For each basis vector, each calibration line is plotted according to its echelle order and x-pixel and colored by the value of the basis vector for that line.
+Principal components beyond the sixth one become steadily more dominated by noise.}
 \label{fig:pcLfc}
@@
 \includegraphics[width=.5\textwidth]{Figures/kvals_all.pdf}
-\caption{Spread in residuals of returned wavelengths for different values of $K$.  There is a dotted, vertical line at $K=6$, our chosen $K$ value.  The dotted, horizontal line is the RMS for $K=6$.}
+\caption{Per line RMS of returned wavelength models for different values of $K$.
+Per line RMS is defined in Equation \ref{eg:rms}.
+There is a dotted, vertical line at $K=6$, our chosen $K$ value.
+The dotted, horizontal line is the RMS for $K=6$.
+The improvement at around $K=6$ plateaus.}
 \label{fig:kvals}
@@
 \centering
-\includegraphics[width=0.56\textwidth]{Figures/pcA_lfc.png}
-\includegraphics[width=0.42\textwidth]{Figures/pcAs_byDay.png}
-\caption{Amplitude of the first two principal components shown as a function of time (left) or fraction of a day (right).
+\includegraphics[width=\textwidth]{Figures/pcAs_byDay.png}
+\caption{Amplitude of the first two principal components shown as a function of time (left) or hour from midnight (right).
 The top row of plots shows the amplitudes for the first principal component while the bottom row shows the amplitudes for the second principal component.
-Points mark the times at which we have calibration data.
+Points mark the times at which calibration exposures were taken.
 Lines show the result of a linear interpolation.
-In the right plot, the principal component amplitudes have been artificially offset by the median amplitude for each day so that all days are roughly on the same scale.
-Here, points are colored by the MJD of each exposure.}
+In the right plot, the principal component amplitudes have been artificially offset by the median amplitude for each night.
+All days are therefore roughly on the same scale, but the y-axis is different from the left plots.
+In the right column, points are colored by the time of each exposure.}
 \label{fig:nightlyVariation}
@@
 \includegraphics[width=\textwidth]{Figures/intpx_tests.png}
-\caption{Residuals from different interpolation schemes over pixels in a single order.
-ThAr lines, shown as black circles, are used to construct a wavelength solution that is then evaluated at each LFC line, shown as black lines.
+\caption{Residuals from different interpolation schemes over pixels in echelle order 100.
+ThAr lines, shown as blue circles, are used to construct a wavelength solution that is then evaluated at each LFC line, shown as blue lines.
 The residuals of each wavelength solution for a subset of the order is shown on the left.
@@
 \includegraphics[width=0.5\textwidth]{Figures/lfcDensityTest.pdf}
-\caption{Scatter in residuals as a function of spacing between lines.
-The residuals are the difference between the\name\ wavelengths and each line's theoretical wavelength in units of $\mps$.
+\caption{Per line RMS as a function of spacing between lines in units of resolution element.
 The average line spacing is calculated using the average distance between LFC lines across its wavelength range.}

### 2020-05-26 paper/paper.tex
New Figures, Significant Re-Phrasing

@@
 \newcommand{\code}[1]{\texttt{#1}}
+\newcommand{\target}{HD $\bigstar$}
 
@@
 \name\ is a hierarchical, non-parametric framework for wavelength-calibrating spectrographs.
-This method capitalizes on recent hardware improvements in \mb{highly-stabilized, extreme-precision radial-velocity spectrographs.
-With stabilized spectrographs, the full optical system and detectors can only vary within a tiny range of configurations.
-This allows us to use all calibration data to determine the entire space of possible calibrations for a stabilized spectrograph.
-A wavelength solution then need only pinpoint where in this low-dimensional calibration space the spectrograph exists.}  \name\ also takes advantage of the development of laser-frequency combs or etalons, which generate a dense set of stable calibration points.
-This density of points permits the freedom of a non-parametric wavelength solution that can match or adapt to any instrument or detector oddities, more so than any parametric model.
+This method is designed with the needs of extreme-precision radial velocity (\eprv) in mind, which require that instruments be calibrated or stabilized to better than $10^{-4}$ of a pixel.
+Instruments are likely to experience only a few degrees of freedom, especially \eprv\ instruments with highly stabilized optical systems and detectors.
+\name\ takes advantage of this property by using all calibration data to construct a low-dimensional space of all possible calibrations for an instrument.
+A wavelength model need only pinpoint where in this low-dimensional calibration space the spectrograph exists.
+\name\ also takes advantage of the development of laser-frequency combs or etalons, which generate a dense set of stable calibration points.
+This density of calibration lines permits the freedom of a non-parametric wavelength solution that can match or adapt to any instrument or detector oddities, more so than any parametric model.
 We demonstrate the success of this method with data from the EXtreme PREcision Spectrograph, \expres, which includes a Menlo Systems laser frequency comb.
@@
 
-This recommends a calibration methodology that is \emph{non-parametric}:  If every point in the spectrograph detector is surrounded by nearby calibration lines, the wavelength solution can, for example, be made simply as an interpolation of the calibration data.
+This recommends a calibration methodology that is \emph{non-parametric}:  If every point in the spectrograph detector is sufficiently surrounded by nearby calibration lines, the wavelength solution can, for example, be made simply as an interpolation of the calibration data.
 The density of lines removes the need to enforce any functional form for the wavelength solution (such as a continuous ninth-order polynomial, for example).
@@
 
-\mb{The second simple idea follows from the observation that contemporary \eprv\ instruments are incredibly stable.
-Temperature-controlled, fiber-fed spectrographs vary only slightly over the night or season.
-This suggests that such stabilized spectrographs vary along only a small number of axes in what you might call ``calibration space'', or the (very high-dimensional) space of all possible wavelength solutions.
-That is, not only are the spectrographs stable, but they also have few environmentally accessible degrees of freedom.
+The second simple idea follows from the observation that most physical systems have only a few dominant degrees of freedom, meaning most spectrographs vary along only a small number of axes in ``calibration space'', or the (very high-dimensional) space of all possible wavelength solutions.
+This is particularly likely with contemporary \eprv\ instruments, which have been stabilized in many directions.
+That is, spectrographs, especially stabilized ones, should have few environmentally accessible degrees of freedom.
 This renders it inadvisable to fit each calibration exposure or calibrate each science exposure independently.
 Instead, all the calibration data (or all the data) should be used to determine the space in which the instrument can and does vary.
-Subsequent calibration work then need only determine where in the accessible part of calibration space the spectrograph was located for each exposure.}
+Subsequent calibration work then need only determine where in the accessible part of calibration space the spectrograph was located for each exposure.
 
+\lz{Hogg: I think it's stuff like this you wanted to audit for consistency with statistical literature?}
 In the context of probabilistic models, this structure is \emph{hierarchical}:  The calibration data are used not just to determine the wavelength solution at one moment, but also to determine the possible \emph{space} of wavelength solutions at all moments.
@@
 By being hierarchical, it restricts that freedom tremendously, but it does so appropriately for the empirically determined variations in the stabilized spectrograph.
+
 \name\ is designed for temperature-controlled, fiber-fed spectrographs with good calibration sources, such as laser-frequency combs, or etalons.
-We have in mind \eprv\ instruments and \eprv\ science cases primarily because the need for good wavelength calibration in this field is so great.
+We have in mind \eprv\ instruments and \eprv\ science cases, primarily because the need for good wavelength calibration in this field is so great.
 Irregardless, we expect \name\ to have applications for other kinds of spectrographs in other contexts.
@@
 The \name\ method is designed to take a series of calibration lines with known wavelengths and well-fit detector positions, and de-noise and interpolate them into a full wavelength model applicable to all exposures taken with the instrument.
-\name\ operates on the two core ideas that the wavelength solution should be given enormous flexibility, but that it lives in a very low-dimensional space, where the degrees of freedom are set by the limited kinematics of the stabilized spectrograph hardware.
+\name\ operates on the two core ideas that the wavelength solution should be given enormous flexibility, but that it lives in a very low-dimensional space, where the degrees of freedom are set by the limited kinematics of the spectrograph hardware.
+\name\ therefore assumes that the space of possible calibration states for an instrument is low-dimensional.
 
-\name\ assumes dense enough calibration line coverage with well-fit line centers to provide some constraint on an interpolated wavelength solution across an echelle order.
-Additionally, \name\ assumes the instrument hardware is stabilized enough that the space of possible calibration states is low-dimensional.
+\name\ also assumes dense enough calibration line coverage with well-fit line centers to provide sufficient constraint on an interpolated wavelength solution across an echelle order.
+Upstream errors in line center positions may propagate through \name\ wavelength models.
+The needed line density is dependent on the required precision of the returned wavelength model; larger spacing between lines offer less constraint and are likely to return worse wavelengths.
 We revisit and quantify these conditions in Section \ref{sec:others}.
@@
 \end{equation}
-Here the y-dependence is dropped as that dependence is carried by spectral order $m$.
+Here, the y-dependence is dropped as that dependence is carried by spectral order $m$.
 The line position information is therefore uniquely identified by echelle order $m$ and pixel in the dispersion direction, $x$.
-The coefficients $c_{nij}$ are interpolated to time $t_n$ of exposure $n$ by a third-order polynomial of time.
-This third-order polynomial is evaluated at the time of non-calibration exposures to re-construct a 2D, 9th-order polynomial wavelength solution for that exposure.
+The coefficients $c_{nij}$ are interpolated from the time of calibration exposures to time $t_n$ of a science exposure $n$ by a third-order polynomial of time.
+This third-order polynomial is evaluated at the time of non-calibration, science exposures to re-construct a 2D, 9th-order polynomial wavelength solution for that exposure.
 Each calibration image obtains its $c_{nij}$ coefficients independently.
 
-Given a stabilized instrument, however, the spectrograph should experience only low-degree variability, meaning the calibration of any image can and should be informed by the calibration of every other image.
-The calibration data themselves can be used to develop a low-dimensional basis for expressing the space of possible calibrations for a stabilized spectrograph.
+Given a stabilized instrument with low degrees of freedom, however,  the calibration of any image can and should be informed by the calibration of every other image.
+The calibration data themselves can be used to develop a low-dimensional basis for expressing the space of all possible calibrations for a spectrograph with few degrees of freedom.
 
@@
 and the $g_k(\lambda,m)$ are basis functions expressing the ``directions'' in calibration space that the spectrograph can depart from the fiducial calibration.
-When this calibration structure is used to deliver a wavelength solution, we will have to invert $x(\lambda,m,n)$ into $\lambda(x,m,n)$.
+The resultant $x(\lambda,m,n)$, a list of calibration line positions for a given exposure, can be regarded as the calibration state of the spectrograph for that exposure.
+When this calibration structure is used to deliver a wavelength solution, $x(\lambda,m,n)$ will be inverted back into $\lambda(x,m,n)$ to recover wavelengths for each detector position $x$ and echelle order $m$..
 
@@
 We enlarge variations in measured line position, changes in calibration space, and interpolation deviations for clarity.
-In step one, dimensionality reduction and denoising (\textsection \ref{sec:denoising}), the complete set of line positions for all exposures is analyzed to return a set of basis vectors, $g_n$, which represent different ways the spectrograph calibration changes.
-These basis vectors span the possible calibration space of the spectrograph.
+In step one, dimensionality reduction and denoising (\textsection \ref{sec:denoising}), the complete set of line positions for all exposures is analyzed to return a set of $K$ basis vectors, $g_n$, which represent different ways the spectrograph calibration changes.
+These basis vectors span the $K$-dimensional calibration space of the spectrograph, which includes all possible wavelength solutions.
 In step two (\textsection \ref{sec:interp_time}), \name\ interpolates the amplitudes of each basis vector, $a_n,k$, to return the calibration state for a specific science exposure, returned as a set of de-noised calibration lines.
 The assigned wavelengths of these de-noised line positions are then interpolated onto other pixels in step three (\textsection \ref{sec:interp_wsol}) to construct a full wavelength model that returns wavelength as a function of detector position $x$ and echelle order $m$.}
-\label{fig:testHists}
+\label{fig:cartoon}
 \end{figure*}
@@
 \subsection{Dimensionality Reduction: De-Noising of Calibration Frames} \label{sec:denoising}
-Within a given range of stability, \name\ will use calibration images to  1) determine the space in which the instrument varies and 2) where in the accessible calibration space the spectrograph existed for each exposure.
+\name\ will use calibration images to  1) determine the complete calibration space in which an instrument varies and 2) where in the accessible calibration space the spectrograph existed for each exposure.
 For each calibration exposure, $n$, \name\ requires a full list of lines, $(\lambda,m)$ that are expected in each calibration exposure.
@@
 There are many strategies for identifying calibration line positions and matching them to their assigned wavelengths; this problem is left out of scope for this work.
+However, we warn that systematic errors or large uncertainties in fitting line positions can propagate through to the wavelength models returned by \name.
 
@@
 The mean of measured line position over the set of calibration exposures represents the fiducial, or standard calibration of the spectrograph, $g_0(\lambda,m)$.
-In this implementation of \name\, principal component analysis is performed on the difference between this fiducial calibration and each individual line position.
+In this implementation of \name, principal component analysis is performed on the difference between this fiducial calibration and each individual line position.
 The returned principal components serve as basis functions,  $g_k(\lambda,m)$, expressing the possible deviations of the spectrograph from this fiducial calibration.
@@
 This process can be repeated on line centers deemed as outliers by some metric, to account for lines that may have been mis-identified or mis-fit.
-The principal components from the final iteration are used to define the spectrograph's calibration space while the associated magnitudes pinpoint where in that calibration space the spectrograph is located for each calibration exposure.
+The principal components from the final iteration are used to define the spectrograph's calibration space, while the associated amplitudes for each component pinpoint where in that calibration space the spectrograph is located for each calibration exposure.
 
@@
 \KwData{line positions $x(\lambda,m,n)$ for each exposure $n$, with wavelengths $\lambda$ and echelle orders $m$}
-\KwResult{Principle vectors of low-dimensional calibration space $g_k(\lambda,m)$, location of exposures in this space $a_{n,k}$}
+\KwResult{Basis vectors of the low-dimensional calibration space $g_k(\lambda,m)$, location of exposures in this space $a_{n,k}$}
 \While{change in missing or outlier line centers $>$ 0.01\%}
@@
 	let $a_{n,k} = U\cdot \Sigma$ and $g_k(\lambda,m) = V$\;
-	$x(\lambda,m,n) = g_0(\lambda,m) + \sum_{k=1}^K a_{nk}\,g_k(\lambda,m)$ where $K=6$ for $x(\lambda,m,n)$ = \code{NaN}
+	$x(\lambda,m,n) = g_0(\lambda,m) + \sum_{k=1}^K a_{nk}\,g_k(\lambda,m)$ for $x(\lambda,m,n)$ = \code{NaN} where $K$ is a a small integer
 	}
@@
  \label{sec:interp_time}
-The magnitude of each principal component can be interpolated with respect to housekeeping data, such as time, temperature, telescope position, etc.
-to determine different calibration states of the spectrograph.
-The choice of what to interpolate against depends on the dominant contribution to variation in the instrument.
+\name\ then interpolates the amplitude, $a_{n,k}$, of each principal component to determine the calibration state of the spectrograph.
+For example, the amplitude can be interpolated with respect to time to recreate the calibration state of the spectrograph at different times.
+The choice of what to interpolate against depends on the dominant contribution to variation in the calibration of the instrument.
 
-In the implementation of \name\ presented here, the magnitudes of the principal components are interpolated linearly with respect to time.
-\name\ finds new magnitudes $a_{n'k}$ for an exposure $n'$ at time $t_{n'}$, which are used to construct the calibration state of the spectrograph for that exposure.
+In the implementation of \name\ presented here, the amplitudes of the principal components are interpolated linearly with respect to time.
+\name\ finds new magnitudes $a_{n'k}$ for a science exposure $n'$ at time $t_{n'}$, which are used to construct the calibration state of the spectrograph for that point in time.
 Using interpolated magnitudes, $a_{n,k}$, and the basis vectors, $g_k(\lambda,m)$, returned by the de-noising process, \name\ can construct a set of calibration lines, $x(\lambda,m,n') $, for any exposure as prescribed in Equation \ref{eq:excl_wsol}.
@@
 From the de-noising step, \name\ can now construct a set of calibration lines, $x(\lambda,m,n')$ for any exposure $n'$.
-To construct a wavelength solution, \name\ then interpolates the known wavelengths over the exposure specific line centers.
+To construct a wavelength solution, \name\ interpolates the known wavelengths over the exposure specific line centers.
 For instance, interpolating the known wavelengths vs.
@@
 A cubic spline allows for more flexibility than a continuous function, while the enforced monotonicity prevents large deviations that may befall a cubic spline.
-Choice in interpolation scheme and other tests are further discussed in Section \ref{sec:choice_wvp}.
+Choice in interpolation scheme, $K$, and other tests are further discussed in Section \ref{sec:choice_wvp}.
 
@@
 \section{Data} \label{sec:data}
-We tested \name\ results using data from \expres\, the EXtreme PRecison Spectrograph.
+We tested \name\ using data from \expres, the EXtreme PRecison Spectrograph.
 \expres\ is an environmentally-stabilized, fiber-fed, $R\sim137,000$, optical spectrograph \citep{jurgenson2016, blackman2020}.
@@
 ThAr exposures are taken at the beginning and end of each night.
-All calibration data is taken through the science fiber, so that they down the same optical pathway and are projected onto the same pixels as all science observations.
+All calibration data is taken through the science fiber, so that calibration light travels down the same optical pathway and is projected onto the same pixels as all science observations.
 
-LFC lines cover 40 echelle orders, which contain 19203 lines.
+LFC lines cover echelle orders 84-124, which contain 19203 calibration lines.
 Though the results are primarily based on work with LFC data, there is some discussion of applications to arc lamps.
-ThAr lines cover all 86 extracted orders of \expres, which includes 5295 lines.
+ThAr lines cover all 86 extracted orders of \expres\ (echelle orders 75-160), which include 5295 lines.
 For both the LFC and ThAr data, lines that appeared in less than 60\% of exposures were not included in the analysis.
@@
 
- \name\ requests as input a list of echelle orders, line wavelengths, and pixel positions of each line for each exposure.
+ \name\ inputs a list of echelle orders $m$, line wavelengths $\lambda$, and pixel positions $x$ of each line for each exposure.
 We, therefore, use line positions generated by the pre-existing \expres\ pipeline from the following procedure \citep{petersburg2020}.
@@
 Lines are matched manually when there is a significant change in the instrument, but is otherwise done using forward-modeling.
-Once each line's position is identified and matched to a wavelength from the line atlas, a sixth-order, 2D polynomial is fit over pixel location $x$, echelle order $m$, and scaled wavelength $m\lambda$ (in order to distinguish lines that may appear in more than one order).
+Once each line's position is identified and matched to a wavelength from the line atlas, a sixth-order, 2D polynomial is fit over pixel location $x$, echelle order $m$, and scaled wavelength $m\lambda$ (wavelengths are scaled in order to distinguish lines that may appear in more than one order).
 
@@
 \expres\ exposures are separated into different instrumental epochs that correspond to atypical hardware changes---changes that altered the position of the echellogram on the detector, the shape of the instrumental PSF, or the calibration sources.
-Significant instrumental changes break the assumption of a stabilized instrument with only low-dimensional variations.
+Significant instrumental changes break the assumption that an  instrument will experience only low-dimensional variations, i.e.
+that the calibration space of all possible wavelength solutions is low-dimensional.
 Each instrument epoch is therefore treated independently.
-In the results presented here, we use 1227 LFC exposures and 78 ThAr exposures taken between October 14 and December 18, 2020 on 29 unique nights.
+In the results presented here, we use 1227 LFC exposures and 78 ThAr exposures taken between October 14 and December 18, 2019 on 29 unique nights.
 
@@
 To assess the classic, polynomial-driven method of wavelength calibration, we take each LFC exposure and separate the lines into even- and odd-indexed lines.
-We then construct a wavelength solution using only the odd lines and use it to predict wavelengths for the even lines; i.e.
-a polynomial is fit to just the odd lines and then evaluated at the detector position of the even lines (see Equation \ref{eq:poly_wsol}).
-We then generate a wavelength solution using only the even lines and use it to predict wavelengths for the odd lines.
+We then construct a wavelength solution using only the odd-indexed lines and use it to predict wavelengths for the even-indexed lines; i.e.
+a polynomial is fit to just the odd-indexed lines and then evaluated at the detector position of the even-indexed lines (see Equation \ref{eq:poly_wsol}).
+We then generate a wavelength solution using only the even-indexed lines and use it to predict wavelengths for the odd-indexed lines.
 
-To test the interpolation step of \name\ (\textsection \ref{sec:interp_wsol}), we employed \name\ on all LFC exposures with odd lines removed.
-The resultant basis vectors, $g_k(x,y,m)$,  and weights, $a_{nk}$, are therefore only informed by the even lines of each LFC exposure.
-We then predict wavelengths for the odd lines that had been excluded and compare these predictions to their assigned wavelengths.
+To test the interpolation step of \name\ (\textsection \ref{sec:interp_wsol}), we employed \name\ on all LFC exposures with odd-indexed lines removed.
+The resultant basis vectors, $g_k(x,y,m)$,  and amplitudes, $a_{nk}$, are therefore only informed by the even-indexed lines of each LFC exposure.
+We then predict wavelengths for the odd-indexed lines that had been excluded and compare these predictions to their assigned wavelengths.
 This allows us to test how accurately an interpolated wavelength solution can predict wavelengths.
@@
 
-Histograms of the residuals for each of the above described polynomial, interpolation, and denoising tests (respectively) are shown in Figure \ref{fig:testHists}.
+Resultant wavelengths are compared on a line-by-line basis.
+The residuals of a wavelength solution represent the difference between the wavelength solution evaluated at the line position of a calibration line, and the assigned theoretical wavelength (i.e.
+from Equation \ref{eq:lfc} for LFC lines and from a line atlas for ThAr lines) for each line in every exposure.
+The reported RMS of a wavelength solution is therefore the per-line RMS, i.e.
+\begin{equation}
+RMS/line \: [m\,s^{-1}] = \sqrt{\sum_{n=1}^N\sum_{p=1}^P\frac{[ \frac{(\lambda_{n,p,predicted} - \lambda_{p,theoretical})}{\lambda_{p,theoretical}} \times c ]^2}{N \times P}}
+\label{eq:rms}
+\end{equation}
+where $\lambda_{p,theoretical}$ is the theoretical wavelength for line $p$, $\lambda_{n,p,predicted}$ is the wavelength predicted by the constructed wavelength solution for line $p$ in exposure $n$, and residuals from all $P$ lines from all $N$ exposures are used, for a total of $N \times P$ lines.
+The difference in wavelength is converted to units of $\mps$, a more intuitive metric for \eprv\ work.
+
+Histograms of the per-line residuals for each of the above described polynomial, interpolation, and denoising tests (respectively) are shown in Figure \ref{fig:testHists}.
 Note that the spread in residuals is much smaller for both the denoising and interpolation tests relative to the results of the polynomial wavelength solution.
@@
 \caption{Difference in predicted and theoretical wavelengths for the wavelength calibration tests described in Section \ref{sec:tests}.
-The standard deviation of the per line residuals for each test, or the per line RMS as defined in Equation \ref{eq:rms} is given in the top-right corner in each method's corresponding color.
+The per line RMS as defined in Equation \ref{eq:rms} is given in the top-right corner in each method's corresponding color.
 \name\ returns the smallest spread in residuals.}
@@
 \name -generated wavelengths also exhibit less structure in the returned residuals.
-For a randomly selected example LFC exposure, Figure \ref{fig:resid2d} plots each line with respect to its echelle order (y-axis) and x-pixel on the detector (x-axis) colored by the difference between the predicted and assigned wavelength for that line in units of $\mps$.
+For a randomly selected example LFC exposure, Figure \ref{fig:resid2d} plots each line with respect to its echelle order (y-axis) and x-pixel on the detector (x-axis) colored by the difference between the predicted and theoretical wavelength for that line in units of $\mps$.
 
@@
 %\includegraphics[width=0.5\textwidth]{Figures/lineResids2D_col.pdf}
-\caption{Residuals for a single LFC exposure plotted with respect to detector position (as defined by echelle order and x-pixel) for parametric and non-parametric wavelength calibration methods.
+\caption{Residuals of a single LFC exposure plotted with respect to detector position (as defined by echelle order and x-pixel) for parametric and non-parametric wavelength calibration methods.
 Each line is colored by the difference between the predicted wavelength and the theoretical wavelength for each line, given in units of $\mps$.
-Residuals to a classic, polynomial fit are shown in the top plot.
-The bottom plot shows residuals to an interpolated wavelength solution.
+\textbf{Top}: residuals to a classic, polynomial fit.
+\textbf{Bottom}: residuals to an interpolated wavelength solution.
 High-order structure, i.e.
@@
 There is a lot of vertical structure and some hints of a periodic, diagonal structure as well.
-The residuals of the interpolation test for the same exposure is shown in the center plot of Figure \ref{fig:resid2d}.
+The residuals of the interpolation test for the same exposure is shown in the bottom plot of Figure \ref{fig:resid2d}.
 There is no coherent structure here and smaller residuals.
 
-This shows how the flexibility of an interpolated model can account for high-order detector defects, which emerged as structure in the residuals of the classic, smooth, polynomial-driven wavelength solution.
-In the \expres\ data shown here, for example, the interpolated model accounted for \lz{detector effect STA will explain to us}.
+This shows how the flexibility of an interpolated model can account for high-order instrument or detector defects, which emerged as structure in the residuals of the classic, smooth, polynomial-driven wavelength solution.
 This same flexibility should similarly allow interpolated wavelength solutions to account for position errors in pixel image blocks for different detectors \citep{fischer2016, milakovic2020}.
-
-The bottom plot of Figure \ref{fig:resid2d} shows the residuals when ThAr exposures, which have much fewer and less regular lines than LFC exposures, are run through \name\ and used to predict wavelengths for the completely independent LFC exposures taken during the same range of time.
-Over-plotted in black are the positions of the ThAr lines.
-
-Note that running \name\ informed by only ThAr lines cannot be regarded as a direct comparison to the LFC runs, as the increased uncertainty and variability in ThAr line positions alone makes the resultant wavelength predictions an order of magnitude worse, hence the different scale of the colorbar.
-All the same, the residuals are relatively worse where lines are further spaced (for example, in redder echelle orders) than where lines are denser.
-
-Though the interpolated wavelength solution returns lower, less-structured residuals than the polynomial wavelength solution when guided by LFC lines, we see in the bottom plot of Figure \ref{fig:resid2d} that an interpolated wavelength solution can return much worse results in the space between widely separated calibration lines.
-Figure \ref{fig:waveResids} plots the residuals for a subset of one order for both a polynomial-based method and a PCHIP-based method guided by either ThAr lines or LFC lines.
-The PCHIP model with ThAr lines (blue) curve returns huge residuals between two widely-spaced ThAr lines.
-The classic, polynomial fit exhibits similar residuals in both amplitude and shape regardless of whether the set of ThAr lines or LFC lines are used.
-An interpolated wavelength solution using LFC lines (orange curve) exhibits the lowest residuals.
-
-The move to an interpolated wavelength solution is driven by the assumption that a high density of calibration lines allows for more freedom in the resultant wavelength solution.
-This freedom allows the wavelength solution to more accurately ascribe wavelengths.
-This flexibility, however, is no longer justified in the regime where there are large spaces between calibration lines, as is the case with a classic ThAr lamp.
 
@@
 \caption{Residuals when using ThAr lines to predict wavelengths for LFC lines.
-\textbf{Right}: residuals for a single LFC exposure plotted with respect to detector position as in Figure \ref{fig:resid2d}.
-Each line is colored by the difference between the predicted wavelength and the theoretical wavelength for each line, given in units of $\mps$.
+\textbf{Right}: residuals for a single LFC exposure plotted with respect to detector position and colored by residual, as in Figure \ref{fig:resid2d}.
 The positions of ThAr lines are over-plotted in yellow.
 In general, residuals are greater between ThAr lines with greater separation.
-\textbf{Left}: Comparison of polynomial-fit and interpolated wavelength solutions using either just ThAr lines or LFC lines.
-For a subset of echelle order 94, the difference in Angstroms between the assigned and model wavelengths for each LFC line is plotted with respect to pixel.
-The large residuals between two widely-spaced ThAr lines when implementing an interpolated wavelength solution with just ThAr lines (orange curve) extends out of frame.
-Note how the shape of the residuals from a polynomial fit are similar whether using ThAr lines or LFC lines, suggesting that little is added with the denser LFC lines.
+\textbf{Left}: Comparison of polynomial-fit and interpolated wavelength solutions using either just ThAr lines or LFC lines for a subset of echelle order 94.
+The shape of the residuals from a polynomial fit are similar whether using ThAr lines or LFC lines, suggesting that little is added with the denser LFC lines.
 The LFC lines only serve to make the residuals more centered around zero.
@@
 
+Though the interpolated wavelength solution returns lower, less-structured residuals than the polynomial wavelength solution when guided by LFC lines, the flexibility of an interpolated wavelength solution can result in much worse residuals when not properly constrained, for example in the space between widely separated calibration lines.
+The left plot of Figure \ref{fig:waveResids} shows the residuals when ThAr calibration lines, which are much fewer and less regularly spaced lines than LFC lines, are run through \name\ and used to predict wavelengths for the (completely independent) LFC exposures taken during the same range of time.
+Over-plotted in yellow are the positions of the ThAr lines.
+
+Note that running \name\ informed by only ThAr lines cannot be regarded as a direct comparison to the LFC runs, as the increased uncertainty and variability in ThAr line positions alone makes the resultant wavelength predictions an order of magnitude worse, hence the different scale of the colorbar in the left plot of Figure \ref{fig:waveResids} as compared to Figure \ref{fig:resid2d}.
+All the same, the residuals are in general worse where lines are further spaced (for example, in redder echelle orders) than where lines are denser.
+
+Figure \ref{fig:waveResids} plots the residuals for a subset of order 94 for both a polynomial-based method and a PCHIP-based method guided by either ThAr lines or LFC lines.
+The PCHIP model with ThAr lines (orange curve) returns huge residuals between two widely-spaced ThAr lines that extends out of frame.
+The classic, polynomial fit exhibits similar residuals in both amplitude and shape regardless of whether the set of ThAr lines or LFC lines are used.
+An interpolated wavelength solution using LFC lines (black curve) exhibits the lowest residuals.
+
+The move to an interpolated wavelength solution is driven by the assumption that a high density of calibration lines allows for more freedom in the resultant wavelength solution.
+This freedom allows the wavelength solution to more accurately ascribe wavelengths.
+This flexibility, however, is no longer justified in the regime where there are large spaces between calibration lines, which no longer provide sufficient constraint on the interpolated wavelength solution, as is the case in some regions of a classic ThAr lamp.
+
 \subsection{Impact on Radial Velocity Measurements}\label{sec:test-rv}
-\lz{Awaiting latest extraction run.  Was supposed to take a weekend, but I'm pretty sure I heard God laugh when RP said that.}
+We tested the effect of \name -generated wavelengths on RV measurements using \expres\ observations of \target.
+\target\ is a BLANK star and also SOME STATEMENT ON SUSPECTED ACTIVITY(?).
+We used BLANK observations of \target\ taken between BLANK and BLANK with an average signal-to-noise ratio of BLANK PROBABLY 250.
+Radial velocity mesurements are derived using the forward-modeling scheme employed by \expres's pipeline \citep{petersburg2020}.
+
+Figure \ref{fig:rvs} compares the resultant RVs when using a classic, ninth-degree polynomial wavelength solution and an \name -generated wavelength model.
+\lz{Awaiting latest extraction run...}
 
 \section{Choose Your Own Implementation} \label{sec:choices}
-We have described and test only one implementation of \name.
+We have described and tested only one implementation of \name.
 Using PCA and an interpolated wavelength solution is a fairly trivial step towards a complete hierarchical and non-parametric wavelength model.
@@
 
-We settled on a $K$ value of 6.
-Figure \ref{fig:pcLfc} shows the first 9 principal components constructed using LFC lines.
-There is clear structure in the first and second principal components.
-\lz{The diagonal-banding structure seen in components 3, 5, and 6 we believe to be capturing detector fabrication defects.
-in \expres's detector or other optical elements}  Components 3 and 4 show aberrant behavior on the edges of the two bluest echelle orders, where lower signal results in more variation in the line fits.
-Later principal components become dominated by noise over coherent structure.
+\begin{figure}[t]
+\centering
+\includegraphics[width=.65\textwidth]{Figures/kvals_all.pdf}
+\caption{Per-line RMS of returned wavelength models for different values of $K$.
+Per-line RMS is defined in Equation \ref{eq:rms} and provides a measure of the accuracy of a wavelength model.
+There is a dotted, vertical line at $K=6$, our chosen $K$ value.
+The dotted, horizontal line is the RMS for $K=6$.
+The improvement at around $K=6$ plateaus.}
+\label{fig:kvals}
+\end{figure}
 
@@
 the mean line positions for each line throughout the epoch of stability.
-The following plots show the first nine principal components constructed using LFC lines.
+The following $3 \times 3$ grid of plots show the first nine principal components constructed using LFC lines.
 These principal components represent the basis vectors along which the calibration of the spectrograph can deviate from the fiducial calibration.
-For each basis vector, each calibration line is plotted according to its echelle order and x-pixel and colored by the value of the basis vector for that line.
+For each principal component, or basis vector, each calibration line is plotted according to its echelle order and x-pixel and colored by the value of the basis vector for that line.
 Principal components beyond the sixth one become steadily more dominated by noise.}
@@
 
-In deciding a K value, we ran denoising tests, as described in Section \ref{sec:tests}, for K values spanning from 2 to 512.
-One would expect the returned wavelengths to get increasingly more accurate with larger K until components that represent only noise are incorporated.
-Residuals might then get worse before ultimately starting to get better again with large K, which marks the model over-fitting the data.
+In deciding a $K$ value, we ran denoising tests, as described in Section \ref{sec:tests}, for $K$ values spanning from 2 to 512.
+The resultant per-line RMS for each test is plotted in Figure \ref{fig:kvals}.
+One would expect the returned wavelengths to get increasingly more accurate with larger $K$ until components that represent only noise are incorporated.
+Residuals might then get worse before ultimately starting to get better again with large $K$, which marks the model over-fitting the data.
 Though the returned RMS never gets worse, we find that the improvement plateaus between $K=6$ and $K=128$.
@@
 a $K=32$ model show significant differences in less than 10 bluer lines, which are known to have greater error and variance in their measured line positions.
-A visual inspection of the principal components, such as the one displayed in Figure \ref{fig:pcLfc}, additionally show most of the structure is being captured in the first six principal components with later components dominated by noise.
+We therefore settled on a $K$ value of six.
 
-\begin{figure}[t]
-\centering
-\includegraphics[width=.5\textwidth]{Figures/kvals_all.pdf}
-\caption{Per line RMS of returned wavelength models for different values of $K$.
-Per line RMS is defined in Equation \ref{eg:rms}.
-There is a dotted, vertical line at $K=6$, our chosen $K$ value.
-The dotted, horizontal line is the RMS for $K=6$.
-The improvement at around $K=6$ plateaus.}
-\label{fig:kvals}
-\end{figure}
+Figure \ref{fig:pcLfc} shows the fiducial calibration state and the first 9 principal components constructed using LFC lines, which represent deviations from the fiducial calibration state.
+There is clear structure in the first and second principal components.
+Components three through six show smaller or more localized structure.
+Components three and four have aberrant behavior on the edges of the two bluest echelle orders, where lower signal results in more variation in the line fits.
+Later principal components become dominated by noise and show less coherent structure.
 
@@
 
-\begin{figure*}[h!]
+\begin{figure*}[b]
 \centering
@@
 The top row of plots shows the amplitudes for the first principal component while the bottom row shows the amplitudes for the second principal component.
-Points mark the times at which calibration exposures were taken.
 Lines show the result of a linear interpolation.
@@
 All days are therefore roughly on the same scale, but the y-axis is different from the left plots.
-In the right column, points are colored by the time of each exposure.}
+In the right column, points are colored by the MJD of each exposure.}
 \label{fig:nightlyVariation}
@@
 
-The amplitudes $a_{n,k}$ can also be interpolated with respect to any housekeeping data, not just time.
-Best results will come from interpolating with respect to whichever data is most strongly correlated with the calibration state of the spectrograph.
+The amplitudes $a_{n,k}$ can also be interpolated with respect to any good housekeeping data, not just time.
+Best results will come from interpolating with respect to whatever is most strongly correlated with the calibration state of the spectrograph.
 For example, with \expres, which is passively temperature controlled, the returned amplitudes $a_{n,k}$ were extremely correlated with the enclosure temperature, suggesting it would also be possible to interpolate with respect to temperature.
@@
 \label{sec:choice_wvp}
-Interpolation of wavelengths over pixel  is done order by order using a Piecewise Cubic Hermite Interpolating Polynomial (PCHIP) interpolator.
+In the implementation described here, interpolation of wavelengths over pixel  is done order by order using a Piecewise Cubic Hermite Interpolating Polynomial (PCHIP) interpolator.
 This interpolation incorporates the flexibility needed to model the changing dispersion of the spectrograph across an echelle order along with any detector defects while also enforcing monotonicity, which we know must be true across any one echelle order.
@@
 A more classic cubic spline interpolation can run into issues with arc lines, which are irregularly spaced or even blended.
-Close lines appearing at virtually the same pixel location but with different wavelengths could coerce a cubic spline into a very high slope, as with the blue curve in Figure \ref{fig:xinterp}.
-A line blend appears at approximately pixel 5300 causing the spline to twist to nearly vertical to account for both points.
+Close lines appearing at virtually the same pixel location but with different wavelengths could coerce a cubic spline into a very high slope, as with the green curve in Figure \ref{fig:xinterp}.
+A line blend appears at approximately pixel 5300, causing the spline to twist to nearly vertical to account for both points.
 This leads to huge deviations from the correct wavelengths around this line blend as the extraneously high slope of the spline is accounted for.
@@
 These huge digressions can be avoided by allowing for some smoothing in the interpolation.
-In Figure \ref{fig:xinterp}, we show an example in green using SciPy's implementation of a univarate spline.
+In Figure \ref{fig:xinterp}, we show an example in orange using SciPy's implementation of a univarate spline.
 While the result appears to follow the calibration lines much better, the smoothing ultimately causes larger residuals that are spatially correlated (Fig.
@@
 The resultant wavelength solution is underestimating the curvature of the pixel-wavelength relation, giving rise to similar issues as with an inflexible, parametric wavelength solution.
-Introducing this smoothing parameter is enforcing a smoothness we have no reason to believe is true of the data.
+Introducing this smoothing parameter is enforcing a smoothness we have no reason to believe is true of the data, thereby re-introducing one of the problems with parametric models.
 
 We instead turn to the PCHIP interpolator, which damps down huge deviations in the traditional cubic spline by requiring the resulting interpolated function be monotonic.
-This is a constraint we know is true for any one echelle order.
+Monotonicity is a constraint we know must be true for any one echelle order.
 Though the PCHIP interpolator shows a similar issue as a classic cubic spline around the ThAr line blend at pixel 5300, the effect is much smaller and effects fewer pixels.
@@
 To determine the needed line density, we systematically removed LFC lines from the \expres\ data and calculated the RMS of the returned wavelengths for the removed lines.
-Figure \ref{fig:density} plots the scatter of residuals of the wavelengths returned by \name\ as a function of separation between lines.
+Figure \ref{fig:density} plots the scatter of residuals of the wavelengths returned by \name\ as a function of separation between lines in units of resolution element.
 It gives an approximate estimation of the expected error between lines of different separation.
 Note, however, that the stability of the lines and their measured line centers quickly becomes the dominant error source in calibration lamps over the separation between lines.
+The needed line density depends on the resolution of a given spectrograph and the needed precision of the returned wavelength solution.
 
@@
 \includegraphics[width=0.5\textwidth]{Figures/lfcDensityTest.pdf}
-\caption{Per line RMS as a function of spacing between lines in units of resolution element.
+\caption{Per-line RMS as a function of spacing between lines in units of resolution element.
 The average line spacing is calculated using the average distance between LFC lines across its wavelength range.}
@@
 
+\name\ can easily incorporate simultaneous calibration data, where a calibration source is taken through a reference fiber during science exposures, especially if the simultaneous reference is also used when taking calibration images.
+The position of calibration lines taken through the reference fiber can simply be concatenated to the array of line positions taken through the science fiber.
+Both sets of lines when then be used when the low-dimensional basis is constructed.
+This allows the simultaneous calibration information to contribute to constructing the complete calibration space of the spectrograph and pinpoint where the spectrograph is in that calibration space for any exposure.
+
+The calibration for any science exposure with a simultaneous reference can be determined by finding the amplitude of each basis vector that most closely recreates the calibration line positions through the reference fiber.
+These amplitudes can then be used to recreate the calibration line positions through the science fiber as well.
+This replaces the need to interpolate the basis vector amplitudes from calibration exposures to science exposures, something that is done with respect to time in the example implementation described in this paper.
+The result is likely to be even more precise, as this method incorporates more data.
+
 It is also possible to apply \name\ to etalon data with some modifications.
-If the free spectral range (FSR) and therefor wavelength of each line is well known, then the etalon lines can be treated just like LFC lines with known wavelengths and orders.
-Alternatively, the FSR can be included as a free parameter that is included when constructing the low-dimensional basis.
-\lz{Hogg, I think more detail is needed here that I don't understand but would like to; can we discuss?}
+The simplest implementation is if the free spectral range (FSR) and therefore wavelength of each line of the etalon is well known.
+The etalon data can be interpolated onto a set of fiducial lines with set wavelengths.
+These fiducial lines would therefore be identifiable by echelle order and wavelength alone, with only their line positions varying with different exposures, thereby returning to the same framework as developed for the case of an LFC.
+This marks the simplest implementation of \name\ on etalon data, as the uncertainty of a line's wavelength is upstream of the model rather than built in.
 
-In terms of stability, \name\ should be adaptable to a wide range of instrument designs.
-\mb{In truth, the amplitude of the variation matters little as long as all axes of variation can be captured by PCA or another dimensionality-reducing model.
-Indeed, larger variations from less stabilized instruments might even make it easier for the PCA to identify these axes.
-(Is this true?)}
+\lz{Hogg: The following paragraph might need some work.
+I struggled to keep this discussion general (i.e.
+not referring to our specific implementation of \name) while also being helpful.
+However, I might have failed or just made all this very confusing.
+What do you think?}
 
-Most physical systems should have only a few dominant axis along which they vary, simply by what is possible from general mechanical engineering.
-With PCA, this can be tested by plotting the power of the basis vectors describing the calibration space, which should fall quickly after a few vectors.
-This only provides a measure of the variance in the PCA space, however, not an explicit test of the end-to-end variation in the resulting model.
+If the FSR varies very little, its value can in principle be captured by a PCA or similar method.
+However, to interpolate this information from calibration exposures to science exposures would necessitate a simultaneous reference or interpolation with respect to some other housekeeping data that contains information on the value of the FSR.
+If the FSR varies a lot, it becomes untenable to capture these changes with a PCA or similar method.
+Instead, the resultant basis vectors of the calibration space would have to be conditioned on the value of the FSR, becoming a non-trivial extension of \name.
+
+In terms of dimensionality reduction, most physical systems should have only a few dominant axis along which they vary, meaning \name\ should be adaptable to a wide range of instrument designs.
+With PCA, this can be tested by plotting the amplitude of the returned principal components, which should fall quickly after a few components.
+However, this only provides a measure of the variance in the PCA space, and is not an explicit test of the end-to-end variation in the resulting model.
 This condition is therefore necessary but not sufficient if implementing \name\ with PCA.
@@
 Here, an era of stability was chosen  manually in order to focus on separating out time-domain intervals in which the data is relatively homogeneous, e.g.
-most exposures show the same lines.
+most exposures show the same calibration lines.
 Homogeneity is, of course, implicitly required when implementing PCA.
 
-Simultaneous calibration data is easily incorporated, especially if the simultaneous reference is also used when taking calibration images.
-The position of calibration lines taken through the simultaneous fiber can be concatenated to the array of line positions when the low-dimensional basis is constructed.
-This not only allows the simultaneous calibration information to contribute to constructing the complete calibration space of the spectrograph, but also allows the simultaneous calibration information to uniquely pinpoint where the spectrograph is in that calibration space for any exposure, thereby removing the need to interpolate the basis vector amplitudes from calibration exposures to science exposures.
+Lastly, we caution that \name\ is extremely sensitive to upstream changes, for example an unstable PSF or slanting slit function, that may effect the line centers.
+For example, PCA is good for detecting variability, but is agnostic to the source of the variability.
+This is why the principal components shown in Figure \ref{fig:pcLfc} exhibit errant values for bluer LFC lines, which are lower signal and therefore have more variable fitted line centers.
+It is essential that the line positions being given to \name\ capture only the changes in the spectrograph's calibration state, not potential errors in the fitted line centers.
 
 \section{Discussion} \label{sec:discussion}
-We show that \name\ returns a lower RMS than classic, parametric methods by over a factor of two (Section \ref{sec:tests}).
+We show that \name\ returns a lower per-line RMS than classic, parametric methods by over a factor of two (Section \ref{sec:tests}).
 Using \name\ wavelengths also reduced the RMS in a planet fit to derived RVs by \lz{some percent hopefully probs} (Section \ref{sec:test-rv}).
+In implementing \name\ on \expres\ data, we have successfully constructed a model of \expres, which truly is an instrument with low-degrees of freedom.
+\name\ does not make any claims about what variability each basis vector represents.
+Those interested in interpreting the variability are encouraged to investigate how the amplitude of the different vectors varies with different housekeeping data to find their source.
+
 The endeavor to reach $0.1\,\mps$ radial-velocity precision necessitates improvements on all levels of spectroscopy, from the instrumentation to the data analysis.
@@
 Starting with a list of calibration lines with known wavelengths and well-fit line centers for each calibration exposure, \name\ will de-noise and interpolate the given lines into a full wavelength solution.
-\name\ levies the incredible stability of contemporary EPRV instruments and high density of lines made available by new calibration sources, such as LFCs and etalons, to achieve more accurate wavelengths.
-\name\ therefore assumes dense enough calibration lines to properly constrain a non-parametric wavelength model, and that the instrument has low degrees of freedom (see Section \ref{sec:others} for further discussion).
+\name\ levies the stability of contemporary EPRV instruments and high density of lines made available by new calibration sources, such as LFCs and etalons, to achieve more accurate wavelengths.
+\name\ therefore assumes dense enough calibration lines to properly constrain a non-parametric wavelength model, and that the instrument has low degrees of freedom.
 
 Denser calibration lines allow us to move to more flexible wavelength models that can account for non-smooth features in the wavelength solution.
-Stabilized spectrograph hardware insures the calibration state of the instrument is low-dimensional.
+Stabilized spectrograph hardware makes it more likely that the calibration state of the instrument is low-dimensional.
 All calibration images in a given generation of stability can therefore be used to constrain the accessible calibration space of the spectrograph as well as where in that calibration space the spectrograph lies for each calibration exposures.
@@
 On the other hand, PCA captures all variance, regardless of source.
-Though \name\ only wants to capture variation in the instrument, the PCA is also sensitive to uncertainties and failures in the upstream line position fitting.
+Though \name\ endeavors to capture only variation in the instrument, the PCA is also sensitive to uncertainties and failures in the upstream line position fitting.
 For example, we have seen that lower-signal lines that are harder to fit faithfully will have greater variety in returned line positions, which is in turn captured by the PCA.
@@
 \name\ is technically robust to localized issues with either the calibration source or the pipeline used to return line positions.
-With an interpolated wavelength model, one errant line position will only effect the resultant wavelength model out to the closest, neighboring lines.
-The effect of an outlier is diminished and kept local.
-In contrast, a parametric fit will change from a single errant line, changing the wavelength solution for the entire echelle order or the entire exposure.
+With an interpolated wavelength model, one errant line position will only effect the resultant wavelength model out to the close, neighboring lines.
+The effect of an outlier is diminished and kept localized.
+In contrast, an entire parametric fit will change from a single errant line, changing the wavelength solution for the entire exposure for a 2D fit.
 Through de-noising and outlier rejection, \name\ adds additional robustness against erroneous line positions.

### 2020-05-28 paper/paper.tex
Debra Comments

@@
 \begin{abstract}\noindent%
-\name\ is a hierarchical, non-parametric framework for wavelength-calibrating spectrographs.
+\name\ is a hierarchical, non-parametric framework for precision wavelength-calibration of spectrographs.
 This method is designed with the needs of extreme-precision radial velocity (\eprv) in mind, which require that instruments be calibrated or stabilized to better than $10^{-4}$ of a pixel.
-Instruments are likely to experience only a few degrees of freedom, especially \eprv\ instruments with highly stabilized optical systems and detectors.
+Instruments are likely to vary along only a few dominant degrees of freedom, especially \eprv\ instruments with highly stabilized optical systems and detectors.
 \name\ takes advantage of this property by using all calibration data to construct a low-dimensional space of all possible calibrations for an instrument.
-A wavelength model need only pinpoint where in this low-dimensional calibration space the spectrograph exists.
 \name\ also takes advantage of the development of laser-frequency combs or etalons, which generate a dense set of stable calibration points.
@@
 We demonstrate the success of this method with data from the EXtreme PREcision Spectrograph, \expres, which includes a Menlo Systems laser frequency comb.
-We find that \name\ returns more accurate wavelengths by greater than a factor of 2 over classic, non-hierarchical, polynomial solutions.
+Using \name\ wavelengths to predict the wavelengths of known calibration lines returns smaller per-line RMS by a factor of 5 when compared to a classic wavelength model.
 Employing \name\ wavelengths on \expres\ RV measurements of HD 3651 reduced the RMS of a planet fit by \lz{blah}.
@@
 \section{Introduction} 
-Extremely precise, radial-velocity (\eprv) programs have been incredibly fruitful in finding and characterizing extra-solar planets \citep{mayor2011, bonfils2013, plavchan2015, butler2017}.
+Precise, radial-velocity (\eprv) programs have been incredibly fruitful in finding and characterizing extra-solar planets \citep{mayor2011, bonfils2013, plavchan2015, butler2017}.
 These programs typically make use of spectrographs with resolutions on the order of $10^5$, which correspond to line widths on the order of $3000\,\mps$.
-Exoplanet science happens at the level of $3\,\mps$ precision, with current instruments aiming to reach $0.1\,\mps$ precision \citep{fischer2016}.
+Exoplanet science happens at the level of $3\,\mps$ precision, which is sensitive enough to discover gas-giant planets around other stars.
+The newest generation of instruments aim  to reach $0.1\,\mps$ precision, the required precision to detect terrestrial worlds \citep{fischer2016}.
 This requires new spectrographs to be calibrated or stabilized to better than $10^{-4}$ of a pixel (assuming that the spectrographs are well sampled).
-Some hardware--software systems, such as \espresso, \expres, and \neid\ were designed to deliver precision at this level \citep{pepe2013,  jurgenson2016, neid}.
-Early data from \expres\ and \espresso\ suggest that the hardware improvements truly propagate through to sub-$\mps$ uncertainties on final, on-sky RV measurements \citep{blackman2020, petersburg2020, pepe2020}.
+Some hardware--software systems, such as \espresso, \expres,  \neid, etc., were designed to deliver precision at this level \citep{pepe2013,  jurgenson2016, neid}.
+Early data from \expres\ and \espresso\ suggest that the hardware improvements truly propagate through to sub-$\mps$ uncertainties on final, on-sky RV measurements \citep{blackman2020, petersburg2020, mascareno2020}.
 
-Traditionally, wavelength solutions are constructed by fitting a polynomial to lines from a calibration source to describe the relationship between wavelength and pixel for each echelle order \citep{lovis2007, cersullo2019}.
-Each calibration image is treated independently, though the entire set of calibration images has been used to vet incongruous lines \citep{coffinet2019}.
-Just recently, work has been done that suggests a segmented polynomial in the dispersion direction, tuned to detector defects, better describes the wavelength solution than a single, continuous polynomial \citep{milakovic2020}.
+Traditionally, wavelength solutions are constructed by fitting a polynomial to lines from a calibration source to describe the relationship between wavelength and pixel for each echelle order \citep{butler1996, lovis2007, cersullo2019}.
+In this framework, each calibration image is treated independently.
+The returned wavelength solutions worked well at the level of 1 $\mps$ precision.
+The move towards 0.1 $\mps$ RV precision, however, necessitates higher-fidelity calibration data and wavelength models.
+These models need to account for high-order spatial variations that can arise from small imperfections in the optics of an instrument and non-uniformity in detector pixel sizes/spacing.
+There has been significant effort in using an entire set of calibration images to vet for incongruous ThAr lines \citep{coffinet2019} or obtain high-resolution Fourier transform spectra of reference cells \citep{wang2020}.
+It has also been found that using a segmented polynomial in the dispersion direction, tuned to detector defects, better describes the wavelength solution than a single, continuous polynomial \citep{milakovic2020}.
 
-Here, we propose to simplify and improve calibration programs for \eprv\ hardware systems with two very simple but innovative ideas.
+Here, we propose to simplify and improve calibration programs for \eprv\ hardware systems with two practical yet innovative ideas.
 The first flows from the fact that calibration sources---which include arc lamps (in some wavelength ranges), etalons, and laser-frequency combs (\lfc s)---illuminate the spectrograph with very stable, very dense sets of lines; almost every location in the spectrograph image plane is surrounded by nearby, useful calibration lines.
 
-This recommends a calibration methodology that is \emph{non-parametric}:  If every point in the spectrograph detector is sufficiently surrounded by nearby calibration lines, the wavelength solution can, for example, be made simply as an interpolation of the calibration data.
+This recommends a calibration methodology that is \emph{non-parametric}, or not defined by a prescribed, analytic function described by a finite number of parameters :  If every point in the spectrograph detector is sufficiently surrounded by nearby calibration lines, the wavelength solution can, for example, be made simply as an interpolation of the calibration data.
 The density of lines removes the need to enforce any functional form for the wavelength solution (such as a continuous ninth-order polynomial, for example).
@@
 The second simple idea follows from the observation that most physical systems have only a few dominant degrees of freedom, meaning most spectrographs vary along only a small number of axes in ``calibration space'', or the (very high-dimensional) space of all possible wavelength solutions.
-This is particularly likely with contemporary \eprv\ instruments, which have been stabilized in many directions.
+This is particularly likely with the latest generation of \eprv\ instruments with stringent environmental stabilizing, likely reducing the axes along which the instrument can vary.
 That is, spectrographs, especially stabilized ones, should have few environmentally accessible degrees of freedom.
 This renders it inadvisable to fit each calibration exposure or calibrate each science exposure independently.
-Instead, all the calibration data (or all the data) should be used to determine the space in which the instrument can and does vary.
+Instead, all the calibration data (or all the data) should be used to determine the calibration space in which the instrument can and does vary.
 Subsequent calibration work then need only determine where in the accessible part of calibration space the spectrograph was located for each exposure.
@@
 \lz{Hogg: I think it's stuff like this you wanted to audit for consistency with statistical literature?}
-In the context of probabilistic models, this structure is \emph{hierarchical}:  The calibration data are used not just to determine the wavelength solution at one moment, but also to determine the possible \emph{space} of wavelength solutions at all moments.
+In the context of probabilistic models, this structure is \emph{hierarchical}:  The calibration data are used not just to determine the wavelength solution at one moment, but also to determine the possible \emph{calibration space} of wavelength solutions at all moments.
 In the statistics literature, this concept is often described as \emph{de-noising}:  Every calibration exposure contains information about every other exposure.
@@
 The \name\ method is designed to take a series of calibration lines with known wavelengths and well-fit detector positions, and de-noise and interpolate them into a full wavelength model applicable to all exposures taken with the instrument.
-\name\ operates on the two core ideas that the wavelength solution should be given enormous flexibility, but that it lives in a very low-dimensional space, where the degrees of freedom are set by the limited kinematics of the spectrograph hardware.
+\name\ operates on the two core ideas that the wavelength solution should be given enormous flexibility, but that it lives in a very low-dimensional calibration space, where the degrees of freedom are set by the limited kinematics of the spectrograph hardware.
 \name\ therefore assumes that the space of possible calibration states for an instrument is low-dimensional.
@@
 There are many strategies for identifying calibration line positions and matching them to their assigned wavelengths; this problem is left out of scope for this work.
-However, we warn that systematic errors or large uncertainties in fitting line positions can propagate through to the wavelength models returned by \name.
+However, we caution that systematic errors or large uncertainties in fitting line positions can propagate through to the wavelength models returned by \name.
 
@@
 \KwData{line positions $x(\lambda,m,n)$ for each exposure $n$, with wavelengths $\lambda$ and echelle orders $m$}
-\KwResult{Basis vectors of the low-dimensional calibration space $g_k(\lambda,m)$, location of exposures in this space $a_{n,k}$}
+\KwResult{Basis vectors of the low-dimensional calibration space $g_k(\lambda,m)$, location of exposures in calibration space $a_{n,k}$}
 \While{change in missing or outlier line centers $>$ 0.01\%}
@@
 \expres\ is an environmentally-stabilized, fiber-fed, $R\sim137,000$, optical spectrograph \citep{jurgenson2016, blackman2020}.
-\expres\ has two different wavelength calibration sources, a ThAr lamp and a Menlo Systems laser frequency comb (LFC, e.g.
-\cite{wilken2012, molaro2013, probst2014}).
+\expres\ has two different wavelength calibration sources, a ThAr lamp and a Menlo Systems laser frequency comb (LFC).
+LFCs are unique in that the wavelengths of their emission lines are stable and exactly known to the order of pico-meters  \citep{wilken2012, molaro2013, probst2014}.
 
-Rather than using a simultaneous calibration fiber, two to three LFC exposures are interspersed between science exposures every 15-30 minutes.
+Rather than using a simultaneous calibration fiber, two to three LFC exposures are interspersed between science exposures roughly every 30 minutes while the telescope is slewing to different targets.
 ThAr exposures are taken at the beginning and end of each night.
@@
  \end{equation}
- where the repetition rate, $f_r$, is known from design of the LFC, and the offset frequency, $f_0$, has been determined experimentally.
+ where the repetition rate, $f_r$, is known from design of the LFC, and the offset frequency, $f_0$, has been determined by Menlo Systems, the manufacturer of the LFC.
 
 In order to satisfy the assumption of only low-order variation needed for \name, we used exposures from only one ``instrumental epoch" of \expres.
-\expres\ exposures are separated into different instrumental epochs that correspond to atypical hardware changes---changes that altered the position of the echellogram on the detector, the shape of the instrumental PSF, or the calibration sources.
+\expres\ exposures are separated into different instrumental epochs that correspond to hardware changes---changes that altered the position of the echellogram on the detector, the shape of the instrumental PSF, or the calibration sources.
 Significant instrumental changes break the assumption that an  instrument will experience only low-dimensional variations, i.e.
@@
 \centering
-\includegraphics[width=.5\textwidth]{Figures/all_results.pdf}
+\includegraphics[width=.6\textwidth]{Figures/all_results.pdf}
 \caption{Difference in predicted and theoretical wavelengths for the wavelength calibration tests described in Section \ref{sec:tests}.
@@
 
-Though the interpolated wavelength solution returns lower, less-structured residuals than the polynomial wavelength solution when guided by LFC lines, the flexibility of an interpolated wavelength solution can result in much worse residuals when not properly constrained, for example in the space between widely separated calibration lines.
-The left plot of Figure \ref{fig:waveResids} shows the residuals when ThAr calibration lines, which are much fewer and less regularly spaced lines than LFC lines, are run through \name\ and used to predict wavelengths for the (completely independent) LFC exposures taken during the same range of time.
+Though the interpolated wavelength solution returns lower, less-structured residuals than the polynomial wavelength solution when guided by LFC lines, the flexibility of an interpolated wavelength solution can result in much worse residuals when not properly constrained, for example in regions between widely separated calibration lines.
+The left plot of Figure \ref{fig:waveResids} shows the residuals when ThAr calibration lines, which are much fewer and are less regular than LFC lines, are run through \name\ and used to predict wavelengths for the (completely independent) LFC exposures taken during the same range of time.
 Over-plotted in yellow are the positions of the ThAr lines.
@@
 Note that running \name\ informed by only ThAr lines cannot be regarded as a direct comparison to the LFC runs, as the increased uncertainty and variability in ThAr line positions alone makes the resultant wavelength predictions an order of magnitude worse, hence the different scale of the colorbar in the left plot of Figure \ref{fig:waveResids} as compared to Figure \ref{fig:resid2d}.
-All the same, the residuals are in general worse where lines are further spaced (for example, in redder echelle orders) than where lines are denser.
+All the same, the residuals are in general worse where lines are further apart (for example, in redder echelle orders) than where lines are denser.
 
 Figure \ref{fig:waveResids} plots the residuals for a subset of order 94 for both a polynomial-based method and a PCHIP-based method guided by either ThAr lines or LFC lines.
-The PCHIP model with ThAr lines (orange curve) returns huge residuals between two widely-spaced ThAr lines that extends out of frame.
+The PCHIP model with ThAr lines (orange curve) returns huge residuals between two widely-separated ThAr lines that extends out of frame.
 The classic, polynomial fit exhibits similar residuals in both amplitude and shape regardless of whether the set of ThAr lines or LFC lines are used.
@@
 This freedom allows the wavelength solution to more accurately ascribe wavelengths.
-This flexibility, however, is no longer justified in the regime where there are large spaces between calibration lines, which no longer provide sufficient constraint on the interpolated wavelength solution, as is the case in some regions of a classic ThAr lamp.
+This flexibility, however, is no longer justified in the regime where there are large separations between calibration lines, as this no longer provide sufficient constraint on the interpolated wavelength solution, as is the case in some regions of a classic ThAr lamp.
 
@@
 We have described and tested only one implementation of \name.
-Using PCA and an interpolated wavelength solution is a fairly trivial step towards a complete hierarchical and non-parametric wavelength model.
+Using PCA and an interpolated wavelength solution is a statisticall straight-forward step towards a complete hierarchical and non-parametric wavelength model.
 It is possible to upgrade both the denoising and wavelength solution steps to true models.
@@
 This means the uncertainty from wavelength calibration could be completely marginalized out.
-Doing so, however, is only worthwhile if the wavelength calibration is a significant fraction of the error budget.
+Doing so will have the largest impact if the wavelength calibration is a significant fraction of the error budget.
 
@@
 \centering
-\includegraphics[width=.65\textwidth]{Figures/kvals_all.pdf}
+\includegraphics[width=.5\textwidth]{Figures/kvals_all.pdf}
 \caption{Per-line RMS of returned wavelength models for different values of $K$.
@@
 Lines show the result of a linear interpolation.
+In the top, right plot, the temperature of the optical bench is also plotted in orange.
 In the right plot, the principal component amplitudes have been artificially offset by the median amplitude for each night.
@@
 Best results will come from interpolating with respect to whatever is most strongly correlated with the calibration state of the spectrograph.
-For example, with \expres, which is passively temperature controlled, the returned amplitudes $a_{n,k}$ were extremely correlated with the enclosure temperature, suggesting it would also be possible to interpolate with respect to temperature.
+For example, with \expres, which is passively temperature controlled, the returned amplitudes $a_{n,k}$ were extremely correlated with the optical bench temperature, as shown in Figure \ref{fig:nightlyVariation}, suggesting it would also be possible to interpolate the amplitudes with respect to temperature.
 
@@
 \label{sec:others}
-We use \expres\ data here not because it is the optimal system for a method like \name, but because there is a strong need for good wavelength calibration in the EPRV community.
+We focus on an \eprv\ use-case here, because there is a strong need for wavelength calibration in the \eprv\ community.
+\expres\ is representative of the newest generation of \eprv\ spectrographs, and an LFC provides stable, dense calibration lines with known wavelengths, ideal for \name.
 The applicability of \name\ to any one instrument is a detailed question of the kind of variance experienced by the spectrograph and calibration sources available, but we hope to provide some approximate benchmarks and diagnostics.
@@
 
+\lz{I completely made up the below paragraph...}
+
+The number of calibration exposures needed is linked to the amount of variability in the instrument.
+Calibration exposures must span the range of calibration space accessible by the spectrograph, or at least accessed by the science calibrations that require wavelength solutions.
+Otherwise, the dimensionality-reduction method will not have the information required to construct a basis vector that spans the entire accessible calibration space.
+More observations are always better to help de-noise the calibration data.
+
 For the results presented here, the data was broken up into different eras of stability based on where the principal component amplitudes showed huge deviations.
@@
 \section{Discussion} \label{sec:discussion}
-We show that \name\ returns a lower per-line RMS than classic, parametric methods by over a factor of two (Section \ref{sec:tests}).
+We show that \name\ returns a lower per-line RMS than classic, parametric methods by a factor of 5 (Section \ref{sec:tests}).
+The residuals were also smoother, showing less spatial correlation.
 Using \name\ wavelengths also reduced the RMS in a planet fit to derived RVs by \lz{some percent hopefully probs} (Section \ref{sec:test-rv}).
@@
 Instead of using an interpolator at all, there might be better results from implementing something more sophisticated, such as a kernel method or a Gaussian process with a kernel adapted for the specifics of an instrument.
-There is in principle an enormous space of non-parametric methods to explore, which we leave outside the scope of this paper.
+There is in principle an enormous number of non-parametric methods to explore, which we leave outside the scope of this paper.
 

### 2020-05-29 paper/paper.tex
Added Needed Number of Exposures

@@
 
-\begin{figure*}[h]
+\begin{figure*}[t]
 \centering
@@
 
-Histograms of the per-line residuals for each of the above described polynomial, interpolation, and denoising tests (respectively) are shown in Figure \ref{fig:testHists}.
-Note that the spread in residuals is much smaller for both the denoising and interpolation tests relative to the results of the polynomial wavelength solution.
-
-\begin{figure}[h]
+\begin{figure}[t]
 \centering
@@
 
-\name -generated wavelengths also exhibit less structure in the returned residuals.
-For a randomly selected example LFC exposure, Figure \ref{fig:resid2d} plots each line with respect to its echelle order (y-axis) and x-pixel on the detector (x-axis) colored by the difference between the predicted and theoretical wavelength for that line in units of $\mps$.
+Histograms of the per-line residuals for each of the above described polynomial, interpolation, and denoising tests (respectively) are shown in Figure \ref{fig:testHists}.
+Note that the spread in residuals is much smaller for both the denoising and interpolation tests relative to the results of the polynomial wavelength solution.
 
@@
 \end{figure}
+
+\name -generated wavelengths also exhibit less structure in the returned residuals.
+For a randomly selected example LFC exposure, Figure \ref{fig:resid2d} plots each line with respect to its echelle order (y-axis) and x-pixel on the detector (x-axis) colored by the difference between the predicted and theoretical wavelength for that line in units of $\mps$.
 
@@
 
-\begin{figure}[t]
+\begin{figure}[h]
 \centering
@@
 Later principal components become dominated by noise and show less coherent structure.
-
-\subsection{Interpolation of Calibration State with Respect to Time}
+\pagebreak
+\subsection{Interpolation of Calibration State to Science Exposures}
 \label{sec:choice_avt}
@@
 
-\begin{figure*}[b]
+\begin{figure*}[t]
 \centering
@@
 
-To determine the needed line density, we systematically removed LFC lines from the \expres\ data and calculated the RMS of the returned wavelengths for the removed lines.
+Implementing \name\ will require an existing library of calibration exposures that span the range of calibration space accessible by the spectrograph, or at least the space accessed by the science exposures being calibrated.
+A new set of calibration exposures would be needed for any changes to the instrument that could significantly change the calibration space.
+While there is no exact cutoff for how many exposures are needed, the number is certainly much larger than $K$, the dimensionality of the calibration space.
+More calibration exposures will help with denoising .
+The number of calibration exposures required throughout a night will depend on how much the instrument varies throughout the night, and the chosen interpolation scheme, as mentioned in Section \ref{sec:choice_avt}.
+
+As an empirical assessment of the needed line density, we systematically removed LFC lines from the \expres\ data and calculated the per-line RMS of the returned wavelengths for the removed lines.
 Figure \ref{fig:density} plots the scatter of residuals of the wavelengths returned by \name\ as a function of separation between lines in units of resolution element.
@@
 \lz{Hogg: My notes said you might want to add more things here?}
-
-\lz{I completely made up the below paragraph...}
-
-The number of calibration exposures needed is linked to the amount of variability in the instrument.
-Calibration exposures must span the range of calibration space accessible by the spectrograph, or at least accessed by the science calibrations that require wavelength solutions.
-Otherwise, the dimensionality-reduction method will not have the information required to construct a basis vector that spans the entire accessible calibration space.
-More observations are always better to help de-noise the calibration data.
 

### 2020-07-16 paper/paper.tex
Pg1 Hogg Edits

@@
   A Non-Parametric, Hierarchical \\
-  Wavelength-Calibration Model for a Precision Spectrograph}
+  Wavelength-Calibration Method for a Precision Spectrograph}
 
@@
 \begin{abstract}\noindent%
-\name\ is a hierarchical, non-parametric framework for precision wavelength-calibration of spectrographs.
+\name\ is a non-parametric, hierarchical framework for precision wavelength-calibration of spectrographs.
 This method is designed with the needs of extreme-precision radial velocity (\eprv) in mind, which require that instruments be calibrated or stabilized to better than $10^{-4}$ of a pixel.
@@
 This density of calibration lines permits the freedom of a non-parametric wavelength solution that can match or adapt to any instrument or detector oddities, more so than any parametric model.
-We demonstrate the success of this method with data from the EXtreme PREcision Spectrograph, \expres, which includes a Menlo Systems laser frequency comb.
-Using \name\ wavelengths to predict the wavelengths of known calibration lines returns smaller per-line RMS by a factor of 5 when compared to a classic wavelength model.
+We demonstrate the success of this method with data from the \textsl{EXtreme PREcision Spectrograph}, \expres, which includes a laser frequency comb.
+Using \name\ wavelengths to predict the wavelengths of known calibration lines returns smaller per-line RMS by a factor of 5 when compared to a classic wavelength solution that is constructed using polynomials and treats each exposure individually .
 Employing \name\ wavelengths on \expres\ RV measurements of HD 3651 reduced the RMS of a planet fit by \lz{blah}.
@@
 \section{Introduction} 
-Precise, radial-velocity (\eprv) programs have been incredibly fruitful in finding and characterizing extra-solar planets \citep{mayor2011, bonfils2013, plavchan2015, butler2017}.
+Precise, radial-velocity (\eprv) programs have been incredibly fruitful in finding and characterizing extra-solar planets \citep[e.g.][]{mayor2011, bonfils2013, plavchan2015, butler2017}.
 These programs typically make use of spectrographs with resolutions on the order of $10^5$, which correspond to line widths on the order of $3000\,\mps$.
@@
 We used BLANK observations of \target\ taken between BLANK and BLANK with an average signal-to-noise ratio of BLANK PROBABLY 250.
-Radial velocity mesurements are derived using the forward-modeling scheme employed by \expres's pipeline \citep{petersburg2020}.
+Radial velocity mesurements are derived using a chunk-by-chunk, forward-modeling algorithm employed by \expres's pipeline \citep{petersburg2020}.
 
@@
 It is possible to upgrade both the denoising and wavelength solution steps to true models.
+It is also possible, of course, to implement either step individually.
+A hierarchical framework can be used to simply denoise the lines before they are fit to a parametric model, or use a non-parametric model on lines that haven't been denoised.
 

### 2020-07-16 paper/paper.tex
Pg 2-3 Hogg Edits + Figures

@@
 These models need to account for high-order spatial variations that can arise from small imperfections in the optics of an instrument and non-uniformity in detector pixel sizes/spacing.
-There has been significant effort in using an entire set of calibration images to vet for incongruous ThAr lines \citep{coffinet2019} or obtain high-resolution Fourier transform spectra of reference cells \citep{wang2020}.
-It has also been found that using a segmented polynomial in the dispersion direction, tuned to detector defects, better describes the wavelength solution than a single, continuous polynomial \citep{milakovic2020}.
+There has been significant effort in using an entire set of calibration images to identify incongruous ThAr lines \citep{coffinet2019} or obtain high-resolution Fourier transform spectra of reference cells \citep{wang2020}.
+It has also been found that using a segmented polynomial in the dispersion direction, tuned to capture detector defects, better describes the wavelength solution than a single, continuous polynomial \citep{milakovic2020}.
 
@@
 In some ways, this is a generalization of the recent work that demonstrated the efficacy of constructing a wavelength solution as multiple, segmented polynomials \citep{milakovic2020}.
-Going non-parametric will improve calibration accuracy by not forcing the choice of a parametric form that may bias the calibration, especially when the chosen function is inappropriate (as polynomials are at detector edges).
+Going non-parametric will improve calibration accuracy by not forcing the choice of a parametric form that may bias the calibration, especially when the chosen function is inappropriate (as, for example, polynomials are at detector edges).
 
 The second simple idea follows from the observation that most physical systems have only a few dominant degrees of freedom, meaning most spectrographs vary along only a small number of axes in ``calibration space'', or the (very high-dimensional) space of all possible wavelength solutions.
-This is particularly likely with the latest generation of \eprv\ instruments with stringent environmental stabilizing, likely reducing the axes along which the instrument can vary.
+This is particularly likely with the latest generation of \eprv\ instruments that are equipped with stringent environmental stabilizing, likely reducing the number and extent of the axes along which the instrument can vary.
 That is, spectrographs, especially stabilized ones, should have few environmentally accessible degrees of freedom.
@@
 Instead, all the calibration data (or all the data) should be used to determine the calibration space in which the instrument can and does vary.
-Subsequent calibration work then need only determine where in the accessible part of calibration space the spectrograph was located for each exposure.
+Subsequent calibration work then need only determine where in the small, accessible part of calibration space the spectrograph was located for each exposure.
 
-\lz{Hogg: I think it's stuff like this you wanted to audit for consistency with statistical literature?}
 In the context of probabilistic models, this structure is \emph{hierarchical}:  The calibration data are used not just to determine the wavelength solution at one moment, but also to determine the possible \emph{calibration space} of wavelength solutions at all moments.
-In the statistics literature, this concept is often described as \emph{de-noising}:  Every calibration exposure contains information about every other exposure.
-Thus every exposure can be improved (i.e.
-de-noised) with information from every other exposure.
+In the statistics literature, this concept is often described as \emph{de-noising}:  we can improve calibration by recognizing that every calibration exposure contains information about every other calibration exposure.
+Thus every exposure can be improved (i.e., de-noised) with information from every other exposure.
 
@@
 \name\ is designed for temperature-controlled, fiber-fed spectrographs with good calibration sources, such as laser-frequency combs, or etalons.
-We have in mind \eprv\ instruments and \eprv\ science cases, primarily because the need for good wavelength calibration in this field is so great.
+We have in mind \eprv\ instruments and \eprv\ science cases, primarily because the need for good wavelength calibration is so great in this field.
 Irregardless, we expect \name\ to have applications for other kinds of spectrographs in other contexts.
@@
 The \name\ method is designed to take a series of calibration lines with known wavelengths and well-fit detector positions, and de-noise and interpolate them into a full wavelength model applicable to all exposures taken with the instrument.
-\name\ operates on the two core ideas that the wavelength solution should be given enormous flexibility, but that it lives in a very low-dimensional calibration space, where the degrees of freedom are set by the limited kinematics of the spectrograph hardware.
-\name\ therefore assumes that the space of possible calibration states for an instrument is low-dimensional.
+\name\ operates on the two core ideas; The wavelength solution should be given enormous flexibility, but it lives in a very low-dimensional calibration space, where the degrees of freedom are set by the limited kinematics of the spectrograph hardware.
+\name\ therefore assumes that the space of possible calibration states for an instrument is low-dimensional while assuming very little about the forms of those states.
 

### 2020-08-22 paper/paper.tex
hogg affils
Completed Draft

@@
 \newcommand{\code}[1]{\texttt{#1}}
-\newcommand{\target}{HD $\bigstar$}
+\newcommand{\target}{HD 34411}
 
@@
 \affil{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, New York, NY 10003, USA}
+\affil{Flatiron Institute, Simons Foundation, 162 Fifth Avenue, New York, NY 10010, USA}
 \affil{Center for Data Science, New York University, 60 Fifth Avenue, New York, NY 10011, USA}
 \affil{Max-Planck-Institut für Astronomie, Königstuhl 17, D-69117 Heidelberg, Germany}
-\affil{Flatiron Institute, Simons Foundation, 162 Fifth Avenue, New York, NY 10010, USA}
 
@@
 This method is designed with the needs of extreme-precision radial velocity (\eprv) in mind, which require that instruments be calibrated or stabilized to better than $10^{-4}$ of a pixel.
-Instruments are likely to vary along only a few dominant degrees of freedom, especially \eprv\ instruments with highly stabilized optical systems and detectors.
-\name\ takes advantage of this property by using all calibration data to construct a low-dimensional space of all possible calibrations for an instrument.
-\name\ also takes advantage of the development of laser-frequency combs or etalons, which generate a dense set of stable calibration points.
-This density of calibration lines permits the freedom of a non-parametric wavelength solution that can match or adapt to any instrument or detector oddities, more so than any parametric model.
-We demonstrate the success of this method with data from the \textsl{EXtreme PREcision Spectrograph}, \expres, which includes a laser frequency comb.
-Using \name\ wavelengths to predict the wavelengths of known calibration lines returns smaller per-line RMS by a factor of 5 when compared to a classic wavelength solution that is constructed using polynomials and treats each exposure individually .
-Employing \name\ wavelengths on \expres\ RV measurements of HD 3651 reduced the RMS of a planet fit by \lz{blah}.
+Instruments vary along only a few dominant degrees of freedom, especially \eprv\ instruments with highly stabilized optical systems and detectors.
+\name\ takes advantage of this property by using all calibration data to construct a low-dimensional representation of all possible calibration states for an instrument.
+\name\ also takes advantage of laser-frequency combs or etalons, which generate a dense set of stable calibration points.
+This density permits the use of a non-parametric wavelength solution that can adapt to any instrument or detector oddities, more than any parametric model.
+We demonstrate the success of this method with data from the \textsl{EXtreme PREcision Spectrograph} (\expres), which includes a laser frequency comb.
+\name\ wavelengths predicts the wavelengths of known calibration lines at smaller per-line RMS by a factor of $\sim$5 when compared to a classic, polynomial wavelength calibration that treats each exposure individually.
+For ten months of radial-velocity measurements of HD 34411, using \name\ wavelengths reduced the RMS of the complete data set from $1.22\, \mps$ to $1.05\, \mps$; this improvement is equivalent to reducing the wavelength term in the error budget by $0.62\,\mps$.
 \end{abstract}
@@
 In the context of probabilistic models, this structure is \emph{hierarchical}:  The calibration data are used not just to determine the wavelength solution at one moment, but also to determine the possible \emph{calibration space} of wavelength solutions at all moments.
-In the statistics literature, this concept is often described as \emph{de-noising}:  we can improve calibration by recognizing that every calibration exposure contains information about every other calibration exposure.
+In statistics, this concept is often described as \emph{de-noising}:  we can improve calibration by recognizing that every calibration exposure contains information about every other calibration exposure.
 Thus every exposure can be improved (i.e., de-noised) with information from every other exposure.
@@
 \caption{A cartoon representation of the \name\ method, as described in Section \ref{sec:method}.
-We enlarge variations in measured line position, changes in calibration space, and interpolation deviations for clarity.
+We exaggerate variations in measured line position, changes in calibration space, and interpolation deviations for clarity.
 In step one, dimensionality reduction and denoising (\textsection \ref{sec:denoising}), the complete set of line positions for all exposures is analyzed to return a set of $K$ basis vectors, $g_n$, which represent different ways the spectrograph calibration changes.
@@
 \subsection{Impact on Radial Velocity Measurements}\label{sec:test-rv}
-We tested the effect of \name -generated wavelengths on RV measurements using \expres\ observations of \target.
-\target\ is a BLANK star and also SOME STATEMENT ON SUSPECTED ACTIVITY(?).
-We used BLANK observations of \target\ taken between BLANK and BLANK with an average signal-to-noise ratio of BLANK PROBABLY 250.
-Radial velocity mesurements are derived using a chunk-by-chunk, forward-modeling algorithm employed by \expres's pipeline \citep{petersburg2020}.
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=\textwidth]{Figures/rvs.pdf}
+\caption{HD 34411 RVs with \expres.
+RVs derived using the classic, polynomial-based wavelength solution are plotted in blue.
+RVs derived using the implementation of \name\ presented in this paper are shown in orange.
+The mean of each data set is shown as a horizontal line while $\pm$ one standard deviation from this mean is shaded in each data set's respective color.}
+\label{fig:rvs}
+\end{figure*}
+
+We tested \name -generated wavelengths on RV measurements using \expres\ observations of \target.
+\target\ is a G0V star.
+It is a $4.8\, Gyr$ star and relatively quiet ($log R'_{HK} = -5.09$) \citep{brewer2016}.
+We used 114 observations of \target\ taken between Oct.
+8, 2019 and Aug.
+13, 2020 with SNR 250.
+Radial-velocity measurements were derived using the chunk-by-chunk, forward-modeling algorithm employed by \expres's pipeline \citep{petersburg2020}.
 
 Figure \ref{fig:rvs} compares the resultant RVs when using a classic, ninth-degree polynomial wavelength solution and an \name -generated wavelength model.
-\lz{Awaiting latest extraction run...}
+Using \name -generated wavelengths reduces the RMS of the entire data set from $1.22\, \mps$ with the classic wavelength solution to $1.05\, \mps$.
+This is equivalent to removing an independent, additive noise component of $0.62\, \mps \,(=\sqrt{1.22^2-1.05^2})$.
+Other data sets showed a reduction in scatter between $0.62-0.80\, \mps$.
 
@@
 Later principal components become dominated by noise and show less coherent structure.
-\pagebreak
+
 \subsection{Interpolation of Calibration State to Science Exposures}
@@
 For example, with the \expres\ data shown here, it is clear that nightly calibration images are needed, but for a linear trend, only two calibration images throughout the night are strictly required.
+
+We also tested an implementation of \name\ where the $K$ principal components within a night were fit to a cubic.
+This emulates the current, polynomial-based wavelength solution implemented in the \expres\ pipline, where polynomial fits to calibration files are interpolated to science exposure by fitting the polynomial coefficients to a cubic.
+We found that this method returns comparable RV RMS for most targets, though appears to do better when a night has sparse calibration data.
 
@@
 \label{sec:others}
-We focus on an \eprv\ use-case here, because there is a strong need for wavelength calibration in the \eprv\ community.
+We focus on an \eprv\ use-case here because there is a strong need for wavelength calibration in the \eprv\ community.
 \expres\ is representative of the newest generation of \eprv\ spectrographs, and an LFC provides stable, dense calibration lines with known wavelengths, ideal for \name.
@@
 Note, however, that the stability of the lines and their measured line centers quickly becomes the dominant error source in calibration lamps over the separation between lines.
-The needed line density depends on the resolution of a given spectrograph and the needed precision of the returned wavelength solution.
+Additionally, in this assessment, the lines remain uniformly spaced.
+The needed line density depends on the resolution of a given spectrograph.
+the needed precision of the returned wavelength solution, and the chosen interpolation scheme.
 
@@
 
-\name\ can easily incorporate simultaneous calibration data, where a calibration source is taken through a reference fiber during science exposures, especially if the simultaneous reference is also used when taking calibration images.
-The position of calibration lines taken through the reference fiber can simply be concatenated to the array of line positions taken through the science fiber.
-Both sets of lines when then be used when the low-dimensional basis is constructed.
+This implementation of \name\ to \expres\ data interpolates calibration information to the science data using surrounding calibration exposures.
+Simultaneous calibration data can be used to reinforce the determined calibration state of an instrument at the time of science exposures.
+This simultaneous calibration data can come from calibration light shined through a reference fiber or any meta-data (e.g.
+temperature, pressure, time, etc.) that correlates with the calibration.
+For example, we have seen that the calibration state of \expres\ is correlated with the temperature of the optical bench.
+
+With a simultaneous reference fiber, the position of calibration lines taken through the reference fiber can simply be concatenated to the array of line positions taken through the science fiber.
+Both sets of lines will then be used when the low-dimensional basis is constructed.
 This allows the simultaneous calibration information to contribute to constructing the complete calibration space of the spectrograph and pinpoint where the spectrograph is in that calibration space for any exposure.
@@
 
-\bibliography{paper.bib}
+\bibliography{paper}
 

### 2020-08-22 paper/paper.tex
Discusion of Interpolation in Pixel

@@
 \label{sec:choice_wvp}
-In the implementation described here, interpolation of wavelengths over pixel  is done order by order using a Piecewise Cubic Hermite Interpolating Polynomial (PCHIP) interpolator.
+In the implementation described here, interpolation of wavelengths over pixel  is done order-by-order using a Piecewise Cubic Hermite Interpolating Polynomial (PCHIP) interpolator.
 This interpolation incorporates the flexibility needed to model the changing dispersion of the spectrograph across an echelle order along with any detector defects while also enforcing monotonicity, which we know must be true across any one echelle order.
@@
 Figure \ref{fig:xinterp} shows that using the PCHIP interpolator returns the lowest residuals.
+
+There likely exists an even more fitting model between an overly-constrained polynomial fit or a completely free spline interpolation.
+For example, there has been success interpolating wavelengths with respect to pixel using a segmented polynomial in the dispersion direction, especially when tuned to known detector defects \citep{milakovic2020}.
+Stiffer, more constrained splines or carefully chosen knot position may afford the perfect marriage of freedom and constraint that will better describe wavelength changes with pixel.
 
@@
 
-\acknowledgements{We thank the EXPRES team for their work on the instrument, software, and pipeline.
+\acknowledgements
+We thank the EXPRES team for their work on the instrument, software, and pipeline.
 We thank Dan Foreman-Mackey for illuminating discussion.
@@
 LLZ gratefully acknowledges support from the National Science Foundation Graduate Research Fellowship under Grant No.
-DGE1122492.}
+DGE1122492.
 

### 2020-08-25 paper/paper.tex
Debra Comments and RV edits

@@
 \newcommand{\project}[1]{\textsl{#1}}
-\newcommand{\name}{\project{Excalibur}}
+\newcommand{\name}{\project{excalibur}}
+\newcommand{\Name}{\project{Excalibur}}
 \newcommand{\acronym}[1]{{\small{#1}}}
@@
 \begin{document}
-\title{\name:
+\title{\Name:
   A Non-Parametric, Hierarchical \\
@@
 \affil{Center for Data Science, New York University, 60 Fifth Avenue, New York, NY 10011, USA}
-\affil{Max-Planck-Institut für Astronomie, Königstuhl 17, D-69117 Heidelberg, Germany}
+\affil{Max-Planck-Institut f\"ur Astronomie, Königstuhl 17, D-69117 Heidelberg, Germany}
 
@@
 \begin{abstract}\noindent%
-\name\ is a non-parametric, hierarchical framework for precision wavelength-calibration of spectrographs.
-This method is designed with the needs of extreme-precision radial velocity (\eprv) in mind, which require that instruments be calibrated or stabilized to better than $10^{-4}$ of a pixel.
-Instruments vary along only a few dominant degrees of freedom, especially \eprv\ instruments with highly stabilized optical systems and detectors.
-\name\ takes advantage of this property by using all calibration data to construct a low-dimensional representation of all possible calibration states for an instrument.
-\name\ also takes advantage of laser-frequency combs or etalons, which generate a dense set of stable calibration points.
-This density permits the use of a non-parametric wavelength solution that can adapt to any instrument or detector oddities, more than any parametric model.
+\Name\ is a non-parametric, hierarchical framework for precision wavelength-calibration of spectrographs.
+It is designed with the needs of extreme-precision radial velocity (\eprv) in mind, which require that instruments be calibrated or stabilized to better than $10^{-4}$ pixels.
+Instruments vary along only a few dominant degrees of freedom, especially \eprv\ instruments, which feature highly stabilized optical systems and detectors.
+\Name\ takes advantage of this property by using all calibration data to construct a low-dimensional representation of all possible calibration states for an instrument.
+\Name\ also takes advantage of laser-frequency combs or etalons, which generate a dense set of stable calibration points.
+This density permits the use of a non-parametric wavelength solution that can adapt to any instrument or detector oddities, better than any parametric model, such as a polynomial.
 We demonstrate the success of this method with data from the \textsl{EXtreme PREcision Spectrograph} (\expres), which includes a laser frequency comb.
-\name\ wavelengths predicts the wavelengths of known calibration lines at smaller per-line RMS by a factor of $\sim$5 when compared to a classic, polynomial wavelength calibration that treats each exposure individually.
-For ten months of radial-velocity measurements of HD 34411, using \name\ wavelengths reduced the RMS of the complete data set from $1.22\, \mps$ to $1.05\, \mps$; this improvement is equivalent to reducing the wavelength term in the error budget by $0.62\,\mps$.
+When wavelengths are assigned to laser comb lines using \name, the RMS of the residuals is about five times lower than wavelengths assigned using polynomial fits to individual exposures.
+In a typical test case, radial-velocity measurements of HD~34411 showed a reduction in RMS scatter over a 10-month time baseline from $1.22$ to $1.05\, \mps$.
+This improvement is equivalent to removing an independent, additive error of $0.62\, \mps$.
 \end{abstract}
@@
 \keywords{instrumentation: spectrographs -- instrumentation: detectors -- techniques: spectroscopic -- techniques: radial velocities -- methods: data analysis -- methods: statistical}
-
+\section*{~}\pagebreak
 \section{Introduction} 
-Precise, radial-velocity (\eprv) programs have been incredibly fruitful in finding and characterizing extra-solar planets \citep[e.g.][]{mayor2011, bonfils2013, plavchan2015, butler2017}.
-These programs typically make use of spectrographs with resolutions on the order of $10^5$, which correspond to line widths on the order of $3000\,\mps$.
-Exoplanet science happens at the level of $3\,\mps$ precision, which is sensitive enough to discover gas-giant planets around other stars.
+Precise, radial-velocity (\eprv) programs have been fruitful in finding and characterizing extra-solar planets \citep[e.g.][]{mayor2011, bonfils2013, plavchan2015, butler2017}.
+These programs typically make use of spectrographs with resolutions on the order of $50,000-100,000$, which correspond to line widths on the order of $3000\,\mps$.
+The state of the art RV precision reached $1\, \mps$ in 2016.
 The newest generation of instruments aim  to reach $0.1\,\mps$ precision, the required precision to detect terrestrial worlds \citep{fischer2016}.
 This requires new spectrographs to be calibrated or stabilized to better than $10^{-4}$ of a pixel (assuming that the spectrographs are well sampled).
-Some hardware--software systems, such as \espresso, \expres,  \neid, etc., were designed to deliver precision at this level \citep{pepe2013,  jurgenson2016, neid}.
-Early data from \expres\ and \espresso\ suggest that the hardware improvements truly propagate through to sub-$\mps$ uncertainties on final, on-sky RV measurements \citep{blackman2020, petersburg2020, mascareno2020}.
+Both \expres\ and \espresso\ have been commissioned for more than a year and are demonstrating $<0.1 \mps$ instrumental errors and $\sim 0.2\, \mps$ errors on stars \citep{pepe2013, jurgenson2016, blackman2020, peterburg2020, brewer2020, mascareno2020}.
 
@@
 The first flows from the fact that calibration sources---which include arc lamps (in some wavelength ranges), etalons, and laser-frequency combs (\lfc s)---illuminate the spectrograph with very stable, very dense sets of lines; almost every location in the spectrograph image plane is surrounded by nearby, useful calibration lines.
-
-This recommends a calibration methodology that is \emph{non-parametric}, or not defined by a prescribed, analytic function described by a finite number of parameters :  If every point in the spectrograph detector is sufficiently surrounded by nearby calibration lines, the wavelength solution can, for example, be made simply as an interpolation of the calibration data.
+This enables use of a calibration methodology that is \emph{non-parametric}, or not defined by a prescribed, analytic function described by a finite number of parameters :  If every point in the spectrograph detector is sufficiently surrounded by nearby calibration lines, the wavelength solution can, for example, be made simply as an interpolation of the calibration data.
 The density of lines removes the need to enforce any functional form for the wavelength solution (such as a continuous ninth-order polynomial, for example).
-In some ways, this is a generalization of the recent work that demonstrated the efficacy of constructing a wavelength solution as multiple, segmented polynomials \citep{milakovic2020}.
-Going non-parametric will improve calibration accuracy by not forcing the choice of a parametric form that may bias the calibration, especially when the chosen function is inappropriate (as, for example, polynomials are at detector edges).
+In some ways, this is a generalization of recent work, which has demonstrated the efficacy of constructing a wavelength solution as multiple, segmented polynomials \citep{milakovic2020}.
+A non-parametric approach will improve calibration accuracy by not forcing the choice of a parametric form that may bias the calibration, especially when the chosen function is inappropriate (as, for example, polynomials are at detector edges).
 
 The second simple idea follows from the observation that most physical systems have only a few dominant degrees of freedom, meaning most spectrographs vary along only a small number of axes in ``calibration space'', or the (very high-dimensional) space of all possible wavelength solutions.
-This is particularly likely with the latest generation of \eprv\ instruments that are equipped with stringent environmental stabilizing, likely reducing the number and extent of the axes along which the instrument can vary.
+This is particularly true of \eprv\ instruments, which are equipped with stringent environmental stabilizing, likely reducing the number and extent of the axes along which the instrument can vary.
 That is, spectrographs, especially stabilized ones, should have few environmentally accessible degrees of freedom.
@@
 
-\name\ is designed for temperature-controlled, fiber-fed spectrographs with good calibration sources, such as laser-frequency combs, or etalons.
+The method \name\ is designed for temperature-controlled, fiber-fed spectrographs with good calibration sources, such as laser-frequency combs, or etalons.
 We have in mind \eprv\ instruments and \eprv\ science cases, primarily because the need for good wavelength calibration is so great in this field.
 Irregardless, we expect \name\ to have applications for other kinds of spectrographs in other contexts.
-\name\ should be applicable to almost every astronomical spectrograph, though the precision of the returned wavelengths will depend on the available calibration sources (more discussion in Section \ref{sec:others} below) .
+\Name\ should be applicable to other spectrograph with low-dimensional variability, though the precision of the returned wavelengths will depend on the available calibration sources (more discussion in Section \ref{sec:others} below) .
 
 \section{Method} \label{sec:method}
-The \name\ method is designed to take a series of calibration lines with known wavelengths and well-fit detector positions, and de-noise and interpolate them into a full wavelength model applicable to all exposures taken with the instrument.
-\name\ operates on the two core ideas; The wavelength solution should be given enormous flexibility, but it lives in a very low-dimensional calibration space, where the degrees of freedom are set by the limited kinematics of the spectrograph hardware.
-\name\ therefore assumes that the space of possible calibration states for an instrument is low-dimensional while assuming very little about the forms of those states.
+The \name\ method is designed to take many calibration images, each containing a series of calibration lines with known wavelengths and well-fit detector positions, and de-noise and interpolate them into a full wavelength model applicable to all exposures taken with the instrument.
+It operates on the two core ideas; The wavelength solution should be allowed flexibility, but it lives in a very low-dimensional calibration space, where the degrees of freedom are set by the limited kinematics of the spectrograph hardware.
+\Name\ therefore assumes that the space of possible calibration states for an instrument is low-dimensional while assuming very little about the forms of those states.
 
-\name\ also assumes dense enough calibration line coverage with well-fit line centers to provide sufficient constraint on an interpolated wavelength solution across an echelle order.
+\Name\ assumes dense-enough calibration line coverage with well-fit line centers to provide sufficient constraints on an interpolated wavelength solution across an echelle order.
 Upstream errors in line center positions may propagate through \name\ wavelength models.
-The needed line density is dependent on the required precision of the returned wavelength model; larger spacing between lines offer less constraint and are likely to return worse wavelengths.
+The required line density is dependent on the required precision of the returned wavelength model; larger spacing between lines offer less constraint and are likely to return worse wavelengths.
 We revisit and quantify these conditions in Section \ref{sec:others}.
 
-Wavelength calibration is usually posed in the following way.  Given an exposure $n$, and echelle order $m$, there is a relationship between
+Wavelength calibration is usually posed in the following way: Given an exposure $n$, and echelle order $m$, there is a relationship between
 the two-dimensional $(x,y)$-position on the detector and the
@@
 Classically, pipelines employ polynomials to construct smooth wavelength solutions for each exposure.
-For example, the \expres\ pipeline sets the function $f(x,y,m;\theta_{n})$ from Equation \ref{eq:wsol} to be a 2D, 9\textsuperscript{th}-order polynomial where $\theta_{n}$ represents the polynomial coefficients, $c_{nij}$, unique to each exposure $n$ \citep{petersburg2020}.
+For example, the \expres\ pipeline sets the function $f(x,y,m;\theta_{n})$ from Equation \ref{eq:wsol} to a 2D, 9\textsuperscript{th}-order polynomial, where $\theta_{n}$ represents the polynomial coefficients, $c_{nij}$, unique to each exposure $n$ \citep{petersburg2020}.
 \begin{equation}
@@
 \end{equation}
-Here, the y-dependence is dropped as that dependence is carried by spectral order $m$.
+Here, the y-dependence is dropped as, in our application, this dependence is carried by spectral order $m$.
 The line position information is therefore uniquely identified by echelle order $m$ and pixel in the dispersion direction, $x$.
@@
 
-Given a stabilized instrument with low degrees of freedom, however,  the calibration of any image can and should be informed by the calibration of every other image.
+Given a stabilized instrument with low degrees of freedom, however, the calibration of any image can be reliably informed by the calibration of every other image.
 The calibration data themselves can be used to develop a low-dimensional basis for expressing the space of all possible calibrations for a spectrograph with few degrees of freedom.
@@
 The resultant $x(\lambda,m,n)$, a list of calibration line positions for a given exposure, can be regarded as the calibration state of the spectrograph for that exposure.
-When this calibration structure is used to deliver a wavelength solution, $x(\lambda,m,n)$ will be inverted back into $\lambda(x,m,n)$ to recover wavelengths for each detector position $x$ and echelle order $m$..
+When this calibration structure is used to deliver a wavelength solution, $x(\lambda,m,n)$ can be inverted back into $\lambda(x,m,n)$ to recover wavelengths for each detector position $x$ and echelle order $m$..
 
-The challenge is to learn these basis functions from the data and get the $K$ amplitudes, $a_{nk}$, for every exposure $n$.
-There are many ways to discern these basis functions.
+The challenge is to learn these basis functions, $g_k(\lambda,m)$, from the data and get the $K$ amplitudes, $a_{nk}$, for every exposure $n$.
+There are many ways to discern the basis functions.
 In this paper, we present a model using principal component analysis (PCA) \citep{wiki_pca}.
-A PCA is justifiable \lz{(Hogg)} in the limit where there is very high signal-to-nose ratio, as is usually the case with typical calibration images.
+A PCA is justifiable in the limit where there is very high signal-to-nose ratio, as is usually the case with typical calibration images.
 There are many alternatives to PCA for this dimensionality reduction; we return to this point in Section \ref{sec:choices} below.
@@
 These basis vectors span the $K$-dimensional calibration space of the spectrograph, which includes all possible wavelength solutions.
-In step two (\textsection \ref{sec:interp_time}), \name\ interpolates the amplitudes of each basis vector, $a_n,k$, to return the calibration state for a specific science exposure, returned as a set of de-noised calibration lines.
+In step two (\textsection \ref{sec:interp_time}), the amplitude of each basis vector, $a_{n,k}$, is interpolated to return the calibration state for a specific science exposure, returned as a set of de-noised calibration lines.
 The assigned wavelengths of these de-noised line positions are then interpolated onto other pixels in step three (\textsection \ref{sec:interp_wsol}) to construct a full wavelength model that returns wavelength as a function of detector position $x$ and echelle order $m$.}
@@
 \subsection{Dimensionality Reduction: De-Noising of Calibration Frames} \label{sec:denoising}
-\name\ will use calibration images to  1) determine the complete calibration space in which an instrument varies and 2) where in the accessible calibration space the spectrograph existed for each exposure.
+\Name\ will use calibration images to determine 1) the space in which an instrument varies and 2) where in the accessible calibration space the spectrograph existed for each exposure.
 For each calibration exposure, $n$, \name\ requires a full list of lines, $(\lambda,m)$ that are expected in each calibration exposure.
@@
 There are many strategies for identifying calibration line positions and matching them to their assigned wavelengths; this problem is left out of scope for this work.
-However, we caution that systematic errors or large uncertainties in fitting line positions can propagate through to the wavelength models returned by \name.
+That is, \name\ assumes that this has been done correctly.
+We caution, though, that systematic errors or large uncertainties in fitting line positions can propagate through to errors or biases in the wavelength models returned by \name.
 
-For each exposure, $n$, every line, $(\lambda,m)$, has an associated fitted detector position, $x(\lambda,m,n)$, for example x-pixel in an 2D extracted echelle order.
+For each exposure, $n$, every line, $(\lambda,m)$, has an associated fitted (measured) detector position, $x(\lambda,m,n)$, for example x-pixel in an 2D extracted echelle order.
 Fitted line centers that are missing from an exposure (e.g.
@@
 Let there be $P$ lines per exposure.
-\name\ reads in a $N \times P$ matrix of line positions for each of $P$ lines for each of $N$ exposures.
+\Name\ reads in a $N \times P$ matrix of line positions for each of $P$ lines for each of $N$ exposures.
 
@@
 The magnitude of each principal component for each exposure, $a_{nk}$, represents the scalar amplitude of these deviations for each exposure.
-\name\ then uses a small number, $K$, of principal components to reconstruct a de-noised version of the line positions as formulated in Equation \ref{eq:excl_wsol}.
+\Name\ then uses a small number, $K$, of principal components to reconstruct a de-noised version of the line positions as formulated in Equation \ref{eq:excl_wsol}.
 
@@
 \KwData{line positions $x(\lambda,m,n)$ for each exposure $n$, with wavelengths $\lambda$ and echelle orders $m$}
-\KwResult{Basis vectors of the low-dimensional calibration space $g_k(\lambda,m)$, location of exposures in calibration space $a_{n,k}$}
+\KwResult{Basis vectors of the low-dimensional calibration space $g_k(\lambda,m)$, location of exposures in calibration space expressed by amplitudes $a_{n,k}$}
 \While{change in missing or outlier line centers $>$ 0.01\%}
@@
 	}
-\caption{Dimension Reduction and De-Noising}
+\caption{Dimensionality Reduction and De-Noising}
 \end{algorithm}
@@
  \label{sec:interp_time}
-\name\ then interpolates the amplitude, $a_{n,k}$, of each principal component to determine the calibration state of the spectrograph.
+In \name\, the amplitude, $a_{n,k}$, of each principal component is then interpolated to determine the calibration state of the spectrograph.
 For example, the amplitude can be interpolated with respect to time to recreate the calibration state of the spectrograph at different times.
@@
 In the implementation of \name\ presented here, the amplitudes of the principal components are interpolated linearly with respect to time.
-\name\ finds new magnitudes $a_{n'k}$ for a science exposure $n'$ at time $t_{n'}$, which are used to construct the calibration state of the spectrograph for that point in time.
-Using interpolated magnitudes, $a_{n,k}$, and the basis vectors, $g_k(\lambda,m)$, returned by the de-noising process, \name\ can construct a set of calibration lines, $x(\lambda,m,n') $, for any exposure as prescribed in Equation \ref{eq:excl_wsol}.
+These new, linearly interpolated magnitudes, $a_{n\,k}$ for a science exposure $n'$ at time $t_{n'}$ are then used to construct the calibration state of the spectrograph for that point in time.
+Using interpolated magnitudes, $a_{n,k}$, and the basis vectors, $g_k(\lambda,m)$, returned by the de-noising process, a new set of calibration lines, $x(\lambda,m,n')$ can be constructed
+ for any exposure as prescribed in Equation \ref{eq:excl_wsol}.
 
@@
 From the de-noising step, \name\ can now construct a set of calibration lines, $x(\lambda,m,n')$ for any exposure $n'$.
-To construct a wavelength solution, \name\ interpolates the known wavelengths over the exposure specific line centers.
+To construct a wavelength solution, we invert $x(\lambda,m,n')$ to $\lambda(x,m,n')$ by interpolating known wavelengths of the calibration lines over detector position.
 For instance, interpolating the known wavelengths vs.
@@
 We found that a cubic-spline interpolation that enforces monotonicity, such as a Piecewise Cubic Hermite Interpolating Polynomial (PCHIP) interpolator, works well for interpolating wavelengths onto pixels.
-A cubic spline allows for more flexibility than a continuous function, while the enforced monotonicity prevents large deviations that may befall a cubic spline.
+A cubic spline allows for more flexibility than a continuous function, while the enforced monotonicity allows the wavelength solution, $\lambda(x,m,n')$, to be invertible and prevents large deviations that may befall a cubic spline.
 Choice in interpolation scheme, $K$, and other tests are further discussed in Section \ref{sec:choice_wvp}.
@@
 We tested \name\ using data from \expres, the EXtreme PRecison Spectrograph.
-\expres\ is an environmentally-stabilized, fiber-fed, $R\sim137,000$, optical spectrograph \citep{jurgenson2016, blackman2020}.
+\expres\ is an environmentally-stabilized, fiber-fed optical spectrograph with a median resolving power $R=\lambda/\delta_{\lambda}=\sim137,000$, over the wavelength range of $390-780\, nm$ \citep{jurgenson2016, blackman2020}.
 \expres\ has two different wavelength calibration sources, a ThAr lamp and a Menlo Systems laser frequency comb (LFC).
@@
 
-Rather than using a simultaneous calibration fiber, two to three LFC exposures are interspersed between science exposures roughly every 30 minutes while the telescope is slewing to different targets.
+Rather than using a simultaneous calibration fiber, two to three LFC exposures are obtained roughly every 30 minutes while the telescope is slewing to the next target.
 ThAr exposures are taken at the beginning and end of each night.
@@
 
-LFC lines cover echelle orders 84-124, which contain 19203 calibration lines.
+LFC lines cover echelle orders 84-124, which contain approximately 19200 calibration lines.
 Though the results are primarily based on work with LFC data, there is some discussion of applications to arc lamps.
-ThAr lines cover all 86 extracted orders of \expres\ (echelle orders 75-160), which include 5295 lines.
+ThAr lines cover all 86 extracted orders of \expres\ (echelle orders 75-160), which include approximately 5300 lines.
 For both the LFC and ThAr data, lines that appeared in less than 60\% of exposures were not included in the analysis.
 Similarly, files with more than 60\% of expected lines missing were cut from the analysis.
+A list of echelle orders $m$, line wavelengths $\lambda$, and pixel positions $x$, which have been calculated by the EXPRES pipeline \citep{petersburg2020}, are read into \name.
 
- \name\ inputs a list of echelle orders $m$, line wavelengths $\lambda$, and pixel positions $x$ of each line for each exposure.
-We, therefore, use line positions generated by the pre-existing \expres\ pipeline from the following procedure \citep{petersburg2020}.
-
- A ThAr wavelength solution is generated from each ThAr exposure using the IDL code \texttt{thid.pro}, developed by Jeff Valenti.
+Line positions from the \expres\ pipeline are generated as follows.
+A thorium argon (ThAr) wavelength solution is generated from each ThAr exposure using the IDL code \texttt{thid.pro}, developed by Jeff Valenti.
 This code identifies ThAr lines by matching lines in an exposure against a line atlas.
-Lines are matched manually when there is a significant change in the instrument, but is otherwise done using forward-modeling.
+Line matching is carried out in an automated and unsupervised way with a Levenburg-Marquardt minimization routine.
 Once each line's position is identified and matched to a wavelength from the line atlas, a sixth-order, 2D polynomial is fit over pixel location $x$, echelle order $m$, and scaled wavelength $m\lambda$ (wavelengths are scaled in order to distinguish lines that may appear in more than one order).
@@
 Each peak in an echelle order is then fit with a Gaussian.
-The mean of the fitted Gaussian is taken to be the center of the line.
+The mean of this fitted Gaussian to a single peak is taken to be the center of the line.
 For each line, the ThAr wavelength solution is used to estimate the mode number of a line.
@@
 
-In order to satisfy the assumption of only low-order variation needed for \name, we used exposures from only one ``instrumental epoch" of \expres.
-\expres\ exposures are separated into different instrumental epochs that correspond to hardware changes---changes that altered the position of the echellogram on the detector, the shape of the instrumental PSF, or the calibration sources.
-Significant instrumental changes break the assumption that an  instrument will experience only low-dimensional variations, i.e.
-that the calibration space of all possible wavelength solutions is low-dimensional.
-Each instrument epoch is therefore treated independently.
+In order to satisfy the assumption that there exists only low-order variation, which is needed for \name, we used exposures from after the photonic crystal fiber was changed in Summer 2019 and the LFC stabilized.
 In the results presented here, we use 1227 LFC exposures and 78 ThAr exposures taken between October 14 and December 18, 2019 on 29 unique nights.
@@
 The per line RMS as defined in Equation \ref{eq:rms} is given in the top-right corner in each method's corresponding color.
-\name\ returns the smallest spread in residuals.}
+\Name\ returns the smallest spread in residuals.}
 \label{fig:testHists}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{Figures/lineResids2D_col.png}
+\includegraphics[width=0.55\textwidth]{Figures/lineResids2D_col.png}
 %\includegraphics[width=0.5\textwidth]{Figures/lineResids2D_col.pdf}
@@
 High-order structure, i.e.
-vertical stripes and patchiness, is apparent in the residuals to a polynomial wavelength solution, which assumes smoothness.
-The increased flexibility of the interpolated wavelength solution can account for this high-order structure, resulting in much smaller residuals with less spatial coherence.}
+vertical stripes and patchiness, is apparent in the residuals to a polynomial wavelength solution, which assumes smoothness.}
 \label{fig:resid2d}
@@
 
-\name -generated wavelengths also exhibit less structure in the returned residuals.
+\Name -generated wavelengths also exhibit less structure in the returned residuals.
 For a randomly selected example LFC exposure, Figure \ref{fig:resid2d} plots each line with respect to its echelle order (y-axis) and x-pixel on the detector (x-axis) colored by the difference between the predicted and theoretical wavelength for that line in units of $\mps$.
@@
 This shows how the flexibility of an interpolated model can account for high-order instrument or detector defects, which emerged as structure in the residuals of the classic, smooth, polynomial-driven wavelength solution.
-This same flexibility should similarly allow interpolated wavelength solutions to account for position errors in pixel image blocks for different detectors \citep{fischer2016, milakovic2020}.
+This same flexibility may similarly allow interpolated wavelength solutions to account for position errors in pixel image blocks for different detectors \citep{fischer2016, milakovic2020}.
 
@@
 \includegraphics[width=\textwidth]{Figures/rvs.pdf}
-\caption{HD 34411 RVs with \expres.
-RVs derived using the classic, polynomial-based wavelength solution are plotted in blue.
-RVs derived using the implementation of \name\ presented in this paper are shown in orange.
-The mean of each data set is shown as a horizontal line while $\pm$ one standard deviation from this mean is shaded in each data set's respective color.}
+\caption{HD 34411 RVs measured with \expres.
+RVs derived using the implementation of \name\ presented in this paper are plotted in blue.
+The nightly-binned RVs are over-plotted as orange squares.
+The RMS of each data set is given in the top-right corner.}
 \label{fig:rvs}
@@
 
-We tested \name -generated wavelengths on RV measurements using \expres\ observations of \target.
+We tested \name -generated wavelengths on RV measurements using \expres\ observations of \target, which are presented in Table \ref{tab:rv}.
 \target\ is a G0V star.
@@
 13, 2020 with SNR 250.
-Radial-velocity measurements were derived using the chunk-by-chunk, forward-modeling algorithm employed by \expres's pipeline \citep{petersburg2020}.
+Radial-velocity measurements were derived using the chunk-by-chunk, forward-modeling algorithm ran by the \expres\ team \citep{petersburg2020}.
 
@@
 This is equivalent to removing an independent, additive noise component of $0.62\, \mps \,(=\sqrt{1.22^2-1.05^2})$.
-Other data sets showed a reduction in scatter between $0.62-0.80\, \mps$.
+A reduction in RMS was observed universally in all data sets tested; other data sets showed a reduction in scatter between $0.62-0.80\, \mps$.
+
+\begin{deluxetable}{c c c}[ht!]
+\tabletypesize{\scriptsize}
+\tablecaption{\expres\ RVs using \name\ Wavelengths \label{tab:rv}}
+\tablenum{2}
+\tablehead{
+\colhead{JD-2440000} & \colhead{RV $\mps$ } & \colhead{Error $\mps$ }
+}
+\startdata
+ 18764.4771  & 3.139  & 0.335 \\
+ 18764.4791  & 1.035  & 0.332 \\ 
+ 18764.4810  & 3.074  & 0.324 \\
+ 18771.4179  &-1.927  & 0.342 \\
+          & \(\vdots\)  & \\
+\enddata
+\tablecomments{The full data set is available online}
+\end{deluxetable}
 
@@
 We have described and tested only one implementation of \name.
-Using PCA and an interpolated wavelength solution is a statisticall straight-forward step towards a complete hierarchical and non-parametric wavelength model.
+Using PCA and an interpolated wavelength solution is a statistically straight-forward step towards a complete hierarchical and non-parametric wavelength model.
 It is possible to upgrade both the denoising and wavelength solution steps to true models.
@@
 
-\subsection{Value of K}
+\subsection{Dimensionality of the Calibration Space, $K$}
 \label{sec:choice_k}
-The value of $K$ represents the dimensionality of the calibration space within which the spectrograph lives.
+The dimensionality of the calibration space within which the spectrograph linves is represented by $K$.
 In practice, it is the number of principal components, or basis vectors, used to reconstruct the de-noised line centers.
@@
 temperature, pressure, time, etc.) that correlates with the calibration.
-For example, we have seen that the calibration state of \expres\ is correlated with the temperature of the optical bench.
+For example, we have seen that the calibration state of \expres\ is correlated with the temperature of the optical bench, even with the bench varying by less than one degree.
+This correlation is not seen in the RVs, suggesting the changes with temperature are calibrated out at the wavelength stage.
 
@@
 
-Lastly, we caution that \name\ is extremely sensitive to upstream changes, for example an unstable PSF or slanting slit function, that may effect the line centers.
+Lastly, we caution that \name\ is extremely sensitive to upstream changes that may effect the line centers.
 For example, PCA is good for detecting variability, but is agnostic to the source of the variability.
@@
 In implementing \name\ on \expres\ data, we have successfully constructed a model of \expres, which truly is an instrument with low-degrees of freedom.
-\name\ does not make any claims about what variability each basis vector represents.
+\Name\ does not make any claims about what variability each basis vector represents.
 Those interested in interpreting the variability are encouraged to investigate how the amplitude of the different vectors varies with different housekeeping data to find their source.
@@
 The endeavor to reach $0.1\,\mps$ radial-velocity precision necessitates improvements on all levels of spectroscopy, from the instrumentation to the data analysis.
-We believe implementing \name, a hierarchical, non-parametric model for wavelength calibration, is a step towards reducing the error contribution from data reduction pipelines.
+Implementing \name, a hierarchical, non-parametric model for wavelength calibration, is a step towards reducing the error contribution from data reduction pipelines.
 
 Starting with a list of calibration lines with known wavelengths and well-fit line centers for each calibration exposure, \name\ will de-noise and interpolate the given lines into a full wavelength solution.
-\name\ levies the stability of contemporary EPRV instruments and high density of lines made available by new calibration sources, such as LFCs and etalons, to achieve more accurate wavelengths.
-\name\ therefore assumes dense enough calibration lines to properly constrain a non-parametric wavelength model, and that the instrument has low degrees of freedom.
+\Name\ leverages the stability of contemporary EPRV instruments and high density of lines made available by new calibration sources, such as LFCs and etalons, to achieve more accurate wavelengths.
+\Name\ therefore assumes dense enough calibration lines to properly constrain a non-parametric wavelength model, and that the instrument has low degrees of freedom.
 
@@
 Compared to traditional methods, which involve fitting high-order polynomials, \name\ has several useful statistical properties.
-\name\ is technically robust to localized issues with either the calibration source or the pipeline used to return line positions.
+\Name\ is technically robust to localized issues with either the calibration source or the pipeline used to return line positions.
 With an interpolated wavelength model, one errant line position will only effect the resultant wavelength model out to the close, neighboring lines.
@@
 
-\name\ can be applied to any data that contains information about the calibration state of the spectrograph (see Section \ref{sec:others}).
+\Name\ can be applied to any data that contains information about the calibration state of the spectrograph (see Section \ref{sec:others}).
 For example, though LFC and ThAr exposures are used as an example in this paper, \name\ would work similarly for an etalon or any other arc lamp with a list of lines and assigned wavelengths.
@@
 
-\name\ is designed and optimized for extreme-precision, radial-velocity spectrographs.
-In the battle to construct a high-fidelity data pipeline for extreme-precision radial-velocity measurements, we believe \name\ represents a step towards mitigating the error from wavelength calibration, as demonstrated by tests using \expres\ data (Section \ref{sec:tests}).
+\Name\ is designed and optimized for extreme-precision, radial-velocity spectrographs.
+In the battle to construct a high-fidelity data pipeline for extreme-precision radial-velocity measurements, we have shown that \name\ represents a step towards mitigating the error from wavelength calibration, as demonstrated by tests using \expres\ data (Section \ref{sec:tests}).
 Though the focus was on \eprv\ instruments here, \name\ should be largely applicable to nearly any astronomical spectrograph.

### 2020-08-25 paper/paper.tex
All Comments Integrated
I even re-read the paper myself.

@@
 
-\documentclass[modern]{aastex63}
+\documentclass[twocolumn,table,xcolor]{aastex63}
 \usepackage[ruled,vlined,norelsize]{algorithm2e}
@@
 \addtolength{\textheight}{0.70in}
-\renewcommand{\twocolumngrid}{} % hahaha!
+%\renewcommand{\twocolumngrid}{} % hahaha!
 \setlength{\parindent}{1.1\baselineskip}
@@
 Instruments vary along only a few dominant degrees of freedom, especially \eprv\ instruments, which feature highly stabilized optical systems and detectors.
-\Name\ takes advantage of this property by using all calibration data to construct a low-dimensional representation of all possible calibration states for an instrument.
+\Name\ takes advantage of this property by using all calibration data to construct a low-dimensional representation of all accessible calibration states for an instrument.
 \Name\ also takes advantage of laser-frequency combs or etalons, which generate a dense set of stable calibration points.
-This density permits the use of a non-parametric wavelength solution that can adapt to any instrument or detector oddities, better than any parametric model, such as a polynomial.
+This density permits the use of a non-parametric wavelength solution that can adapt to any instrument or detector oddities better than any parametric model, such as a polynomial.
 We demonstrate the success of this method with data from the \textsl{EXtreme PREcision Spectrograph} (\expres), which includes a laser frequency comb.
@@
 In a typical test case, radial-velocity measurements of HD~34411 showed a reduction in RMS scatter over a 10-month time baseline from $1.22$ to $1.05\, \mps$.
-This improvement is equivalent to removing an independent, additive error of $0.62\, \mps$.
 \end{abstract}
@@
 \keywords{instrumentation: spectrographs -- instrumentation: detectors -- techniques: spectroscopic -- techniques: radial velocities -- methods: data analysis -- methods: statistical}
-\section*{~}\pagebreak
+%\section*{~}\pagebreak
 \section{Introduction} 
@@
 This requires new spectrographs to be calibrated or stabilized to better than $10^{-4}$ of a pixel (assuming that the spectrographs are well sampled).
-Both \expres\ and \espresso\ have been commissioned for more than a year and are demonstrating $<0.1 \mps$ instrumental errors and $\sim 0.2\, \mps$ errors on stars \citep{pepe2013, jurgenson2016, blackman2020, peterburg2020, brewer2020, mascareno2020}.
+Two next-generation spectrographs, \expres\ and \espresso, have been commissioned for more than a year and are demonstrating $<0.1 \mps$ instrumental errors and $\sim 0.2\, \mps$ errors on stars \citep{pepe2013, jurgenson2016, blackman2020, petersburg2020, brewer2020, mascareno2020}.
 
@@
 The returned wavelength solutions worked well at the level of 1 $\mps$ precision.
-The move towards 0.1 $\mps$ RV precision, however, necessitates higher-fidelity calibration data and wavelength models.
+
+The move towards 0.1 $\mps$ RV precision, necessitates higher-fidelity calibration data and wavelength models.
 These models need to account for high-order spatial variations that can arise from small imperfections in the optics of an instrument and non-uniformity in detector pixel sizes/spacing.
@@
 \section{Method} \label{sec:method}
-The \name\ method is designed to take many calibration images, each containing a series of calibration lines with known wavelengths and well-fit detector positions, and de-noise and interpolate them into a full wavelength model applicable to all exposures taken with the instrument.
+\begin{figure*}[t]
+\centering
+\includegraphics[width=\textwidth]{Figures/methodCartoon.pdf}
+\caption{A cartoon representation of the \name\ method, as described in Section \ref{sec:method}.
+We exaggerate variations in measured line position, changes in calibration space, and interpolation deviations for clarity.
+In step one, dimensionality reduction and denoising (\textsection \ref{sec:denoising}), the complete set of line positions for all exposures is analyzed to return a set of $K$ basis vectors, $g_n$, which represent different ways the spectrograph calibration changes.
+These basis vectors span the $K$-dimensional calibration space of the spectrograph, which includes all possible wavelength solutions.
+In step two (\textsection \ref{sec:interp_time}), the amplitude of each basis vector, $a_{n,k}$, is interpolated to return the calibration state for a specific science exposure, returned as a set of de-noised calibration lines.
+The assigned wavelengths of these de-noised line positions are then interpolated onto other pixels in step three (\textsection \ref{sec:interp_wsol}) to construct a full wavelength model that returns wavelength as a function of detector position $x$ and echelle order $m$.}
+\label{fig:cartoon}
+\end{figure*}
+
+The \name\ method is designed to take many calibration images, each containing a series of calibration lines with known wavelengths and well-fit detector positions, and de-noise and interpolate this information into a full wavelength model applicable to all exposures taken with the instrument.
 It operates on the two core ideas; The wavelength solution should be allowed flexibility, but it lives in a very low-dimensional calibration space, where the degrees of freedom are set by the limited kinematics of the spectrograph hardware.
@@
 \end{equation}
-Here, the y-dependence is dropped as, in our application, this dependence is carried by spectral order $m$.
-The line position information is therefore uniquely identified by echelle order $m$ and pixel in the dispersion direction, $x$.
+Here, the y-dependence is dropped as, in our framing, this dependence is carried by spectral order $m$.
+The line position on the detector is therefore uniquely identified by echelle order $m$ and pixel in the dispersion direction, $x$.
 The coefficients $c_{nij}$ are interpolated from the time of calibration exposures to time $t_n$ of a science exposure $n$ by a third-order polynomial of time.
-This third-order polynomial is evaluated at the time of non-calibration, science exposures to re-construct a 2D, 9th-order polynomial wavelength solution for that exposure.
+This third-order polynomial is evaluated at the time of non-calibration, science exposures to re-construct the coefficients for a 2D, 9th-order polynomial wavelength solution for that exposure.
 Each calibration image obtains its $c_{nij}$ coefficients independently.
@@
 The resultant $x(\lambda,m,n)$, a list of calibration line positions for a given exposure, can be regarded as the calibration state of the spectrograph for that exposure.
-When this calibration structure is used to deliver a wavelength solution, $x(\lambda,m,n)$ can be inverted back into $\lambda(x,m,n)$ to recover wavelengths for each detector position $x$ and echelle order $m$..
+When this calibration structure is used to deliver a wavelength solution, $x(\lambda,m,n)$ can be inverted back into $\lambda(x,m,n)$ to recover wavelengths for each detector position $x$ and echelle order $m$.
 
@@
 There are many ways to discern the basis functions.
-In this paper, we present a model using principal component analysis (PCA) \citep{wiki_pca}.
+In this paper, we present one implementation of \expres\ using principal component analysis (PCA) \citep{wiki_pca}.
 A PCA is justifiable in the limit where there is very high signal-to-nose ratio, as is usually the case with typical calibration images.
 There are many alternatives to PCA for this dimensionality reduction; we return to this point in Section \ref{sec:choices} below.
-
-\begin{figure*}[t]
-\centering
-\includegraphics[width=\textwidth]{Figures/methodCartoon.pdf}
-\caption{A cartoon representation of the \name\ method, as described in Section \ref{sec:method}.
-We exaggerate variations in measured line position, changes in calibration space, and interpolation deviations for clarity.
-In step one, dimensionality reduction and denoising (\textsection \ref{sec:denoising}), the complete set of line positions for all exposures is analyzed to return a set of $K$ basis vectors, $g_n$, which represent different ways the spectrograph calibration changes.
-These basis vectors span the $K$-dimensional calibration space of the spectrograph, which includes all possible wavelength solutions.
-In step two (\textsection \ref{sec:interp_time}), the amplitude of each basis vector, $a_{n,k}$, is interpolated to return the calibration state for a specific science exposure, returned as a set of de-noised calibration lines.
-The assigned wavelengths of these de-noised line positions are then interpolated onto other pixels in step three (\textsection \ref{sec:interp_wsol}) to construct a full wavelength model that returns wavelength as a function of detector position $x$ and echelle order $m$.}
-\label{fig:cartoon}
-\end{figure*}
 
@@
  \label{sec:interp_time}
-In \name\, the amplitude, $a_{n,k}$, of each principal component is then interpolated to determine the calibration state of the spectrograph.
+In \name, the amplitude, $a_{n,k}$, of each principal component is interpolated to determine the calibration state of the spectrograph.
 For example, the amplitude can be interpolated with respect to time to recreate the calibration state of the spectrograph at different times.
@@
 In the implementation of \name\ presented here, the amplitudes of the principal components are interpolated linearly with respect to time.
-These new, linearly interpolated magnitudes, $a_{n\,k}$ for a science exposure $n'$ at time $t_{n'}$ are then used to construct the calibration state of the spectrograph for that point in time.
-Using interpolated magnitudes, $a_{n,k}$, and the basis vectors, $g_k(\lambda,m)$, returned by the de-noising process, a new set of calibration lines, $x(\lambda,m,n')$ can be constructed
- for any exposure as prescribed in Equation \ref{eq:excl_wsol}.
+This is discussed more in Section \ref{sec:choice_avt}.
+These new, linearly interpolated magnitudes, $a_{n\,k}$ for a science exposure $n'$ at time $t_{n'}$, are then used to construct the calibration state of the spectrograph for that point in time.
+Using interpolated amplitudes, $a_{n,k}$, and the basis vectors, $g_k(\lambda,m)$, returned by the de-noising process, a new set of calibration lines, $x(\lambda,m,n')$ can be constructed
+ for any exposure as formulated in Equation \ref{eq:excl_wsol}.
 
@@
 For instance, interpolating the known wavelengths vs.
-line centers onto every integer x will generate wavelengths for each pixel in an echelle order.
+line centers onto every integer $x$ will generate wavelengths for each pixel in an echelle order.
 
-We found that a cubic-spline interpolation that enforces monotonicity, such as a Piecewise Cubic Hermite Interpolating Polynomial (PCHIP) interpolator, works well for interpolating wavelengths onto pixels.
-A cubic spline allows for more flexibility than a continuous function, while the enforced monotonicity allows the wavelength solution, $\lambda(x,m,n')$, to be invertible and prevents large deviations that may befall a cubic spline.
-Choice in interpolation scheme, $K$, and other tests are further discussed in Section \ref{sec:choice_wvp}.
+After experiments, we found that a cubic-spline interpolation that enforces monotonicity, such as a Piecewise Cubic Hermite Interpolating Polynomial (PCHIP) interpolator, works well for interpolating wavelengths onto pixels.
+A cubic spline allows for more flexibility than a parameterized function, while the enforced monotonicity allows the wavelength solution, $\lambda(x,m,n')$, to be invertible and prevents spurious deviations that may befall a cubic spline.
+Choices in interpolation scheme, $K$, and other tests are further discussed in Section \ref{sec:choice_wvp}.
 
@@
 We tested \name\ using data from \expres, the EXtreme PRecison Spectrograph.
-\expres\ is an environmentally-stabilized, fiber-fed optical spectrograph with a median resolving power $R=\lambda/\delta_{\lambda}=\sim137,000$, over the wavelength range of $390-780\, nm$ \citep{jurgenson2016, blackman2020}.
-\expres\ has two different wavelength calibration sources, a ThAr lamp and a Menlo Systems laser frequency comb (LFC).
-LFCs are unique in that the wavelengths of their emission lines are stable and exactly known to the order of pico-meters  \citep{wilken2012, molaro2013, probst2014}.
+\expres\ is an environmentally-stabilized, fiber-fed optical spectrograph with a median resolving power $R=\lambda/\delta_{\lambda}=\sim137,000$, over a wavelength range of $390-780\, nm$ \citep{jurgenson2016, blackman2020}.
+\expres\ has two different wavelength calibration sources, a thorium argon (ThAr) lamp and a Menlo Systems laser frequency comb (LFC).
+LFCs are unique in that the wavelengths of their emission lines are stable and exactly known at pico-meter accuracy \citep{wilken2012, molaro2013, probst2014}.
 
-Rather than using a simultaneous calibration fiber, two to three LFC exposures are obtained roughly every 30 minutes while the telescope is slewing to the next target.
+Rather than using a simultaneous calibration fiber, two to three LFC exposures are obtained roughly every 30 minutes while the telescope is slewing to new targets.
 ThAr exposures are taken at the beginning and end of each night.
-All calibration data is taken through the science fiber, so that calibration light travels down the same optical pathway and is projected onto the same pixels as all science observations.
+All calibration data are taken through the science fiber, so that calibration light travels down the same optical pathway and is projected onto the same pixels as the science observations.
+Light passes through a pupil slicer and double scrambler before being injected into a rectangular fiber, which is fed through a mechanical agitator to ensure modal mixing\citep{petersburg2018}.
 
 LFC lines cover echelle orders 84-124, which contain approximately 19200 calibration lines.
-Though the results are primarily based on work with LFC data, there is some discussion of applications to arc lamps.
+Though our results are primarily based on work with LFC data, there will be some discussion of applications to arc lamps below.
 ThAr lines cover all 86 extracted orders of \expres\ (echelle orders 75-160), which include approximately 5300 lines.
 For both the LFC and ThAr data, lines that appeared in less than 60\% of exposures were not included in the analysis.
-Similarly, files with more than 60\% of expected lines missing were cut from the analysis.
-A list of echelle orders $m$, line wavelengths $\lambda$, and pixel positions $x$, which have been calculated by the EXPRES pipeline \citep{petersburg2020}, are read into \name.
+Similarly, exposures with more than 60\% of expected lines missing were cut from the analysis.
+A list of echelle orders $m$, line wavelengths $\lambda$, and pixel positions $x$, which have been calculated by the EXPRES pipeline \citep{petersburg2020} for every line of every exposure, is read into \name.
 
 Line positions from the \expres\ pipeline are generated as follows.
-A thorium argon (ThAr) wavelength solution is generated from each ThAr exposure using the IDL code \texttt{thid.pro}, developed by Jeff Valenti.
+A ThAr wavelength solution is generated from each ThAr exposure using the IDL code \texttt{thid.pro}, developed by Jeff Valenti.
 This code identifies ThAr lines by matching lines in an exposure against a line atlas.
@@
  \end{equation}
- where the repetition rate, $f_r$, is known from design of the LFC, and the offset frequency, $f_0$, has been determined by Menlo Systems, the manufacturer of the LFC.
+ where the repetition rate, $f_r$, is known from the design of the LFC, and the offset frequency, $f_0$, has been determined by Menlo Systems, the manufacturer of the LFC.
 
-In order to satisfy the assumption that there exists only low-order variation, which is needed for \name, we used exposures from after the photonic crystal fiber was changed in Summer 2019 and the LFC stabilized.
+In order to satisfy the assumption that there exists only low-order variation, which is needed for \name, we used exposures from after the photonic crystal fiber within the LFC was changed in Summer 2019, and the LFC stabilized.
 In the results presented here, we use 1227 LFC exposures and 78 ThAr exposures taken between October 14 and December 18, 2019 on 29 unique nights.
@@
 \section{Tests}\label{sec:tests}
-We perform a series of tests to validate the performance of \name\ and benchmark \name -generated wavelengths against wavelengths generated by a classic, parametric method.
-For all tests, we leave out a subset of calibration lines with known wavelengths as a ``validation'' sample, generate wavelengths for these lines using the remaining data, and compare the predicted wavelength to the assigned wavelength of each line.
+We perform a series of tests to validate the performance of \name\ and benchmark \name -generated wavelengths against wavelengths generated by a classic, non-hierarchical, parametric method.
+To implement training-validation tests, we leave out a subset of calibration lines with known wavelengths as a ``validation'' sample, generate wavelengths for these lines using the remaining data, and compare the predicted wavelength to the assigned wavelength of each line.
 This inherently folds in errors in the measured line center of each calibration line, but this contribution to the residuals will be the same across all tests.
@@
 To assess the classic, polynomial-driven method of wavelength calibration, we take each LFC exposure and separate the lines into even- and odd-indexed lines.
-We then construct a wavelength solution using only the odd-indexed lines and use it to predict wavelengths for the even-indexed lines; i.e.
-a polynomial is fit to just the odd-indexed lines and then evaluated at the detector position of the even-indexed lines (see Equation \ref{eq:poly_wsol}).
+We then construct a wavelength solution using only the odd-indexed lines and use  that wavelength solution to predict wavelengths for the even-indexed lines; i.e.
+a polynomial is fit to just the odd-indexed lines and then evaluated at the detector positions of the even-indexed lines (see Equation \ref{eq:poly_wsol}).
 We then generate a wavelength solution using only the even-indexed lines and use it to predict wavelengths for the odd-indexed lines.
@@
 \begin{equation}
-RMS/line \: [m\,s^{-1}] = \sqrt{\sum_{n=1}^N\sum_{p=1}^P\frac{[ \frac{(\lambda_{n,p,predicted} - \lambda_{p,theoretical})}{\lambda_{p,theoretical}} \times c ]^2}{N \times P}}
+RMS/line \: [m\,s^{-1}] = \sqrt{\sum_{n=1}^N\sum_{p=1}^P\frac{[ \frac{(\lambda_{n,p,pred.} - \lambda_{p,theory})}{\lambda_{p,theory}} \times c ]^2}{N \times P}}
 \label{eq:rms}
 \end{equation}
-where $\lambda_{p,theoretical}$ is the theoretical wavelength for line $p$, $\lambda_{n,p,predicted}$ is the wavelength predicted by the constructed wavelength solution for line $p$ in exposure $n$, and residuals from all $P$ lines from all $N$ exposures are used, for a total of $N \times P$ lines.
+where $\lambda_{p,theory}$ is the theoretical wavelength for line $p$, $\lambda_{n,p,pred.}$ is the wavelength predicted by the constructed wavelength solution for line $p$ in exposure $n$, and residuals from all $P$ lines from all $N$ exposures are used, for a total of $N \times P$ lines.
 The difference in wavelength is converted to units of $\mps$, a more intuitive metric for \eprv\ work.
+
+Histograms of the per-line residuals for each of the above described polynomial, interpolation, and denoising tests (respectively) are shown in Figure \ref{fig:testHists}.
+Note that the spread in residuals is much smaller for both the denoising and interpolation tests relative to the results of the polynomial wavelength solution.
 
@@
 \centering
-\includegraphics[width=.6\textwidth]{Figures/all_results.pdf}
+\includegraphics[width=.45\textwidth]{Figures/all_results.pdf}
 \caption{Difference in predicted and theoretical wavelengths for the wavelength calibration tests described in Section \ref{sec:tests}.
@@
 
-Histograms of the per-line residuals for each of the above described polynomial, interpolation, and denoising tests (respectively) are shown in Figure \ref{fig:testHists}.
-Note that the spread in residuals is much smaller for both the denoising and interpolation tests relative to the results of the polynomial wavelength solution.
-
-\begin{figure}[t]
+\begin{figure}[h]
 \centering
-\includegraphics[width=0.55\textwidth]{Figures/lineResids2D_col.png}
+\includegraphics[width=0.455\textwidth]{Figures/lineResids2D_col.png}
 %\includegraphics[width=0.5\textwidth]{Figures/lineResids2D_col.pdf}
@@
 
+\begin{figure*}[t]
+\centering
+\includegraphics[width=\textwidth]{Figures/lineDensity.png}
+\caption{Residuals when using ThAr lines to predict wavelengths for LFC lines.
+\textbf{Right}: residuals for a single LFC exposure plotted with respect to detector position and colored by residual, as in Figure \ref{fig:resid2d}.
+The positions of ThAr lines are over-plotted in yellow.
+In general, residuals are greater between ThAr lines with greater separation.
+\textbf{Left}: Comparison of polynomial and interpolated wavelength solutions using either just ThAr lines or LFC lines for a subset of echelle order 94.
+The shape of the residuals from a polynomial fit are similar whether using ThAr lines or LFC lines.
+A PCHIP interpolated wavelength model guided by LFC lines returns the smallest residuals.}
+\label{fig:waveResids}
+\end{figure*}
+
 The residuals of the classic, polynomial wavelength solution is shown in the top plot of Figure \ref{fig:resid2d}.
@@
 This shows how the flexibility of an interpolated model can account for high-order instrument or detector defects, which emerged as structure in the residuals of the classic, smooth, polynomial-driven wavelength solution.
-This same flexibility may similarly allow interpolated wavelength solutions to account for position errors in pixel image blocks for different detectors \citep{fischer2016, milakovic2020}.
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=\textwidth]{Figures/lineDensity.png}
-\caption{Residuals when using ThAr lines to predict wavelengths for LFC lines.
-\textbf{Right}: residuals for a single LFC exposure plotted with respect to detector position and colored by residual, as in Figure \ref{fig:resid2d}.
-The positions of ThAr lines are over-plotted in yellow.
-In general, residuals are greater between ThAr lines with greater separation.
-\textbf{Left}: Comparison of polynomial-fit and interpolated wavelength solutions using either just ThAr lines or LFC lines for a subset of echelle order 94.
-The shape of the residuals from a polynomial fit are similar whether using ThAr lines or LFC lines, suggesting that little is added with the denser LFC lines.
-The LFC lines only serve to make the residuals more centered around zero.
-A PCHIP interpolated wavelength model guided by LFC lines returns the smallest residuals.}
-\label{fig:waveResids}
-\end{figure}
+This same flexibility may similarly allow interpolated wavelength solutions to account for position errors in pixel image blocks for different detectors depending on how the interpolation is framed \citep{fischer2016, milakovic2020}.
 
 Though the interpolated wavelength solution returns lower, less-structured residuals than the polynomial wavelength solution when guided by LFC lines, the flexibility of an interpolated wavelength solution can result in much worse residuals when not properly constrained, for example in regions between widely separated calibration lines.
-The left plot of Figure \ref{fig:waveResids} shows the residuals when ThAr calibration lines, which are much fewer and are less regular than LFC lines, are run through \name\ and used to predict wavelengths for the (completely independent) LFC exposures taken during the same range of time.
+The left plot of Figure \ref{fig:waveResids} shows the residuals when ThAr calibration lines, which are much fewer and less regularly-spaced than LFC lines, are run through \name\ and used to predict wavelengths for the (completely independent) LFC exposures taken during the same range of time.
 Over-plotted in yellow are the positions of the ThAr lines.
@@
 Figure \ref{fig:waveResids} plots the residuals for a subset of order 94 for both a polynomial-based method and a PCHIP-based method guided by either ThAr lines or LFC lines.
-The PCHIP model with ThAr lines (orange curve) returns huge residuals between two widely-separated ThAr lines that extends out of frame.
+The PCHIP model with ThAr lines (orange, dashed curve) returns huge residuals between two widely-separated ThAr lines that extends out of frame.
 The classic, polynomial fit exhibits similar residuals in both amplitude and shape regardless of whether the set of ThAr lines or LFC lines are used.
-An interpolated wavelength solution using LFC lines (black curve) exhibits the lowest residuals.
+An interpolated wavelength solution using LFC lines (black, solid curve) exhibits the lowest residuals.
 
@@
 The nightly-binned RVs are over-plotted as orange squares.
-The RMS of each data set is given in the top-right corner.}
+The RMS of the data set and the RMS of the binned values is given in the top-right corner.}
 \label{fig:rvs}
@@
 \target\ is a G0V star.
-It is a $4.8\, Gyr$ star and relatively quiet ($log R'_{HK} = -5.09$) \citep{brewer2016}.
-We used 114 observations of \target\ taken between Oct.
-8, 2019 and Aug.
-13, 2020 with SNR 250.
-Radial-velocity measurements were derived using the chunk-by-chunk, forward-modeling algorithm ran by the \expres\ team \citep{petersburg2020}.
+It is $4.8\, Gyr$ old and relatively quiet ($log R'_{HK} = -5.09$) \citep{brewer2016}.
+We use 114 observations of \target\ taken between October 8, 2019 and August 13, 2020 with SNR 250.
+Radial-velocity measurements were derived using a chunk-by-chunk, forward-modeling algorithm ran by the \expres\ team \citep{petersburg2020}.
 
@@
 This is equivalent to removing an independent, additive noise component of $0.62\, \mps \,(=\sqrt{1.22^2-1.05^2})$.
-A reduction in RMS was observed universally in all data sets tested; other data sets showed a reduction in scatter between $0.62-0.80\, \mps$.
+A reduction in RMS was observed universally in all data sets tested; other data sets showed a reduction in RMS between $0.2-0.5\, \mps$.
 
-\begin{deluxetable}{c c c}[ht!]
+\begin{deluxetable}{c c c}[t!]
 \tabletypesize{\scriptsize}
 \tablecaption{\expres\ RVs using \name\ Wavelengths \label{tab:rv}}
-\tablenum{2}
 \tablehead{
@@
 It is also possible, of course, to implement either step individually.
-A hierarchical framework can be used to simply denoise the lines before they are fit to a parametric model, or use a non-parametric model on lines that haven't been denoised.
+A hierarchical framework can be used to simply denoise the lines before they are fit to a parametric model, and a non-parametric model can be used on lines that have not been denoised.
 
-\lz{Hogg: please edit/add to these methods}
-
-For the dimensionality reduction and denoising, the PCA could be replaced by a probabilistic PCA model or other probabilistic linear reduction methods.
-It is also possible to move to non-linear reduction methods, like...?
+For the dimensionality reduction and denoising, the PCA could be replaced by a probabilistic PCA model or other probabilistic linear reduction methods, such as heteroscedastic matrix factorization (HMF)\citep{tsalmantza2012}.
+It is also possible to move to non-linear reduction methods, like an auto-encoder or a normalizing flow \citep[e.g.][]{kramer1991,woodbridge2020}
 
@@
 \subsection{Dimensionality of the Calibration Space, $K$}
-\label{sec:choice_k}
-The dimensionality of the calibration space within which the spectrograph linves is represented by $K$.
-In practice, it is the number of principal components, or basis vectors, used to reconstruct the de-noised line centers.
-$K$ needs to be large enough so that all variability in the spectrograph is captured.
-Too large, however, and the reconstruction will begin to incorporate noise, thereby defeating the purpose.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=.5\textwidth]{Figures/kvals_all.pdf}
-\caption{Per-line RMS of returned wavelength models for different values of $K$.
-Per-line RMS is defined in Equation \ref{eq:rms} and provides a measure of the accuracy of a wavelength model.
-There is a dotted, vertical line at $K=6$, our chosen $K$ value.
-The dotted, horizontal line is the RMS for $K=6$.
-The improvement at around $K=6$ plateaus.}
-\label{fig:kvals}
-\end{figure}
-
 \begin{figure*}[t]
@@
 
-In deciding a $K$ value, we ran denoising tests, as described in Section \ref{sec:tests}, for $K$ values spanning from 2 to 512.
-The resultant per-line RMS for each test is plotted in Figure \ref{fig:kvals}.
-One would expect the returned wavelengths to get increasingly more accurate with larger $K$ until components that represent only noise are incorporated.
-Residuals might then get worse before ultimately starting to get better again with large $K$, which marks the model over-fitting the data.
-Though the returned RMS never gets worse, we find that the improvement plateaus between $K=6$ and $K=128$.
-Comparisons of wavelengths returned by a $K=6$ model vs.
-a $K=32$ model show significant differences in less than 10 bluer lines, which are known to have greater error and variance in their measured line positions.
-We therefore settled on a $K$ value of six.
+\begin{figure}[b]
+\centering
+\includegraphics[width=.45\textwidth]{Figures/kvals_all.pdf}
+\caption{Per-line RMS of returned wavelength models for different values of $K$.
+Per-line RMS, as defined in Equation \ref{eq:rms}, provides a measure of the accuracy of a wavelength model.
+There is a dotted, vertical line at $K=6$, and a dotted, horizontal line is the RMS for $K=6$.
+The improvement at around $K=6$ plateaus.}
+\label{fig:kvals}
+\end{figure}
+\label{sec:choice_k}
+The dimensionality of the calibration space within which the spectrograph linves is represented by $K$.
+In practice, it is the number of principal components, or basis vectors, used to reconstruct the de-noised line centers.
+$K$ needs to be large enough so that all variability in the spectrograph is captured.
+Too large, however, and the reconstruction will begin to incorporate noise, thereby defeating the purpose.
 
@@
 
-\subsection{Interpolation of Calibration State to Science Exposures}
+In deciding a $K$ value, we ran denoising tests, as described in Section \ref{sec:tests}, for $K$ values spanning from 2 to 512.
+The resultant per-line RMS for each test is plotted in Figure \ref{fig:kvals}.
+One would expect the returned wavelengths to get increasingly more accurate with larger $K$ until components that represent only noise are incorporated.
+Residuals might then get worse before ultimately starting to get better again with large $K$, which marks when the model starts over-fitting the data.
+Though the returned RMS never gets worse, we find that the improvement plateaus between $K=6$ and $K=128$.
+Comparisons of wavelengths returned by a $K=6$ model vs.
+a $K=32$ model show significant differences in less than 10 bluer lines, which are known to have greater error and variance in their measured line positions.
+We therefore settled on a $K$ value of six.
+
 \label{sec:choice_avt}
-Figure \ref{fig:nightlyVariation} shows the amplitude of the first and second principal component with respect to time on the left.
-Though there exists a complex overall shape to the amplitudes with respect to time, a clear linear trend exists within each night.
-This is shown by the right plots in Figure  \ref{fig:nightlyVariation}.
-As the beginning-of-night and end-of-night calibration sets always include LFC exposures, we use a simple linear interpolation to interpolate principal component amplitudes with respect to time.
-
 \begin{figure*}[t]
@@
 In the top, right plot, the temperature of the optical bench is also plotted in orange.
-In the right plot, the principal component amplitudes have been artificially offset by the median amplitude for each night.
+In the right plots, the principal component amplitudes for each night have been artificially offset by the median amplitude.
 All days are therefore roughly on the same scale, but the y-axis is different from the left plots.
@@
 \end{figure*}
+
+\subsection{Interpolation of Calibration State to Science Exposures}
+Figure \ref{fig:nightlyVariation} shows the amplitude of the first and second principal component with respect to time on the left.
+Though there exists a complex overall shape to the amplitudes with respect to time, a clear linear trend exists within each night.
+This is shown by the right plots in Figure  \ref{fig:nightlyVariation}.
+As the beginning-of-night and end-of-night calibration sets always include LFC exposures, we use a simple linear interpolation to interpolate principal component amplitudes with respect to time.
 
@@
 
+\begin{figure*}[t]
+\centering
+\includegraphics[width=\textwidth]{Figures/intpx_tests.png}
+\caption{Residuals from different interpolation schemes over pixels in echelle order 100.
+ThAr lines, shown as blue circles, are used to construct a wavelength solution that is then evaluated at each LFC line, shown as blue lines.
+The residuals of each wavelength solution for a subset of the order is shown on the left.
+Histograms of the residuals for each method for the complete order is shown on the right.
+Here, the y-axis is zoomed in to show the shape of each histogram more clearly.
+Note: there is a blended ThAr line at approximately pixel 5300, the last ThAr line plotted.}
+\label{fig:xinterp}
+\end{figure*}
+
 \subsection{Interpolation of Wavelengths with Respect to Pixel}
 \label{sec:choice_wvp}
-In the implementation described here, interpolation of wavelengths over pixel  is done order-by-order using a Piecewise Cubic Hermite Interpolating Polynomial (PCHIP) interpolator.
+In the implementation described and tested by this paper, interpolation of wavelengths over pixel  is done order-by-order using a Piecewise Cubic Hermite Interpolating Polynomial (PCHIP) interpolator.
 This interpolation incorporates the flexibility needed to model the changing dispersion of the spectrograph across an echelle order along with any detector defects while also enforcing monotonicity, which we know must be true across any one echelle order.
@@
 
-\begin{figure*}[t]
-\centering
-\includegraphics[width=\textwidth]{Figures/intpx_tests.png}
-\caption{Residuals from different interpolation schemes over pixels in echelle order 100.
-ThAr lines, shown as blue circles, are used to construct a wavelength solution that is then evaluated at each LFC line, shown as blue lines.
-The residuals of each wavelength solution for a subset of the order is shown on the left.
-Histograms of the residuals for each method for the complete order is shown on the right.
-Here, the y-axis is zoomed in to show the shape of each histogram more clearly.
-Note: there is a blended ThAr line at approximately pixel 5300, the last ThAr line plotted.}
-\label{fig:xinterp}
-\end{figure*}
-
 These huge digressions can be avoided by allowing for some smoothing in the interpolation.
@@
 Monotonicity is a constraint we know must be true for any one echelle order.
-Though the PCHIP interpolator shows a similar issue as a classic cubic spline around the ThAr line blend at pixel 5300, the effect is much smaller and effects fewer pixels.
-Figure \ref{fig:xinterp} shows that using the PCHIP interpolator returns the lowest residuals.
+Though the PCHIP interpolator shows a similar issue as a classic cubic spline around the ThAr line blend at pixel 5300, the effect is much smaller and affects fewer pixels.
+Figure \ref{fig:xinterp}, right, shows that using the PCHIP interpolator returns the lowest residuals.
 
@@
 Implementing \name\ will require an existing library of calibration exposures that span the range of calibration space accessible by the spectrograph, or at least the space accessed by the science exposures being calibrated.
-A new set of calibration exposures would be needed for any changes to the instrument that could significantly change the calibration space.
+A new set of calibration exposures would be needed for any changes to the instrument that could significantly change the accessible calibration space.
 While there is no exact cutoff for how many exposures are needed, the number is certainly much larger than $K$, the dimensionality of the calibration space.
 More calibration exposures will help with denoising .
-The number of calibration exposures required throughout a night will depend on how much the instrument varies throughout the night, and the chosen interpolation scheme, as mentioned in Section \ref{sec:choice_avt}.
+The number of calibration exposures required throughout a night will depend on how much the instrument varies throughout the night as well as the chosen interpolation scheme, as mentioned in Section \ref{sec:choice_avt}.
 
@@
 \centering
-\includegraphics[width=0.5\textwidth]{Figures/lfcDensityTest.pdf}
+\includegraphics[width=0.45\textwidth]{Figures/lfcDensityTest.pdf}
 \caption{Per-line RMS as a function of spacing between lines in units of resolution element.
@@
 
-\lz{Hogg: The following paragraph might need some work.
-I struggled to keep this discussion general (i.e.
-not referring to our specific implementation of \name) while also being helpful.
-However, I might have failed or just made all this very confusing.
-What do you think?}
-
 If the FSR varies very little, its value can in principle be captured by a PCA or similar method.
@@
 If the FSR varies a lot, it becomes untenable to capture these changes with a PCA or similar method.
-Instead, the resultant basis vectors of the calibration space would have to be conditioned on the value of the FSR, becoming a non-trivial extension of \name.
+Instead, the resultant basis vectors of the calibration space would have to be conditioned on the value of the FSR.
+This would require a non-trivial extension of \name.
 
@@
 With PCA, this can be tested by plotting the amplitude of the returned principal components, which should fall quickly after a few components.
-However, this only provides a measure of the variance in the PCA space, and is not an explicit test of the end-to-end variation in the resulting model.
+It should be noted that this only provides a measure of the variance in the PCA space, and is not an explicit test of the end-to-end variation in the resulting model.
 This condition is therefore necessary but not sufficient if implementing \name\ with PCA.
@@
 Within such a framework, \name\ may still deliver good results.
-\lz{Hogg: My notes said you might want to add more things here?}
 
@@
 For example, PCA is good for detecting variability, but is agnostic to the source of the variability.
-This is why the principal components shown in Figure \ref{fig:pcLfc} exhibit errant values for bluer LFC lines, which are lower signal and therefore have more variable fitted line centers.
-It is essential that the line positions being given to \name\ capture only the changes in the spectrograph's calibration state, not potential errors in the fitted line centers.
+This is why the principal components shown in Figure \ref{fig:pcLfc} exhibit errant values for bluer LFC lines, which are lower signal and therefore exhibit more variation in their fitted line centers.
+It is essential that the line positions being fed to \name\ capture only the changes in the spectrograph's calibration state, not potential errors in the fitted line centers.
 
@@
 We show that \name\ returns a lower per-line RMS than classic, parametric methods by a factor of 5 (Section \ref{sec:tests}).
-The residuals were also smoother, showing less spatial correlation.
-Using \name\ wavelengths also reduced the RMS in a planet fit to derived RVs by \lz{some percent hopefully probs} (Section \ref{sec:test-rv}).
-In implementing \name\ on \expres\ data, we have successfully constructed a model of \expres, which truly is an instrument with low-degrees of freedom.
+The residuals were also smoother, exhibiting less spatial correlation.
+Using \name\ wavelengths reduced the RMS in RVs of HD 34411 from $1.22\, \mps$ to $1.05\, \mps$ (Section \ref{sec:test-rv}).
+
+In implementing \name\ on \expres\ data, we have successfully constructed a model of \expres, confirming that \expres\ truly is an instrument with low-degrees of freedom.
 \Name\ does not make any claims about what variability each basis vector represents.
@@
 
-Denser calibration lines allow us to move to more flexible wavelength models that can account for non-smooth features in the wavelength solution.
-Stabilized spectrograph hardware makes it more likely that the calibration state of the instrument is low-dimensional.
+Denser calibration lines allow us to move to more flexible wavelength models, which can then account for non-smooth features in the wavelength solution.
+Stabilized spectrograph hardware makes it more likely that the calibration space of the instrument is low-dimensional.
 All calibration images in a given generation of stability can therefore be used to constrain the accessible calibration space of the spectrograph as well as where in that calibration space the spectrograph lies for each calibration exposures.
@@
 
-An advantage of this implementation, where PCA is applied to line positions from all LFC exposures, is the ability to isolate exposures that exhibit errant variation, which is typically associated with flawed exposures.
+An advantage of this implementation of \name, where PCA is applied to line positions from all LFC exposures, is the ability to isolate exposures that exhibit errant variation, which is typically associated with flawed exposures.
 This allowed us to quickly vet for problematic LFC exposures, which otherwise would have required visual inspection of all 1200+ LFC exposures.
-In a classic framework where each calibration exposure is treated independently, these aberrant exposures would likely have persisted undetected and would likely have swayed the resultant wavelength solutions for at least an entire night.
+In a classic framework where each calibration exposure is treated independently, these aberrant exposures would likely have persisted undetected and are liable to sway the resultant wavelength solutions for at least an entire night.
 
 On the other hand, PCA captures all variance, regardless of source.
-Though \name\ endeavors to capture only variation in the instrument, the PCA is also sensitive to uncertainties and failures in the upstream line position fitting.
+Though \name\ endeavors to capture only variation in the instrument, the PCA is also sensitive to uncertainties and failures in the upstream line-position fitting.
 For example, we have seen that lower-signal lines that are harder to fit faithfully will have greater variety in returned line positions, which is in turn captured by the PCA.
@@
 
-\lz{Hogg: I'm still not sure I completely grasp the issue, can this be made clearer?}
 Along those lines, we caution that with any wavelength solution, there is a perfect degeneracy between what is defined as the ``position of the line'' and the resultant wavelength solution.
-If, for example, a cross correlation method is used to extract RVs from the data, a systematic difference may be introduced depending on the definition of what exactly is the line position, whether it be the center of a Gaussian fit, the edge of a complicated PSF, etc.
+If, for example, a cross correlation method is used to extract RVs from the data, a systematic difference may be introduced depending on what exactly is defined to be the line position, whether it be the mode of a Gaussian fit, the first moment of a complicated PSF, or the output of some particular peak-finding algorithm, etc.
 In principle, the best way to mitigate this uncertainty would be to use a calibration source that looks like a star.
@@
 Compared to traditional methods, which involve fitting high-order polynomials, \name\ has several useful statistical properties.
-\Name\ is technically robust to localized issues with either the calibration source or the pipeline used to return line positions.
+\Name\ is technically robust to localized issues that arise from either the calibration source or the pipeline used to return line positions.
 With an interpolated wavelength model, one errant line position will only effect the resultant wavelength model out to the close, neighboring lines.
 The effect of an outlier is diminished and kept localized.
-In contrast, an entire parametric fit will change from a single errant line, changing the wavelength solution for the entire exposure for a 2D fit.
+In contrast, an entire parametric fit will be affected by a single errant line, changing the wavelength solution for the entire exposure for a 2D fit.
 Through de-noising and outlier rejection, \name\ adds additional robustness against erroneous line positions.

### 2020-10-08 paper/paper.tex
Referee Report 1 Draft 1

@@
 
-\documentclass[twocolumn,table,xcolor]{aastex63}
+\documentclass[twocolumn,table,xcolor,trackchanges]{aastex63}
 \usepackage[ruled,vlined,norelsize]{algorithm2e}
@@
 These programs typically make use of spectrographs with resolutions on the order of $50,000-100,000$, which correspond to line widths on the order of $3000\,\mps$.
-The state of the art RV precision reached $1\, \mps$ in 2016.
-The newest generation of instruments aim  to reach $0.1\,\mps$ precision, the required precision to detect terrestrial worlds \citep{fischer2016}.
+The state of the art RV precision reached $1\, \mps$ in 2016 \added{\citep{fischer2016}}.
+The newest generation of instruments aim  to reach $0.1\,\mps$ precision, the required precision to detect terrestrial worlds.
 This requires new spectrographs to be calibrated or stabilized to better than $10^{-4}$ of a pixel (assuming that the spectrographs are well sampled).
@@
 The second simple idea follows from the observation that most physical systems have only a few dominant degrees of freedom, meaning most spectrographs vary along only a small number of axes in ``calibration space'', or the (very high-dimensional) space of all possible wavelength solutions.
-This is particularly true of \eprv\ instruments, which are equipped with stringent environmental stabilizing, likely reducing the number and extent of the axes along which the instrument can vary.
-That is, spectrographs, especially stabilized ones, should have few environmentally accessible degrees of freedom.
+\replaced{This is particularly true of \eprv\ instruments, which are equipped with stringent environmental stabilizing, likely reducing the number and extent of the axes along which the instrument can vary.}{This is particularly true of \eprv\ instruments, equipped with stringent environmental stabilizing.
+The thermomechanical stability of these instruments reduces the variations they experience to something that can be represented by a low-dimensional framework.}  That is, spectrographs, especially stabilized ones, should have few environmentally accessible degrees of freedom.
 This renders it inadvisable to fit each calibration exposure or calibrate each science exposure independently.
@@
 There are many ways to discern the basis functions.
-In this paper, we present one implementation of \expres\ using principal component analysis (PCA) \citep{wiki_pca}.
+In this paper, we present one implementation of \expres\ using principal component analysis (PCA) \added{\citep{pearson1901, jolliffe2016}}.
 A PCA is justifiable in the limit where there is very high signal-to-nose ratio, as is usually the case with typical calibration images.
@@
 There are many strategies for identifying calibration line positions and matching them to their assigned wavelengths; this problem is left out of scope for this work.
-That is, \name\ assumes that this has been done correctly.
-We caution, though, that systematic errors or large uncertainties in fitting line positions can propagate through to errors or biases in the wavelength models returned by \name.
+
+\replaced{That is, \name\ assumes that this has been done correctly.
+We caution, though, that systematic errors or large uncertainties in fitting line positions can propagate through to errors or biases in the wavelength models returned by \name.}{\Name\ assumes that line positions have been identified ``correctly,'' as in the position of a calibration line is determined the same way as the position of a stellar line when extracting RVs.
+This also inherently assumes that the calibration lines are not subject to an effect that the science exposures are not, for example differences in charge transfer inefficiency, non-linearities in the PSF, etc.
+We caution that systemic errors or large uncertainties in fitting line positions easily propagate through to biases in the wavelength models returned by \name.
+For more discussion, see Section \ref{sec:discussion}.}
 
@@
 	$g_0(\lambda,m) = \overline{x(\lambda,m,n)}$\;
-	find $U, \Sigma, V$ s.t. $U\Sigma V^* = (x(\lambda,m,n)-g_0(\lambda,m))$\;
+	\added{using Singular-Value Decomposition,} find $U, \Sigma, V$ s.t. $U\Sigma V^* = (x(\lambda,m,n)-g_0(\lambda,m))$\;
 	let $a_{n,k} = U\cdot \Sigma$ and $g_k(\lambda,m) = V$\;
@@
 In \name, the amplitude, $a_{n,k}$, of each principal component is interpolated to determine the calibration state of the spectrograph.
-For example, the amplitude can be interpolated with respect to time to recreate the calibration state of the spectrograph at different times.
+For example, the amplitude can be interpolated with respect to time to recreate the calibration state of the spectrograph at different times,\added{$t_{n'}$}.
 The choice of what to interpolate against depends on the dominant contribution to variation in the calibration of the instrument.
@@
 This is discussed more in Section \ref{sec:choice_avt}.
-These new, linearly interpolated magnitudes, $a_{n\,k}$ for a science exposure $n'$ at time $t_{n'}$, are then used to construct the calibration state of the spectrograph for that point in time.
-Using interpolated amplitudes, $a_{n,k}$, and the basis vectors, $g_k(\lambda,m)$, returned by the de-noising process, a new set of calibration lines, $x(\lambda,m,n')$ can be constructed
- for any exposure as formulated in Equation \ref{eq:excl_wsol}.
+\added{Let a prime denote values related to a science exposure $n'$ for which we want wavelengths.}  We use linearly interpolated magnitudes, $a_{n',k}$ at time $t_{n'}$ to construct the calibration state of the spectrograph for that point in time.
+Using interpolated amplitudes, $a_{n',k}$, and the basis vectors, $g_k(\lambda,m)$, returned by the de-noising process, a new set of calibration lines, $x'(\lambda,m,n')$ can be constructed for any exposure as formulated in Equation \ref{eq:excl_wsol}.
 
 \subsection{Interpolating a Wavelength Solution} \label{sec:interp_wsol}
-From the de-noising step, \name\ can now construct a set of calibration lines, $x(\lambda,m,n')$ for any exposure $n'$.
-To construct a wavelength solution, we invert $x(\lambda,m,n')$ to $\lambda(x,m,n')$ by interpolating known wavelengths of the calibration lines over detector position.
+From the de-noising step, \name\ can now construct a set of calibration lines, $x'(\lambda,m,n')$ for any exposure $n'$.
+To construct a wavelength solution, we invert $x'(\lambda,m,n')$ to $\lambda(x',m,n')$ by interpolating known wavelengths of the calibration lines over detector position.
 For instance, interpolating the known wavelengths vs.
@@
 After experiments, we found that a cubic-spline interpolation that enforces monotonicity, such as a Piecewise Cubic Hermite Interpolating Polynomial (PCHIP) interpolator, works well for interpolating wavelengths onto pixels.
-A cubic spline allows for more flexibility than a parameterized function, while the enforced monotonicity allows the wavelength solution, $\lambda(x,m,n')$, to be invertible and prevents spurious deviations that may befall a cubic spline.
+A cubic spline allows for more flexibility than a parameterized function, while the enforced monotonicity allows the wavelength solution, $\lambda(x',m,n')$, to be invertible and prevents spurious deviations that may befall a cubic spline.
 Choices in interpolation scheme, $K$, and other tests are further discussed in Section \ref{sec:choice_wvp}.
@@
 \KwData{the fiducial calibration of the spectrograph $g_0(\lambda,m)$; magnitudes of the principal components for each exposure $a_{n,k}$; basis vectors spanning the calibration space of the spectrograph $g_k(\lambda,m)$; }
-\KwResult{Wavelengths for detector positions $x'(m,n')$ of exposure $n'$ with time $t_{n'}$}
+\KwResult{Wavelengths for detector positions $x'(m,n')$ of exposure $n'$ with time $t_{n'}$, \added{where the prime is used to denote values relevant to this new exposure}}
 
 Find $a_{n',k}$ by interpolating $a_{n,k}$ with respect to $t_{n'}$\;
-$x(\lambda,m,n') = g_0(\lambda,m) + \sum_{k=1}^K a_{n'k}\,g_k(\lambda,m)$ where $K=6$\;
+$x'(\lambda,m,n') = g_0(\lambda,m) + \sum_{k=1}^K a_{n'k}\,g_k(\lambda,m)$ where $K=6$\;
 \For{each unique $m$}{
-	interpolate $\lambda$ with respect to $x(\lambda,m,n')$ onto pixels $x'(m,n')$\;
+	interpolate $\lambda$ with respect to $x'(\lambda,m,n')$ onto pixels $x'(m,n')$\;
 	}
@@
 
-In order to satisfy the assumption that there exists only low-order variation, which is needed for \name, we used exposures from after the photonic crystal fiber within the LFC was changed in Summer 2019, and the LFC stabilized.
+In order to \added{comfortably} satisfy the assumption that there exists only low-order variation, which is needed for \name, \replaced{we used exposures from after the photonic crystal fiber within the LFC was changed in Summer 2019, and the LFC stabilized}{we used exposures from after the LFC stabilized following servicing in summer 2019, where the photonic crystal fiber was replaced and the polarization was changed to shift the wavelength range of the LFC redwards}.
 In the results presented here, we use 1227 LFC exposures and 78 ThAr exposures taken between October 14 and December 18, 2019 on 29 unique nights.
@@
 
+\added{The polynomial and interpolation tests remove the same 50\% of lines from each exposure while the denoising test completely removes a randomly selected 10\% of calibration exposures and their associated line position measurements.
+Errors from interpolation will be localized, extending only to neighboring lines.
+We therefore aggressively remove every other line to ensure we are capturing these local effects.
+The PCA denoising, on the other hand, folds in information of all lines from all exposures.
+Here, it is sufficient to completely remove 10\% of exposures, a traditional training/validation fraction.
+Since the information being removed varies between each test depending on its focus, we present the results per line, treating each line as an independent test.}
+
 Resultant wavelengths are compared on a line-by-line basis.
@@
 
+\subsection{Results}
 Histograms of the per-line residuals for each of the above described polynomial, interpolation, and denoising tests (respectively) are shown in Figure \ref{fig:testHists}.
 Note that the spread in residuals is much smaller for both the denoising and interpolation tests relative to the results of the polynomial wavelength solution.
+
+\added{The per-line residuals from the denoising test also exhibit smaller spread than interpolation alone.
+This suggests that the spectrograph truly is accurately represented by a low-dimensional model, and recreating line positions using this model gives better line position estimates than treating each exposure independently.
+The low-dimensional model does not incorporate noise from individual line measurements.
+Returning more precise, denoised line positions results in smaller per-line residuals.}
 
@@
 \centering
-\includegraphics[width=0.455\textwidth]{Figures/lineResids2D_col.png}
-%\includegraphics[width=0.5\textwidth]{Figures/lineResids2D_col.pdf}
+\includegraphics[width=0.455\textwidth]{Figures/lineResids2D_col.pdf}
 \caption{Residuals of a single LFC exposure plotted with respect to detector position (as defined by echelle order and x-pixel) for parametric and non-parametric wavelength calibration methods.
@@
 \centering
-\includegraphics[width=\textwidth]{Figures/lineDensity.png}
+\includegraphics[width=\textwidth]{Figures/lineDensity.pdf}
 \caption{Residuals when using ThAr lines to predict wavelengths for LFC lines.
-\textbf{Right}: residuals for a single LFC exposure plotted with respect to detector position and colored by residual, as in Figure \ref{fig:resid2d}.
+\textbf{Left}: residuals for a single LFC exposure plotted with respect to detector position and colored by residual, as in Figure \ref{fig:resid2d}.
 The positions of ThAr lines are over-plotted in yellow.
 In general, residuals are greater between ThAr lines with greater separation.
-\textbf{Left}: Comparison of polynomial and interpolated wavelength solutions using either just ThAr lines or LFC lines for a subset of echelle order 94.
+\textbf{Right}: Comparison of polynomial and interpolated wavelength solutions using either just ThAr lines or LFC lines for a subset of echelle order 94.
 The shape of the residuals from a polynomial fit are similar whether using ThAr lines or LFC lines.
@@
 \centering
-\includegraphics[width=\textwidth]{Figures/rvs.pdf}
+\includegraphics[width=\textwidth]{Figures/dm9_rvs.pdf}
 \caption{HD 34411 RVs measured with \expres.
-RVs derived using the implementation of \name\ presented in this paper are plotted in blue.
-The nightly-binned RVs are over-plotted as orange squares.
-The RMS of the data set and the RMS of the binned values is given in the top-right corner.}
+All RVs are shown in blue; nightly binned RVs are over-plotted as orange squares.
+The RMS of the data set and the RMS of the binned values are given in the top-left corner.
+\textbf{Left}: RVs derived using a polynomial-based wavelength solution.
+\textbf{Right}: RVs derived using wavelengths from the implementation of \name\ presented in this paper.}
 \label{fig:rvs}
@@
 It is $4.8\, Gyr$ old and relatively quiet ($log R'_{HK} = -5.09$) \citep{brewer2016}.
-We use 114 observations of \target\ taken between October 8, 2019 and August 13, 2020 with SNR 250.
+We use 114 observations of \target\ taken between October 8, 2019 and March 5, 2020 with SNR 250.
 Radial-velocity measurements were derived using a chunk-by-chunk, forward-modeling algorithm ran by the \expres\ team \citep{petersburg2020}.
@@
 Figure \ref{fig:rvs} compares the resultant RVs when using a classic, ninth-degree polynomial wavelength solution and an \name -generated wavelength model.
-Using \name -generated wavelengths reduces the RMS of the entire data set from $1.22\, \mps$ with the classic wavelength solution to $1.05\, \mps$.
-This is equivalent to removing an independent, additive noise component of $0.62\, \mps \,(=\sqrt{1.22^2-1.05^2})$.
-A reduction in RMS was observed universally in all data sets tested; other data sets showed a reduction in RMS between $0.2-0.5\, \mps$.
+Using \name -generated wavelengths reduces the RMS of the entire data set from $1.17\, \mps$ with the classic wavelength solution to $1.05\, \mps$.
+This is equivalent to removing an independent, additive noise component of $0.52\, \mps \,(=\sqrt{1.17^2-1.05^2})$.
+\deleted{A reduction in RMS was observed universally in all data sets tested; other data sets showed a reduction in RMS between $0.2-0.5\, \mps$}.
 
@@
 For the dimensionality reduction and denoising, the PCA could be replaced by a probabilistic PCA model or other probabilistic linear reduction methods, such as heteroscedastic matrix factorization (HMF)\citep{tsalmantza2012}.
-It is also possible to move to non-linear reduction methods, like an auto-encoder or a normalizing flow \citep[e.g.][]{kramer1991,woodbridge2020}
+It is also possible to move to non-linear reduction methods, like an auto-encoder or a normalizing flow \citep[e.g.][]{kramer1991,woodbridge2020}.
+\added{Careful choice in the denoising model could enable \name\ to capture large-scale changes as well as small variations in calibration state.
+This would bridge data broken up by significant format changes (e.g.
+replacing hardware, opening the spectrograph chamber, etc.).}
 
@@
 \centering
-\includegraphics[width=\textwidth]{Figures/pcsLfc9.png}
-%\includegraphics[width=\textwidth]{Figures/pcsLfc6.pdf}
+\includegraphics[width=\textwidth]{Figures/pcsLfc9.pdf}
 \caption{\textbf{Top}:  The fiducial calibration of the spectrograph, i.e.
@@
 \centering
-\includegraphics[width=\textwidth]{Figures/pcAs_byDay.png}
+\includegraphics[width=\textwidth]{Figures/pcAs_byDay.pdf}
 \caption{Amplitude of the first two principal components shown as a function of time (left) or hour from midnight (right).
@@
 The choice in interpolation method can help guide how many wavelength calibration images are truly needed.
-It is unnecessary to take calibration images at times where the same information can be constructed by a well-chosen interpolation scheme.
+It is unnecessary to take calibration images at times where the same information can be \added{reconstructed at the desired precision} by a well-chosen interpolation scheme.
 For example, with the \expres\ data shown here, it is clear that nightly calibration images are needed, but for a linear trend, only two calibration images throughout the night are strictly required.
 
-We also tested an implementation of \name\ where the $K$ principal components within a night were fit to a cubic.
-This emulates the current, polynomial-based wavelength solution implemented in the \expres\ pipline, where polynomial fits to calibration files are interpolated to science exposure by fitting the polynomial coefficients to a cubic.
-We found that this method returns comparable RV RMS for most targets, though appears to do better when a night has sparse calibration data.
+We also tested an implementation of \name\ where the $K$ principal components within a night were fit to a cubic \added{with respect to time rather than linearly interpolated}.
+This emulates the current, polynomial-based wavelength solution implemented in the \expres\ pipeline, where polynomial fits to calibration files are interpolated to science exposure by fitting polynomial coefficients with respect to time to a cubic.
+We found that \replaced{this method}{using a cubic in place of linear interpolation} returns comparable RV RMS for most targets, though appears to do better when a night has sparse calibration data.
+\added{This suggests that the nightly behavior of \expres\ with respect to time is well described by a cubic function, but LFC exposures are typically taken with enough frequency that a linear interpolation provides a good approximation (see Figure \ref{fig:nightlyVariation}}
 
@@
 \centering
-\includegraphics[width=\textwidth]{Figures/intpx_tests.png}
+\includegraphics[width=\textwidth]{Figures/intpx_tests.pdf}
 \caption{Residuals from different interpolation schemes over pixels in echelle order 100.
@@
 Histograms of the residuals for each method for the complete order is shown on the right.
-Here, the y-axis is zoomed in to show the shape of each histogram more clearly.
-Note: there is a blended ThAr line at approximately pixel 5300, the last ThAr line plotted.}
+\deleted{Here, the y-axis is zoomed in to show the shape of each histogram more clearly.
+}Note: there is a blended ThAr line at approximately pixel 5300, the last ThAr line plotted.}
 \label{fig:xinterp}
@@
 
-A more classic cubic spline interpolation can run into issues with arc lines, which are irregularly spaced or even blended.
-Close lines appearing at virtually the same pixel location but with different wavelengths could coerce a cubic spline into a very high slope, as with the green curve in Figure \ref{fig:xinterp}.
-A line blend appears at approximately pixel 5300, causing the spline to twist to nearly vertical to account for both points.
+\added{Though less of an issue with LFC lines,} a more classic cubic spline interpolation can run into issues with arc lines, which are irregularly spaced or even blended.
+Close lines appearing at virtually the same pixel location but with different wavelengths could coerce a cubic spline into a very high slope.
+\added{This is demonstrated by the green line in Figure \ref{fig:xinterp}, which shows the results of interpolating between ThAr lines rather than LFC lines.}  A line blend appears at approximately pixel 5300, causing the spline to twist to nearly vertical to account for both points.
 This leads to huge deviations from the correct wavelengths around this line blend as the extraneously high slope of the spline is accounted for.
@@
 
-As an empirical assessment of the needed line density, we systematically removed LFC lines from the \expres\ data and calculated the per-line RMS of the returned wavelengths for the removed lines.
+As an empirical assessment of the needed line density, we removed LFC lines from the \expres\ data and calculated the per-line RMS of the returned wavelengths for the removed lines.
+\added{These tests are similar to the interpolation test explained in Section \ref{sec:tests}, in that a fraction of lines are systematically removed from the analysis.
+An increasing fraction of lines are removed to simulate different line densities.
+For these line density tests, though, we are also implementing denoising unlike the pure interpolation test of Section \ref{sec:tests}.}
+
 Figure \ref{fig:density} plots the scatter of residuals of the wavelengths returned by \name\ as a function of separation between lines in units of resolution element.
@@
 The result is likely to be even more precise, as this method incorporates more data.
+\added{This method, as with all analysis involving a simultaneous reference fiber, will work only as well as the reference fiber's ability to trace changes in the main science fiber.}
 
@@
 The etalon data can be interpolated onto a set of fiducial lines with set wavelengths.
-These fiducial lines would therefore be identifiable by echelle order and wavelength alone, with only their line positions varying with different exposures, thereby returning to the same framework as developed for the case of an LFC.
+These fiducial lines would therefore be identifiable by echelle order and wavelength alone, with only their line positions varying with different exposures.
+This returns us to the same framework as developed for the case of an LFC.
 This marks the simplest implementation of \name\ on etalon data, as the uncertainty of a line's wavelength is upstream of the model rather than built in.
 
-If the FSR varies very little, its value can in principle be captured by a PCA or similar method.
+\deleted{If the FSR varies very little, its value can in principle be captured by a PCA or similar method.
 However, to interpolate this information from calibration exposures to science exposures would necessitate a simultaneous reference or interpolation with respect to some other housekeeping data that contains information on the value of the FSR.
@@
 Instead, the resultant basis vectors of the calibration space would have to be conditioned on the value of the FSR.
-This would require a non-trivial extension of \name.
+This would require a non-trivial extension of \name.}
+
+\added{Incorporating the FSR as part of the \name\ model will require introducing a free parameter to capture changes in the FSR independent of variation in an instrument's calibration state.
+The calibration state can then be described with respect to mode number, which will be used to uniquely identify a calibration line across exposures rather than wavelength.
+The FSR is then used to determine how the mode number of each line maps to wavelength for a given exposure.
+The FSR must not vary so much that the change in this mode-number-to-wavelength mapping becomes non-linear.
+This model would require a simultaneous reference or other housekeeping data that can be used to determine the FSR for every exposure.}
 
@@
 Homogeneity is, of course, implicitly required when implementing PCA.
+\added{Different denoising models will be able to account for different amounts of stability or lack thereof.}
 
@@
 If \name\ was updated to a full probabilistic model, the PCA along with the interpolation model would have to be upgraded to something with better probabilistic properties.
-Further discussion of other implementations of \name\ can be found in Section \ref{sec:choices}.
+\added{Other denoising methods may be more robust to large changes, allowing all calibration images ever taken with an instrument to be used to construct the accessible calibration state regardless of hardware adjustments.}  Further discussion of other implementations of \name\ can be found in Section \ref{sec:choices}.
 
@@
 Furthermore, all exposures taken with the spectrograph in principle contains information about the calibration state of the spectrograph.
-Theoretically, tellurics, lines in science exposures, or just the trace positions themselves could also be used to determine an instrument's calibration state, negating the need for any additional wavelength-calibration source beyond defining the calibration space.
+Theoretically, tellurics, lines in science exposures, or just the trace positions themselves could also be used to determine an instrument's calibration state, \replaced{negating the need for any additional wavelength-calibration source beyond defining the calibration space.}{thereby providing free simultaneous calibration information.}
 

### 2020-10-19 paper/paper.tex
Final(-ish) Referee Response 1

@@
 It is designed with the needs of extreme-precision radial velocity (\eprv) in mind, which require that instruments be calibrated or stabilized to better than $10^{-4}$ pixels.
-Instruments vary along only a few dominant degrees of freedom, especially \eprv\ instruments, which feature highly stabilized optical systems and detectors.
+Instruments vary along only a few dominant degrees of freedom, especially \eprv\ instruments that feature highly stabilized optical systems and detectors.
 \Name\ takes advantage of this property by using all calibration data to construct a low-dimensional representation of all accessible calibration states for an instrument.
-\Name\ also takes advantage of laser-frequency combs or etalons, which generate a dense set of stable calibration points.
-This density permits the use of a non-parametric wavelength solution that can adapt to any instrument or detector oddities better than any parametric model, such as a polynomial.
-We demonstrate the success of this method with data from the \textsl{EXtreme PREcision Spectrograph} (\expres), which includes a laser frequency comb.
+\Name\ also takes advantage of laser frequency combs or etalons, which generate a dense set of stable calibration points.
+This density permits the use of a non-parametric wavelength solution that can adapt to any instrument or detector oddities better than parametric models, such as a polynomial.
+We demonstrate the success of this method with data from the \textsl{EXtreme PREcision Spectrograph} (\expres), which uses a laser frequency comb.
 When wavelengths are assigned to laser comb lines using \name, the RMS of the residuals is about five times lower than wavelengths assigned using polynomial fits to individual exposures.
-In a typical test case, radial-velocity measurements of HD~34411 showed a reduction in RMS scatter over a 10-month time baseline from $1.22$ to $1.05\, \mps$.
+Radial-velocity measurements of HD~34411 showed a reduction in RMS scatter over a 10-month time baseline from $1.22$ to $1.05\, \mps$.
 \end{abstract}
@@
 \section{Introduction} 
-Precise, radial-velocity (\eprv) programs have been fruitful in finding and characterizing extra-solar planets \citep[e.g.][]{mayor2011, bonfils2013, plavchan2015, butler2017}.
+Precise, radial-velocity programs have been fruitful in finding and characterizing extra-solar planets \citep[e.g.][]{mayor2011, bonfils2013, plavchan2015, butler2017}.
 These programs typically make use of spectrographs with resolutions on the order of $50,000-100,000$, which correspond to line widths on the order of $3000\,\mps$.
@@
 
-Traditionally, wavelength solutions are constructed by fitting a polynomial to lines from a calibration source to describe the relationship between wavelength and pixel for each echelle order \citep{butler1996, lovis2007, cersullo2019}.
+Traditionally, wavelength solutions are constructed by fitting a polynomial to lines from a calibration source in order to describe the relationship between wavelength and pixel for each echelle order \citep{butler1996, lovis2007, cersullo2019}.
 In this framework, each calibration image is treated independently.
@@
 There has been significant effort in using an entire set of calibration images to identify incongruous ThAr lines \citep{coffinet2019} or obtain high-resolution Fourier transform spectra of reference cells \citep{wang2020}.
-It has also been found that using a segmented polynomial in the dispersion direction, tuned to capture detector defects, better describes the wavelength solution than a single, continuous polynomial \citep{milakovic2020}.
+It has also been found that using multiple polynomials in the dispersion direction, tuned to capture detector defects, better describes the wavelength solution than a single, continuous polynomial \citep{milakovic2020}.
 
 Here, we propose to simplify and improve calibration programs for \eprv\ hardware systems with two practical yet innovative ideas.
-The first flows from the fact that calibration sources---which include arc lamps (in some wavelength ranges), etalons, and laser-frequency combs (\lfc s)---illuminate the spectrograph with very stable, very dense sets of lines; almost every location in the spectrograph image plane is surrounded by nearby, useful calibration lines.
-This enables use of a calibration methodology that is \emph{non-parametric}, or not defined by a prescribed, analytic function described by a finite number of parameters :  If every point in the spectrograph detector is sufficiently surrounded by nearby calibration lines, the wavelength solution can, for example, be made simply as an interpolation of the calibration data.
-The density of lines removes the need to enforce any functional form for the wavelength solution (such as a continuous ninth-order polynomial, for example).
-In some ways, this is a generalization of recent work, which has demonstrated the efficacy of constructing a wavelength solution as multiple, segmented polynomials \citep{milakovic2020}.
+The first flows from the fact that calibration sources---which include arc lamps (in some wavelength ranges), etalons, and laser-frequency combs (\lfc s)---illuminate the spectrograph with very stable, very dense sets of lines--almost every location in the spectrograph image plane is surrounded by nearby, useful calibration lines.
+This enables use of a calibration methodology that is \emph{non-parametric}, or not defined by a prescribed, analytic function described by a finite number of parameters:  If every point in the spectrograph detector is sufficiently surrounded by nearby calibration lines, the wavelength solution can, for example, be made simply as an interpolation of the calibration data.
+The density of lines removes the need to enforce any functional form for the wavelength solution (such as a continuous ninth-order polynomial).
+In some ways, this is a generalization of recent work that has demonstrated the efficacy of constructing a wavelength solution as multiple, segmented polynomials \citep{milakovic2020}.
 A non-parametric approach will improve calibration accuracy by not forcing the choice of a parametric form that may bias the calibration, especially when the chosen function is inappropriate (as, for example, polynomials are at detector edges).
@@
 The second simple idea follows from the observation that most physical systems have only a few dominant degrees of freedom, meaning most spectrographs vary along only a small number of axes in ``calibration space'', or the (very high-dimensional) space of all possible wavelength solutions.
-\replaced{This is particularly true of \eprv\ instruments, which are equipped with stringent environmental stabilizing, likely reducing the number and extent of the axes along which the instrument can vary.}{This is particularly true of \eprv\ instruments, equipped with stringent environmental stabilizing.
-The thermomechanical stability of these instruments reduces the variations they experience to something that can be represented by a low-dimensional framework.}  That is, spectrographs, especially stabilized ones, should have few environmentally accessible degrees of freedom.
+This is particularly true of \eprv\ instruments, which are equipped with stringent environmental stabilizing.
+\replaced{ likely reducing the number and extent of the axes along which the instrument can vary.}{The thermomechanical stability of these instruments reduces the variations they experience to something that can be represented by a low-dimensional framework.}  That is, spectrographs, especially stabilized ones, should have few environmentally accessible degrees of freedom.
 This renders it inadvisable to fit each calibration exposure or calibrate each science exposure independently.
@@
 In statistics, this concept is often described as \emph{de-noising}:  we can improve calibration by recognizing that every calibration exposure contains information about every other calibration exposure.
-Thus every exposure can be improved (i.e., de-noised) with information from every other exposure.
+Thus, every exposure can be improved (i.e., de-noised) with information from every other exposure.
 
@@
 By being non-parametric, it delivers enormous freedom to the wavelength solution to match or adapt to any instrument or detector oddities.
-By being hierarchical, it restricts that freedom tremendously, but it does so appropriately for the empirically determined variations in the stabilized spectrograph.
+By being hierarchical, it restricts that freedom tremendously, but it does so appropriately for the empirically determined variations in a spectrograph.
 
@@
 The \name\ method is designed to take many calibration images, each containing a series of calibration lines with known wavelengths and well-fit detector positions, and de-noise and interpolate this information into a full wavelength model applicable to all exposures taken with the instrument.
-It operates on the two core ideas; The wavelength solution should be allowed flexibility, but it lives in a very low-dimensional calibration space, where the degrees of freedom are set by the limited kinematics of the spectrograph hardware.
-\Name\ therefore assumes that the space of possible calibration states for an instrument is low-dimensional while assuming very little about the forms of those states.
+It operates on two core ideas; The wavelength solution should be allowed flexibility, but it lives in a very low-dimensional calibration space where the degrees of freedom are set by the limited kinematics of the spectrograph hardware.
+\Name\ therefore assumes that the space of possible calibration states for an instrument is low-dimensional but assumes very little about the forms of those states.
 
-\Name\ assumes dense-enough calibration line coverage with well-fit line centers to provide sufficient constraints on an interpolated wavelength solution across an echelle order.
+\Name\ also assumes dense-enough calibration line coverage with well-fit line centers to provide sufficient constraints on an interpolated wavelength solution across an echelle order.
 Upstream errors in line center positions may propagate through \name\ wavelength models.
@@
 The line position on the detector is therefore uniquely identified by echelle order $m$ and pixel in the dispersion direction, $x$.
-The coefficients $c_{nij}$ are interpolated from the time of calibration exposures to time $t_n$ of a science exposure $n$ by a third-order polynomial of time.
+The coefficients $c_{nij}$ are interpolated from the time of calibration exposures to time $t_n$ of a science exposure $n$ by a third-order polynomial with respect to time.
 This third-order polynomial is evaluated at the time of non-calibration, science exposures to re-construct the coefficients for a 2D, 9th-order polynomial wavelength solution for that exposure.
@@
 $g_0(\lambda,m)$ is the fiducial or mean or standard calibration of the spectrograph,
-the $a_{nk}$ are $K$ scalar amplitudes for each exposure $n$,
-and the $g_k(\lambda,m)$ are basis functions expressing the ``directions'' in calibration space that the spectrograph can depart from the fiducial calibration.
+the $a_{nk}$ are $K$ scalar amplitudes for each exposure $n$, and the $g_k(\lambda,m)$ are basis functions expressing the ``directions'' in calibration space that the spectrograph can depart from the fiducial calibration.
 The resultant $x(\lambda,m,n)$, a list of calibration line positions for a given exposure, can be regarded as the calibration state of the spectrograph for that exposure.
-When this calibration structure is used to deliver a wavelength solution, $x(\lambda,m,n)$ can be inverted back into $\lambda(x,m,n)$ to recover wavelengths for each detector position $x$ and echelle order $m$.
+When this calibration structure is used to deliver a wavelength solution, $x(\lambda,m,n)$ can be inverted back into $\lambda(x,m,n)$ to recover wavelengths for each detector position $x$ and echelle order $m$ (see Figure \ref{fig:cartoon}).
 
@@
 There are many ways to discern the basis functions.
-In this paper, we present one implementation of \expres\ using principal component analysis (PCA) \added{\citep{pearson1901, jolliffe2016}}.
-A PCA is justifiable in the limit where there is very high signal-to-nose ratio, as is usually the case with typical calibration images.
+In this paper, we present one implementation of \name\ using principal component analysis (PCA) \added{\citep{pearson1901, jolliffe2016}}.
+A PCA is justifiable in the limit where exposures have very high signal-to-nose ratio, as is usually the case with typical calibration images.
 There are many alternatives to PCA for this dimensionality reduction; we return to this point in Section \ref{sec:choices} below.
@@
 \Name\ will use calibration images to determine 1) the space in which an instrument varies and 2) where in the accessible calibration space the spectrograph existed for each exposure.
-For each calibration exposure, $n$, \name\ requires a full list of lines, $(\lambda,m)$ that are expected in each calibration exposure.
+For each calibration exposure, $n$, \name\ requires a full list of lines, $(\lambda,m)$ that are expected to appear in each calibration exposure.
 Each line is uniquely defined by a combination of echelle order, $m$, and ``true'' or theoretical wavelength, $\lambda$.
@@
 Fitted line centers that are missing from an exposure (e.g.
-because the fit failed due to noise, the line is not in its usual echelle order, etc.) can be assigned a \code{NaN} (hardware not-a-number) for that echelle order instead.
+because the fit failed due to noise, the line is not in its usual echelle order, etc.) can be assigned a \code{NaN} (hardware not-a-number) for that exposure instead.
 Let there be $P$ lines per exposure.
@@
 \KwData{line positions $x(\lambda,m,n)$ for each exposure $n$, with wavelengths $\lambda$ and echelle orders $m$}
-\KwResult{Basis vectors of the low-dimensional calibration space $g_k(\lambda,m)$, location of exposures in calibration space expressed by amplitudes $a_{n,k}$}
+\KwResult{Basis vectors of the low-dimensional calibration space $g_k(\lambda,m)$ and location of exposures in calibration space expressed by amplitudes $a_{n,k}$}
 \While{change in missing or outlier line centers $>$ 0.01\%}
@@
 In \name, the amplitude, $a_{n,k}$, of each principal component is interpolated to determine the calibration state of the spectrograph.
-For example, the amplitude can be interpolated with respect to time to recreate the calibration state of the spectrograph at different times,\added{$t_{n'}$}.
+For example, the amplitude can be interpolated with respect to time to recreate the calibration state of the spectrograph at different times.
 The choice of what to interpolate against depends on the dominant contribution to variation in the calibration of the instrument.
@@
 Choices in interpolation scheme, $K$, and other tests are further discussed in Section \ref{sec:choice_wvp}.
+\added{Our implementation of \name\ is hosted on GitHub\footnote{\url{https://www.github.com/davidwhogg/EPRVCalibration}}.}
 
@@
 Similarly, exposures with more than 60\% of expected lines missing were cut from the analysis.
-A list of echelle orders $m$, line wavelengths $\lambda$, and pixel positions $x$, which have been calculated by the EXPRES pipeline \citep{petersburg2020} for every line of every exposure, is read into \name.
+A list of echelle orders $m$, line wavelengths $\lambda$, and pixel positions $x$ were calculated by the EXPRES pipeline \citep{petersburg2020} for every line of every exposure and read into \name.
 
@@
 
+\begin{figure}[t]
+\centering
+\includegraphics[width=.45\textwidth]{Figures/all_results.pdf}
+\caption{Difference in predicted and theoretical wavelengths for the wavelength calibration tests described in Section \ref{sec:tests}.
+The per line RMS as defined in Equation \ref{eq:rms} is given in the top-right corner in each method's corresponding color.
+Incorporating denoising returns the smallest spread in residuals.}
+\label{fig:testHists}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.455\textwidth]{Figures/lineResids2D_col.pdf}
+\caption{Residuals of a single LFC exposure plotted with respect to detector position (as defined by echelle order and x-pixel) for parametric (top) and non-parametric (bottom) wavelength calibration methods.
+Each line is colored by the difference between the predicted wavelength and the theoretical wavelength for each line, given in units of $\mps$.
+High-order structure, i.e.
+vertical stripes and patchiness, is apparent in the residuals to a polynomial wavelength solution, which assumes smoothness.}
+\label{fig:resid2d}
+\end{figure}
+
 To test the denoising step of \name\ (\textsection \ref{sec:denoising}, \textsection\ref{sec:interp_time}), we employed \name\ on a randomly selected 90\% of all LFC exposures.
@@
 
-Resultant wavelengths are compared on a line-by-line basis.
 The residuals of a wavelength solution represent the difference between the wavelength solution evaluated at the line position of a calibration line, and the assigned theoretical wavelength (i.e.
-from Equation \ref{eq:lfc} for LFC lines and from a line atlas for ThAr lines) for each line in every exposure.
+from Equation \ref{eq:lfc} for LFC lines) on a line-by-line basis in every exposure.
 The reported RMS of a wavelength solution is therefore the per-line RMS, i.e.
@@
 \added{The per-line residuals from the denoising test also exhibit smaller spread than interpolation alone.
-This suggests that the spectrograph truly is accurately represented by a low-dimensional model, and recreating line positions using this model gives better line position estimates than treating each exposure independently.
+This suggests that the spectrograph truly is accurately represented by a low-dimensional model.
+Recreating line positions using this model gives better line position estimates than treating each exposure independently.
 The low-dimensional model does not incorporate noise from individual line measurements.
 Returning more precise, denoised line positions results in smaller per-line residuals.}
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=.45\textwidth]{Figures/all_results.pdf}
-\caption{Difference in predicted and theoretical wavelengths for the wavelength calibration tests described in Section \ref{sec:tests}.
-The per line RMS as defined in Equation \ref{eq:rms} is given in the top-right corner in each method's corresponding color.
-\Name\ returns the smallest spread in residuals.}
-\label{fig:testHists}
-\end{figure}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.455\textwidth]{Figures/lineResids2D_col.pdf}
-\caption{Residuals of a single LFC exposure plotted with respect to detector position (as defined by echelle order and x-pixel) for parametric and non-parametric wavelength calibration methods.
-Each line is colored by the difference between the predicted wavelength and the theoretical wavelength for each line, given in units of $\mps$.
-\textbf{Top}: residuals to a classic, polynomial fit.
-\textbf{Bottom}: residuals to an interpolated wavelength solution.
-High-order structure, i.e.
-vertical stripes and patchiness, is apparent in the residuals to a polynomial wavelength solution, which assumes smoothness.}
-\label{fig:resid2d}
-\end{figure}
 
@@
 
-Figure \ref{fig:waveResids} plots the residuals for a subset of order 94 for both a polynomial-based method and a PCHIP-based method guided by either ThAr lines or LFC lines.
+Figure \ref{fig:waveResids} (right) plots the residuals for a subset of order 94 for both a polynomial-based method and a PCHIP-based method guided by either ThAr lines or LFC lines.
 The PCHIP model with ThAr lines (orange, dashed curve) returns huge residuals between two widely-separated ThAr lines that extends out of frame.
@@
 \caption{HD 34411 RVs measured with \expres.
+We zoom in on all but 8 of the exposures, which were taken over 150 days later.
 All RVs are shown in blue; nightly binned RVs are over-plotted as orange squares.
@@
 the mean line positions for each line throughout the epoch of stability.
-The following $3 \times 3$ grid of plots show the first nine principal components constructed using LFC lines.
+\textbf{The following $3 \times 3$ grid of plots} show the first nine principal components constructed using LFC lines.
 These principal components represent the basis vectors along which the calibration of the spectrograph can deviate from the fiducial calibration.
@@
 \label{sec:choice_k}
-The dimensionality of the calibration space within which the spectrograph linves is represented by $K$.
+The dimensionality of the calibration space within which the spectrograph lives is represented by $K$.
 In practice, it is the number of principal components, or basis vectors, used to reconstruct the de-noised line centers.
@@
 In the top, right plot, the temperature of the optical bench is also plotted in orange.
-In the right plots, the principal component amplitudes for each night have been artificially offset by the median amplitude.
+In the right plots, the principal component amplitudes for each night have been artificially offset by the median amplitude per night.
 All days are therefore roughly on the same scale, but the y-axis is different from the left plots.
@@
 Best results will come from interpolating with respect to whatever is most strongly correlated with the calibration state of the spectrograph.
-For example, with \expres, which is passively temperature controlled, the returned amplitudes $a_{n,k}$ were extremely correlated with the optical bench temperature, as shown in Figure \ref{fig:nightlyVariation}, suggesting it would also be possible to interpolate the amplitudes with respect to temperature.
+For example, with \expres, which is passively temperature controlled, the returned amplitudes $a_{n,k}$ were extremely correlated with the optical bench temperature, as shown in the top-left plot of Figure \ref{fig:nightlyVariation}, suggesting it would also be possible to interpolate the amplitudes with respect to temperature.
 
@@
 \caption{Residuals from different interpolation schemes over pixels in echelle order 100.
-ThAr lines, shown as blue circles, are used to construct a wavelength solution that is then evaluated at each LFC line, shown as blue lines.
+ThAr lines, shown as blue circles, are used to construct a wavelength solution that is then evaluated at each LFC line, shown as blue vertical lines.
 The residuals of each wavelength solution for a subset of the order is shown on the left.
@@
 \deleted{Here, the y-axis is zoomed in to show the shape of each histogram more clearly.
-}Note: there is a blended ThAr line at approximately pixel 5300, the last ThAr line plotted.}
+}Note: there is a blended ThAr line at approximately pixel 5300, the right most ThAr lines plotted.}
 \label{fig:xinterp}
@@
 A simple linear interpolation would give erroneously low values everywhere.
-Due to the dispersion intrinsic to echelle spectrographs, the wavelength change between pixels grows greater with greater wavelengths.
+Due to the dispersion intrinsic to echelle spectrographs, the wavelength change between pixels grows greater with greater wavelengths, \added{even within an order}.
 This means that the function of wavelength vs.
@@
 
-We instead turn to the PCHIP interpolator, which damps down huge deviations in the traditional cubic spline by requiring the resulting interpolated function be monotonic.
+We instead turn to the PCHIP algorithm, which damps down huge deviations in the traditional cubic spline by requiring the resulting interpolated function be monotonic.
 Monotonicity is a constraint we know must be true for any one echelle order.
 Though the PCHIP interpolator shows a similar issue as a classic cubic spline around the ThAr line blend at pixel 5300, the effect is much smaller and affects fewer pixels.
-Figure \ref{fig:xinterp}, right, shows that using the PCHIP interpolator returns the lowest residuals.
+Figure \ref{fig:xinterp}, right, shows that using the PCHIP interpolator returns the lowest spread residuals.
 
@@
 \expres\ is representative of the newest generation of \eprv\ spectrographs, and an LFC provides stable, dense calibration lines with known wavelengths, ideal for \name.
-The applicability of \name\ to any one instrument is a detailed question of the kind of variance experienced by the spectrograph and calibration sources available, but we hope to provide some approximate benchmarks and diagnostics.
+The applicability of \name\ to any one instrument is a detailed question of the kind of variance experienced by the spectrograph and calibration sources available, but we hope to provide some approximate benchmarks and diagnostics here.
 
@@
 Additionally, in this assessment, the lines remain uniformly spaced.
-The needed line density depends on the resolution of a given spectrograph.
-the needed precision of the returned wavelength solution, and the chosen interpolation scheme.
+The needed line density depends on the resolution of a given spectrograph, the needed precision of the returned wavelength solution, and the chosen interpolation scheme.
 
@@
 temperature, pressure, time, etc.) that correlates with the calibration.
-For example, we have seen that the calibration state of \expres\ is correlated with the temperature of the optical bench, even with the bench varying by less than one degree.
+For example, we have seen that the calibration state of \expres\ is correlated with the temperature of the optical bench, even with the optical bench temperature varying by less than one degree.
 This correlation is not seen in the RVs, suggesting the changes with temperature are calibrated out at the wavelength stage.
@@
 
-It could still be possible to run \name\ on a spectrograph that has a high-dimensional calibration space, meaning a large number of basis vectors is required to capture all the variance in the spectrograph.
+It could still be possible to run \name\ on a spectrograph that has a high-dimensional calibration space, meaning a large number of basis vectors are required to capture all the variance in the spectrograph.
 In this regime, there is always the risk of over-fitting.
@@
 We show that \name\ returns a lower per-line RMS than classic, parametric methods by a factor of 5 (Section \ref{sec:tests}).
-The residuals were also smoother, exhibiting less spatial correlation.
-Using \name\ wavelengths reduced the RMS in RVs of HD 34411 from $1.22\, \mps$ to $1.05\, \mps$ (Section \ref{sec:test-rv}).
+The residuals were also smoother, exhibiting less spatial correlation (Figure \ref{fig:resid2d}).
+Using \name\ wavelengths reduced the RMS in RVs of HD 34411 from $1.17\, \mps$ to $1.05\, \mps$ (Section \ref{sec:test-rv}).
 
-In implementing \name\ on \expres\ data, we have successfully constructed a model of \expres, confirming that \expres\ truly is an instrument with low-degrees of freedom.
+In implementing \name\ on \expres\ data, we have successfully constructed a model of \expres's accessible calibration space, confirming that \expres\ truly is an instrument with low-degrees of freedom.
 \Name\ does not make any claims about what variability each basis vector represents.
@@
 
-The endeavor to reach $0.1\,\mps$ radial-velocity precision necessitates improvements on all levels of spectroscopy, from the instrumentation to the data analysis.
-Implementing \name, a hierarchical, non-parametric model for wavelength calibration, is a step towards reducing the error contribution from data reduction pipelines.
+\deleted{The endeavor to reach $0.1\,\mps$ radial-velocity precision necessitates improvements on all levels of spectroscopy, from the instrumentation to the data analysis.
+Implementing \name, a hierarchical, non-parametric model for wavelength calibration, is a step towards reducing the error contribution from data reduction pipelines.}
 
@@
 Stabilized spectrograph hardware makes it more likely that the calibration space of the instrument is low-dimensional.
-All calibration images in a given generation of stability can therefore be used to constrain the accessible calibration space of the spectrograph as well as where in that calibration space the spectrograph lies for each calibration exposures.
+All calibration images in a given generation of stability can therefore be used to constrain the accessible calibration space of the spectrograph as well as where in that calibration space the spectrograph lies.
 We have described only one, fairly simplistic implementation of \name\ here.

### 2020-10-24 paper/paper.tex
Submitted Referee Report

@@
 When wavelengths are assigned to laser comb lines using \name, the RMS of the residuals is about five times lower than wavelengths assigned using polynomial fits to individual exposures.
-Radial-velocity measurements of HD~34411 showed a reduction in RMS scatter over a 10-month time baseline from $1.22$ to $1.05\, \mps$.
+Radial-velocity measurements of HD~34411 showed a reduction in RMS scatter over a 10-month time baseline from $1.17$ to $1.05\, \mps$.
 \end{abstract}
@@
 These programs typically make use of spectrographs with resolutions on the order of $50,000-100,000$, which correspond to line widths on the order of $3000\,\mps$.
-The state of the art RV precision reached $1\, \mps$ in 2016 \added{\citep{fischer2016}}.
+The state of the art RV precision reached $1\, \mps$ by 2016 \citep{fischer2016}.
 The newest generation of instruments aim  to reach $0.1\,\mps$ precision, the required precision to detect terrestrial worlds.
@@
 This is particularly true of \eprv\ instruments, which are equipped with stringent environmental stabilizing.
-\replaced{ likely reducing the number and extent of the axes along which the instrument can vary.}{The thermomechanical stability of these instruments reduces the variations they experience to something that can be represented by a low-dimensional framework.}  That is, spectrographs, especially stabilized ones, should have few environmentally accessible degrees of freedom.
+The thermomechanical stability of these instruments reduces the variations they experience to something that can be represented by a low-dimensional framework.
+That is, spectrographs, especially stabilized ones, should have few environmentally accessible degrees of freedom.
 This renders it inadvisable to fit each calibration exposure or calibrate each science exposure independently.
@@
 There are many ways to discern the basis functions.
-In this paper, we present one implementation of \name\ using principal component analysis (PCA) \added{\citep{pearson1901, jolliffe2016}}.
+In this paper, we present one implementation of \name\ using principal component analysis (PCA) \citep{pearson1901, jolliffe2016}.
 A PCA is justifiable in the limit where exposures have very high signal-to-nose ratio, as is usually the case with typical calibration images.
@@
 
-\replaced{That is, \name\ assumes that this has been done correctly.
-We caution, though, that systematic errors or large uncertainties in fitting line positions can propagate through to errors or biases in the wavelength models returned by \name.}{\Name\ assumes that line positions have been identified ``correctly,'' as in the position of a calibration line is determined the same way as the position of a stellar line when extracting RVs.
+\Name\ assumes that line positions have been identified ``correctly,'' as in the position of a calibration line is determined the same way as the position of a stellar line when extracting RVs.
 This also inherently assumes that the calibration lines are not subject to an effect that the science exposures are not, for example differences in charge transfer inefficiency, non-linearities in the PSF, etc.
 We caution that systemic errors or large uncertainties in fitting line positions easily propagate through to biases in the wavelength models returned by \name.
-For more discussion, see Section \ref{sec:discussion}.}
+For more discussion, see Section \ref{sec:discussion}.
 
@@
 	$g_0(\lambda,m) = \overline{x(\lambda,m,n)}$\;
-	\added{using Singular-Value Decomposition,} find $U, \Sigma, V$ s.t. $U\Sigma V^* = (x(\lambda,m,n)-g_0(\lambda,m))$\;
+	using Singular-Value Decomposition, find $U, \Sigma, V$ s.t. $U\Sigma V^* = (x(\lambda,m,n)-g_0(\lambda,m))$\;
 	let $a_{n,k} = U\cdot \Sigma$ and $g_k(\lambda,m) = V$\;
@@
 This is discussed more in Section \ref{sec:choice_avt}.
-\added{Let a prime denote values related to a science exposure $n'$ for which we want wavelengths.}  We use linearly interpolated magnitudes, $a_{n',k}$ at time $t_{n'}$ to construct the calibration state of the spectrograph for that point in time.
+Let a prime denote values related to a science exposure $n'$ for which we want wavelengths.
+We use linearly interpolated magnitudes, $a_{n',k}$ at time $t_{n'}$ to construct the calibration state of the spectrograph for that point in time.
 Using interpolated amplitudes, $a_{n',k}$, and the basis vectors, $g_k(\lambda,m)$, returned by the de-noising process, a new set of calibration lines, $x'(\lambda,m,n')$ can be constructed for any exposure as formulated in Equation \ref{eq:excl_wsol}.
@@
 Choices in interpolation scheme, $K$, and other tests are further discussed in Section \ref{sec:choice_wvp}.
-\added{Our implementation of \name\ is hosted on GitHub\footnote{\url{https://www.github.com/davidwhogg/EPRVCalibration}}.}
 
@@
 \KwData{the fiducial calibration of the spectrograph $g_0(\lambda,m)$; magnitudes of the principal components for each exposure $a_{n,k}$; basis vectors spanning the calibration space of the spectrograph $g_k(\lambda,m)$; }
-\KwResult{Wavelengths for detector positions $x'(m,n')$ of exposure $n'$ with time $t_{n'}$, \added{where the prime is used to denote values relevant to this new exposure}}
+\KwResult{Wavelengths for detector positions $x'(m,n')$ of exposure $n'$ with time $t_{n'}$, where the prime is used to denote values relevant to this new exposure}
 
@@
 \end{algorithm}
+
+The implementation of \name\ described here is hosted on GitHub\footnote{\url{https://www.github.com/lilyling27/excalibur}}.
 
@@
 
-In order to \added{comfortably} satisfy the assumption that there exists only low-order variation, which is needed for \name, \replaced{we used exposures from after the photonic crystal fiber within the LFC was changed in Summer 2019, and the LFC stabilized}{we used exposures from after the LFC stabilized following servicing in summer 2019, where the photonic crystal fiber was replaced and the polarization was changed to shift the wavelength range of the LFC redwards}.
+In order to comfortably satisfy the assumption that there exists only low-order variation, which is needed for \name, we used exposures from after the LFC stabilized following servicing in summer 2019, where the photonic crystal fiber was replaced and the polarization was changed to shift the wavelength range of the LFC redwards.
 In the results presented here, we use 1227 LFC exposures and 78 ThAr exposures taken between October 14 and December 18, 2019 on 29 unique nights.
@@
 
-\begin{figure}[t]
-\centering
-\includegraphics[width=.45\textwidth]{Figures/all_results.pdf}
-\caption{Difference in predicted and theoretical wavelengths for the wavelength calibration tests described in Section \ref{sec:tests}.
-The per line RMS as defined in Equation \ref{eq:rms} is given in the top-right corner in each method's corresponding color.
-Incorporating denoising returns the smallest spread in residuals.}
-\label{fig:testHists}
-\end{figure}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.455\textwidth]{Figures/lineResids2D_col.pdf}
-\caption{Residuals of a single LFC exposure plotted with respect to detector position (as defined by echelle order and x-pixel) for parametric (top) and non-parametric (bottom) wavelength calibration methods.
-Each line is colored by the difference between the predicted wavelength and the theoretical wavelength for each line, given in units of $\mps$.
-High-order structure, i.e.
-vertical stripes and patchiness, is apparent in the residuals to a polynomial wavelength solution, which assumes smoothness.}
-\label{fig:resid2d}
-\end{figure}
-
 To test the denoising step of \name\ (\textsection \ref{sec:denoising}, \textsection\ref{sec:interp_time}), we employed \name\ on a randomly selected 90\% of all LFC exposures.
@@
 
-\added{The polynomial and interpolation tests remove the same 50\% of lines from each exposure while the denoising test completely removes a randomly selected 10\% of calibration exposures and their associated line position measurements.
+The polynomial and interpolation tests remove the same 50\% of lines from each exposure while the denoising test completely removes a randomly selected 10\% of calibration exposures and their associated line position measurements.
 Errors from interpolation will be localized, extending only to neighboring lines.
@@
 Here, it is sufficient to completely remove 10\% of exposures, a traditional training/validation fraction.
-Since the information being removed varies between each test depending on its focus, we present the results per line, treating each line as an independent test.}
+Since the information being removed varies between each test depending on its focus, we present the results per line, treating each line as an independent test.
 
@@
 
-\added{The per-line residuals from the denoising test also exhibit smaller spread than interpolation alone.
+\begin{figure}[b]
+\centering
+\includegraphics[width=.45\textwidth]{Figures/all_results.pdf}
+\caption{Difference in predicted and theoretical wavelengths for the wavelength calibration tests described in Section \ref{sec:tests}.
+The per line RMS as defined in Equation \ref{eq:rms} is given in the top-right corner in each method's corresponding color.
+Incorporating denoising returns the smallest spread in residuals.}
+\label{fig:testHists}
+\end{figure}
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=\textwidth]{Figures/lineResids2D.pdf}
+\caption{Residuals of a single LFC exposure plotted with respect to detector position (as defined by echelle order and x-pixel) for parametric (top) and non-parametric (bottom) wavelength calibration methods.
+Each line is colored by the difference between the predicted wavelength and the theoretical wavelength for each line, given in units of $\mps$.
+High-order structure, i.e.
+vertical stripes and patchiness, is apparent in the residuals to a polynomial wavelength solution, which assumes smoothness.}
+\label{fig:resid2d}
+\end{figure*}
+
+The per-line residuals from the denoising test also exhibit smaller spread than interpolation alone.
 This suggests that the spectrograph truly is accurately represented by a low-dimensional model.
@@
 The low-dimensional model does not incorporate noise from individual line measurements.
-Returning more precise, denoised line positions results in smaller per-line residuals.}
+Returning more precise, denoised line positions results in smaller per-line residuals.
 
@@
 For a randomly selected example LFC exposure, Figure \ref{fig:resid2d} plots each line with respect to its echelle order (y-axis) and x-pixel on the detector (x-axis) colored by the difference between the predicted and theoretical wavelength for that line in units of $\mps$.
-
-\begin{figure*}[t]
-\centering
-\includegraphics[width=\textwidth]{Figures/lineDensity.pdf}
-\caption{Residuals when using ThAr lines to predict wavelengths for LFC lines.
-\textbf{Left}: residuals for a single LFC exposure plotted with respect to detector position and colored by residual, as in Figure \ref{fig:resid2d}.
-The positions of ThAr lines are over-plotted in yellow.
-In general, residuals are greater between ThAr lines with greater separation.
-\textbf{Right}: Comparison of polynomial and interpolated wavelength solutions using either just ThAr lines or LFC lines for a subset of echelle order 94.
-The shape of the residuals from a polynomial fit are similar whether using ThAr lines or LFC lines.
-A PCHIP interpolated wavelength model guided by LFC lines returns the smallest residuals.}
-\label{fig:waveResids}
-\end{figure*}
 
@@
 All the same, the residuals are in general worse where lines are further apart (for example, in redder echelle orders) than where lines are denser.
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=\textwidth]{Figures/lineDensity.pdf}
+\caption{Residuals when using ThAr lines to predict wavelengths for LFC lines.
+\textbf{Left}: residuals for a single LFC exposure plotted with respect to detector position and colored by residual, as in Figure \ref{fig:resid2d}.
+The positions of ThAr lines are over-plotted in yellow.
+In general, residuals are greater between ThAr lines with greater separation.
+\textbf{Right}: Comparison of polynomial and interpolated wavelength solutions using either just ThAr lines or LFC lines for a subset of echelle order 94.
+The shape of the residuals from a polynomial fit are similar whether using ThAr lines or LFC lines.
+A PCHIP interpolated wavelength model guided by LFC lines returns the smallest residuals.}
+\label{fig:waveResids}
+\end{figure*}
 
@@
 It is $4.8\, Gyr$ old and relatively quiet ($log R'_{HK} = -5.09$) \citep{brewer2016}.
+Because \target\ has no known planets and should have a smaller contribution from stellar signals, it is a good star to test the effects of the wavelength calibration on the RMS of the returned RVs.
 We use 114 observations of \target\ taken between October 8, 2019 and March 5, 2020 with SNR 250.
@@
 This is equivalent to removing an independent, additive noise component of $0.52\, \mps \,(=\sqrt{1.17^2-1.05^2})$.
-\deleted{A reduction in RMS was observed universally in all data sets tested; other data sets showed a reduction in RMS between $0.2-0.5\, \mps$}.
+
+We conducted a direct test of a classically-generated wavelength solution with \name -wavelengths on four other data sets.
+All targets showed a reduction in or comparable RV RMS.
+The results from these data sets can not be interpreted as directly as with \target, though, due to larger contributions from stellar variability, known planets, etc.
+As completely mitigating these different effects is out of scope for this paper, we focus here on the results with \target.
 
@@
  18771.4179  &-1.927  & 0.342 \\
+ 18771.4196 & 2.688 &	0.357 \\
           & \(\vdots\)  & \\
@@
 For the dimensionality reduction and denoising, the PCA could be replaced by a probabilistic PCA model or other probabilistic linear reduction methods, such as heteroscedastic matrix factorization (HMF)\citep{tsalmantza2012}.
-It is also possible to move to non-linear reduction methods, like an auto-encoder or a normalizing flow \citep[e.g.][]{kramer1991,woodbridge2020}.
-\added{Careful choice in the denoising model could enable \name\ to capture large-scale changes as well as small variations in calibration state.
-This would bridge data broken up by significant format changes (e.g.
-replacing hardware, opening the spectrograph chamber, etc.).}
+It is also possible to move to non-linear reduction methods, like a Guassian process latent-variable model, an auto-encoder, a normalizing flow \citep[e.g.][]{kramer1991,woodbridge2020}.
+Using a non-linear denoising model could enable \name\ to capture large-scale changes as well as small variations in calibration state.
 
@@
 The choice in interpolation method can help guide how many wavelength calibration images are truly needed.
-It is unnecessary to take calibration images at times where the same information can be \added{reconstructed at the desired precision} by a well-chosen interpolation scheme.
+It is unnecessary to take calibration images at times where the same information can be reconstructed at the desired precision by a well-chosen interpolation scheme.
 For example, with the \expres\ data shown here, it is clear that nightly calibration images are needed, but for a linear trend, only two calibration images throughout the night are strictly required.
 
-We also tested an implementation of \name\ where the $K$ principal components within a night were fit to a cubic \added{with respect to time rather than linearly interpolated}.
+We also tested an implementation of \name\ where the $K$ principal components within a night were fit to a cubic with respect to time rather than linearly interpolated.
 This emulates the current, polynomial-based wavelength solution implemented in the \expres\ pipeline, where polynomial fits to calibration files are interpolated to science exposure by fitting polynomial coefficients with respect to time to a cubic.
-We found that \replaced{this method}{using a cubic in place of linear interpolation} returns comparable RV RMS for most targets, though appears to do better when a night has sparse calibration data.
-\added{This suggests that the nightly behavior of \expres\ with respect to time is well described by a cubic function, but LFC exposures are typically taken with enough frequency that a linear interpolation provides a good approximation (see Figure \ref{fig:nightlyVariation}}
+We found that using a cubic in place of linear interpolation returns comparable RV RMS for most targets, though appears to do better when a night has sparse calibration data.
+This suggests that the nightly behavior of \expres\ with respect to time is well described by a cubic function, but LFC exposures are typically taken with enough frequency that a linear interpolation provides a good approximation (see Figure \ref{fig:nightlyVariation})
 
@@
 Histograms of the residuals for each method for the complete order is shown on the right.
-\deleted{Here, the y-axis is zoomed in to show the shape of each histogram more clearly.
-}Note: there is a blended ThAr line at approximately pixel 5300, the right most ThAr lines plotted.}
+Note: there is a blended ThAr line at approximately pixel 5300, the right most ThAr lines plotted.}
 \label{fig:xinterp}
@@
 A simple linear interpolation would give erroneously low values everywhere.
-Due to the dispersion intrinsic to echelle spectrographs, the wavelength change between pixels grows greater with greater wavelengths, \added{even within an order}.
+Due to the dispersion intrinsic to echelle spectrographs, the wavelength change between pixels grows greater with greater wavelengths, even within an order.
 This means that the function of wavelength vs.
@@
 
-\added{Though less of an issue with LFC lines,} a more classic cubic spline interpolation can run into issues with arc lines, which are irregularly spaced or even blended.
+Though less of an issue with LFC lines, a more classic cubic spline interpolation can run into issues with arc lines, which are irregularly spaced or even blended.
 Close lines appearing at virtually the same pixel location but with different wavelengths could coerce a cubic spline into a very high slope.
-\added{This is demonstrated by the green line in Figure \ref{fig:xinterp}, which shows the results of interpolating between ThAr lines rather than LFC lines.}  A line blend appears at approximately pixel 5300, causing the spline to twist to nearly vertical to account for both points.
+This is demonstrated by the green line in Figure \ref{fig:xinterp}, which shows the results of interpolating between ThAr lines rather than LFC lines.
+A line blend appears at approximately pixel 5300, causing the spline to twist to nearly vertical to account for both points.
 This leads to huge deviations from the correct wavelengths around this line blend as the extraneously high slope of the spline is accounted for.
@@
 As an empirical assessment of the needed line density, we removed LFC lines from the \expres\ data and calculated the per-line RMS of the returned wavelengths for the removed lines.
-\added{These tests are similar to the interpolation test explained in Section \ref{sec:tests}, in that a fraction of lines are systematically removed from the analysis.
+These tests are similar to the interpolation test explained in Section \ref{sec:tests}, in that a fraction of lines are systematically removed from the analysis.
 An increasing fraction of lines are removed to simulate different line densities.
-For these line density tests, though, we are also implementing denoising unlike the pure interpolation test of Section \ref{sec:tests}.}
+For these line density tests, though, we are also implementing denoising unlike the pure interpolation test of Section \ref{sec:tests}.
+
+\begin{figure}[b]
+\centering
+\includegraphics[width=0.45\textwidth]{Figures/lfcDensityTest.pdf}
+\caption{Per-line RMS as a function of spacing between lines in units of resolution element.
+The average line spacing is calculated using the average distance between LFC lines across its wavelength range.}
+\label{fig:density}
+\end{figure}
 
@@
 The needed line density depends on the resolution of a given spectrograph, the needed precision of the returned wavelength solution, and the chosen interpolation scheme.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.45\textwidth]{Figures/lfcDensityTest.pdf}
-\caption{Per-line RMS as a function of spacing between lines in units of resolution element.
-The average line spacing is calculated using the average distance between LFC lines across its wavelength range.}
-\label{fig:density}
-\end{figure}
 
@@
 The result is likely to be even more precise, as this method incorporates more data.
-\added{This method, as with all analysis involving a simultaneous reference fiber, will work only as well as the reference fiber's ability to trace changes in the main science fiber.}
+This method, as with all analysis involving a simultaneous reference fiber, will work only as well as the reference fiber's ability to trace changes in the main science fiber.
 
@@
 
-\deleted{If the FSR varies very little, its value can in principle be captured by a PCA or similar method.
-However, to interpolate this information from calibration exposures to science exposures would necessitate a simultaneous reference or interpolation with respect to some other housekeeping data that contains information on the value of the FSR.
-If the FSR varies a lot, it becomes untenable to capture these changes with a PCA or similar method.
-Instead, the resultant basis vectors of the calibration space would have to be conditioned on the value of the FSR.
-This would require a non-trivial extension of \name.}
-
-\added{Incorporating the FSR as part of the \name\ model will require introducing a free parameter to capture changes in the FSR independent of variation in an instrument's calibration state.
+Incorporating the FSR as part of the \name\ model will require introducing a free parameter to capture changes in the FSR independent of variation in an instrument's calibration state.
 The calibration state can then be described with respect to mode number, which will be used to uniquely identify a calibration line across exposures rather than wavelength.
@@
 The FSR must not vary so much that the change in this mode-number-to-wavelength mapping becomes non-linear.
-This model would require a simultaneous reference or other housekeeping data that can be used to determine the FSR for every exposure.}
+This model would require a simultaneous reference or other housekeeping data that can be used to determine the FSR for every exposure.
 
@@
 Homogeneity is, of course, implicitly required when implementing PCA.
-\added{Different denoising models will be able to account for different amounts of stability or lack thereof.}
+Different denoising models will be able to account for different amounts of stability or lack thereof.
 
@@
 Those interested in interpreting the variability are encouraged to investigate how the amplitude of the different vectors varies with different housekeeping data to find their source.
-
-\deleted{The endeavor to reach $0.1\,\mps$ radial-velocity precision necessitates improvements on all levels of spectroscopy, from the instrumentation to the data analysis.
-Implementing \name, a hierarchical, non-parametric model for wavelength calibration, is a step towards reducing the error contribution from data reduction pipelines.}
 
@@
 We chose to implement \name\ using PCA here for simplicity and computational tractability.
+PCA is a good option because the instrument changes here are small enough that a linear model is an approrpiate representation of the changes.
 If \name\ was updated to a full probabilistic model, the PCA along with the interpolation model would have to be upgraded to something with better probabilistic properties.
-\added{Other denoising methods may be more robust to large changes, allowing all calibration images ever taken with an instrument to be used to construct the accessible calibration state regardless of hardware adjustments.}  Further discussion of other implementations of \name\ can be found in Section \ref{sec:choices}.
+Other, nonlinear denoising methods may be more robust to large changes, allowing all calibration images ever taken with an instrument to be used to construct the accessible calibration state regardless of hardware adjustments.
+Further discussion of other implementations of \name\ can be found in Section \ref{sec:choices}.
 
@@
 Furthermore, all exposures taken with the spectrograph in principle contains information about the calibration state of the spectrograph.
-Theoretically, tellurics, lines in science exposures, or just the trace positions themselves could also be used to determine an instrument's calibration state, \replaced{negating the need for any additional wavelength-calibration source beyond defining the calibration space.}{thereby providing free simultaneous calibration information.}
+Theoretically, tellurics, lines in science exposures, or just the trace positions themselves could also be used to determine an instrument's calibration state, thereby providing free simultaneous calibration information.
 
@@
 \acknowledgements
+We  gratefully  acknowledge  the  anonymous  referee whose  insightful  and  well thought out  comments  significantly improved the paper.
 We thank the EXPRES team for their work on the instrument, software, and pipeline.
